### 2020-12-12 README.Rmd
add

---
output: github_document
---

# `peacesciencer`: Various Tools and Data for Quantitative Peace Science

```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
```

```{r, echo = FALSE, results='asis'}
cat(
	badge_cran_release("peacesciencer", "green"),
	badge_cran_download("peacesciencer", "grand-total", "green"),
	badge_cran_download("peacesciencer", "last-month", "green"),
	badge_cran_download("peacesciencer", "last-week", "green")
)
```

<img src="http://svmiller.com/images/peacesciencer-hexlogo.png" alt="peacesciencer  hexlogo" align="right" width="200" style="padding: 0 15px; float: right;"/>

`peacesciencer` is an R package including various functions and data sets to allow easier analyses in the field of quantitative peace science. The package is very much a work in progress. The goal is to provide an R package that reasonably approximates what made [EUGene](https://journals.sagepub.com/doi/abs/10.1177/0738894211413055) so attractive to scholars working in the field of quantitative peace science in the early 2000s. EUGene shined because it encouraged replications of conflict models while having the user also generate data from scratch. Likewise, this R package will offer tools to approximate what EUGene did within the R environment (i.e. not requiring Windows for installation).

### 2020-12-12 R/rd-cow_states.R

@@ -1,0 +1,25 @@
+#' @importFrom tibble tibble
+NULL
+
+#' Correlates of War State System Membership Data (1816-2016)
+#'
+#' These are the Correlates of War state system membership data.
+#'
+#' @format A data frame with 243 observations on the following 10 variables.
+#' \describe{
+#' \item{\code{stateabb}}{a character vector for the state abbrevation}
+#' \item{\code{ccode}}{a numeric vector for the Correlates of War country code}
+#' \item{\code{styear}}{the start year in the system}
+#' \item{\code{stmonth}}{the start month in the system}
+#' \item{\code{stday}}{the start day in the system}
+#' \item{\code{endyear}}{the end year in the system}
+#' \item{\code{endmonth}}{the end month in the system}
+#' \item{\code{endday}}{the end day in the system}
+#' \item{\code{version}}{a version identifier}
+#' }
+#'
+#' @details Data are provided "as-is" with no additional re-cleaning before inclusion into this data set.
+#'
+#' @references Correlates of War Project. 2017. "State System Membership List, v2016." Online, \url{https://correlatesofwar.org/data-sets/state-system-membership}
+#'
+"cow_states"

### 2020-12-12 R/rd-gw_states.R

@@ -1,0 +1,22 @@
+#' @importFrom tibble tibble
+NULL
+
+#' Gleditsch-Ward (Independent States) System Membership Data (1816-2017)
+#'
+#' These are the independent states in Gleditsch and Ward's data.
+#'
+#' @format A data frame with 216 observations on the following 5 variables.
+#' \describe{
+#' \item{\code{gwcode}}{a numeric vector for the Gleditsch-Ward country code}
+#' \item{\code{stateabb}}{a character vector for state abbreviation}
+#' \item{\code{statename}}{a character vector for the state name}
+#' \item{\code{startdate}}{the start date in the data}
+#' \item{\code{enddate}}{the end date in the data}
+#' }
+#'
+#' @details Data originally provided by Gleditsch with no column names. Column names were added before some light re-cleaning in order
+#' to generate these data.
+#'
+#' @references \url{http://ksgleditsch.com/data-4.html}
+#'
+"gw_states"

### 2020-12-12 README.Rmd
afdsf

@@
 `peacesciencer` is an R package including various functions and data sets to allow easier analyses in the field of quantitative peace science.
-The package is very much a work in progress.
 The goal is to provide an R package that reasonably approximates what made [EUGene](https://journals.sagepub.com/doi/abs/10.1177/0738894211413055) so attractive to scholars working in the field of quantitative peace science in the early 2000s.
@@
 not requiring Windows for installation).
+
+The package is very much a work in progress. Right now, it has the following functions:
+
+- `make_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
+
+It also has the following data sets:
+
+- `cow_states`: Correlates of War state system membership data (version: 2016)
+- `gw_states`: Gleditsch-Ward independent state system data (version: 2017)

### 2020-12-13 R/create_dyadyears.R
add proto version of dyad-year function

@@ -1,0 +1,125 @@
+#' Create dyad-years from state system membership data
+#'
+#' @description \code{create_dyadyears()} allows you to dyad-year data from
+#' either the Correlates of War (\code{CoW}) state system membership data or the
+#' Gleditsch-Ward (\code{gw}) system membership data. The function leans on internal
+#' data provided in the package.
+#'
+#' @return \code{create_dyadyears()} takes state system membership data provided
+#' by either Correlates of War or Gleditsch-Ward and returns a dyad-year data frame.
+#'
+#' @author Steven V. Miller
+#'
+#' @references Miller, Steven V. 2019. ``Create Country-Year and (Non)-Directed Dyad-Year Data With Just a Few Lines in R''
+#' \url{http://svmiller.com/blog/2019/01/create-country-year-dyad-year-from-country-data/}
+#'
+#' @param system a character specifying whether the user wants Correlates of War
+#' state-years ("cow") or Gleditsch-Ward ("gw") state-years. Correlates of War is the
+#' default.
+#' @param mry optional, defaults to TRUE. If TRUE, the function extends the script
+#' beyond the most recent system membership updates to include observation to the
+#' most recently concluded calendar year. For example, the Gleditsch-Ward data extend
+#' to the end of 2017. When \code{mry == TRUE}, the function returns more recent years
+#' (e.g. 2018, 2019) under the assumption that states alive at the end of 2017 are still alive
+#' today. Use with some care.
+#' @param directed optional, defaults to TRUE. If TRUE, the function returns so-called
+#' "directed" dyad-year data. In directed dyad-year data, France-Germany (220-255) and
+#' Germany-France (255-220) are observationally different. If FALSE, the function returns
+#' non-directed data. In non-directed data, France-Germany and Germany-France in the same year
+#' are the same observation. The standard here is to drop cases where the country code for the
+#' second observation is less than the country code for the first observation.
+#'
+#' @examples
+#' # CoW is default, will include years beyond 2016 (most recent CoW update)
+#' create_dyadyears()
+#'
+#' # Gleditsch-Ward, include most recent years
+#' create_dyadyears(system="gw")
+#'
+#' # Gleditsch-Ward, don't include most recent years
+#' create_stateyears(system="gw", mry=FALSE)
[…]

### 2020-12-13 R/create_stateyears.R

@@ -6,7 +6,8 @@
 #' data provided in the package.
 #'
 #' @return \code{create_stateyears()} takes state system membership data provided
-#' by either Correlates of War or Gleditsch-Ward and returns a simple coun
+#' by either Correlates of War or Gleditsch-Ward and returns a simple state-year
+#' data frame.
 #'
 #' @author Steven V. Miller
 #'
@@ -35,6 +36,11 @@
 #'
 #'
 create_stateyears <- function(system = "cow", mry = TRUE) {
+  require(dplyr)
+  require(magrittr)
+  require(dplyr)
+  require(tidyr)
+  require(lubridate)
   if (system == "cow") {
     if (mry == TRUE) {
       mry <- as.numeric(format(Sys.Date(), "%Y"))-1
@@ -53,7 +59,7 @@ create_stateyears <- function(system = "cow", mry = TRUE) {
   } else if(system == "gw") {
     if (mry == TRUE) {
       mry <- as.numeric(format(Sys.Date(), "%Y"))-1
-      gw_states$endyear = ifelse(year(gw_states$enddate) == max(year(gw_states$enddate)), 2019, year(gw_states$enddate))
+      gw_states$endyear = ifelse(year(gw_states$enddate) == max(year(gw_states$enddate)), mry, year(gw_states$enddate))
     } else {
       gw_states$endyear <- year(gw_states$enddate)
     }

### 2020-12-13 README.Rmd
add proto version of dyad-year function
this and that, these and those

@@
 
-- `make_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
+- `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
+- `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format.
 

### 2020-12-13 README.Rmd
new features

@@
 
+- `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year data.
 - `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
@@
 
+- `capitals`: a list of capitals and capital transition for Correlates of War state system members
+- `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members
 - `cow_states`: Correlates of War state system membership data (version: 2016)

### 2020-12-13 R/add_capital_distance.R

@@ -1,0 +1,56 @@
+#' Add capital-to-capital distance to a dyad-year data frame
+#'
+#' @description \code{add_capital_distance()} allows you to add capital-to-capital
+#' distance to a dyad-year data frame. The capitals are coded in the \code{capitals}
+#' data frame, along with their latitudes and longitudes. The distance variable that
+#' emerges \code{capdist} is calculated using the "Vincenty" method (i.e. "as the crow
+#' flies") and is expressed in kilometers.
+#'
+#' @return \code{add_capital_distance()} takes a dyad-year data frame and adds the
+#' capital-to-capital distance between the first state and the second state. A minor
+#' note about this function: cases of capital transition are recorded in the
+#' \code{capitals} data but, in the conversion to capital-years (and eventual
+#' merging into a dyad-year data frame), the Jan. 1 capital is used for calculating
+#' distances.
+#'
+#' @author Steven V. Miller
+#'
+#' @param data a dyad-year data frame (either "directed" or "non-directed")
+#'
+#' @examples
+#'
+#' library(magrittr)
+#' cow_ddy %>% add_capital_distance()
+#'
+#'
+
+add_capital_distance <- function(data) {
+  require(dplyr)
+  require(magrittr)
+  require(dplyr)
+  require(tidyr)
+  require(lubridate)
+  require(geosphere)
+  capitals %>% rowwise() %>%
+    mutate(year = list(seq(styear, endyear))) %>%
+    unnest(year) %>%
+    select(ccode, year, lat, lng) %>%
+    # There will be duplicates for when the country moved.
+    # Under those conditions, the first capital should be first. It's basically the Jan. 1 capital, if you will.
+    group_by(ccode, year) %>% slice(1) %>% ungroup() -> capital_years
+
+  data %>%
+    left_join(., capital_years %>% select(ccode, year, lat,lng), by=c("ccode1"="ccode","year"="year")) %>%
+    rename(lat1 = lat,
+           lng1 = lng) %>%
+    left_join(., capital_years %>% select(ccode, year, lat,lng), by=c("ccode2"="ccode","year"="year")) %>%
+    rename(lat2 = lat,
+           lng2 = lng) -> data
+
+  latlng1 <- data %>% select(lng1, lat1)
+  latlng2 <- data %>% select(lng2, lat2)
+  data$capdist <- distVincentySphere(latlng1, latlng2) / 1000
+  data %>% select(-lat1, -lng1, -lat2, -lng2) -> data
+  return(data)
+
+}

### 2020-12-13 data-raw/cow_ddy.R

@@ -1,0 +1,3 @@
+cow_ddy <- create_dyadyears()
+
+save(cow_ddy, file= "data/cow_ddy.rda")

### 2020-12-13 README.Rmd
change README

--- 2020-12-13 README.Rmd
@@
 
+- `capitals`: a list of capitals and capital transition for Correlates of War state system members
+- `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members
+- `cow_majors`: Correlates of War major powers data (version: 2016)
 - `cow_states`: Correlates of War state system membership data (version: 2016)

### 2020-12-13 README.Rmd
bah

--- 2020-12-13 README.Rmd
@@
 
+- `capitals`: a list of capitals and capital transitions for Correlates of War state system members
+- `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members
+- `cow_majors`: Correlates of War major powers data (version: 2016)
 - `cow_states`: Correlates of War state system membership data (version: 2016)

### 2020-12-13 README.Rmd
add Maoz global/regional power data.

--- 2020-12-13 README.Rmd
@@
 
-- `capitals`: a list of capitals and capital transition for Correlates of War state system members
+- `capitals`: a list of capitals and capital transitions for Correlates of War state system members
 - `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members
+- `cow_majors`: Correlates of War major powers data (version: 2016)
 - `cow_states`: Correlates of War state system membership data (version: 2016)
 - `gw_states`: Gleditsch-Ward independent state system data (version: 2017)
+- `maoz_powers`: Zeev Maoz' global/regional power data.

### 2020-12-13 README.Rmd
add

--- 2020-12-13 README.Rmd
@@
 
-- `capitals`: a list of capitals and capital transition for Correlates of War state system members
+- `capitals`: a list of capitals and capital transitions for Correlates of War state system members
+- `cow_contdir`: Correlates of War Direct Contiguity Data (v. 3.2)
 - `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members
@@
 - `gw_states`: Gleditsch-Ward independent state system data (version: 2017)
+- `maoz_powers`: Zeev Maoz' global/regional power data.

### 2020-12-13 data-raw/cow_contdir.R

@@ -1,0 +1,21 @@
+library(tidyverse)
+
+haven::read_dta("~/Dropbox/data/cow/contiguity/3.2/contdir.dta") %>%
+  select(statelno, statehno, conttype, begin, end) %>%
+  rename(ccode1 = statelno,
+         ccode2 = statehno) -> cow_contdir
+
+cow_contdir %>%
+  rename(ccode1 = ccode2,
+         ccode2 = ccode1) %>%
+  bind_rows(cow_contdir, .) -> cow_contdir
+
+save(cow_contdir, file="data/cow_contdir.rda")
+
+
+# cow_contdir %>%
+#   mutate(styear = as.numeric(str_sub(begin, 1, 4)),
+#          endyear = as.numeric(str_sub(end, 1, 4))) %>%
+#   rowwise() %>%
+#   mutate(year = list(seq(styear, endyear))) %>%
+#   unnest(year)

### 2020-12-13 R/add_cow_majors.R
add

@@ -1,0 +1,45 @@
+#' Add Correlates of War major power information to a dyad-year data frame
+#'
+#' @description \code{add_cow_majors()} allows you to add Correlates of War major power variables
+#' to a dyad-year data frame.
+#'
+#' @return \code{add_cow_majors()} takes a dyad-year data frame and adds two columns for whether
+#' the first state (i.e. \code{ccode1}) or the second state (i.e. \code{ccode2}) are major powers
+#' in the given year, according to the Correlates of War. 1 = is a major power. 0 = is not a major
+#' power.
+#'
+#' @author Steven V. Miller
+#'
+#' @param data a dyad-year data frame (either "directed" or "non-directed")
+#'
+#' @references Correlates of War Project. 2017. "State System Membership List, v2016." Online, \url{https://correlatesofwar.org/data-sets/state-system-membership}
+#'
+#' @examples
+#'
+#' library(magrittr)
+#' cow_ddy %>% add_cow_majors()
+#'
+#'
+
+add_cow_majors <- function(data) {
+  require(dplyr)
+  require(magrittr)
+  require(dplyr)
+  require(tidyr)
+  require(stringr)
+
+  cow_majors %>%
+    select(ccode, styear, endyear) %>%
+    rowwise() %>%
+    mutate(year = list(seq(styear, endyear))) %>% unnest(year) %>%
+    select(-styear, -endyear) %>%
+    mutate(cowmaj = 1) -> major_years
+
+  cow_ddy %>% left_join(., major_years, by=c("ccode1"="ccode","year"="year")) %>%
+    rename(cowmaj1 = cowmaj) %>%
+    left_join(., major_years, by=c("ccode2"="ccode","year"="year")) %>%
+    rename(cowmaj2 = cowmaj) %>%
+    mutate_at(vars("cowmaj1", "cowmaj2"), ~ifelse(is.na(.), 0, .)) -> data
+
+  return(data)
+}

### 2020-12-13 R/add_cow_majors.R
add

@@ -3,6 +3,7 @@
 #' @description \code{add_cow_majors()} allows you to add Correlates of War major power variables
 #' to a dyad-year data frame.
 #'
+#'
 #' @return \code{add_cow_majors()} takes a dyad-year data frame and adds two columns for whether
 #' the first state (i.e. \code{ccode1}) or the second state (i.e. \code{ccode2}) are major powers
 #' in the given year, according to the Correlates of War. 1 = is a major power. 0 = is not a major

### 2020-12-14 README.Rmd
add
change README, add some more state-year equivalents.

@@
 
-- `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year data.
+- `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year or state-year data.
+- `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year data.
+- `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data.
 - `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
 - `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format.
+- `filter_prd()`: filters dyad-year data frame to just "politically relevant" dyads.
 

### 2020-12-14 R/add_capital_distance.R

@@ -1,26 +1,33 @@
-#' Add capital-to-capital distance to a dyad-year data frame
+#' Add capital-to-capital distance to a dyad-year or state-year data frame
 #'
 #' @description \code{add_capital_distance()} allows you to add capital-to-capital
-#' distance to a dyad-year data frame. The capitals are coded in the \code{capitals}
+#' distance to a dyad-year or state-year data frame. The capitals are coded in the \code{capitals}
 #' data frame, along with their latitudes and longitudes. The distance variable that
 #' emerges \code{capdist} is calculated using the "Vincenty" method (i.e. "as the crow
 #' flies") and is expressed in kilometers.
 #'
-#' @return \code{add_capital_distance()} takes a dyad-year data frame and adds the
-#' capital-to-capital distance between the first state and the second state. A minor
+#' @return \code{add_capital_distance()} takes a dyad-year or state-year data frame and adds the
+#' capital-to-capital distance between the first state and the second state (in dyad-year data) or the minimum
+#' capital-to-capital distance for a given state in a given year. A minor
 #' note about this function: cases of capital transition are recorded in the
 #' \code{capitals} data but, in the conversion to capital-years (and eventual
 #' merging into a dyad-year data frame), the Jan. 1 capital is used for calculating
 #' distances.
 #'
+#' @details The function leans on attributes of the data that are provided by the \code{create_dyadyear()} or
+#' \code{create_stateyear()} function. Make sure that function (or data created by that function) appear at the top
+#' of the proverbial pipe.
+#'
 #' @author Steven V. Miller
 #'
-#' @param data a dyad-year data frame (either "directed" or "non-directed")
+#' @param data a dyad-year data frame (either "directed" or "non-directed") or state-year data frame
 #'
 #' @examples
 #'
 #' library(magrittr)
 #' cow_ddy %>% add_capital_distance()
+#'
+#' create_stateyears() %>% add_capital_distance()
 #'
 #'
 
@@ -39,18 +46,46 @@ add_capital_distance <- function(data) {
     # Under those conditions, the first capital should be first. It's basically the Jan. 1 capital, if you will.
[…]

### 2020-12-14 R/add_cow_majors.R

@@ -56,8 +56,7 @@ add_cow_majors <- function(data) {
       mutate(cowmaj = ifelse(is.na(cowmaj), 0, 1)) -> data
 
   } else  {
-      stop("add_cow_majors() requires a data/tibble with attributes$ps_data_type of state_year or dyad_year. Try running
-           create_dyadyears() or create_stateyears() at the start of the pipe.")
+      stop("add_cow_majors() requires a data/tibble with attributes$ps_data_type of state_year or dyad_year. Try running create_dyadyears() or create_stateyears() at the start of the pipe.")
     }
 
   return(data)

### 2020-12-14 R/filter_prd.R

@@ -41,6 +41,8 @@ filter_prd <- function(data) {
   require(tidyr)
   require(stringr)
 
+  if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
+
   data %>%
     mutate(prd = case_when(
       conttype <= 5 ~ 1,
@@ -49,5 +51,13 @@ filter_prd <- function(data) {
       TRUE ~ 0
     )) %>%
     filter(prd == 1) -> data
+
+  } else  if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "state_year") {
+
+    stop("filter_prd() only makes sense in the context of dyad-year data.")
+
+  } else {
+    stop("{peacesciencer} functions require attributes(data)$ps_data_type (either dyad-year or state-year for a given function). Try running create_dyadyears() or create_stateyears() at the start of the pipe.")
+  }
   return(data)
 }

### 2020-12-14 R/add_contiguity.R
allow for contiguity to be state-year.

@@ -1,13 +1,16 @@
-#' Add Correlates of War direct contiguity information to a dyad-year data frame
+#' Add Correlates of War direct contiguity information to a dyad-year or state-year data frame
 #'
 #' @description \code{add_contiguity()} allows you to add Correlates of War contiguity
-#' data to a dyad-year data frame.
+#' data to a dyad-year or state-year data frame.
 #'
 #' @return \code{add_contiguity()} takes a dyad-year data frame and adds information
 #' about the contiguity relationship based on the "master records" for the
-#' Correlates of War direct contiguity data (v. 3.2).
+#' Correlates of War direct contiguity data (v. 3.2). If the data are dyad-year, the function
+#' returns the lowest contiguity type observed in the dyad-year (if contiguity is observed at all).
+#' If the data are state-year, the data return the total number of land and sea borders calculated
+#' from these master records.
 #'
-#' @details The contiguity codes range from 1 to 6. 1 = direct land contiguity. 2 =
+#' @details The contiguity codes in the dyad-year data range from 1 to 6. 1 = direct land contiguity. 2 =
 #' separated by 12 miles of water or fewer (a la Stannis Baratheon). 3 = separated by
 #' 24 miles of water or fewer (but more than 12 miles). 4 = separated by 150 miles
 #' of water or fewer (but more than 24 miles). 5 = separated by 400 miles of water
@@ -21,7 +24,7 @@
 #'
 #' @author Steven V. Miller
 #'
-#' @param data a dyad-year data frame (either "directed" or "non-directed")
+#' @param data a dyad-year data frame (either "directed" or "non-directed") or a state-year data frame
 #'
 #' @references Stinnett, Douglas M., Jaroslav Tir, Philip Schafer, Paul F. Diehl, and Charles Gochman
 #' (2002). "The Correlates of War Project Direct Contiguity Data, Version 3." Conflict
@@ -32,6 +35,8 @@
 #' library(magrittr)
 #' cow_ddy %>% add_contiguity()
 #'
+#' create_stateyear() %>% add_contiguity()
+#'
 #'
 add_contiguity <- function(data) {
   require(dplyr)
@@ -40,6 +45,7 @@ add_contiguity <- function(data) {
   require(tidyr)
   require(stringr)
 
[…]

### 2020-12-14 README.Rmd
allow for contiguity to be state-year.
democracy stuff

@@
 - `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year or state-year data.
-- `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year data.
+- `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year or state-year data.
 - `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data.
+- `add_democracy()`: adds estimates of democracy/levels of democracy to dyad-year or state-year data.
 - `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
@@
 - `capitals`: a list of capitals and capital transitions for Correlates of War state system members
+- `ccode_democracy`: Correlates of War state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores").
 - `cow_contdir`: Correlates of War Direct Contiguity Data (v. 3.2)

### 2020-12-14 R/add_democracy.R

@@ -1,0 +1,70 @@
+#' Add democracy information to dyad-year or state-year data.
+#'
+#' @description \code{add_democracy()} allows you to add estimates of democracy to either dyad-year or state-year data.
+#'
+#'
+#' @return \code{add_democracy()} takes a dyad-year data frame or state-year data frame and adds information
+#' about the level of democracy for the state or two states in the dyad in a given year. If the data are dyad-year, the function
+#' adds six total columns for the first state (i.e. \code{ccode1}) and the second state (i.e. \code{ccode2}) about the level
+#' of democracy measured by the Varieties of Democracy project (\code{v2x_polyarchy}), the Polity project (\code{polity2}), and
+#' Xavier Marquez' \code{QuickUDS} extensions/estimates. If the data are state-year, the function returns three additional columns
+#' to the original data that contain that same information for a given state in a given year.
+#'
+#' @details The function leans on attributes of the data that are provided by the \code{create_dyadyear()} or
+#' \code{create_stateyear()} function. Make sure that function (or data created by that function) appear at the top
+#' of the proverbial pipe.
+#'
+#' @author Steven V. Miller
+#'
+#' @param data a dyad-year data frame (either "directed" or "non-directed") or a state-year data frame.
+#'
+#' @references
+#'
+#' Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg, Jan Teorell, David Altman, Michael Bernhard, M. Steven Fish, Adam Glynn, Allen Hicken, Anna Luhrmann, Kyle L. Marquardt, Kelly McMann, Pamela Paxton, Daniel Pemstein, Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Agnes Cornell, Lisa Gastaldi, Haakon Gjerløw, Valeriya Mechkova, Johannes von Römer, Aksel Sundtröm, Eitan Tzelgov, Luca Uberti, Yi-ting Wang, Tore Wig, and Daniel Ziblatt. 2020. ”V-Dem Codebook v10” Varieties of Democracy (V-Dem) Project.
+#'
+#' Marshall, Monty G., Ted Robert Gurr, and Keith Jaggers. 2017. "Polity IV Project: Political Regime CHaracteristics and Transitions, 1800-2017." Center for Systemic Peace.
+#'
[…]

### 2020-12-14 R/rd-ccode_democracy.R

@@ -1,0 +1,32 @@
+#' @importFrom tibble tibble
+NULL
+
+#' Democracy data for all Correlates of War states
+#'
+#' These are democracy data for all Correlates of War state system members.
+#'
+#' @format A data frame with 16536 observations on the following 5 variables.
+#' \describe{
+#' \item{\code{ccode}}{the Correlates of War system code}
+#' \item{\code{year}}{a numeric vector for the year}
+#' \item{\code{v2x_polyarchy}}{the Varieties of Democracy "polyarchy" estimate}
+#' \item{\code{polity2}}{the the \code{polity2} score from the Polity project}
+#' \item{\code{xm_qudsest}}{an extension of the Unified Democracy Scores (UDS) estimates, made possibly by the \code{QuickUDS} package from Xavier Marquez.}
+#' }
+#'
+#' @details Missing data connote data that are unavailable for various reasons. Either there is no democracy data to code or, in the case of the Polity project, the state
+#' system member is outright not evaluated for the variable.
+#'
+#' The Polity data are from 2017. The Varieties of Democracy data are version 10. Xavier Marquez' \code{QuickUDS} estimates (i.e. extensions of Pemstein et al. (2010)) come from a package Marquez makes available on his Github (\url{https://github.com/xmarquez/QuickUDS}).
+#'
+#' @references
+#'
+#' Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg, Jan Teorell, David Altman, Michael Bernhard, M. Steven Fish, Adam Glynn, Allen Hicken, Anna Luhrmann, Kyle L. Marquardt, Kelly McMann, Pamela Paxton, Daniel Pemstein, Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Agnes Cornell, Lisa Gastaldi, Haakon Gjerløw, Valeriya Mechkova, Johannes von Römer, Aksel Sundtröm, Eitan Tzelgov, Luca Uberti, Yi-ting Wang, Tore Wig, and Daniel Ziblatt. 2020. ”V-Dem Codebook v10” Varieties of Democracy (V-Dem) Project.
+#'
+#' Marshall, Monty G., Ted Robert Gurr, and Keith Jaggers. 2017. "Polity IV Project: Political Regime CHaracteristics and Transitions, 1800-2017." Center for Systemic Peace.
+#'
[…]

### 2020-12-14 data-raw/add_something.R

@@ -1,0 +1,20 @@
+# This is a skeleton...
+
+add_something <- function(data) {
+  require(dplyr)
+  require(magrittr)
+  require(dplyr)
+  require(tidyr)
+  require(stringr)
+
+  if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
+
+
+  } else if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "state_year") {
+
+  } else  {
+    stop("add_something() requires a data/tibble with attributes$ps_data_type of state_year or dyad_year. Try running create_dyadyears() or create_stateyears() at the start of the pipe.")
+  }
+
+  return(data)
+}

### 2020-12-15 README.Rmd
add

@@
 - `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members
+- `cow_gw_years`: a yearly data frame including information about Correlates of War and Gleditsch-Ward states.
 - `cow_majors`: Correlates of War major powers data (version: 2016)

### 2020-12-15 README.Rmd
add

@@
 - `cow_states`: Correlates of War state system membership data (version: 2016)
+- `gw_ddy`: a full directed dyad-year data frame of Gleditsch-Ward state system members
 - `gw_states`: Gleditsch-Ward independent state system data (version: 2017)

### 2020-12-15 R/rd-gw_ddy.R

@@ -1,0 +1,18 @@
+#' @importFrom tibble tibble
+NULL
+
+#' A directed dyad-year data frame of Gleditsch-Ward state system members
+#'
+#' This is a complete directed dyad-year data frame of Gleditsch-Ward
+#' state system members. I offer it here as a shortcut for various other functions.
+#'
+#' @format A data frame with 1999558 observations on the following 4 variables.
+#' \describe{
+#' \item{\code{gwcode1}}{a numeric vector for the Correlates of War state code for the first state}
+#' \item{\code{gwcode2}}{a numeric vector for the Correlates of War state code for the second state}
+#' \item{\code{year}}{a numeric vector for the year}
+#' }
+#'
+#' @details Data are a quick generation from the \code{create_dyadyears(system="gw")} function in this package.
+#'
+"gw_ddy"

### 2020-12-15 data-raw/cow_ddy.R

@@ -1,3 +1,8 @@
 cow_ddy <- create_dyadyears()
 
 save(cow_ddy, file= "data/cow_ddy.rda")
+
+
+gw_ddy <- create_dyadyears(system="gw")
+
+save(gw_ddy, file= "data/gw_ddy.rda")

### 2020-12-15 README.Rmd
add

@@
 - `add_democracy()`: adds estimates of democracy/levels of democracy to dyad-year or state-year data.
+- `add_gwcode_to_cow():` adds Gleditsch-Ward state codes to dyad-year or state-year data with Correlates of War state codes.
 - `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
@@
 - `ccode_democracy`: Correlates of War state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores").
+- `cow_alliance`: directed dyad-year alliance formation data from the Correlates of War
 - `cow_contdir`: Correlates of War Direct Contiguity Data (v. 3.2)

### 2020-12-15 README.Rmd
add usage
add/fix/whatever

@@
 - `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year or state-year data.
+- `add_cow_alliance()`: adds Correlates of War major power information to dyad-year data.
 - `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data.
@@
 - `maoz_powers`: Zeev Maoz' global/regional power data.
+
+# Usage
+
+The workflow is going to look something like this.
+
+```{r, cache=F, message=F}
+
+library(tidyverse, quietly = TRUE)
+library(peacesciencer)
+library(tictoc)
+
+tic()
+create_dyadyears() %>%
+  # Add Gleditsch-Ward codes
+  add_gwcode_to_cow() %>%
+  # Add contiguity information
+  add_contiguity() %>%
+  # Add major power data
+  add_cow_majors() %>%
+  # Add democracy variables
+  add_democracy() %>%
+  # add alliance data from Correlates of War
+  add_cow_alliance() %>%
+  # you should probably filter to politically relevant dyads earlier than later...
+  # Or not, it's your time and computer processor...
+  filter_prd()
+toc()
+
+```

### 2020-12-15 R/add_cow_alliance.R

@@ -1,0 +1,47 @@
+#' Add Correlates of War alliance data to a dyad-year data frame
+#'
+#' @description \code{add_cow_alliance()} allows you to add Correlates of War alliance
+#' data to a dyad-year data frame
+#'
+#' @return \code{add_cow_alliance()} takes a dyad-year data frame and adds information about
+#' the alliance pledge in that given dyad-year. These include whether there was an alliance
+#' with a defense pledge, neutrality pledge, non-aggression pledge, or pledge for consultation
+#' in time of crisis (entente).
+#'
+#' @details Duplicates in the original directed dyad-year alliance data were pre-processed. Check
+#' \code{cow_alliance} for more information.
+#'
+#' @author Steven V. Miller
+#'
+#' @param data a dyad-year data frame (either "directed" or "non-directed")
+#'
+#' @references
+#'
+#' Gibler, Douglas M. 2009. \emph{International Military Alliances, 1648-2008}. Congressional Quarterly Press.
+#'
+#' @examples
+#'
+#' library(magrittr)
+#' cow_ddy %>% add_cow_alliance()
+#'
+
+add_cow_alliance <- function(data) {
+  require(dplyr)
+  require(magrittr)
+
+  if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
+    cow_alliance %>%
+      left_join(data, .) %>%
+      mutate_at(vars("defense", "neutrality", "nonaggression", "entente"), ~ifelse(is.na(.), 0, .)) -> data
+
+
+  } else if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "state_year") {
+
+    stop("Right now, there is only support for dyad-year data.")
+
+  } else  {
+    stop("add_something() requires a data/tibble with attributes$ps_data_type of dyad_year. Try running create_dyadyears() at the start of the pipe.")
+  }
+
+  return(data)
+}

### 2020-12-15 R/rd-cow_alliance.R

@@ -5,13 +5,11 @@ NULL
 #'
 #' These are version 4.1 of the Correlates of War directed dyad-year alliance data.
 #'
-#' @format A data frame with 148258 observations on the following 9 variables.
+#' @format A data frame with 120784 observations on the following 7 variables.
 #' \describe{
 #' \item{\code{ccode1}}{a numeric vector for the Correlates of War state code for the first state}
 #' \item{\code{ccode2}}{a numeric vector for the Correlates of War state code for the second state}
 #' \item{\code{year}}{a numeric vector for the year}
-#' \item{\code{left_censor}}{a numeric vector that equals 1 if the alliance started before Jan. 1, 1816}
-#' \item{\code{right_censor}}{a numeric vector that equals 1 if the alliance was still active after Dec. 31, 2012}
 #' \item{\code{defense}}{a numeric vector that equals 1 if the alliance included a defense pledge}
 #' \item{\code{neutrality}}{a numeric vector that equals 1 if the alliance included a neutrality pledge}
 #' \item{\code{nonaggression}}{a numeric vector that equals 1 if the alliance included a non-aggression pledge}
@@ -19,7 +17,11 @@ NULL
 #' }
 #'
 #' @details The directed dyad-year alliance data are for alliance initiations, not straight dyad-years, "per se." This suggests
-#' the presence of duplicate directed dyad-years. These are processed in various ways in \code{add_cow_alliances()}.
+#' the presence of duplicate directed dyad-years. For computing ease, given the intended use, I take care of these duplicate
+#' dyad-years behind the scenes. Consider the case of the U.S. and Canada in 1958. Therein, there were apparently two separate
+#' alliance initiations  that included defense pledges. My behind-the-scenes cleaning process groups by \code{ccode1},
+#' \code{ccode2}, and \code{year} and summarizes those alliance pledge variables. I then replace any value greater than 1
+#' with 1. This indicates the presence or absence of a defense pledge in a given directed dyad-year.
 #'
 #' @references
 #'

### 2020-12-15 data-raw/alliances.R

@@ -1,0 +1,17 @@
+library(tidyverse)
+
+read_csv("~/Dropbox/data/cow/alliance/4.1/alliance_v4.1_by_directed_yearly.csv") %>%
+  select(ccode1, ccode2, year, left_censor, right_censor, defense:entente) -> cow_alliance
+
+cow_alliance %>%
+  select(-left_censor, -right_censor) %>%
+  group_by(ccode1, ccode2, year) %>%
+  # There will be duplicates.
+  summarize_at(vars("defense","neutrality","nonaggression", "entente"), ~sum(.)) -> cow_alliance
+
+cow_alliance %>%
+  ungroup() %>%
+  mutate_at(vars("defense","neutrality","nonaggression", "entente"), ~ifelse(. > 1, 1, .)) -> cow_alliance
+
+save(cow_alliance, file="data/cow_alliance.rda")
+

### 2020-12-16 README.Rmd
update README. Still working on this.

--- 2020-12-15 README.Rmd
@@
 - `maoz_powers`: Zeev Maoz' global/regional power data.
+
+# Usage
+
+The workflow is going to look something like this.
+
+```{r, cache=F, message=F}
+
+library(tidyverse, quietly = TRUE)
+library(peacesciencer)
+library(tictoc)
+
+tic()
+create_dyadyears() %>%
+  # Add Gleditsch-Ward codes
+  add_gwcode_to_cow() %>%
+  # Add capital-to-capital distance
+  add_capital_distance() %>%
+  # Add contiguity information
+  add_contiguity() %>%
+  # Add major power data
+  add_cow_majors() %>%
+  # Add democracy variables
+  add_democracy() %>%
+  # add alliance data from Correlates of War
+  add_cow_alliance() %>%
+  # you should probably filter to politically relevant dyads earlier than later...
+  # Or not, it's your time and computer processor...
+  filter_prd()
+toc()
+
+```

### 2020-12-16 README.Rmd
update README
meh
finally took care of those binding variable issues.

@@
 
+# Installation
+
+You will ideally soon be able to install this on CRAN, as follows:
+
+```r
+install.packages("peacesciencer")
+
+```
+
+Until then, you can install the development version of this package through the `devtools` package.
+
+```r
+devtools::install_github("svmiller/stevetemplates")
+```
+
+# Usage
+
 The package is very much a work in progress. Right now, it has the following functions:
@@
 
-# Usage
+The workflow is going to look something like this. This is a "tidy"-friendly approach to a data-generating process in quantitative peace science.
 
-The workflow is going to look something like this.
+First, start with one of two processes to create either dyad-year or state-year data.
+The dyad-year data are created with the `create_dyadyears()` function.
+It has a few optional parameters with hidden defaults.
+The user can specify what kind of state system (`system`) data they want to use---either Correlates of War (`"cow"`) or Gledtisch-Ward (`"gw"`), whether they want to extend the data to the most recently concluded calendar year (`mry`) (i.e.
+Correlates of War state system membership data are current as of Dec.
+31, 2016 and the script can extend that to the end of 2019), and whether the user wants directed or non-directed dyad-year data (`directed`).
+
+The `create_stateyears()` works much the same way, though "directed" and "non-directed" make no sense in the state-year context.
+Both functions default to Correlates of War state system membership data to the most recently concluded calendar year.
+
+Thereafter, the user can specify what additional variables they want added to these dyad-year or state-year data.
+Do note: the additional functions lean primarily on Correlates of War state code identifiers.
+Indeed, the bulk of the quantitative peace science data ecosystem is built around the Correlates of War project.
+The variables the user wants are added in [a "pipe"](https://r4ds.had.co.nz/pipes.html) in a process like this.
+Do note that the user may want to break up the data-generating process into a few manageable "chunks" (e.g.
+first generating dyad-year data and saving to an object, adding to it piece by piece).
+
+All told, the process will look something like this.
 
@@
 
+# state-years now...
+
+tic()
+create_stateyears() %>%
+  add_gwcode_to_cow() %>%
+  add_capital_distance() %>%
+  add_contiguity() %>%
+  add_cow_majors() %>%
+  add_democracy()
+toc()
+
 ```

### 2020-12-16 R/add_capital_distance.R

@@ -1,3 +1,5 @@
+globalVariables(c('.', 'capitals', 'cow_ddy'))
+
 #' Add capital-to-capital distance to a dyad-year or state-year data frame
 #'
 #' @description \code{add_capital_distance()} allows you to add capital-to-capital
@@ -32,55 +34,59 @@
 #'
 #' @importFrom rlang .data
 #' @importFrom rlang .env
+#'
+
+
+
 
 add_capital_distance <- function(data) {
-  # require(dplyr)
-  # require(magrittr)
-  # require(dplyr)
-  # require(tidyr)
-  # require(lubridate)
-  # require(geosphere)
+  # require dplyr)
+  # require magrittr)
+  # require dplyr)
+  # require tidyr)
+  # require lubridate)
+  # require geosphere)
   capitals %>% rowwise() %>%
     mutate(year = list(seq(.data$styear, .data$endyear))) %>%
     unnest(.data$year) %>%
     select(.data$ccode, .data$year, .data$lat, .data$lng) %>%
     # There will be duplicates for when the country moved.
     # Under those conditions, the first capital should be first. It's basically the Jan. 1 capital, if you will.
-    group_by(ccode, year) %>% slice(1) %>% ungroup() -> capital_years
+    group_by(.data$ccode, .data$year) %>% slice(1) %>% ungroup() -> capital_years
 
   if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
     data %>%
-      left_join(., capital_years %>% select(ccode, year, lat,lng), by=c("ccode1"="ccode","year"="year")) %>%
-      rename(lat1 = lat,
-             lng1 = lng) %>%
-      left_join(., capital_years %>% select(ccode, year, lat,lng), by=c("ccode2"="ccode","year"="year")) %>%
-      rename(lat2 = lat,
-             lng2 = lng) -> data
+      left_join(., capital_years %>% select(.data$ccode, .data$year, .data$lat, .data$lng), by=c("ccode1"="ccode","year"="year")) %>%
+      rename(lat1 = .data$lat,
+             lng1 = .data$lng) %>%
+      left_join(., capital_years %>% select(.data$ccode, .data$year, .data$lat, .data$lng), by=c("ccode2"="ccode","year"="year")) %>%
+      rename(lat2 = .data$lat,
[…]

### 2020-12-16 R/zzz.R

@@ -1,0 +1,10 @@
+# CRAN Note avoidance. I hate this check.
+# Inspiration: https://github.com/HughParsonage/grattan/blob/master/R/zzz.R
+# Also:
+if(getRversion() >= "2.15.1")
+  utils::globalVariables(
+    # underlying data used for behind-the-scenes handsomeness
+    c("capitals", "ccode_democracy", "cow_alliance", "cow_contdir",
+      "cow_ddy", "cow_gw_years", "cow_majors", "cow_states",
+      "gml_dirdisp", "gw_ddy", "gw_states", "maoz_powers")
+  )

### 2020-12-16 README.Rmd
batch upload

@@
 
-# `peacesciencer`: Various Tools and Data for Quantitative Peace Science
+# `peacesciencer`: Tools and Data for Quantitative Peace Science
 
@@
 - `add_gwcode_to_cow():` adds Gleditsch-Ward state codes to dyad-year or state-year data with Correlates of War state codes.
+- `add_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data.
+- `add_nmc()`: adds estimates of national material capabilities (from Correlates of War) to dyad-year or state-year data.
 - `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
@@
 - `cow_majors`: Correlates of War major powers data (version: 2016)
+- `cow_nmc`: Correlates of War National Material Capabilities data (version 5.0)
 - `cow_states`: Correlates of War state system membership data (version: 2016)
+- `gml_dirdisp`: directed dispute-year data from version 2.1.1 of the Gibler-Miller-Little inter-state dispute data.
 - `gw_ddy`: a full directed dyad-year data frame of Gleditsch-Ward state system members
@@
   add_gwcode_to_cow() %>%
+  # Add GML MIDs 
+  add_mids %>%
   # Add capital-to-capital distance
@@
   add_democracy() %>%
+  # Add National Material Capabilities
+  add_nmc() %>%
   # add alliance data from Correlates of War
@@
   add_cow_majors() %>%
-  add_democracy()
+  add_democracy() %>%
+  add_nmc()
 toc()

### 2020-12-16 R/add_mids.R

@@ -1,0 +1,121 @@
+#' Add Gibler-Miller-Little (GML) Militarized Interstate Dispute (MID) data to dyad-year data frame
+#'
+#' @description \code{add_mids()} merges in GML's MID data to a dyad-year data frame. The current version
+#' of the GML MID data is 2.1.1.
+#'
+#' @return \code{add_mids()} takes a dyad-year data frame and adds dyad-year dispute information
+#' from the GML MID data.
+#'
+#' @details Dyads are capable of having multiple disputes in a given year, which can create a problem
+#' for merging into a complete dyad-year data frame. Consider the case of France and Italy in 1860, which
+#' had three separate dispute onsets that year (MID#0112, MID#0113, MID#0306), as illustrative of the problem.
+#' This merging process employs the following rules to whittle down these duplicate dispute-year observations. It
+#' first selects on MID onsets, then selecting highest fatality level, then highest hostility level, then
+#' the longest estimating minimum dispute duration, and finally, in the event of duplicates still outstanding,
+#' selecting the MID that came first. This is how GML present their full directed and non-directed dyad-year data.
+#'
+#' @author Steven V. Miller
+#'
+#' @param data a dyad-year data frame (either "directed" or "non-directed")
+#' @param keep an optional parameter, specified as a character vector, passed to the function in a \code{select(one_of(.))} wrapper. This
+#' allows the user to discard unwanted columns from the directed dispute data so that the output does not consume
+#' too much space in memory. Note: the Correlates of War system codes (\code{ccode1}, \code{ccode2}), the observation year
+#' (\code{year}), the presence or absence of an ongoing MID (\code{midongoing}), and the presence or absence of a unique
+#' MID onset (\code{midonset}) are *always* returned. It would be foolish and self-defeating to eliminate those observations.
+#' The user is free to keep or discard anything else they see fit.
+#'
+#' If \code{keep} is not specified in the function, the ensuing output returns everything.
+#'
+#' @references
+#'
[…]

### 2020-12-16 R/add_nmc.R

@@ -1,0 +1,70 @@
+#' Add Correlates of War National Military Capabilities Data
+#'
+#' @description \code{add_nmc()} allows you to add the Correlates of War National Material Capabilities data to dyad-yer or state-year data.
+#'
+#'
+#' @return \code{add_nmc()} takes a dyad-year data frame or state-year data frame and adds information
+#' about the national material capabilities for the state or two states in the dyad in a given year. If the data are dyad-year, the function
+#' adds 12 total columns for the first state (i.e. \code{ccode1}) and the second state (i.e. \code{ccode2}) for all estimates of national
+#' military capabilities provided by the Correlates of War project. If the data are state-year, the function returns six additional columns
+#' to the original data that contain that same information for a given state in a given year.
+#'
+#' @details The function leans on attributes of the data that are provided by the \code{create_dyadyear()} or
+#' \code{create_stateyear()} function. Make sure that function (or data created by that function) appear at the top
+#' of the proverbial pipe.
+#'
+#' @author Steven V. Miller
+#'
+#' @param data a dyad-year data frame (either "directed" or "non-directed") or a state-year data frame.
+#'
+#' @references
+#'
+#' Singer, J. David, Stuart Bremer, and John Stuckey. (1972). "Capability Distribution, Uncertainty, and Major Power War, 1820-1965." in Bruce Russett (ed) Peace, War, and Numbers, Beverly Hills: Sage, 19-48.
+#'
+#' Singer, J. David. 1987. "Reconstructing the Correlates of War Dataset on Material Capabilities of States, 1816-1985" International Interactions, 14: 115-32.
+#'
+#'
+#' @examples
+#'
+#' library(magrittr)
+#' cow_ddy %>% add_nmc()
+#'
+#' create_stateyears() %>% add_nmc()
+#'
+#'
+add_nmc <- function(data) {
+
+  if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
+
+    data %>% left_join(., cow_nmc, by=c("ccode1"="ccode","year"="year")) %>%
+      rename(milex1 = .data$milex,
+             milper1 = .data$milper,
+             irst1 = .data$irst,
+             pec1 = .data$pec,
+             tpop1 = .data$tpop,
+             upop1 = .data$upop,
+             cinc1 = .data$cinc) %>%
[…]

### 2020-12-16 R/data.R

@@ -355,11 +355,11 @@ NULL
 #'
 "maoz_powers"
 
-# Correlates of War National Military Capabilities Data
-
-# These are version 5.0 of the Correlates of War National Military Capabilities data. Data omit
-# the state abbreviation and version identifier for consideration.
-
+#'  Correlates of War National Military Capabilities Data
+#'
+#' These are version 5.0 of the Correlates of War National Military Capabilities data. Data omit
+#' the state abbreviation and version identifier for consideration.
+#'
 #' @format A data frame with 15171 observations on the following 9 variables.
 #' \describe{
 #' \item{\code{ccode}}{a numeric vector for the Correlates of War country code}

### 2020-12-16 data-raw/disputes/gml_disputes.R

@@ -1,0 +1,6 @@
+library(tidyverse)
+
+gml_dirdisp <- read_csv("data-raw/disputes/gml-ddy-disputes-2.1.1.csv")
+
+
+save(gml_dirdisp, file="data/gml_dirdisp.rda")

### 2020-12-16 data-raw/resave.R

@@ -1,0 +1,1 @@
+tools::resaveRdaFiles("data")

### 2020-12-16 tests/testthat/test_nodamnduplicates.R

@@ -11,6 +11,7 @@ test_that("Dyad-year additions do not create duplicates", {
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_gwcode_to_cow()))
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_mids()))
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_alliance()))
+  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_nmc()))
 
 })
 
@@ -21,4 +22,5 @@ test_that("State-year additions do not create duplicates", {
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_cow_majors()))
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_democracy()))
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_gwcode_to_cow()))
+  expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_nmc()))
 })

### 2020-12-17 README.Rmd
oops

@@
 ```r
-devtools::install_github("svmiller/stevetemplates")
+devtools::install_github("svmiller/peacesciencer")
 ```
@@
   # Add GML MIDs 
-  add_mids %>%
+  add_mids() %>%
   # Add capital-to-capital distance

### 2020-12-17 data-raw/cow_igo.R
Lot of things: 1) add IGOs data (no function yet), 2) compress data, 3) add citations data frame

@@ -1,0 +1,37 @@
+library(tidyverse)
+library(qs)
+
+# This is just the dyadic_formatv3.csv, but compressed because Jesus this file is huge.
+QIGO <- qs::qread("~/Dropbox/data/cow/igo/dyadic_formatv3.qs") %>%
+  # If it's negative, it's either missing or IGO not in existence
+  mutate_at(vars(AALCO:ncol(.)), ~ifelse(. < 0, NA, .))
+
+QIGO  %>%
+  select(AALCO:ncol(.)) -> the_igos
+
+QIGO$dyadigos <- rowSums(the_igos[,1:ncol(the_igos)], na.rm=TRUE)
+
+QIGO %>% select(ccode1, ccode2, year, dyadigos) -> cow_igo_ndy
+
+save(cow_igo_ndy, file="data/cow_igo_ndy.rda")
+
+read_csv("~/Dropbox/data/cow/igo/state_year_formatv3.csv") %>%
+  # If it's negative, it's either missing or IGO not in existence
+  mutate_at(vars(AAAID:ncol(.)), ~ifelse(. < 0, NA, .)) %>%
+  select(-state) %>%
+  group_by(ccode, year) %>%
+  gather(igo, member_code, AAAID:ncol(.)) %>%
+  mutate(full = ifelse(member_code == 1, 1, 0),
+         associate = ifelse(member_code == 2, 1, 0),
+         observer = ifelse(member_code == 3, 1, 0),
+         anytype = ifelse(member_code >= 1, 1, 0)) -> SY_IGO
+
+SY_IGO %>%
+  summarize(sum_igo_full = sum(full, na.rm=T),
+            sum_igo_associate = sum(associate, na.rm=T),
+            sum_igo_observer = sum(observer, na.rm=T),
+            sum_igo_anytype = sum(anytype, na.rm=T)) %>%
+  ungroup() -> cow_igo_sy
+
+
+save(cow_igo_sy, file="data/cow_igo_sy.rda")

### 2020-12-19 README.Rmd
Lot of things: 1) add IGOs data (no function yet), 2) compress data, 3) add citations data frame
oops

@@
 - `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year or state-year data.
-- `add_cow_alliance()`: adds Correlates of War major power information to dyad-year data.
+- `add_cow_alliance()`: adds Correlates of War major alliance information to dyad-year data.
 - `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data.
@@
 - `cow_gw_years`: a yearly data frame including information about Correlates of War and Gleditsch-Ward states.
+- `cow_igo_ndy`: non-directed dyad-year data for Correlates of War intergovernmental organizations data.
+- `cow_igo_sy`: state-year data for Correlates of War intergovernmental organizations data.
 - `cow_majors`: Correlates of War major powers data (version: 2016)

### 2020-12-19 README.Rmd
fix some things, add create_statedays()

@@
 - `add_nmc()`: adds estimates of national material capabilities (from Correlates of War) to dyad-year or state-year data.
+- `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format
+- `create_statedays()`: converts Correlates of War or Gleditsch-Ward state membership data into state-day format..
 - `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
-- `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format.
 - `filter_prd()`: filters dyad-year data frame to just "politically relevant" dyads.

### 2020-12-20 README.Rmd
add IGOs function, fix a few things

@@
 - `add_gwcode_to_cow():` adds Gleditsch-Ward state codes to dyad-year or state-year data with Correlates of War state codes.
+- `add_igos()`: adds Correlates of War International Governmental Organizations (IGOs) data to dyad-year or state-year data.
 - `add_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data.
@@
   add_democracy() %>%
+  # Add IGOs data
+  add_igos() %>%
   # Add National Material Capabilities
@@
   add_democracy() %>%
+  add_igos() %>%
   add_nmc()

### 2020-12-20 R/add_igos.R

@@ -1,0 +1,59 @@
+#' Add Correlates of War international governmental organizations (IGOs) data to dyad-year or state-year data.
+#'
+#' @description \code{add_igos()} allows you to add information from the Correlates oF War International
+#' Governmental Organizations data to dyad-year or state-year data, matching on Correlates of War system codes.
+#'
+#'
+#' @return \code{add_igos()} takes a dyad-year data frame or state-year data frame and adds information available from
+#' the Correlates of War International Governmental Organizations data. If the data are dyad-year, the function returns
+#' the original data with just one additional column for the total number of mutual IGOs for which both members of the dyad
+#' are full members. If the data are state-year, the function returns the original data with four additional columns. These are
+#' the number oF IGOs for which the state is a full member, the number of IGOs for which the state is an associate member, the
+#' number of IGOs for which the state is an observer, and the number of IGOs for which the state is involved in any way (i.e. the
+#' sum of the other three columns).
+#'
+#' @details The function leans on attributes of the data that are provided by the \code{create_dyadyear()} or
+#' \code{create_stateyear()} function. Make sure that function (or data created by that function) appear at the top
+#' of the proverbial pipe.
+#'
+#' @author Steven V. Miller
+#'
+#' @param data a dyad-year data frame (either "directed" or "non-directed") or a state-year data frame.
+#'
+#' @references
+#'
+#' Pevehouse, Jon C.W., Timothy Nordstron, Roseanne W McManus, Anne Spencer Jamison, “Tracking Organizations in the World: The Correlates of War IGO Version 3.0 datasets”, Journal of Peace Research 57(3): 492-503.
+#'
+#' Wallace, Michael, and J. David Singer. 1970. "International Governmental Organization in the Global System, 1815-1964." International Organization 24: 239-87.
+#'
+#' @examples
+#'
+#' \dontrun{
+#' library(magrittr)
+#' cow_ddy %>% add_igos()
+#'
+#' create_stateyears() %>% add_igos()
+#' }
+#'
+
+add_igos <- function(data) {
+  # The earliest missing is 1990, so we're going to treat these as just missing or unavailable.
+  if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
+
+    cow_igo_ndy %>%
+      rename(ccode1 = .data$ccode2,
+             ccode2 = .data$ccode1) %>%
[…]

### 2020-12-20 README.Rmd
add mindist

@@
 - `cow_majors`: Correlates of War major powers data (version: 2016)
+- `cow_mindist`: the minimum distance between Correlates of War states, in kilometers, in non-directed dyad-year format (1946-2015)
 - `cow_nmc`: Correlates of War National Material Capabilities data (version 5.0)

### 2020-12-20 README.Rmd
add stuff

@@
 - `gw_ddy`: a full directed dyad-year data frame of Gleditsch-Ward state system members
+- `gw_mindist`: the minimum distance between  Gleditsch-Ward states, in kilometers, in non-directed dyad-year format (1946-2015)
 - `gw_states`: Gleditsch-Ward independent state system data (version: 2017)

### 2020-12-20 R/data.R

@@ -495,3 +495,38 @@ NULL
 #' Weidmann, Nils B. and Kristian Skrede Gleditsch. 2010. "Mapping and Measuring Country Shapes: The \code{cshapes} Package." \emph{The R Journal} 2(1): 18-24
 
 "cow_mindist"
+
+#' The Minimum Distance Between States in the Gleditsch-Ward System, 1946-2015
+#'
+#' These are non-directed dyad-year data for the minimum distance between states in the Gleditsch-Ward state system from
+#' 1946 to 2015. The data are generated from the \code{cshapes} package.
+#'
+#' @format A data frame with 868813 observations on the following 4 variables.
+#' \describe{
+#' \item{\code{gwode1}}{the Gleditsch-Ward state system code for the first state}
+#' \item{\code{gwode2}}{the Gleditsch-Ward state system code for the second state}
+#' \item{\code{year}}{the year}
+#' \item{\code{mindist}}{the minimum distance between states on Dec. 31 of the year, in kilometers}
+#' }
+#'
+#' @details The data are generated from the \code{cshapes} package. The package authors purport that the data
+#' are generated to be compatible with the Gleditsch-Ward system. I trust them on this; indeed, Gleditsch is one of the
+#' authors of the \code{cshapes} package. However, I'm not sure how exhaustive the coverage is. For example,
+#' Tibet is missing in these data and it should not be. I do not use Gleditsch-Ward codes for my own research, so my
+#' quality control here for functions using these data will be minimal. I can only confirm there are no duplicates in the
+#' data.
+#'
+#' Data are automatically generated (by default) as directed dyad-years. I elect to make them non-directed for space
+#' considerations. Making non-directed dyad-year data into directed dyad-year data isn't too difficult in R. It just
+#' looks weird to see the code that does it.
+#'
+#' Most of the data I prove elsewhere in this package are to be understood as the data as they were at the *start* of
+#' the year. This is how I process, for example, the \code{capitals} data as they get merged in the \code{add_capital_distance()}
+#' function. However, the script that generates these data are set at Dec. 31 of the year and not Jan. 1. I do this for concerns
+#' of maximizing data coverage. If you wanted the same effect, just lag the data a year.
+#'
+#' @references
+#'
+#' Weidmann, Nils B. and Kristian Skrede Gleditsch. 2010. "Mapping and Measuring Country Shapes: The \code{cshapes} Package." \emph{The R Journal} 2(1): 18-24
+
+"gw_mindist"

### 2021-03-04 README.Rmd
a few new things

@@
 - `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data.
+- `add_cow_trade()`: adds Correlates of War trade data to dyad-year or state-year data.
 - `add_democracy()`: adds estimates of democracy/levels of democracy to dyad-year or state-year data.
@@
 - `cow_states`: Correlates of War state system membership data (version: 2016)
+- `cow_trade_sy`: Correlates of War state-year trade data (version 4.0)
 - `gml_dirdisp`: directed dispute-year data from version 2.1.1 of the Gibler-Miller-Little inter-state dispute data.
@@
   add_mids() %>%
+  # Add trade data
+  add_cow_trade() %>%
   # Add capital-to-capital distance
@@
   add_cow_majors() %>%
+  add_cow_trade() %>%
   add_democracy() %>%

### 2021-03-04 data-raw/citations.R

@@ -19,7 +19,9 @@ tribble(
   'politically relevant dyads', 'Weede, Erich. 1976. "Overwhelming preponderance as a pacifying condition among contiguous Asian dyads." Journal of Conflict Resolution 20: 395-411.',
   'politically relevant dyads', 'Lemke, Douglas and William Reed. 2001. "The Relevance of Politically Relevant Dyads." Journal of Conflict Resolution 45(1): 126-144.',
   'states (CoW)','Correlates of War Project. 2017. "State System Membership List, v2016." Online, https://correlatesofwar.org/data-sets/state-system-membership',
-  'states (GW)', 'Gleditsch, Kristian S. and Michael D. Ward. 1999. "A Revised List of Independent States since the Congress of Vienna" 25(4): 393--413'
+  'states (GW)', 'Gleditsch, Kristian S. and Michael D. Ward. 1999. "A Revised List of Independent States since the Congress of Vienna" 25(4): 393--413',
+  'trade', 'Barbieri, Katherine and Omar M.G. Keshk. 2016. Correlates of War Project Trade Data Set Cdebook, Version 4.0. Online: https://correlatesofwar.org',
+  'trade', 'Barbieri, Katherine, Omar M.G. Keshk, and Brian Pollins. 2009. "TRADING DATA: Evaluating Our Assumptions and Coding Rules." Conflict Management and Peace Science, 26(5): 471-491.'
 ) -> citations
 
 

### 2021-03-04 data-raw/cow_trade.R

@@ -23,6 +23,20 @@ saveRDS(cow_trade_ndy, file="data/cow_trade_ndy.rds")
 # save(cow_trade_ndy, file="data/cow_trade_ndy.rda")
 # ^ I tried...
 
+# I'mma try the DDY version now...
+
+cow_trade_ndy %>%
+  rename(ccode1 = ccode2,
+         ccode2 = ccode1,
+         flow1 = flow2,
+         flow2 = flow1,
+         smoothflow1 = smoothflow2,
+         smoothflow2 = smoothflow1) %>%
+  bind_rows(., cow_trade_ndy) ->cow_trade_ddy
+
+saveRDS(cow_trade_ddy, file="~/Dropbox/svmiller.github.io/R/peacesciencer/cow_trade_ddy.rds")
+
+
 
 cow_trade_sy <- read_csv("~/Dropbox/data/cow/trade/National_COW_4.0.csv") %>%
   # select just what we want

### 2021-03-23 README.Rmd
add minimum distance

@@
 - `add_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data.
+- `add_minimum_distance()`: adds minimum distance (in kilometers) to dyad-year or state-year data.
 - `add_nmc()`: adds estimates of national material capabilities (from Correlates of War) to dyad-year or state-year data.
@@
   add_cow_alliance() %>%
+  # add minimum distance
+  add_minimum_distance() %>%
   # you should probably filter to politically relevant dyads earlier than later...
@@
   add_igos() %>%
+  add_minimum_distance() %>%
   add_nmc()

### 2021-03-26 README.Rmd
Okay, I think we're up to speed now.

@@
 
+- `add_atop_alliance()`: adds ATOP alliance information to dyad-year data.
 - `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year or state-year data.
@@
 
+- `atop_alliance:` directed dyad-year alliance data from ATOP
 - `capitals`: a list of capitals and capital transitions for Correlates of War state system members
@@
   add_cow_alliance() %>%
-  # add minimum distance
-  add_minimum_distance() %>%
+  # add alliance data from ATOP
+  add_atop_alliance() %>%
+  # add minimum distance. No default, must specify "cow" or "gw"
+  add_minimum_distance(system = "cow") %>%
   # you should probably filter to politically relevant dyads earlier than later...
@@
   add_igos() %>%
-  add_minimum_distance() %>%
+  add_minimum_distance(system = "cow") %>%
   add_nmc()

### 2021-03-26 R/add_mids.R

@@ -44,11 +44,15 @@
 
 
 add_mids <- function(data, keep) {
-  # require(dplyr)
-  # require(magrittr)
-  # require(tidyr)
 
   if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
+
+    if (!all(i <- c("ccode1", "ccode2") %in% colnames(data))) {
+
+      stop("add_mids() merges on two Correlates of War codes (ccode1, ccode2), which your data don't have right now. Make sure to run create_dyadyears() at the top of the pipe. You'll want the default option, which returns Correlates of War codes.")
+
+
+    } else {
 
     gml_dirdisp %>%
       group_by(.data$ccode1, .data$ccode2, .data$year)  %>%
@@ -108,6 +112,10 @@ add_mids <- function(data, keep) {
     dirdisp %>%
       left_join(data, .) %>%
       mutate_at(vars("midongoing", "midonset"), ~ifelse(is.na(.), 0, .)) -> data
+
+    return(data)
+
+    }
 
 
 

### 2021-03-26 README.Rmd
update

@@
 
+- `add_archigos()`: adds some summary variables from Archigos about political leaders for state-year or dyad-year data.
 - `add_atop_alliance()`: adds ATOP alliance information to dyad-year data.
@@
 
+- `archigos`: an abbreviated version of the Archigos data, used internally
 - `atop_alliance:` directed dyad-year alliance data from ATOP
@@
   add_minimum_distance(system = "cow") %>%
+  # add Archigos data
+  add_archigos() %>%
   # you should probably filter to politically relevant dyads earlier than later...
@@
   add_minimum_distance(system = "cow") %>%
+  add_archigos() %>%
   add_nmc()

### 2021-03-26 R/add_archigos.R

@@ -53,7 +53,7 @@ add_archigos <- function(data) {
            dec31leadid = last(.data$leadid),
            leadertransition = ifelse(.data$jan1leadid != .data$dec31leadid, 1, 0),
            n_leaders = n_distinct(.data$leadid),
-           irregular = ifelse(leadertransition == 1 & any(.data$exit == "Irregular"), 1, 0)) %>%
+           irregular = ifelse(.data$leadertransition == 1 & any(.data$exit == "Irregular"), 1, 0)) %>%
     group_by(.data$ccode, .data$year) %>%
     select(.data$ccode, .data$year, .data$leadertransition, .data$irregular, .data$n_leaders, .data$jan1leadid, .data$dec31leadid) %>%
     group_by(.data$ccode, .data$year) %>%

### 2021-03-26 R/add_minimum_distance.R

@@ -79,10 +79,10 @@ add_minimum_distance <- function(data, system) {
     if (system == "cow") {
 
       cow_mindist %>%
-        group_by(ccode1, year) %>%
-        summarize(minmindist = min(mindist, na.rm = TRUE)) %>%
+        group_by(.data$ccode1, .data$year) %>%
+        summarize(minmindist = min(.data$mindist, na.rm = TRUE)) %>%
         ungroup()  %>%
-        rename(ccode = ccode1) %>%
+        rename(ccode = .data$ccode1) %>%
         left_join(data, .) -> data
 
       return(data)
@@ -90,10 +90,10 @@ add_minimum_distance <- function(data, system) {
     } else if (system == "gw") {
 
       gw_mindist %>%
-        group_by(gwcode1, year) %>%
-        summarize(minmindist = min(mindist, na.rm = TRUE)) %>%
+        group_by(.data$gwcode1, .data$year) %>%
+        summarize(minmindist = min(.data$mindist, na.rm = TRUE)) %>%
         ungroup()  %>%
-        rename(gwcode = gwcode1) %>%
+        rename(gwcode = .data$gwcode1) %>%
         left_join(data, .) -> data
 
       return(data)

### 2021-03-29 README.Rmd
whole bunch of stuff

@@
 - `add_nmc()`: adds estimates of national material capabilities (from Correlates of War) to dyad-year or state-year data.
+- `add_sdp_gdp()`: adds estimates of (gross and surplus) domestic product and population size to dyad-year or state-year data
 - `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format
@@
 - `cow_nmc`: Correlates of War National Material Capabilities data (version 5.0)
+- `cow_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States
 - `cow_states`: Correlates of War state system membership data (version: 2016)
@@
 - `gw_mindist`: the minimum distance between  Gleditsch-Ward states, in kilometers, in non-directed dyad-year format (1946-2015)
+- `gw_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States
 - `gw_states`: Gleditsch-Ward independent state system data (version: 2017)
@@
   add_archigos() %>%
+  # add gross and surplus GDP data
+  add_sdp_gdp(system = "cow") %>%
   # you should probably filter to politically relevant dyads earlier than later...
@@
   add_archigos() %>%
-  add_nmc()
+  add_nmc() %>%
+  add_sdp_gdp(system = "cow") 
 toc()

### 2021-03-29 R/data.R

@@ -536,7 +536,7 @@ NULL
 #' These are state-year-level data for national trade from the Correlates of War project.
 #'
 #'
-#' @format A data frame with 14410 on the following four variables.
+#' @format A data frame with 14410 observations on the following four variables.
 #' \describe{
 #' \item{\code{ccode}}{the Correlates of War state system code}
 #' \item{\code{year}}{the year}
@@ -559,7 +559,7 @@ NULL
 #' These are directed dyad-year-level data for alliance obligations and provisions from the ATOP project
 #'
 #'
-#' @format A data frame with 272,046 on the following eight variables.
+#' @format A data frame with 272,046 observations on the following eight variables.
 #' \describe{
 #' \item{\code{ccode1}}{a numeric vector for the Correlates of War state code for the first state}
 #' \item{\code{ccode2}}{a numeric vector for the Correlates of War state code for the second state}
@@ -586,7 +586,7 @@ NULL
 #' based on these data.
 #'
 #'
-#' @format A data frame with 3409 on the following seven variables.
+#' @format A data frame with 3409 observations on the following seven variables.
 #' \describe{
 #' \item{\code{ccode}}{a numeric vector for the Correlates of War state code}
 #' \item{\code{leadid}}{the unique leader identifier}
@@ -605,3 +605,55 @@ NULL
 #' \emph{Journal of Peace Research} 46(2): 269--83.
 
 "archigos"
+
+
+#' (Surplus and Gross) Domestic Product for Correlates of War States
+#'
+#' These are state-year level data for surplus and gross domestic product for Correlates of War state system members. Data also
+#' include population estimates for per capita standardization.
+#'
+#'
+#' @format A data frame with 27753 observations on the following five variables.
+#' \describe{
+#' \item{\code{ccode}}{a numeric vector for the Correlates of War state code}
+#' \item{\code{year}}{a numeric vector for the year}
+#' \item{\code{wbgdp2011est}}{a numeric vector for the estimated natural log of GDP in 2011 USD (log-transformed)}
+#' \item{\code{sdpest}}{a numeric vector for the estimated surplus domestic product (log-transformed)}
+#' }
+#' @details These were extracted from the actual replication files from \emph{International Studies Quarterly}. Because these
+#' data are ultimately being simulated, a user can expect some slight differences between the Correlates of War version of these data
[…]

### 2021-03-29 R/zzz.R

@@ -8,5 +8,5 @@ if(getRversion() >= "2.15.1")
       "cow_ddy", "cow_gw_years", "cow_majors", "cow_states",
       "gml_dirdisp", "gw_ddy", "gw_states", "maoz_powers", "cow_nmc",
       "cow_igo_ndy", "cow_igo_sy", "cow_trade_sy", "gw_mindist", "cow_mindist",
-      "archigos", "atop_alliance")
+      "archigos", "atop_alliance", "gw_sdp_gdp", "cow_sdp_gdp")
   )

### 2021-03-29 data-raw/surplus_gdp.R

@@ -1,0 +1,40 @@
+library(tidyverse)
+
+cow_sdp_gdp <- readRDS("~/Dropbox/data/surplus-gdp/master.rds") %>%
+  select(ccode, year, WorldBank_gdp_2011_ppp_estimate, WorldBank_pop_estimate, gdp_surplus1095_truncated) %>%
+  rename(wbgdp2011est = WorldBank_gdp_2011_ppp_estimate,
+         wbpopest = WorldBank_pop_estimate,
+         sdpest = gdp_surplus1095_truncated) %>%
+  # IIRC, R seems to get squeamish with file sizes for really large nominal numbers, like these. So +1 and log
+  # That'll standardize with the gw_* data frame next as well.
+  mutate_at(vars(wbgdp2011est, wbpopest, sdpest), ~log(.+1)) %>%
+  # I'm *pretty sure* Anders et al. are including years for which the state is not in the CoW system.
+  # That'd explain why there are more rows in here than they should be.
+  # The ultimate test for this for when left_join creates duplicates.
+  # I don't think they will here, but that's why I test for that.
+  print()
+
+save(cow_sdp_gdp, file="data/cow_sdp_gdp.rda")
+
+gw_sdp_gdp <- read_csv("~/Dropbox/data/surplus-gdp/gdp_pop_sdp_estimates.csv") %>%
+  rename(gwcode = gwno,
+         wbgdp2011est = WorldBank_gdp_2011_ppp_estimate,
+         wbpopest = WorldBank_pop_estimate,
+         sdpest = sdp) %>%
+  select(gwcode, year, wbgdp2011est,wbpopest,sdpest) %>%
+  print()
+
+save(gw_sdp_gdp, file="data/gw_sdp_gdp.rda")
+
+# fiddling with things here...
+
+create_stateyears() %>%
+  left_join(., cow_surplus_gdp)
+
+create_stateyears(system = "gw") %>%
+  left_join(., gw_surplus_gdp)
+
+cow_surplus_gdp %>%
+  group_by(ccode, year) %>%
+  tally() %>%
+  arrange(-n)

### 2021-03-29 tests/testthat/test_nodamnduplicates.R

@@ -13,6 +13,9 @@ test_that("Dyad-year additions do not create duplicates", {
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_mids()))
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_alliance()))
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_nmc()))
+  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_archigos()))
+  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_minimum_distance(system="cow")))
+  expect_equal(nrow(create_dyadyears(system = "gw")), nrow(create_dyadyears(system = "gw") %>% add_minimum_distance(system="gw")))
 
 })
 
@@ -25,4 +28,7 @@ test_that("State-year additions do not create duplicates", {
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_democracy()))
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_gwcode_to_cow()))
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_nmc()))
+  expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_archigos()))
+  expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_minimum_distance(system="cow")))
+  expect_equal(nrow(create_stateyears(system = "gw")), nrow(create_stateyears(system="gw") %>% add_minimum_distance(system="gw")))
 })

### 2021-03-30 README.Rmd
new data/function, pilot mode

@@
 - `add_sdp_gdp()`: adds estimates of (gross and surplus) domestic product and population size to dyad-year or state-year data
+- `add_ucdp_onsets()`: adds UCDP onsets to state-year data
 - `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format
@@
 - `maoz_powers`: Zeev Maoz' global/regional power data.
+- `ucdp_onsets`: a state-year data set of UCDP armed conflict onsets/episodes.
 
@@
   add_gwcode_to_cow() %>%
+  add_ucdp_onsets() %>%
   add_capital_distance() %>%

### 2021-03-30 R/data.R

@@ -618,6 +618,7 @@ NULL
 #' \item{\code{ccode}}{a numeric vector for the Correlates of War state code}
 #' \item{\code{year}}{a numeric vector for the year}
 #' \item{\code{wbgdp2011est}}{a numeric vector for the estimated natural log of GDP in 2011 USD (log-transformed)}
+#' \item{\code{wbpopest}}{a numeric vector for the estimated population size (log-transformed)}
 #' \item{\code{sdpest}}{a numeric vector for the estimated surplus domestic product (log-transformed)}
 #' }
 #' @details These were extracted from the actual replication files from \emph{International Studies Quarterly}. Because these
@@ -641,9 +642,10 @@ NULL
 #'
 #' @format A data frame with 27387 observations on the following five variables.
 #' \describe{
-#' \item{\code{gwcode}}{a numeric vector for the Correlates of War state code}
+#' \item{\code{gwcode}}{a numeric vector for the Gleditsch-Ward state code}
 #' \item{\code{year}}{a numeric vector for the year}
 #' \item{\code{wbgdp2011est}}{a numeric vector for the estimated natural log of GDP in 2011 USD (log-transformed)}
+#' \item{\code{wbpopest}}{a numeric vector for the estimated population size (log-transformed)}
 #' \item{\code{sdpest}}{a numeric vector for the estimated surplus domestic product (log-transformed)}
 #' }
 #' @details These were provided by Anders on a separate Github repository for this project. Because these
@@ -657,3 +659,36 @@ NULL
 #' \emph{International Studies Quarterly} 64(2): 392--405.
 
 "gw_sdp_gdp"
+
+#' UCDP Onset Data (v. 19.1)
+#'
+#' These are state-year level data for armed conflict onsets provided by the Uppsala Conflict Data Program (UCDP).
+#'
+#'
+#' @format A data frame with 10142 observations on the following eight variables.
+#' \describe{
+#' \item{\code{gwcode}}{a numeric vector for the Gleditsch-Ward state code}
+#' \item{\code{year}}{a numeric vector for the year}
+#' \item{\code{sumnewconf}}{a numeric vector for the sum of new conflicts/conflict-dyads}
+#' \item{\code{sumonset1}}{a numeric vector for the sum of new conflict episodes, whether because this is a new conflict or because there is more than one year since last conflict episode}
+#' \item{\code{sumonset2}}{a numeric vector for the sum of new conflict episodes, whether because this is a new conflict or because there is more than two years since last conflict episode}
[…]

### 2021-03-30 data-raw/citations.R

@@ -20,8 +20,11 @@ tribble(
   'politically relevant dyads', 'Lemke, Douglas and William Reed. 2001. "The Relevance of Politically Relevant Dyads." Journal of Conflict Resolution 45(1): 126-144.',
   'states (CoW)','Correlates of War Project. 2017. "State System Membership List, v2016." Online, https://correlatesofwar.org/data-sets/state-system-membership',
   'states (GW)', 'Gleditsch, Kristian S. and Michael D. Ward. 1999. "A Revised List of Independent States since the Congress of Vienna" 25(4): 393--413',
+  'surplus/gross domestic product', 'Anders, Therese, Christopher J. Fariss, and Jonathan N. Markowitz. 2020. "Bread Before Guns or Butter: Introducing Surplus Domestic Product (SDP)", International Studies Quarterly 64(2): 392--405.',
   'trade', 'Barbieri, Katherine and Omar M.G. Keshk. 2016. Correlates of War Project Trade Data Set Cdebook, Version 4.0. Online: https://correlatesofwar.org',
-  'trade', 'Barbieri, Katherine, Omar M.G. Keshk, and Brian Pollins. 2009. "TRADING DATA: Evaluating Our Assumptions and Coding Rules." Conflict Management and Peace Science, 26(5): 471-491.'
+  'trade', 'Barbieri, Katherine, Omar M.G. Keshk, and Brian Pollins. 2009. "TRADING DATA: Evaluating Our Assumptions and Coding Rules." Conflict Management and Peace Science, 26(5): 471-491.',
+  'ucdp_onsets', 'Gleditsch, Nils Petter; Peter Wallensteen, Mikael Eriksson, Margareta Sollenberg & Håvard Strand (2002) Armed Conflict 1946–2001: A New Dataset. Journal of Peace Research 39(5): 615–637.',
+  'ucdp_onsets', 'Pettersson, Therese; Stina Högbladh & Magnus Öberg (2019). Organized violence, 1989-2018 and peace agreements. Journal of Peace Research 56(4): 589-603.'
 ) -> citations
 
 

### 2021-03-30 data-raw/ucdp_onsets.R

@@ -1,0 +1,56 @@
+# this looks like a simple data set. I'll probably just poke around a little bit, for funsies.
+
+library(tidyverse)
+
+ucdp_onsets <- read_csv("~/Dropbox/data/ucdpprio/onset/onsets_computed-19.1.csv") %>%
+  select(-abc, -name) %>%
+  select(gwno_a, year, everything()) %>%
+  rename(gwcode = gwno_a) %>%
+  print()
+
+# Let me check for duplicates...
+
+ucdp_onsets %>%
+  group_by(gwcode, year) %>%
+  mutate(n = n()) %>%
+  filter(n > 1) %>% tail
+
+# Hmm. What is even going on here?
+# If I understand the codebook correctly, this should not happen. The data are offered as monadic, country-year data.
+# However, the litany of duplicates here suggest the data are extra-counting the country-years with onset episodes.
+# France in 1946 seems illustrative here. There are four France-1946 observations, which would square with onset episodes
+# with Cambodia, Laos, Thailand, and Vietnam. Likewise, Panama in 1989 has two observations, one for the United States
+# and another for the failed coup attempt led by Moisés Giroldi.
+
+# So, under those conditions, I am going to create a new sum variable that gets the total of new onsets.
+# I think, then, in the spirit of simplicity, I'm going to group_by(gwcode, year) and *add* those duplicates
+# together. What this will do is communicate the sum of new onsets with those specified windows from the codebook.
+
+# This might create some other weirdness with the year_prev variable. In a lot of these applications, its 1815, but
+# that's almost beside the point. The problem is, in the case of having to truncate duplicate observations, the year_prev
+# variable can't be informative. I suppose one way around this is to collapse it into a string. I'm tempted to just drop it.
+
+ucdp_onsets %>%
+  group_by(gwcode, year) %>%
+  summarize(sumnewconf = sum(newconf),
+            sumonset1 = sum(onset1),
+            sumonset2 = sum(onset2),
+            sumonset3 = sum(onset3),
+            sumonset5 = sum(onset5),
+            sumonset10 = sum(onset10)) %>%
+  ungroup() -> ucdp_onsets
+
+
+ucdp_onsets %>%
+  group_by(gwcode, year) %>% filter(n() > 1)
+
+# we're good
+
+save(ucdp_onsets, file="data/ucdp_onsets.rda")
+
+
[…]

### 2021-03-30 data-raw/citations.R
alrightie, let 'er rip.

@@ -23,8 +23,8 @@ tribble(
   'surplus/gross domestic product', 'Anders, Therese, Christopher J. Fariss, and Jonathan N. Markowitz. 2020. "Bread Before Guns or Butter: Introducing Surplus Domestic Product (SDP)", International Studies Quarterly 64(2): 392--405.',
   'trade', 'Barbieri, Katherine and Omar M.G. Keshk. 2016. Correlates of War Project Trade Data Set Cdebook, Version 4.0. Online: https://correlatesofwar.org',
   'trade', 'Barbieri, Katherine, Omar M.G. Keshk, and Brian Pollins. 2009. "TRADING DATA: Evaluating Our Assumptions and Coding Rules." Conflict Management and Peace Science, 26(5): 471-491.',
-  'ucdp_onsets', 'Gleditsch, Nils Petter; Peter Wallensteen, Mikael Eriksson, Margareta Sollenberg & Håvard Strand (2002) Armed Conflict 1946–2001: A New Dataset. Journal of Peace Research 39(5): 615–637.',
-  'ucdp_onsets', 'Pettersson, Therese; Stina Högbladh & Magnus Öberg (2019). Organized violence, 1989-2018 and peace agreements. Journal of Peace Research 56(4): 589-603.'
+  'ucdp', 'Gleditsch, Nils Petter; Peter Wallensteen, Mikael Eriksson, Margareta Sollenberg & Håvard Strand (2002) Armed Conflict 1946–2001: A New Dataset. Journal of Peace Research 39(5): 615–637.',
+  'ucdp', 'Pettersson, Therese; Stina Högbladh & Magnus Öberg (2019). Organized violence, 1989-2018 and peace agreements. Journal of Peace Research 56(4): 589-603.'
 ) -> citations
 
 

### 2021-03-30 tests/testthat/test_nodamnduplicates.R

@@ -5,6 +5,7 @@ library(magrittr)
 
 test_that("Dyad-year additions do not create duplicates", {
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_contiguity()))
+  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_gwcode_to_cow()))
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_capital_distance()))
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_majors()))
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_trade()))
@@ -21,6 +22,7 @@ test_that("Dyad-year additions do not create duplicates", {
 
 
 test_that("State-year additions do not create duplicates", {
+  expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_gwcode_to_cow()))
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_contiguity()))
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_capital_distance()))
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_cow_majors()))
@@ -30,5 +32,6 @@ test_that("State-year additions do not create duplicates", {
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_nmc()))
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_archigos()))
   expect_equal(nrow(create_stateyears()), nrow(create_stateyears() %>% add_minimum_distance(system="cow")))
+  expect_equal(nrow(create_stateyears(system = "gw")), nrow(create_stateyears(system = "gw") %>% add_ucdp_onsets()))
   expect_equal(nrow(create_stateyears(system = "gw")), nrow(create_stateyears(system="gw") %>% add_minimum_distance(system="gw")))
 })

### 2021-04-28 README.Rmd
alrightie, let 'er rip.
batch update on stuff

@@
 - `maoz_powers`: Zeev Maoz' global/regional power data.
+- `ucdp_acd`: a (not quite) dyad-year and (not quite) state-year data set on armed conflict episodes
 - `ucdp_onsets`: a state-year data set of UCDP armed conflict onsets/episodes.
@@
 It has a few optional parameters with hidden defaults.
-The user can specify what kind of state system (`system`) data they want to use---either Correlates of War (`"cow"`) or Gledtisch-Ward (`"gw"`), whether they want to extend the data to the most recently concluded calendar year (`mry`) (i.e.
+The user can specify what kind of state system (`system`) data they want to use---either Correlates of War (`"cow"`) or Gleditsch-Ward (`"gw"`), whether they want to extend the data to the most recently concluded calendar year (`mry`) (i.e.
 Correlates of War state system membership data are current as of Dec.

### 2021-04-28 R/add_archigos.R

@@ -24,14 +24,14 @@
 #' \emph{Journal of Peace Research} 46(2): 269--83.
 #'
 #' @examples
-#'
+#' \donttest{
 #' # just call `library(tidyverse)` at the top of the your script
 #' library(magrittr)
 #'
 #' cow_ddy %>% add_archigos()
 #'
 #' create_stateyears() %>% add_archigos()
-#'
+#' }
 #'
 #'
 #' @importFrom rlang .data

### 2021-04-28 R/add_minimum_distance.R

@@ -29,7 +29,7 @@
 #' \donttest{
 #' # just call `library(tidyverse)` at the top of the your script
 #' library(magrittr)
-#' cow_ddy %>% add_minimum_distance()
+#' cow_ddy %>% add_minimum_distance(system="cow")
 #'
 #' create_dyadyears(system = "gw") %>% add_minimum_distance(system = "gw")
 #'

### 2021-05-26 README.Rmd
add directed dispute-year data and such

@@
 - `cow_majors`: Correlates of War major powers data (version: 2016)
+- `cow_mid_dirdisps`: Correlates of War Militarized Interstate Dispute data (directed dyadic dispute-year, v. 5.0) 
 - `cow_mindist`: the minimum distance between Correlates of War states, in kilometers, in non-directed dyad-year format (1946-2015)

### 2021-05-27 README.Rmd
starting v. 0.4.

@@
 - `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data.
+- `add_cow_mids()`: adds Correlates of War (CoW) Militarized Interstate Dispute (MID) data to dyad-year data frame
 - `add_cow_trade()`: adds Correlates of War trade data to dyad-year or state-year data.
@@
 - `cow_majors`: Correlates of War major powers data (version: 2016)
+- `cow_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (CoW-MID, v. 5.0)
 - `cow_mid_dirdisps`: Correlates of War Militarized Interstate Dispute data (directed dyadic dispute-year, v. 5.0) 
+- `cow_mid_disps`: Correlates of War Militarized Interstate Dispute data (abbreviated dispute-level, v. 5.0) 
 - `cow_mindist`: the minimum distance between Correlates of War states, in kilometers, in non-directed dyad-year format (1946-2015)

### 2021-05-27 R/data.R

@@ -776,3 +776,92 @@ NULL
 #'
 
 "cow_mid_dirdisps"
+
+
+#' Abbreviate CoW-MID Dispute-level Data (v. 5.0)
+#'
+#' This is an abbreviated version of the dispute-level CoW-MID data.
+#'
+#'
+#' @format A data frame with 2436 observations on the following 7 variables.
+#' \describe{
+#' \item{\code{dispnum}}{a numeric vector for the CoW-MID dispute number}
+#' \item{\code{outcome}}{a numeric vector for the outcome of the MID}
+#' \item{\code{styear}}{a numeric vector for the start year of the MID}
+#' \item{\code{stmon}}{a numeric vector for the start month of the MID}
+#' \item{\code{settle}}{a numeric vector for the how dispute was settled}
+#' \item{\code{fatality}}{a numeric vector for the fatality level of the dispute}
+#' \item{\code{mindur}}{a numeric vector for the minimum duration of the MID}
+#' \item{\code{maxdur}}{a numeric vector for the maximum duration of the MID}
+#' \item{\code{hiact}}{a numeric vector for the highest action of the MID}
+#' \item{\code{hostlev}}{a numeric vector for the hostility level of the MID}
+#'  \item{\code{recip}}{a numeric vector for whether a MID was reciprocated}
+#' }
+#'
+#' @details These data are purposely light on information; they're not intended to be used for dispute-level analyses, per se.
+#' They're intended to augment the directed dyadic dispute-year data by adding in variables that serve as exclusion rules to whittle
+#' the data from dyadic dispute-year to just dyad-year data.
+#'
+#' @references
+#'
+#' Palmer, Glenn, and Roseanne W. McManus and Vito D'Orazio and Michael R. Kenwick and Mikaela Karstens
+#' and Chase Bloch and Nick Dietrich and Kayla Kahn and Kellan Ritter and Michael J. Soules. 2021.
+#' "The MID5 Dataset, 2011–2014: Procedures, coding rules, and description" \emph{Conflict Management and Peace Science}.
+#'
+
+"cow_mid_disps"
+
+
+
+#' Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (CoW-MID, v. 5.0)
+#'
+#' These are directed dyadic dispute year data derived from the Correlates of War (CoW) Militarized Interstate
+#'  Dispute (MID) project. Data are from version 5.0. These were whittled to where there is no duplicate dyad-years.
[…]

### 2021-05-27 R/zzz.R

@@ -9,5 +9,5 @@ if(getRversion() >= "2.15.1")
       "gml_dirdisp", "gw_ddy", "gw_states", "maoz_powers", "cow_nmc",
       "cow_igo_ndy", "cow_igo_sy", "cow_trade_sy", "gw_mindist", "cow_mindist",
       "archigos", "atop_alliance", "gw_sdp_gdp", "cow_sdp_gdp", "ucdp_onsets",
-      "ucdp_acd", "cow_mid_dirdisps")
+      "ucdp_acd", "cow_mid_dirdisps", "cow_mid_disps", "cow_mid_ddydisps")
   )

### 2021-05-30 README.Rmd
batch update of stuff, which includes things about the stuff

@@
 - `add_gwcode_to_cow():` adds Gleditsch-Ward state codes to dyad-year or state-year data with Correlates of War state codes.
+- `add_gml_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data.
 - `add_igos()`: adds Correlates of War International Governmental Organizations (IGOs) data to dyad-year or state-year data.
-- `add_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data.
+- `add_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data. (DEPRECATED)
 - `add_minimum_distance()`: adds minimum distance (in kilometers) to dyad-year or state-year data.
@@
 - `gml_dirdisp`: directed dispute-year data from version 2.1.1 of the Gibler-Miller-Little inter-state dispute data.
+- `gml_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (Gibler-Miller-Little, v. 2.1.1)
 - `gw_ddy`: a full directed dyad-year data frame of Gleditsch-Ward state system members

### 2021-05-30 R/add_cow_mids.R

@@ -18,8 +18,8 @@
 #' @param keep an optional parameter, specified as a character vector, passed to the function in a \code{select(one_of(.))} wrapper. This
 #' allows the user to discard unwanted columns from the directed dispute data so that the output does not consume
 #' too much space in memory. Note: the Correlates of War system codes (\code{ccode1}, \code{ccode2}), the observation year
-#' (\code{year}), the presence or absence of an ongoing MID (\code{dispongoing}), and the presence or absence of a unique
-#' MID onset (\code{disponset}) are *always* returned. It would be foolish and self-defeating to eliminate those observations.
+#' (\code{year}), the presence or absence of an ongoing MID (\code{cowmidongoing}), and the presence or absence of a unique
+#' MID onset (\code{cowmidonset}) are *always* returned. It would be foolish and self-defeating to eliminate those observations.
 #' The user is free to keep or discard anything else they see fit.
 #'
 #' If \code{keep} is not specified in the function, the ensuing output returns everything.
@@ -35,10 +35,10 @@
 #' \donttest{
 #' # just call `library(tidyverse)` at the top of the your script
 #' library(magrittr)
-#' cow_ddy %>% add_mids()
+#' cow_ddy %>% add_cow_mids()
 #'
 #' # keep just the dispute number and Side A/B identifiers
-#' cow_ddy %>% add_mids(keep=c("dispnum","sidea1", "sidea2"))
+#' cow_ddy %>% add_cow_mids(keep=c("dispnum","sidea1", "sidea2"))
 #' }
 #'
 
@@ -57,12 +57,12 @@ add_cow_mids <- function(data, keep) {
     if (missing(keep)) {
       cow_mid_ddydisps %>% select(everything()) -> dirdisp
     } else {
-      cow_mid_ddydisps %>% select(one_of("ccode1", "ccode2", "year", "disponset", "dispongoing", keep)) -> dirdisp
+      cow_mid_ddydisps %>% select(one_of("ccode1", "ccode2", "year", "cowmidonset", "cowmidongoing", keep)) -> dirdisp
     }
 
     dirdisp %>%
       left_join(data, .) %>%
-      mutate_at(vars("disponset", "dispongoing"), ~ifelse(is.na(.), 0, .)) -> data
+      mutate_at(vars("cowmidonset", "cowmidongoing"), ~ifelse(is.na(.), 0, .)) -> data
 
     return(data)
 

### 2021-05-30 R/add_gml_mids.R

@@ -1,0 +1,80 @@
+#' Add Gibler-Miller-Little (GML) Militarized Interstate Dispute (MID) data to dyad-year data frame
+#'
+#' @description \code{add_gml_mids()} merges in GML's MID data to a dyad-year data frame. The current version
+#' of the GML MID data is 2.1.1.
+#'
+#' @return \code{add_gml_mids()} takes a dyad-year data frame and adds dyad-year dispute information
+#' from the GML MID data.
+#'
+#' @details Dyads are capable of having multiple disputes in a given year, which can create a problem
+#' for merging into a complete dyad-year data frame. Consider the case of France and Italy in 1860, which
+#' had three separate dispute onsets that year (MID#0112, MID#0113, MID#0306), as illustrative of the problem.
+#' This merging process employs several rules to whittle down these duplicate dyad-years for merging into a dyad-year
+#' data frame.
+#'
+#' @author Steven V. Miller
+#'
+#' @param data a dyad-year data frame (either "directed" or "non-directed")
+#' @param keep an optional parameter, specified as a character vector, passed to the function in a \code{select(one_of(.))} wrapper. This
+#' allows the user to discard unwanted columns from the directed dispute data so that the output does not consume
+#' too much space in memory. Note: the Correlates of War system codes (\code{ccode1}, \code{ccode2}), the observation year
+#' (\code{year}), the presence or absence of an ongoing MID (\code{gmlmidongoing}), and the presence or absence of a unique
+#' MID onset (\code{gmlmidonset}) are *always* returned. It would be foolish and self-defeating to eliminate those observations.
+#' The user is free to keep or discard anything else they see fit.
+#'
+#' If \code{keep} is not specified in the function, the ensuing output returns everything.
+#'
+#' @references
+#'
+#' Gibler, Douglas M., Steven V. Miller, and Erin K. Little. 2016. “An Analysis of the Militarized
+#' Interstate Dispute (MID) Dataset, 1816-2001.” International Studies Quarterly 60(4): 719-730.
+#'
+#' @examples
+#'
+#' \donttest{
+#' # just call `library(tidyverse)` at the top of the your script
+#' library(magrittr)
+#' cow_ddy %>% add_gml_mids()
+#'
+#' # keep just the dispute number and Side A/B identifiers
[…]

### 2021-05-30 R/add_mids.R

@@ -46,6 +46,8 @@
 
 add_mids <- function(data, keep) {
 
+  .Deprecated("add_gml_mids")
+
   if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
 
     if (!all(i <- c("ccode1", "ccode2") %in% colnames(data))) {

### 2021-05-30 R/data.R

@@ -826,8 +826,8 @@ NULL
 #' \item{\code{ccode1}}{a numeric vector for the focal state in the dyad}
 #' \item{\code{ccode2}}{a numeric vector for the target state in the dyad}
 #' \item{\code{year}}{a numeric vector for the dispute-year}
-#' \item{\code{dispongoing}}{a numeric vector for whether there was a dispute ongoing in that year}
-#' \item{\code{disponset}}{a numeric vector for whether it was the onset of a new dispute (or new participant-entry into a recurring dispute)}
+#' \item{\code{cowmidongoing}}{a numeric vector for whether there was a dispute ongoing in that year}
+#' \item{\code{cowmidonset}}{a numeric vector for whether it was the onset of a new dispute (or new participant-entry into a recurring dispute)}
 #' \item{\code{sidea1}}{is \code{ccode1} on side A of the dispute?}
 #' \item{\code{sidea2}}{is \code{ccode2} on side A of the dispute?}
 #' \item{\code{fatality1}}{a numeric vector for the overall fatality level of \code{ccode1} in the dispute}

### 2021-05-30 R/zzz.R

@@ -9,5 +9,6 @@ if(getRversion() >= "2.15.1")
       "gml_dirdisp", "gw_ddy", "gw_states", "maoz_powers", "cow_nmc",
       "cow_igo_ndy", "cow_igo_sy", "cow_trade_sy", "gw_mindist", "cow_mindist",
       "archigos", "atop_alliance", "gw_sdp_gdp", "cow_sdp_gdp", "ucdp_onsets",
-      "ucdp_acd", "cow_mid_dirdisps", "cow_mid_disps", "cow_mid_ddydisps")
+      "ucdp_acd", "cow_mid_dirdisps", "cow_mid_disps", "cow_mid_ddydisps",
+      "gml_mid_ddydisps")
   )

### 2021-05-30 data-raw/gml_mid_ddydisps.R

@@ -2,13 +2,13 @@ library(tidyverse)
 library(peacesciencer)
 
 
-cow_mid_dirdisps %>%
+gml_dirdisp %>%
   group_by(ccode1, ccode2, year)  %>%
   mutate(duplicated = ifelse(n() > 1, 1, 0)) %>%
   # Remove anything that's not a unique MID onset
-  mutate(sd = sd(disponset),
+  mutate(sd = sd(midonset),
          sd = ifelse(is.na(sd), 0, sd)) %>%
-  mutate(removeme = ifelse(duplicated == 1 & disponset == 0 & sd > 0, 1, 0)) %>%
+  mutate(removeme = ifelse(duplicated == 1 & midonset == 0 & sd > 0, 1, 0)) %>%
   filter(removeme != 1) %>%
   # remove detritus
   select(-removeme, -sd) %>%
@@ -86,11 +86,11 @@ hold_this %>%
   filter(stmon == min(stmon)) %>%
   arrange(ccode1, ccode2, year) %>%
   # practice safe group_by()
-  ungroup() -> cow_mid_ddydisps
+  ungroup() -> gml_mid_ddydisps
 
-cow_mid_ddydisps %>%
-  rename(cowmidongoing = dispongoing,
-         cowmidonset = disponset) -> cow_mid_ddydisps
+gml_mid_ddydisps %>%
+  rename(gmlmidongoing = midongoing,
+         gmlmidonset = midonset) -> gml_mid_ddydisps
 
-save(cow_mid_ddydisps, file="data/cow_mid_ddydisps.rda")
+save(gml_mid_ddydisps, file="data/gml_mid_ddydisps.rda")
 

### 2021-05-31 README.Rmd
patch

@@
 create_dyadyears() %>%
-  # Add Gleditsch-Ward codes
+  # Add GML MIDs 
+  add_gml_mids(keep=NULL) %>%
+  # Add Correlates of War MIDs 
+  add_cow_mids(keep=NULL) %>%
+  # add peace years
+  add_peace_years() %>%
+  # Add Gleditsch-Ward codes, because we can
   add_gwcode_to_cow() %>%
-  # Add GML MIDs 
-  add_mids() %>%
   # Add trade data

### 2021-05-31 R/add_peace_years.R

@@ -62,6 +62,7 @@ add_peace_years <- function(data, pad = FALSE) {
 
       data <- sbtscs(data, .data$cowmidongoing, .data$year, .data$dyad, pad_ts = pad)
       names(data)[names(data) == "spell"] <- "cowmidspell"
+      attr(data, "ps_data_type") = "dyad_year"
 
     }
 
@@ -69,6 +70,7 @@ add_peace_years <- function(data, pad = FALSE) {
 
       data <- sbtscs(data, .data$gmlmidongoing, .data$year, .data$dyad, pad_ts = pad)
       names(data)[names(data) == "spell"] <- "gmlmidspell"
+      attr(data, "ps_data_type") = "dyad_year"
 
 
     }

### 2021-05-31 R/data.R

@@ -874,7 +874,7 @@ NULL
 #'  Its primary aim here is merging into a dyad-year data frame.
 #'
 #'
-#' @format A data frame with 10234 observations on the following 25 variables.
+#' @format A data frame with 9262 observations on the following 25 variables.
 #' \describe{
 #' \item{\code{dispnum}}{a numeric vector for the dispute number}
 #' \item{\code{ccode1}}{a numeric vector for the focal state in the dyad}
@@ -917,3 +917,39 @@ NULL
 #' Interstate Dispute (MID) Dataset, 1816-2001.” International Studies Quarterly 60(4): 719-730.
 
 "gml_mid_ddydisps"
+
+
+
+
+
+#' Thompson and Dreyer's (2012) Strategic Rivalries, 1494-2010
+#'
+#' A simple summary of all strategic (inter-state) rivalries from Thompson and Dreyer (2012).
+#'
+#'
+#' @format A data frame with 197 observations on the following 10 variables.
+#' \describe{
+#'\item{\code{rivalryno}}{a numeric vector for the rivalry number}
+#'\item{\code{rivalryname}}{a character vector for the rivalry name}
+#'\item{\code{ccode1}}{the Correlates of War state code for the state with the lowest Correlates of War state code in the rivalry}
+#'\item{\code{ccode2}}{the Correlates of War state code for the state with the highest Correlates of War state code in the rivalry}
+#'\item{\code{styear}}{a numeric vector for the start year of the rivalry}
+#'\item{\code{endyear}}{a numeric vector for the end year of the rivalry}
+#'\item{\code{region}}{a character vector for the region of the rivalry, per Thompson and Dreyer (2012)}
+#'\item{\code{type1}}{a character vector for the primary type of the rivalry (spatial, positional, ideological, or interventionary)}
+#'\item{\code{type2}}{a character vector for the secondary type of the rivalry, if applicable (spatial, positional, ideological, or interventionary)}
+#' \item{\code{type3}}{a character vector for the tertiary type of the rivalry, if applicable (spatial, positional, ideological, or interventionary)}
+#' }
+#' @details Information gathered from the appendix of Thompson and Dreyer (2012). Ongoing rivalries are
+#' right-bound at 2010, the date of publication for Thompson and Dreyer's handbook. Users are free to change this if they like. Data are effectively
[…]

### 2021-05-31 R/zzz.R

@@ -10,5 +10,5 @@ if(getRversion() >= "2.15.1")
       "cow_igo_ndy", "cow_igo_sy", "cow_trade_sy", "gw_mindist", "cow_mindist",
       "archigos", "atop_alliance", "gw_sdp_gdp", "cow_sdp_gdp", "ucdp_onsets",
       "ucdp_acd", "cow_mid_dirdisps", "cow_mid_disps", "cow_mid_ddydisps",
-      "gml_mid_ddydisps")
+      "gml_mid_ddydisps", "td_rivalries")
   )

### 2021-05-31 README.Rmd
more stuff

--- 2021-05-30 README.Rmd
@@
 create_dyadyears() %>%
-  # Add Gleditsch-Ward codes
+  # Add GML MIDs 
+  add_gml_mids(keep=NULL) %>%
+  # Add Correlates of War MIDs 
+  add_cow_mids(keep=NULL) %>%
+  # add peace years
+  add_peace_years() %>%
+  # Add Gleditsch-Ward codes, because we can
   add_gwcode_to_cow() %>%
-  # Add GML MIDs 
-  add_mids() %>%
+  # Add strategic rivalry data
+  add_strategic_rivalries() %>%
   # Add trade data

### 2021-05-31 tests/testthat/test_nodamnduplicates.R

@@ -18,6 +18,7 @@ test_that("Dyad-year additions do not create duplicates", {
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_nmc()))
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_archigos()))
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_minimum_distance(system="cow")))
+  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_strategic_rivalries()))
   expect_equal(nrow(create_dyadyears(system = "gw")), nrow(create_dyadyears(system = "gw") %>% add_minimum_distance(system="gw")))
 
 })

### 2021-06-04 README.Rmd
sfsfsfs

@@
 
-You will ideally soon be able to install this on CRAN, as follows:
+You can install this on CRAN, as follows:
 
@@
 
-Until then, you can install the development version of this package through the `devtools` package.
+You can install the development version of this package through the `devtools` package.
+The development version of the package invariably has more goodies, but may or may not be at various levels of stress-testing.
 
@@
 - `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year or state-year data.
+- `add_ccode_to_gw()`: 	Add Correlates of War state system codes to dyad-year or state-year data with Gleditsch-Ward state codes.
 - `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year or state-year data.
@@
 - `add_nmc()`: adds estimates of national material capabilities (from Correlates of War) to dyad-year or state-year data.
+- `add_peace_years()`: Add Peace Years to Your Conflict Data
 - `add_sdp_gdp()`: adds estimates of (gross and surplus) domestic product and population size to dyad-year or state-year data
+- `add_strategic_rivalries()`: Add Thompson and Dreyer's (2012) strategic rivalry data to dyad-year data frame
+- `add_ucdp_acd()`: Add UCDP Armed Conflict Data to state-year data frame
 - `add_ucdp_onsets()`: adds UCDP onsets to state-year data
@@
 - `cow_trade_sy`: Correlates of War state-year trade data (version 4.0)
+- `creg`: Composition of Religious and Ethnic Groups (CREG) Fractionalization/Polarization Estimates
 - `gml_dirdisp`: directed dispute-year data from version 2.1.1 of the Gibler-Miller-Little inter-state dispute data.
@@
 - `gw_states`: Gleditsch-Ward independent state system data (version: 2017)
-- `maoz_powers`: Zeev Maoz' global/regional power data.
+- `hief`: Historical Index of Ethnic Fractionalization data
+- `maoz_powers`: Zeev Maoz' global/regional power data
+- `rugged`: Rugged/Mountainous Terrain Data
+- `td_rivalries`: Thompson and Dreyer's (2012) Strategic Rivalries, 1494-2010
 - `ucdp_acd`: a (not quite) dyad-year and (not quite) state-year data set on armed conflict episodes
@@
   add_sdp_gdp(system = "cow") %>%
+  # add terrain ruggedness data, if you wanted it
+  add_rugged_terrain() %>%
   # you should probably filter to politically relevant dyads earlier than later...

### 2021-06-04 R/add_minimum_distance.R

@@ -17,8 +17,6 @@
 #' @author Steven V. Miller
 #'
 #' @param data a dyad-year data frame (either "directed" or "non-directed") or state-year data frame
-#' @param system a character specifying whether the user wants Correlates of War
-#' state-years ("cow") or Gleditsch-Ward ("gw") state-years.
 #'
 #' @references
 #'
@@ -40,11 +38,11 @@
 #' @importFrom rlang .data
 #' @importFrom rlang .env
 #'
-add_minimum_distance <- function(data, system) {
+add_minimum_distance <- function(data) {
 
   if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
 
-  if (system == "cow") {
+  if (length(attributes(data)$ps_system) > 0 && attributes(data)$ps_system == "cow") {
     # Just to deal with all cases, we're going to create a directed version
     cow_mindist %>%
       rename(ccode1 = .data$ccode2,
@@ -56,7 +54,7 @@ add_minimum_distance <- function(data, system) {
 
     return(data)
 
-  } else if (system == "gw") {
+  } else if (length(attributes(data)$ps_system) > 0 && attributes(data)$ps_system == "gw") {
 
     gw_mindist %>%
       rename(gwcode1 = .data$gwcode2,
@@ -70,12 +68,12 @@ add_minimum_distance <- function(data, system) {
 
   } else {
 
-    stop("add_minimum_distance() requires a declaration of either Correlates of War ('cow') or Gleditsch-Ward ('gw') as system type.")
+    stop("add_minimum_distance() requires either Correlates of War ('cow') or Gleditsch-Ward ('gw') as system type.")
   }
 
   } else if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "state_year") {
 
-    if (system == "cow") {
+    if (length(attributes(data)$ps_system) > 0 && attributes(data)$ps_system == "cow") {
 
       cow_mindist %>%
         group_by(.data$ccode1, .data$year) %>%
@@ -86,7 +84,7 @@ add_minimum_distance <- function(data, system) {
 
       return(data)
 
-    } else if (system == "gw") {
+    } else if (length(attributes(data)$ps_system) > 0 && attributes(data)$ps_system == "gw") {
 
       gw_mindist %>%
         group_by(.data$gwcode1, .data$year) %>%
@@ -99,7 +97,7 @@ add_minimum_distance <- function(data, system) {
 
[…]

### 2021-06-04 README.Rmd
dsfsfs

@@
 - `add_cow_trade()`: adds Correlates of War trade data to dyad-year or state-year data.
+- `add_creg_fractionalization()`: add fractionalization/polarization estimates from CREG to your dyad-year or state-year data
 - `add_democracy()`: adds estimates of democracy/levels of democracy to dyad-year or state-year data.
@@
   add_atop_alliance() %>%
-  # add minimum distance. No default, must specify "cow" or "gw"
-  add_minimum_distance(system = "cow") %>%
+  # add minimum distance
+  add_minimum_distance() %>%
   # add Archigos data
@@
   # add gross and surplus GDP data
-  add_sdp_gdp(system = "cow") %>%
+  add_sdp_gdp() %>%
   # add terrain ruggedness data, if you wanted it
@@
   add_igos() %>%
-  add_minimum_distance(system = "cow") %>%
+  add_minimum_distance() %>%
   add_archigos() %>%
   add_nmc() %>%
-  add_sdp_gdp(system = "cow") 
+  add_sdp_gdp() 
 toc()

### 2021-06-05 README.Rmd
addy mcaddadd

@@
 - `gw_states`: Gleditsch-Ward independent state system data (version: 2017)
+- `gwcode_democracy`: Gleditsch-Ward state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores").
 - `hief`: Historical Index of Ethnic Fractionalization data

### 2021-06-10 user-guide/doc/ms.tex
let 'er rip

% ARTICLE 2 ----
% This is just here so I know exactly what I'm looking at in Rstudio when messing with stuff.
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{cochineal}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}
    }
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: June 10, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
\author{true}
\date{}

% Jesus, okay, everything above this comment is default Pandoc LaTeX template. -----
% ----------------------------------------------------------------------------------
% I think I had assumed beamer and LaTex were somehow different templates.

\usepackage{kantlipsum}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
      {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}
    }
%\fi
\makeatother

\title{\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: June 10, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }

\date{}

\usepackage{titlesec}

% 
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\itshape} % \small\uppercase
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Steven V. Miller (Clemson University)},
             pdfkeywords = {software, statistical analysis, peace science, reproducibility},
            pdftitle={\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

% This will better treat References as a section when using natbib
% https://tex.stackexchange.com/questions/49962/bibliography-title-fontsize-problem-with-bibtex-and-the-natbib-package

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{longtable}
\usepackage{booktabs}
\LTcapwidth=.95\textwidth
\linespread{1.05}
\usepackage{hyperref}
\usepackage{setspace}
\let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
\usepackage{colortbl}
\arrayrulecolor{black}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\newtheorem{hypothesis}{Hypothesis}

% trick for moving figures to back of document
% really wish we'd knock this shit off with moving tables/figures to back of document
% but, alas...

\begin{document}

% \textsf{\textbf{This is sans-serif bold text.}}
% \textbf{\textsf{This is bold sans-serif text.}}

% \maketitle

{% \usefont{T1}{pnc}{m}{n}
\setlength{\parindent}{0pt}
\thispagestyle{plain}
{%\fontsize{18}{20}\selectfont\raggedright
\maketitle  % title \par

}

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12}
   \MakeUppercase{Steven V. Miller}, \small{Clemson University}

}

}

\begin{abstract}

%    \hbox{\vrule height .2pt width 39.14pc}

    \vskip 8.5pt % \small

\noindent \small{This article introduces \texttt{\{peacesciencer\}}, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a recent installation of the R programming language with some of the enhanced functionality of the broadly popular \texttt{\{tidvyerse\}} packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year or state-year analysis (e.g.~democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at \url{http://svmiller.com/peacesciencer}.}

\vskip 8.5pt \noindent \emph{Keywords}: software, statistical analysis, peace science, reproducibility \par

%    \hbox{\vrule height .2pt width 39.14pc}

\end{abstract}

\vskip -8.5pt

 % removetitleabstract

\setlength{\parindent}{16pt}
\setlength{\parskip}{0pt}

% We'll put doublespacing here
% Remember to cut it out later before bib
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

This data feature tackles a recurring problem for researchers in the peace science community. True research reproducibility is achieved creating original dyad-year or state-year data from scratch, though no published guide exists that informs how to do this on their own. Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice. In a related manner, researchers may spend too much time reproducing old code for standard information that goes into any dyad-year or state-year analysis---like contiguity relationships and democracy, among many others---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes. This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well. Graduate students, in particular, face unique challenges associated with these developments. Graduate students beginning their careers in peace science to learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.

\texttt{\{peacesciencer\}} is designed to address these problems. Built around the free and open source R programming language, \texttt{\{peacesciencer\}} contains a suite of data and functions for creating data of interest to researchers. Researchers can use \texttt{\{peacesciencer\}} to create dyad-year, state-year, and even state-day data from scratch. Afterward, they can add a variety of standard information (e.g.~contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. \texttt{\{peacesciencer\}} comes with some data innovations as well, including a comprehensive data set on democracy, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts. All are done with the maximum possible transparency. The project is available for public view of Github (\url{https://github.com/svmiller/peacesciencer}). The \texttt{data-raw} directory on the project's Github contains information and comments about how every data set was created and the \texttt{R} directory informs the user how each function works. The function manuals (\url{http://svmiller.com/peacesciencer/reference/}) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, \texttt{\{peacesciencer\}} not only assists a peace scientist with their research, but it does so in a manner that best conforms to the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

This data feature proceeds in the following fashion. The next section expands what need this package fills for peace scientists in the present. Afterward, it provides an overview of what is included in \texttt{\{peacesciencer\}} to help researchers more quickly conduct the kind of quantitative research they want. Thereafter, it provides a tutorial on how to install and best use \texttt{\{peacesciencer\}} in the R programming language. A more comprehensive tutorial follows, showing how \texttt{\{peacesciencer\}} already has a suite of data and functions that can allow for effective replications of a ``dangerous dyads'' type analysis (Bremer 1992) and standard state-year analyses of civil conflict onset (e.g. Fearon and Laitin 2003). This feature concludes with a discussion of what else \texttt{\{peacesciencer\}} is capable of doing in future updates and how this R package can inform more reasoned design decisions for researchers in peace science.

\hypertarget{why-peacesciencer}{%
\section{\texorpdfstring{Why \texttt{\{peacesciencer\}}?}{Why \{peacesciencer\}?}}\label{why-peacesciencer}}

\texttt{\{peacesciencer\}} is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch. NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars (Bennett, Poast, and Stam 2019). NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade (Bennett and Stam 2000). No matter, \texttt{\{peacesciencer\}} is motivated by the following ideals that led to its creation.

For one, researchers may invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume, for example, a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's primary impetus was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g.~dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit. \texttt{\{peacesciencer\}} is primarily born from this question about how to create dyad-year and state-year data from scratch. The underlying code that produces these data types is available online and are converted into simple functions for the ease of the researcher.

Second, researchers also invest too much time in retracing steps for peace science analyses for new projects. Assume, for example, a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data). Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project. They may then have to change the name of some files, change some code to account for potentially new column names in the newer data, and troubleshoot instances where their old code does not perform as it once did. At its worst, this process may lead to some errors by the researcher. At best, this is tedium that spends the researcher's time they would rather invest in analyzing the data. The lion's share of \texttt{\{peacesciencer\}}'s functionality is both creating the units of analysis for the researcher and merging in different forms of data in wide use in the peace science community so that the researcher can spend less of their time on tedium.

Third, the creation of the data and the analysis of the data is increasingly becoming one continuous process. Not too long ago, it used to be the case that researchers had to download a data set, or create one from scratch (possibly in a spreadsheet or through a program like EUGene). After downloading or constructing the data, the researcher then opened a specialty program for statistical analysis (e.g.~SAS, SPSS, Stata) to recode raw data into a form suitable for analysis before running a statistical model that regresses some outcome on a set of covariates. Current research practices still resemble this process, but the steps between them are no longer as large as they were in the past. Software options exist that allow the researcher to load data, create data, clean data, analyze data, and present the results of the analysis all within one program. \texttt{\{peacesciencer\}}, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and Rstudio (its free-for-use integrated desktop environment {[}IDE{]}).

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer (c.f. Bowers and Voors 2016). This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that \texttt{\{peacesciencer\}} uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet (Arel-Bundock 2021b), scrape \emph{any} type of data from the internet (Boehmke 2016), and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero (Arel-Bundock 2021a). \texttt{\{peacesciencer\}} embraces this. This R programming language package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

Fifth, statistical software should be kind to graduate students in peace science with the aforementioned point in mind. Graduate students just beginning their career in peace science scholarship are tasked to learn the substantive material, the characteristics of the underlying data in wide use in the community, and the quantitative methods and statistical software available to them, effectively all at the same time. A student interested in civil conflict scholarship will need to learn the UCDP-PRIO armed conflict data in some detail at the same time they learn about the theoretical arguments that informed the creation of the data, likely at the same time they take quantitative methods courses that show them how to load and explore data in a choseen statistical software. At the same time that creating data is increasingly blurred with coding and processing data, students are simultaneously learning to explore data for themselves as they read about the analyses of the data. \texttt{\{peacesciencer\}} aims to make this process easier for graduate students using the features of the R programming language. Graduate students can use \texttt{\{peacesciencer\}} to learn about peace science data and computer programming in R at the same time.

Finally, the creation and presentation of data in peace science should be 100\% transparent, which \texttt{\{peacesciencer\}} takes seriously in the following ways. The website for \texttt{\{peacesciencer\}} has several vignettes that describe its processes in some detail. These include how it provides reasonable estimates of democracy that may not be available in the Polity data or the Varieties of Democracy data and how a researcher can whittle dyadic dispute-years into true dyad-years through reasonable case exclusions. The project's Github contains a publicly available \texttt{data-raw} directory that shows how every data set included (and processed) in \texttt{\{peacesciencer\}} was created. The function manuals included \texttt{\{peacesciencer\}} contain ample documentation that clarifies what the function is doing, what it returns to the user, and why it is doing it this way. Users can also use the project's Github to point out bugs, ask for further clarification, and propose additions. \texttt{\{peacesciencer\}} takes seriously the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science and endeavors for maximum transparency, leveraging open source and version control software to inform users of what data it uses and how it uses the data.

\hypertarget{what-is-included-in-peacesciencer}{%
\section{\texorpdfstring{What is Included in \texttt{\{peacesciencer\}}}{What is Included in \{peacesciencer\}}}\label{what-is-included-in-peacesciencer}}

\texttt{\{peacesciencer\}} comes with a battery of built-in functions for generating the most widespread forms of peace science data. Each function uses raw or pre-processed data included in \texttt{\{peacesciencer\}}. For example, \texttt{create\_dyadyears()} transforms the raw state system membership data for either the Correlates of War or Gleditsch-Ward system into the full universe of post-1816 dyad-years. \texttt{add\_gml\_mids()} uses a dyadic dispute-year version of the MID data offered by Gibler, Miller, and Little (2016), which, as the \texttt{data-raw} directory on the project's Github shows, is itself derived from the dispute-level and participant-level versions of the data. \texttt{\{peacesciencer\}} also has some data innovations. It has an original data set on capital-to-capital distances, calculating for distance in kilometers between state capitals using the ``Vincenty'' method (i.e.~``as the crow flies'') that also accounts for instances when capitals moved (e.g.~Brazil in 1960, Burundi in 2018). It also features an innovation of democracy data, providing reasonable estimates of democracy using the Marquez (2016) of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve, and Melton 2010).\footnote{A vignette on the \texttt{\{peacesciencer\}} website explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.} As of writing, \texttt{\{peacesciencer\}} comes with the following functionality.

\begingroup\fontsize{7}{9}\selectfont

\begin{longtable}[t]{>{}l|>{}l|>{}c|>{}c}
\caption{\label{tab:unnamed-chunk-1}Features of \{peacesciencer\}}\\
\hline
\textbf{Data} & \textbf{Description} & \textbf{Type} & \textbf{Function}\\
\hline
(CoW, G-W) Dyad-Year Data & Create dyad-years from state system membership data &  & \ttfamily{create\_dyadyears()}\\
\arrayrulecolor{white}
\hline
(CoW, G-W) State-Year Data & Create state-years from state system membership data &  & \ttfamily{create\_stateyears()}\\
\arrayrulecolor{white}
\hline
(CoW, G-W) State-Day Data & Create state-days from state system membership data &  & \ttfamily{create\_statedays()}\\
\arrayrulecolor{white}
\hline
Archigos (v. 4.1) & Add Archigos political leader information & D, S & \ttfamily{add\_archigos()}\\
\arrayrulecolor{white}
\hline
ATOP Alliances (v. 5.0) & Add ATOP alliance data & D & \ttfamily{add\_atop\_alliance()}\\
\arrayrulecolor{white}
\hline
Capital-to-Capital Distances & Add capital-to-capital distance & D, S & \ttfamily{add\_capital\_distance()}\\
\arrayrulecolor{white}
\hline
Convert CoW codes to G-W codes & Add matching CoW codes for G-W states by year & D, S & \ttfamily{add\_ccode\_to\_gw()}\\
\arrayrulecolor{white}
\hline
Convert G-W codes to CoW codes & Add matching G-W codes for CoW states by year & D, S & \ttfamily{add\_gwcode\_to\_cow()}\\
\arrayrulecolor{white}
\hline
CoW National Capabilities (v. 5.0) & Adds capabilities of states by year & D, S & \ttfamily{add\_nmc()}\\
\arrayrulecolor{white}
\hline
CoW Direct Contiguity (v. 3.2) & Adds contiguity information by year & D, S & \ttfamily{add\_contiguity()}\\
\arrayrulecolor{white}
\hline
CoW Alliances (v. 4.1) & Adds CoW alliance data by year & D & \ttfamily{add\_cow\_alliance()}\\
\arrayrulecolor{white}
\hline
CoW IGOs (v. 3.0) & Adds CoW IGO data by year & D, S & \ttfamily{add\_igos()}\\
\arrayrulecolor{white}
\hline
CoW Major Powers (v. 2016) & Adds CoW major power by year & D, S & \ttfamily{add\_cow\_majors()}\\
\arrayrulecolor{white}
\hline
CoW MIDs (v. 5.0) & Add CoW MID information to data & D & \ttfamily{add\_cow\_mids()}\\
\arrayrulecolor{white}
\hline
CoW Trade Data (v. 4.0) & Add CoW trade information to data & D, S & \ttfamily{add\_cow\_trade()}\\
\arrayrulecolor{white}
\hline
CoW Wars & Add CoW inter-state (v. 4.0) and intra-state (v. 4.1) war data & D & \ttfamily{add\_cow\_wars()}\\
\arrayrulecolor{white}
\hline
CREG (v. 1.02) & Add ethnic/religious fractionalization/polarization by year & D, S & \ttfamily{add\_creg\_fractionalization()}\\
\arrayrulecolor{white}
\hline
Democracy & Add Polity (v. 2017), V-Dem (v. 10), and Marquez' (2016) UDS extensions & D,S & \ttfamily{add\_democracy()}\\
\arrayrulecolor{white}
\hline
GML MIDs (v. 2.1.1) & Add Gibler-Miller-Little MIDs to data & D & \ttfamily{add\_gml\_mids()}\\
\arrayrulecolor{white}
\hline
Minimum Distance (v. 2016) & Adds minimum distance data by year & D, S & \ttfamily{add\_minimum\_distance()}\\
\arrayrulecolor{white}
\hline
Peace Years & Calculate peace years for various disputes & D, S & \ttfamily{add\_peace\_years()}\\
\arrayrulecolor{white}
\hline
Rugged/Mountainous Terrain & Add 'ruggedness'/'mountainous' terrain data by year & D, S & \ttfamily{add\_rugged\_terrain()}\\
\arrayrulecolor{white}
\hline
(Surplus/Gross) Domestic Product & Add Anders et al. (2020) GDP/SDP simulations by year & D, S & \ttfamily{add\_sdp\_gdp()}\\
\arrayrulecolor{white}
\hline
Strategic Rivalries & Add Thompson and Dreyer (2012) strategic rivalry data by year & D, S & \ttfamily{add\_strategic\_rivalries()}\\
\arrayrulecolor{white}
\hline
UCDP Armed Conflicts & Add UCDP armed conflict data by year & S & \ttfamily{add\_ucdp\_acd()}\\
\arrayrulecolor{black}
\hline
\end{longtable}
\endgroup{}

\texttt{\{peacesciencer\}}'s coverage focuses mostly on data that are released as standalone data sets for download, especially those in the Correlates of War or Gleditsch-Ward ecosystem of data. Data that can be obtained from a stable advanced programming interface---like the World Bank, for example---can be obtained through those other means (e.g. Arel-Bundock 2021b). Its coverage will assuredly expand with new additions of interest to the peace science community.

\hypertarget{how-to-install-peacesciencer}{%
\section{\texorpdfstring{How to Install \texttt{\{peacesciencer\}}}{How to Install \{peacesciencer\}}}\label{how-to-install-peacesciencer}}

\texttt{\{peacesciencer\}} is a package for the R programming language. This assumes at least some familiarity with the R programming language. Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.0. \texttt{\{peacesciencer\}} is designed to be as user-friendly as possible. Those proficient in R, those just learning R, and those with no experience in R should be able to pick up its use fairly quickly.

\texttt{\{peacesciencer\}}'s functions work out of the box, though it works best with one additional package. This is \texttt{\{tidyverse\}}, itself a set of packages that share a common form and design (Wickham and Grolemund 2017). \texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}} but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality. A user can open an R session and install both packages as follows.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{install.packages}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"tidyverse"}\NormalTok{, }\StringTok{"peacesciencer"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

R packages once installed need to be loaded with every R session. The user can load both with the \texttt{library()} function in R.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{library}\NormalTok{(peacesciencer)}
\end{Highlighting}
\end{Shaded}

Thereafter, a researcher can start using \texttt{\{peacesciencer\}} to create the kind of data they need.

\hypertarget{a-tutorial-on-how-to-use-peacesciencer}{%
\section{\texorpdfstring{A Tutorial on How to Use \texttt{\{peacesciencer\}}}{A Tutorial on How to Use \{peacesciencer\}}}\label{a-tutorial-on-how-to-use-peacesciencer}}

\texttt{\{peacesciencer\}} has a simple ``pipe''-based workflow. The top of the ``pipe'' is the first function to create the data. These options include \texttt{create\_dyadyears()}, \texttt{create\_stateyears()}, and \texttt{create\_statedays()}. All options include two arguments: \texttt{system} and \texttt{mry}. \texttt{system} takes one of two values---``cow'' or ``gw''---for what state system membership data the user wants. If \texttt{system} is not specified, the function defaults to Correlates of War system membership data given the prominence of the Correlates of War project in the conflict data ecosystem. \texttt{mry}---short for ``most recent year''---determines whether or not the function extends the membership data to the most recently concluded calendar year. The function defaults to \texttt{TRUE} in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017). \texttt{create\_dyadyears()} has a third argument---\texttt{directed}---that determines whether the user wants directed dyad-year data or non-directed dyad-year data. The function defaults to \texttt{TRUE}, returning directed dyad-year data.

The following example creates a full state-year data frame of Gleditsch-Ward states from 1816 to 2020.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{create\_stateyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

This example would create a full non-directed dyad-year data frame of Correlates of War states from 1816 to 2016, the last year of record in the Correlates of War state system membership data.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{directed =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{mry =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

This example would create a full state-day data frame from the Gleditsch-Ward system for all states from Jan.~1, 1816 to Dec.~31, 2020.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{create\_statedays}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

There are not too many data sets in the peace science community with daily data, certainly dating to the 19th century, but \texttt{\{peacesciencer\}} can greatly reduce the costs of creating the kind of data the user may want.

\hypertarget{add-standard-information-to-dyad-year-or-state-year-data}{%
\subsection{Add Standard Information to Dyad-Year or State-Year Data}\label{add-standard-information-to-dyad-year-or-state-year-data}}

The next step in the \texttt{\{peacesciencer\}} ``pipe''-based workflow will also be among the biggest time-savers for researchers using it. Researchers who want to understand, say, the correlates of inter-state conflict onset almost always do the following. They will add the dependent variables from an inter-state conflict data set, like the Gibler-Miller-Little (GML) dispute data set (Gibler, Miller, and Little 2016). They will add several indicators to the dyad-year data that are well-traveled in conflict scholarship, including indicators of things like joint alliance contracts and the contiguity relationship of both states in the dyad. \texttt{\{peacesciencer\}} has great coverage for a lot of these data and can even allow almost full-scale replications with the data it offers.

For example, assume a researcher wanted to replicate something analogous to Bremer's (1992) famous ``dangerous dyads'' analysis about the causes of conflict onset. \texttt{\{peacesciencer\}} can do something like this in the following pipe-based workflow. First, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)} will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016. This is effectively the same function used above, but with two user-specified overrides.

After creating the base data they want, the user can add data to it with the ``pipe'' operator. The \texttt{\{tidyverse\}} represents this pipe operator as \texttt{\%\textgreater{}\%}. Ending each line with this pipe passes forward an object (here: data) into the next function on the next line. Put in other words, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)\ \%\textgreater{}\%\ filter\_prd()} is a command in \texttt{\{tidyverse\}} and \texttt{\{peacesciencer\}} to create all non-directed dyad-years from 1816 to 2016 and then subset those data to only the politically relevant dyad-years. Because politically relevant dyads depend on information about the major power status for both states as well as their contiguity relationship (Lemke and Reed 2001), the function adds in those data (i.e.~quietly executing \texttt{add\_contiguity()} and \texttt{add\_cow\_majors()}) before subsetting the data to just the dyads that have some kind of contiguity relationship and/or have a major power in it.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{directed =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{mry =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter\_prd}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_gml\_mids}\NormalTok{(}\AttributeTok{keep =} \ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_peace\_years}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_nmc}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_democracy}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_cow\_alliance}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_sdp\_gdp}\NormalTok{() }\OtherTok{{-}\textgreater{}}\NormalTok{ Data}
\end{Highlighting}
\end{Shaded}

Next, the pipe-based workflow continues with \texttt{add\_gml\_mids(keep\ =\ NULL)}, which will add information about whether there was an ongoing dispute in the dyad (and whether it was a dispute onset) from the GML dispute data.\footnote{\texttt{keep\ =\ NULL} is optional and removes supporting information about dyadic dispute-participants (e.g.~dispute-level hostility, dispute-level outcome) in order to reduce the size of the data set.} \texttt{add\_peace\_years()} will calculate peace years between ongoing disputes in these politically relevant non-directed dyads.\footnote{\texttt{add\_peace\_years()} will only calculate the peace years---titled, in this case, \texttt{gmlmidspell}---and will leave the temporal dependence adjustment to the taste of the researcher. Importantly, I do not recommend manually creating splines or square/cube terms because it creates more problems in adjusting for temporal dependence in model predictions. In a regression formula in R, you can specify the D. B. Carter and Signorino (2010) approach as \texttt{...\ +\ gmlmidspell\ +\ I(gmlmidspell\^{}2)\ +\ I(gmlmidspell\^{}3)}. The Beck, Katz, and Tucker (1998) cubic splines approach is \texttt{...\ +\ splines::bs(gmlmidspell,\ 4)}. This function includes the spell and three splines (hence the 4 in the command).} \texttt{add\_nmc()} will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity. \texttt{add\_democracy()} will add information about the level of democracy for both states in the dyad using three different indicators. Researchers can select which one they prefer from these indicators. \texttt{add\_cow\_alliance()} will add information about whether there was an alliance in the dyad-year by way of whether ther was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge. Finally, \texttt{add\_sdp\_gdp()} will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by Anders, Fariss, and Markowitz (2020).

Whereas \texttt{add\_sdp\_gdp()} is the last command in the pipe-based workflow, the \texttt{\{peacesciencer\}} call ends by assigning to an object called \texttt{Data}. This type of assignment is done with the ``right hand'' assignment operator (i.e.~\texttt{-\textgreater{}}). If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like \texttt{haven::write\_dta(Data,\ "my-data.dta")} and import it into Stata when they are done.

\texttt{\{peacesciencer\}} has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding \texttt{-9s} in these data to be \texttt{NA} in the National Material Capabilities data. There is only some slight data work to create the desired indicators for a statistical model of conflict onset. Here, briefly, these functions lean on \texttt{\{tidyverse\}} ``verbs'' and base R terminology to do things like creating a dummy variable for whether the dyad was land-contiguous, whether there was a major power on either side of the non-directed dyad, creating some ``weak-link'' indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the Marquez (2016) for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{landcontig =} \FunctionTok{ifelse}\NormalTok{(conttype }\SpecialCharTok{==} \DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cowmajdyad =} \FunctionTok{ifelse}\NormalTok{(cowmaj1 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{|}\NormalTok{ cowmaj2 }\SpecialCharTok{==} \DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# Create estimate of militarization as milper/tpop}
  \CommentTok{\# Then make a weak{-}link}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{milit1 =}\NormalTok{ milper1}\SpecialCharTok{/}\NormalTok{tpop1,}
         \AttributeTok{milit2 =}\NormalTok{ milper2}\SpecialCharTok{/}\NormalTok{tpop2,}
         \AttributeTok{minmilit =} \FunctionTok{ifelse}\NormalTok{(milit1 }\SpecialCharTok{\textgreater{}}\NormalTok{ milit2,}
\NormalTok{                           milit2, milit1)) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# create CINC proportion (lower over higher)}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cincprop =} \FunctionTok{ifelse}\NormalTok{(cinc1 }\SpecialCharTok{\textgreater{}}\NormalTok{ cinc2,}
\NormalTok{                           cinc2}\SpecialCharTok{/}\NormalTok{cinc1, cinc1}\SpecialCharTok{/}\NormalTok{cinc2)) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# create weak{-}link specification using Quick UDS data}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{mindemest =} \FunctionTok{ifelse}\NormalTok{(xm\_qudsest1 }\SpecialCharTok{\textgreater{}}\NormalTok{ xm\_qudsest2,}
\NormalTok{                            xm\_qudsest2, xm\_qudsest1)) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# Create "weak{-}link" measure of jointly advanced economies}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{minwbgdppc =} \FunctionTok{ifelse}\NormalTok{(wbgdppc2011est1 }\SpecialCharTok{\textgreater{}}\NormalTok{ wbgdppc2011est2,}
\NormalTok{                             wbgdppc2011est2, wbgdppc2011est1)) }\OtherTok{{-}\textgreater{}}\NormalTok{ Data}
\end{Highlighting}
\end{Shaded}

Table \ref{tab:tab-dd} is a formatted version of the results of a logistic regression model of conflict onset using these ``dangerous dyads'' indicators. Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but \texttt{\{peacesciencer\}} greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.

\begin{table}

\caption{\label{tab:tab-dd}A "Dangerous Dyads" Analysis of Non-Directed Dyad-Years from \{peacesciencer\}}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{8cm}c}
\toprule
\textbf{ } & \textbf{Model 1}\\
\midrule
Land Contiguity & 1.044*\\
 & (0.057)\\
Dyadic CINC Proportion (Lower/Higher) & 0.444*\\
 & (0.036)\\
CoW Major Power in Dyad & 0.145*\\
 & \vphantom{1} (0.058)\\
Defense Pact & -0.102+\\
 & (0.058)\\
Dyadic Democracy (Weak-Link) & -0.509*\\
 & (0.052)\\
Dyadic GDP per Capita (Weak-Link) & 0.262*\\
 & (0.050)\\
Dyadic Militarization (Minimum) & 0.267*\\
 & (0.023)\\
t & -0.148*\\
 & (0.005)\\
t\textasciicircum{}2 & 0.002*\\
 & \vphantom{1} (0.000)\\
t\textasciicircum{}3 & 0.000*\\
 & (0.000)\\
Intercept & -3.055*\\
 & (0.064)\\
\midrule
Num.Obs. & 106937\\
\bottomrule
\multicolumn{2}{l}{\textsuperscript{} + p $<$ 0.1, * p $<$ 0.05}\\
\end{tabular}
\end{table}

\texttt{\{peacesciencer\}} is also capable of creating data for replications of standard civil conflict analyses. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's (2003) well-cited analysis of civil conflict onset using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with \texttt{create\_stateyears(system\ =\ \textquotesingle{}gw\textquotesingle{})}, creating the full universe of Gleditsch-Ward state years from 1816 to 2020. Whereas the UCDP data included in \texttt{\{peacesciencer\}} are from 1946 to 2019, we can subset (``filter'') the data to just those years with \texttt{filter(year\ \%in\%\ c(1946,\ 2019))}. Next, we can use the \texttt{add\_ucdp\_acd()} function to return information about ongoing UCDP conflicts and onsets for these states. \texttt{add\_ucdp\_acd()} takes three arguments: \texttt{type}, \texttt{issue}, and \texttt{only\_wars}. \texttt{type} is an optional argument for the type of armed conflicts for which the researcher wants information. Options include ``extrasystemic,'' ``interstate,'' ``intrastate,'' and ``II'' (short for ``internationalized intrastate''). If no \texttt{type} is specified, the function returns information about ongoing disputes and onsets for all states for all tyeps of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g.~\texttt{type\ =\ c("intrastate",\ "II")}). \texttt{issue} is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict. Options include ``territory,'' ``government,'' and ``both.'' If no \texttt{issue} is specified, the function returns information for all conflicts irregarding the particular issue. \texttt{only\_wars} is an argument that subsets the data to just those with the intensity levels of ``war'' when \texttt{only\_wars\ =\ TRUE}. The argument defaults to \texttt{FALSE}, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, \texttt{add\_ucdp\_acd(type\ =\ "intrastate",\ only\_wars\ =\ FALSE))} returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.\footnote{The function also returns some background information of interest to the researcher, including the maximum intensity observed and the IDs associated with all ongoing conflicts in the state that year.}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{create\_stateyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(year }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1946}\SpecialCharTok{:}\DecValTok{2019}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_ucdp\_acd}\NormalTok{(}\AttributeTok{type=}\StringTok{"intrastate"}\NormalTok{, }\AttributeTok{only\_wars =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_peace\_years}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_democracy}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_creg\_fractionalization}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_sdp\_gdp}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_rugged\_terrain}\NormalTok{() }\OtherTok{{-}\textgreater{}}\NormalTok{ Data}
\end{Highlighting}
\end{Shaded}

Finally, we can add some covariates of interest to these data. \texttt{add\_peace\_years()} calculates peace spells between ongoing conflicts in the data generated by \texttt{add\_ucdp\_acd()}. \texttt{add\_democracy()} adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' (2016) extension of Pemstein et al's (2010) Unified Democracy Scores). \texttt{add\_creg\_fractionalization()} adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois. \texttt{add\_sdp\_gdp()} will add information about a state's estimated GDP, population, and GDP per capita from the Anders, Fariss, and Markowitz (2020) simulations. Finally, \texttt{add\_rugged\_terrain()} provide two estimates of the ruggedness of a state's terrain. The first is the terrain ruggedness index calculated by Nunn and Puga (2012) and the second is the Gibler and Miller (2014) extension of the natural logged percentage of the state that is mountainous (originally calculated by Fearon and Laitin (2003)). At the end of the pipe, the data returned by \texttt{\{peacesciencer\}} is assigned to an object minimally called \texttt{Data}.

\texttt{\{peacesciencer\}}'s tight integration with the \texttt{\{tidyverse\}} permits wide flexibility for the researcher. For example, assume the researcher wants to discern the estimated effect of the same set of covariates on intra-state conflicts at the threshold of war and those intra-state conflicts at or below the threshold of war. The first call included all conflicts with at least 25 deaths, per the UCDP's inclusion rules, and the peace years were calculated for those as well. If the researcher wants a new set of conflicts with a new set of peace years, it would be a simple matter of repeating the pipe-based workflow, but altering the argument in \texttt{add\_ucdp\_acd()} to be \texttt{only\_wars\ =\ TRUE}. \texttt{\{peacesciencer\}} would then calculate the peace years for those (\texttt{add\_peace\_years()}). To avoid confusion with the overlapping column names, the researcher can use some \texttt{\{tidyverse\}} verbs to rename all those conflict variables to have a distinct prefix of \texttt{war\_} (i.e.~\texttt{rename\_at(vars(ucdpongoing:ucdpspell),\ \textasciitilde{}paste0("war\_",\ .))}) before finally joining these data into the master data frame (i.e.~\texttt{left\_join(Data,\ .)\ -\textgreater{}\ Data}). Table \ref{tab:tab-cw} shows the fruits of the data \texttt{\{peacesciencer\}} generated after some post-processing for lagging important variables.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{create\_stateyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(year }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1946}\SpecialCharTok{:}\DecValTok{2019}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_ucdp\_acd}\NormalTok{(}\AttributeTok{type=}\StringTok{"intrastate"}\NormalTok{, }\AttributeTok{only\_wars =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_peace\_years}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rename\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(ucdpongoing}\SpecialCharTok{:}\NormalTok{ucdpspell), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{paste0}\NormalTok{(}\StringTok{"war\_"}\NormalTok{, .)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{left\_join}\NormalTok{(Data, .) }\OtherTok{{-}\textgreater{}}\NormalTok{ Data}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:tab-cw}A Civil Conflict Analysis of Gleditsch-Ward State-Years in \{peacesciencer\}}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{8cm}cc}
\toprule
\textbf{ } & \textbf{All UCDP Conflicts} & \textbf{Wars Only}\\
\midrule
Population Size (Lagged) & 0.229* & 0.272*\\
 & (0.067) & (0.106)\\
Extended UDS Democracy Score (Lagged) & 0.257 & -0.085\\
 & (0.181) & (0.270)\\
Extended UDS Democracy Score\textasciicircum{}2 (Lagged) & -0.726* & -0.761*\\
 & (0.211) & (0.352)\\
\% Mountainous Terrain (Logged) & 0.055 & 0.342*\\
 & (0.067) & (0.112)\\
Ethnic Fractionalization & 0.442 & 0.333\\
 & (0.358) & (0.554)\\
Religious Fractionalization & -0.389 & -0.281\\
 & (0.402) & (0.593)\\
t & -0.074+ & -0.111*\\
 & (0.039) & (0.056)\\
t\textasciicircum{}2 & 0.004* & 0.005+\\
 & (0.002) & (0.003)\\
t\textasciicircum{}3 & 0.000* & 0.000+\\
 & (0.000) & (0.000)\\
Intercept & -5.097* & -6.590*\\
 & (1.351) & (2.084)\\
\midrule
Num.Obs. & 8192 & 8192\\
\bottomrule
\multicolumn{3}{l}{\textsuperscript{} + p $<$ 0.1, * p $<$ 0.05}\\
\end{tabular}
\end{table}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\texttt{\{peacesciencer\}} is already more than capable of creating the kind of data in high demand in peace science. It can create dyad-year and state-year data. It is also generalizable to the dispute data included in the package, allowing for merging into \emph{dispute-year} data as well. This feature showed how it can effectively approximate two types of analyses in wide use in the peace science community. Surely researchers can and will add more information to these simple analyses after using \texttt{\{peacesciencer\}}, but the package already does a lot of the tedious work for researchers. It also does this in a maximally transparent way that conforms well to the DA-RT initiative across all political science.

This is not to say \texttt{\{peacesciencer\}} does everything, but \texttt{\{peacesciencer\}} can only evolve and expand on what it already does well. For example, \texttt{\{peacesciencer\}} does not right now handle \emph{k}-adic data (c.f. Poast 2010). However, the nature of \emph{k}-ads do not make for necessarily problematic data merges and should not be difficult to add for researchers who want it in future updates. Likewise, \texttt{\{peacesciencer\}} contains the Archigos leader data (Goemans, Gleditsch, and Chiozza 2009) and is able to transform that into interesting data at the dyad-level and state-level, but doing a leader-level analysis with information about leader willingness to use force (e.g. J. Carter and Smith 2020) is not right now available. However, these are possible expansions on what \texttt{\{peacesciencer\}} already does well. Thus, this package can only evolve to meet new analytical demands for the peace science community.

Finally, a skeptical reader should not think that making the process as simple as possible necessarily facilitates poor decision-making by the user. In cases where it is evident what the user wants (e.g.~an estimate of the level of democracy in the state-year), \texttt{\{peacesciencer\}} does the necessary work to provide the user that information. However, the package makes sure to leave important decision-making to the researcher. For example, \texttt{add\_cow\_alliance()} returns information about various types of alliance pledges in the dyad-year---should one exist---but leaves it to the researcher to say whether they want to define the presence of an alliance to be just a defense pledge or any type of alliance pledge. \texttt{add\_contiguity()} returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity threshold. The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g.~the contiguity data are not ordinal and should not be treated as such), how and where data differ (e.g.~how CoW state system data differ from Gleditsch-Ward data), and how researchers should consider optimally using its functionality (e.g.~\texttt{add\_ucdp\_acd()} probably should not lump all forms of conflict together). \texttt{\{peacesciencer\}} does not endeavor to make researchers lazy or sloppy, and it does not ultimately do this. Instead, \texttt{\{peacesciencer\}} reduces the tedium associated with starting quantitative peace science research, achieving this in a quick, robust, and transparent way.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parskip}{6pt}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-andersetal2020bbgb}{}%
Anders, Therese, Christopher J. Fariss, and Jonathan N. Markowitz. 2020. {``Bread Before Guns or Butter: Introducing Surplus Domestic Product (SDP).''} \emph{International Studies Quarterly} 64 (2): 392--405.

\leavevmode\hypertarget{ref-arelbundock2021m}{}%
Arel-Bundock, Vincent. 2021a. \emph{Modelsummary: Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready}. \url{https://CRAN.R-project.org/package=modelsummary}.

\leavevmode\hypertarget{ref-arelbundock2021w}{}%
---------. 2021b. \emph{WDI: World Development Indicators and Other World Bank Data}. \url{https://CRAN.R-project.org/package=WDI}.

\leavevmode\hypertarget{ref-becketal1998tts}{}%
Beck, Nathaniel, Jonathan N. Katz, and Richard Tucker. 1998. {``Taking Time Seriously: Time-Series-Cross-Section Analysis with a Binary Dependent Variable.''} \emph{American Journal of Political Science} 42 (4): 1260--88.

\leavevmode\hypertarget{ref-bennettetal2019n}{}%
Bennett, D. Scott, Paul Poast, and Allan C. Stam. 2019. {``NewGene: An Introduction for Users.''} \emph{Journal of Conflict Resolution} 63 (6): 1579--92.

\leavevmode\hypertarget{ref-bennettstam2000e}{}%
Bennett, D. Scott, and Allan Stam. 2000. {``EUGene: A Conceptual Manual.''} \emph{International Interactions} 26 (2): 179--204.

\leavevmode\hypertarget{ref-boehmke2016dwr}{}%
Boehmke, Bradley C. 2016. \emph{Data Wrangling with {R}}. Springer.

\leavevmode\hypertarget{ref-bowersvoors2016hiyr}{}%
Bowers, Jake, and Maarten Voors. 2016. {``How to Improve Your Relationship with Your Future Self.''} \emph{Revista de Ciencia Politica} 36 (3): 829--48.

\leavevmode\hypertarget{ref-bremer1992ddc}{}%
Bremer, Stuart A. 1992. {``Dangerous Dyads: Conditions Affecting the Likelihood of Interstate War, 1816-1965.''} \emph{Journal of Conflict Resolution} 36 (2): 309--41.

\leavevmode\hypertarget{ref-cartersignorino2010bf}{}%
Carter, David B., and Curtis S. Signorino. 2010. {``Back to the Future: Modeling Time Dependence in Binary Data.''} \emph{Political Analysis} 18 (3): 271--92.

\leavevmode\hypertarget{ref-cartersmith2020fml}{}%
Carter, Jeff, and Charles E. Smith. 2020. {``A Framework for Measuring Leaders' Willingness to Use Force.''} \emph{American Political Science Review} 114 (4): 1352--58.

\leavevmode\hypertarget{ref-fearonlaitin2003eicw}{}%
Fearon, James D., and David D. Laitin. 2003. {``Ethnicity, Insurgency, and Civil War.''} \emph{American Political Science Review} 97 (1): 75--90.

\leavevmode\hypertarget{ref-giblermiller2014etts}{}%
Gibler, Douglas M., and Steven V. Miller. 2014. {``External Territorial Threat, State Capacity, and Civil War.''} \emph{Journal of Peace Research} 51 (5): 634--46.

\leavevmode\hypertarget{ref-gibleretal2016amid}{}%
Gibler, Douglas M., Steven V. Miller, and Erin K. Little. 2016. {``An Analysis of the {M}ilitarized {I}nterstate {D}ispute {(MID)} Dataset, 1816-2001.''} \emph{International Studies Quarterly} 60 (4): 719--30.

\leavevmode\hypertarget{ref-goemansetal2009ia}{}%
Goemans, Henk E., Kristian Skrede Gleditsch, and Giacomo Chiozza. 2009. {``Introducing {A}rchigos: A Dataset on Political Leaders.''} \emph{Journal of Peace Research} 46 (2): 269--83.

\leavevmode\hypertarget{ref-lemkereed2001rprd}{}%
Lemke, Douglas, and William Reed. 2001. {``The Relevance of Politically Relevant Dyads.''} \emph{Journal of Conflict Resolution} 45 (1): 126--44.

\leavevmode\hypertarget{ref-marquez2016qme}{}%
Marquez, Xavier. 2016. {``A Quick Method for Extending the {U}nified {D}emocracy {S}cores.''}

\leavevmode\hypertarget{ref-nunnpuga2012r}{}%
Nunn, Nathan, and Diego Puga. 2012. {``Ruggedness: The Blessing of Bad Geography in {A}frica.''} \emph{Review of Economics and Statistics} 94 (1): 20--36.

\leavevmode\hypertarget{ref-pemsteinetal2010dc}{}%
Pemstein, Daniel, Stephen A. Meserve, and James Melton. 2010. {``Democratic Compromise: A Latent Variable Analysis of Ten Measures of Regime Type.''} \emph{Political Analysis} 18 (4): 426--49.

\leavevmode\hypertarget{ref-poast2010mdd}{}%
Poast, Paul. 2010. {``(Mis)using Dyadic Data to Analyze Multilateral Events.''} \emph{Political Analysis} 18 (4): 403--25.

\leavevmode\hypertarget{ref-wickhamgrolemund2017rds}{}%
Wickham, Hadley, and Garrett Grolemund. 2017. \emph{{R} for Data Science}. Sebastopol, CA: O'Reilly Media, Inc.

\end{CSLReferences}

\end{document}

### 2021-06-10 user-guide/ms.Rmd
let 'er rip

---
title: "`{peacesciencer}`:  A User’s  Guide  for  Quantitative  Peace Science in R"
thanks: "Replication files are available on the author's Github account (https://github.com/svmiller/peacesciencer). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
author:
- name: Steven V. Miller
  affiliation: Clemson University
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`" # This would be more general
# bibliography: inst/refs.bib
abstract: "This article introduces `{peacesciencer}`, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year or state-year analysis (e.g. democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at http://svmiller.com/peacesciencer."
keywords: "software, statistical analysis, peace science, reproducibility"
geometry: margin=1in
mainfont: cochineal
#sansfont: Linux Biolinum O
fontsize: 11pt
endnote: no
# pandocparas: TRUE
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
alwaays_allow_html: TRUE
header-includes:
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
  - \usepackage{colortbl}
  - \arrayrulecolor{black}
#  - \midrule\relax
---

```{r setup, include=FALSE}
is_docx <- knitr::pandoc_to("docx")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# I don't know how Texas Instruments smart this is, but p-sure default DPI is 96.
# That's not a problem for LaTeX, but it looks not-so-great for Word.
# For Word, let's up that to 600. This should allow for cross-referencing 
# in {bookdown} while allowing for conditional DPI

if (is_latex) {
  conditional_dpi <- 96
  return(conditional_dpi)
} else {
  conditional_dpi <- 600
  return(conditional_dpi)
}

knitr::opts_chunk$set(echo = FALSE, cache = FALSE,
                      message=FALSE, warning=FALSE,
                      fig.path='doc/figs/',
                      cache.path = 'doc/_cache/',
                      fig.width = 8.5, dpi = conditional_dpi,
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      })

# library(dplyr)
# library(magrittr)
library(tidyverse)
library(stevemisc)
library(modelsummary)
library(huxtable)
library(kableExtra)
library(peacesciencer)

allrds <- list.files(pattern = ".rds", path = "data") %>%
  stringr::str_remove(., ".rds")

for(i in allrds) {
      filepath <- file.path(paste0("data/",i,".rds"))
      assign(i, readRDS(filepath))
    }

```

# Introduction

This data feature tackles a recurring problem for researchers in the peace science community. True research reproducibility is achieved creating original dyad-year or state-year data from scratch, though no published guide exists that informs how to do this on their own. Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice. In a related manner, researchers may spend too much time reproducing old code for standard information that goes into any dyad-year or state-year analysis---like contiguity relationships and democracy, among many others---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes. This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well. Graduate students, in particular, face unique challenges associated with these developments. Graduate students beginning their careers in peace science to learn how scholarship informs data in peace science and  how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.

`{peacesciencer}` is designed to address these problems. Built around the free and open source R programming language, `{peacesciencer}` contains a suite of data and functions for creating data of interest to researchers. Researchers can use `{peacesciencer}` to create dyad-year, state-year, and even state-day data from scratch. Afterward, they can add a variety of standard information (e.g. contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. `{peacesciencer}` comes with some data innovations as well, including a comprehensive data set on democracy, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts. All are done with the maximum possible transparency. The project is available for public view of Github (https://github.com/svmiller/peacesciencer). The `data-raw` directory on the project's Github contains information and comments about how every data set was created and the `R` directory informs the user how each function works. The function manuals (http://svmiller.com/peacesciencer/reference/) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, `{peacesciencer}` not only assists a peace scientist with their research, but it does so in a manner that best conforms to the  Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

This data feature proceeds in the following fashion. The next section expands what need this package fills for peace scientists in the present. Afterward, it provides an overview of what is included in `{peacesciencer}` to help researchers more quickly conduct the kind of quantitative research they want. Thereafter, it provides a tutorial on how to install and best use `{peacesciencer}` in the R programming language. A more comprehensive tutorial follows, showing how `{peacesciencer}` already has a suite of data and functions that can allow for effective replications of a "dangerous dyads" type analysis [@bremer1992ddc] and standard state-year analyses of civil conflict onset [e.g. @fearonlaitin2003eicw]. This feature concludes with a discussion of what else `{peacesciencer}` is capable of doing in future updates and how this R package can inform more reasoned design decisions for researchers in peace science.

# Why `{peacesciencer}`?

`{peacesciencer}` is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch. NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars [@bennettetal2019n]. NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade [@bennettstam2000e]. No matter, `{peacesciencer}` is motivated by the following ideals that led to its creation.

For one, researchers may invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume, for example, a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's primary impetus was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g. dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit. `{peacesciencer}` is primarily born from this question about how to create dyad-year and state-year data from scratch. The underlying code that produces these data types is available online and are converted into simple functions for the ease of the researcher.

Second, researchers also invest too much time in retracing steps for peace science analyses for new projects. Assume, for example, a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data). Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project. They may then have to change the name of some files, change some code to account for potentially new column names in the newer data, and troubleshoot instances where their old code does not perform as it once did. At its worst, this process may lead to some errors by the researcher. At best, this is tedium that spends the researcher's time they would rather invest in analyzing the data. The lion's share of `{peacesciencer}`'s functionality is both creating the units of analysis for the researcher and merging in different forms of data in wide use in the peace science community so that the researcher can spend less of their time on tedium.

Third, the creation of the data and the analysis of the data is increasingly becoming one continuous process. Not too long ago, it used to be the case that researchers had to download a data set, or create one from scratch (possibly in a spreadsheet or through a program like EUGene). After downloading or constructing the data, the researcher then opened a specialty program for statistical analysis (e.g. SAS, SPSS, Stata) to recode raw data into a form suitable for analysis before running a statistical model that regresses some outcome on a set of covariates. Current research practices still resemble this process, but the steps between them are no longer as large as they were in the past. Software options exist that allow the researcher to load data, create data, clean data, analyze data, and present the results of the analysis all within one program. `{peacesciencer}`, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and Rstudio (its free-for-use integrated desktop environment [IDE]).

<!-- [^arelbundock] -->

<!-- [^arelbundock]: For example, Vincent Arel-Bundock has a set of tools that allow a researcher to get World Bank data from the internet [@arelbundock2021w], sync World Bank codes to Correlates of War codes [@arelbundocketal2018c], and present statistical models of those data [@arelbundock2021m]. -->

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer [c.f. @bowersvoors2016hiyr]. This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that `{peacesciencer}` uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet [@arelbundock2021w], scrape *any* type of data from the internet [@boehmke2016dwr], and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero [@arelbundock2021m]. `{peacesciencer}` embraces this. This R programming language package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

Fifth, statistical software should be kind to graduate students in peace science with the aforementioned point in mind. Graduate students just beginning their career in peace science scholarship are tasked to learn the substantive material, the characteristics of the underlying data in wide use in the community, and the quantitative methods and statistical software available to them, effectively all at the same time. A student interested in civil conflict scholarship will need to learn the UCDP-PRIO armed conflict data in some detail at the same time they learn about the theoretical arguments that informed the creation of the data, likely at the same time they take quantitative methods courses that show them how to load and explore data in a choseen statistical software. At the same time that creating data is increasingly blurred with coding and processing data, students are simultaneously learning to explore data for themselves as they read about the analyses of the data. `{peacesciencer}` aims to make this process easier for graduate students using the features of the R programming language. Graduate students can use `{peacesciencer}` to learn about peace science data and computer programming in R at the same time.

Finally, the creation and presentation of data in peace science should be 100% transparent, which `{peacesciencer}` takes seriously in the following ways. The website for `{peacesciencer}` has several vignettes that describe its processes in some detail. These include how it provides reasonable estimates of democracy that may not be available in the Polity data or the Varieties of Democracy data and how a researcher can whittle dyadic dispute-years into true dyad-years through reasonable case exclusions. The project's Github contains a publicly available `data-raw` directory that shows how every data set included (and processed) in `{peacesciencer}` was created. The function manuals included `{peacesciencer}` contain ample documentation that clarifies what the function is doing, what it returns to the user, and why it is doing it this way. Users can also use the project's Github to point out bugs, ask for further clarification, and propose additions. `{peacesciencer}` takes seriously the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science and endeavors for maximum transparency, leveraging open source and version control software to inform users of what data it uses and how it uses the data.

# What is Included in `{peacesciencer}`

`{peacesciencer}` comes with a battery of built-in functions for generating the most widespread forms of peace science data. Each function uses raw or pre-processed data included in `{peacesciencer}`. For example, `create_dyadyears()` transforms the raw state system membership data for either the Correlates of War or Gleditsch-Ward system into the full universe of post-1816 dyad-years. `add_gml_mids()` uses a dyadic dispute-year version of the MID data offered by @gibleretal2016amid, which, as the `data-raw` directory on the project's Github shows, is itself derived from the dispute-level and participant-level versions of the data. `{peacesciencer}` also has some data innovations. It has an original data set on capital-to-capital distances, calculating for distance in kilometers between state capitals using the "Vincenty" method (i.e. "as the crow flies") that also accounts for instances when capitals moved (e.g. Brazil in 1960, Burundi in 2018). It also features an innovation of democracy data, providing reasonable estimates of democracy using the @marquez2016qme of extending the Unified Democracy Scores (UDS) data [@pemsteinetal2010dc].[^demvig] As of writing, `{peacesciencer}` comes with the following functionality.

[^demvig]: A vignette on the `{peacesciencer}` website explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.

```{r, echo=F, eval=T}
tribble(~data, ~description, ~type, ~func,
        "(CoW, G-W) Dyad-Year Data", "Create dyad-years from state system membership data", "", "create_dyadyears()",
        "(CoW, G-W) State-Year Data", "Create state-years from state system membership data", "", "create_stateyears()",
        "(CoW, G-W) State-Day Data", "Create state-days from state system membership data", "", "create_statedays()",
        "Archigos (v. 4.1)", "Add Archigos political leader information", "D, S", "add_archigos()",
        "ATOP Alliances (v. 5.0)", "Add ATOP alliance data", "D", "add_atop_alliance()",
        "Capital-to-Capital Distances", "Add capital-to-capital distance", "D, S", "add_capital_distance()",
        "Convert CoW codes to G-W codes", "Add matching CoW codes for G-W states by year", "D, S", "add_ccode_to_gw()",
        "Convert G-W codes to CoW codes", "Add matching G-W codes for CoW states by year", "D, S", "add_gwcode_to_cow()",
        "CoW National Capabilities (v. 5.0)", "Adds capabilities of states by year", "D, S", "add_nmc()",
        "CoW Direct Contiguity (v. 3.2)", "Adds contiguity information by year", "D, S", "add_contiguity()",
        "CoW Alliances (v. 4.1)", "Adds CoW alliance data by year", "D", "add_cow_alliance()",
        "CoW IGOs (v. 3.0)", "Adds CoW IGO data by year", "D, S", "add_igos()",
        "CoW Major Powers (v. 2016)", "Adds CoW major power by year", "D, S", "add_cow_majors()",
        "CoW MIDs (v. 5.0)", "Add CoW MID information to data", "D", "add_cow_mids()",
        "CoW Trade Data (v. 4.0)", "Add CoW trade information to data", "D, S", "add_cow_trade()",
        "CoW Wars", "Add CoW inter-state (v. 4.0) and intra-state (v. 4.1) war data", "D", "add_cow_wars()",
        "CREG (v. 1.02)", "Add ethnic/religious fractionalization/polarization by year", "D, S", "add_creg_fractionalization()",
        "Democracy", "Add Polity (v. 2017), V-Dem (v. 10), and Marquez' (2016) UDS extensions", "D,S", "add_democracy()",
        "GML MIDs (v. 2.1.1)", "Add Gibler-Miller-Little MIDs to data", "D", "add_gml_mids()",
        "Minimum Distance (v. 2016)", "Adds minimum distance data by year", "D, S", "add_minimum_distance()",
        "Peace Years", "Calculate peace years for various disputes", "D, S", "add_peace_years()",
        "Rugged/Mountainous Terrain", "Add 'ruggedness'/'mountainous' terrain data by year", "D, S", "add_rugged_terrain()",
        "(Surplus/Gross) Domestic Product", "Add Anders et al. (2020) GDP/SDP simulations by year", "D, S", "add_sdp_gdp()",
        "Strategic Rivalries", "Add Thompson and Dreyer (2012) strategic rivalry data by year", "D, S", "add_strategic_rivalries()",
        "UCDP Armed Conflicts", "Add UCDP armed conflict data by year", "S", "add_ucdp_acd()"
        ) -> included

# tribble(~data, ~description, ~type, ~func,
#         "(CoW, G-W)", "Create dyad-years", "", "stuff") -> included

if (is_latex) {
kable(included, booktabs = FALSE, longtable = TRUE,
    caption = "Features of \\{peacesciencer\\}",
    align = c("l","l","c", "c"),
    col.names = c("Data", "Description", "Type", "Function")) %>%
  kable_styling(font_size = 7) %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(4, monospace = TRUE) %>%
  column_spec(c(1,2,3), border_left = FALSE) %>%
  row_spec((head(c(1:nrow(included)), -1)), extra_latex_after = "\\arrayrulecolor{white}") %>%
  row_spec((included %>% 
             nrow()), extra_latex_after = "\\arrayrulecolor{black}")
} else {
  kable(included,
    caption = "Features of \\{peacesciencer\\}",
    align = c("l","l","c", "c"),
    col.names = c("Data", "Description", "Type", "Function")) %>%
  # kable_styling(font_size = 7) %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(4, monospace = TRUE) 
}

```

`{peacesciencer}`'s coverage focuses mostly on data that are released as standalone data sets for download, especially those in the Correlates of War or Gleditsch-Ward ecosystem of data. Data that can be obtained from a stable advanced programming interface---like the World Bank, for example---can be obtained through those other means [e.g. @arelbundock2021w]. Its coverage will assuredly expand with new additions of interest to the peace science community.

# How to Install `{peacesciencer}`

`{peacesciencer}` is a package for the R programming language. This assumes at least some familiarity with the R programming language. Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.0. `{peacesciencer}` is designed to be as user-friendly as possible. Those proficient in R, those just learning R, and those with no experience in R should be able to pick up its use fairly quickly.

`{peacesciencer}`'s functions work out of the box, though it works best with one additional package.  This is `{tidyverse}`, itself a set of packages that share a common form and design [@wickhamgrolemund2017rds]. `{peacesciencer}` functions make considerable use of the component packages of `{tidyverse}` but installing and loading `{tidyverse}` will allow the researcher to make quicker use of `{peacesciencer}`'s functionality. A user can open an R session and install both packages as follows.

```r
install.packages(c("tidyverse", "peacesciencer"))
```

R packages once installed need to be loaded with every R session. The user can load both with the `library()` function in R.

```r
library(tidyverse)
library(peacesciencer)
```

Thereafter, a researcher can start using `{peacesciencer}` to create the kind of data they need.

# A Tutorial on How to Use `{peacesciencer}`

`{peacesciencer}` has a simple "pipe"-based workflow. The top of the "pipe" is the first function to create the data. These options include `create_dyadyears()`, `create_stateyears()`, and `create_statedays()`. All options include two arguments: `system` and `mry`. `system` takes one of two values---"cow" or "gw"---for what state system membership data the user wants. If `system` is not specified, the function defaults to Correlates of War system membership data given the prominence of the Correlates of War project in the conflict data ecosystem. `mry`---short for "most recent year"---determines whether or not the function extends the membership data to the most recently concluded calendar year. The function defaults to `TRUE` in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017). `create_dyadyears()` has a third argument---`directed`---that determines whether the user wants directed dyad-year data or non-directed dyad-year data. The function defaults to `TRUE`, returning directed dyad-year data.

The following example creates a full state-year data frame of Gleditsch-Ward states from 1816 to 2020.

```{r, echo=T, eval=F}
create_stateyears(system = 'gw')
```

This example would create a full non-directed dyad-year data frame of Correlates of War states from 1816 to 2016, the last year of record in the Correlates of War state system membership data.

```{r, echo=T, eval=F}
create_dyadyears(directed = FALSE, mry = FALSE)
```

This example would create a full state-day data frame from the Gleditsch-Ward system for all states from Jan. 1, 1816 to Dec. 31, 2020. 
```{r, echo=T, eval=F}
create_statedays(system = 'gw')
```

There are not too many data sets in the peace science community with daily data, certainly dating to the 19th century, but `{peacesciencer}` can greatly reduce the costs of creating the kind of data the user may want.

## Add Standard Information to Dyad-Year or State-Year Data

The next step in the `{peacesciencer}` "pipe"-based workflow will also be among the biggest time-savers for researchers using it. Researchers who want to understand, say, the correlates of inter-state conflict onset almost always do the following. They will add the dependent variables from an inter-state conflict data set, like the Gibler-Miller-Little (GML) dispute data set [@gibleretal2016amid]. They will add several indicators to the dyad-year data that are well-traveled in conflict scholarship, including indicators of things like joint alliance contracts and the contiguity relationship of both states in the dyad. `{peacesciencer}` has great coverage for a lot of these data and can even allow almost full-scale replications with the data it offers.

For example, assume a researcher wanted to replicate something analogous to Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis about the causes of conflict onset. `{peacesciencer}` can do something like this in the following pipe-based workflow. First, `create_dyadyears(directed = FALSE, mry = FALSE)` will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016. This is effectively the same function used above, but with two user-specified overrides.

After creating the base data they want, the user can add data to it with the "pipe" operator. The `{tidyverse}` represents this pipe operator as `%>%`. Ending each line with this pipe passes forward an object (here: data) into the next function on the next line. Put in other words, `create_dyadyears(directed = FALSE, mry = FALSE) %>% filter_prd()` is a command in `{tidyverse}` and `{peacesciencer}` to create all non-directed dyad-years from 1816 to 2016 and then subset those data to only the politically relevant dyad-years. Because politically relevant dyads depend on information about the major power status for both states as well as their contiguity relationship [@lemkereed2001rprd], the function adds in those data (i.e. quietly executing `add_contiguity()` and `add_cow_majors()`) before subsetting the data to just the dyads that have some kind of contiguity relationship and/or have a major power in it.

```{r, echo=T, eval=F}
create_dyadyears(directed = FALSE, mry = FALSE) %>%
  filter_prd() %>%
  add_gml_mids(keep = NULL) %>%
  add_peace_years() %>%
  add_nmc() %>%
  add_democracy() %>%
  add_cow_alliance() %>%
  add_sdp_gdp() -> Data
```

Next, the pipe-based workflow continues with `add_gml_mids(keep = NULL)`, which will add information about whether there was an ongoing dispute in the dyad (and whether it was a dispute onset) from the GML dispute data.[^keepnull] `add_peace_years()` will calculate peace years between ongoing disputes in these politically relevant non-directed dyads.[^onpy] `add_nmc()` will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity. `add_democracy()` will add information about the level of democracy for both states in the dyad using three different indicators. Researchers can select which one they prefer from these indicators. `add_cow_alliance()` will add information about whether there was an alliance in the dyad-year by way of whether ther was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge. Finally, `add_sdp_gdp()` will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by @andersetal2020bbgb.

[^keepnull]: `keep = NULL` is optional and removes supporting information about dyadic dispute-participants (e.g. dispute-level hostility, dispute-level outcome) in order to reduce the size of the data set.

[^onpy]: `add_peace_years()` will only calculate the peace years---titled, in this case, `gmlmidspell`---and will leave the temporal dependence adjustment to the taste of the researcher. Importantly, I do not recommend manually creating splines or square/cube terms because it creates more problems in adjusting for temporal dependence in model predictions. In a regression formula in R, you can specify the @cartersignorino2010bf approach as `... + gmlmidspell + I(gmlmidspell^2) + I(gmlmidspell^3)`. The @becketal1998tts cubic splines approach is `... + splines::bs(gmlmidspell, 4)`. This function includes the spell and three splines (hence the 4 in the command).

Whereas `add_sdp_gdp()` is the last command in the pipe-based workflow, the `{peacesciencer}` call ends by assigning to an object called `Data`. This type of assignment is done with the "right hand" assignment operator (i.e. `->`). If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like `haven::write_dta(Data, "my-data.dta")` and import it into Stata when they are done.

`{peacesciencer}` has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding `-9s` in these data to be `NA` in the National Material Capabilities data. There is only some slight data work to create the desired indicators for a statistical model of conflict onset. Here, briefly, these functions lean on `{tidyverse}` "verbs" and base R terminology to do things like creating a dummy variable for whether the dyad was land-contiguous, whether there was a major power on either side of the non-directed dyad, creating some "weak-link" indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the @marquez2016qme for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.

```{r, echo=T, eval=F}
Data %>%
  mutate(landcontig = ifelse(conttype == 1, 1, 0)) %>%
  mutate(cowmajdyad = ifelse(cowmaj1 == 1 | cowmaj2 == 1, 1, 0)) %>%
  # Create estimate of militarization as milper/tpop
  # Then make a weak-link
  mutate(milit1 = milper1/tpop1,
         milit2 = milper2/tpop2,
         minmilit = ifelse(milit1 > milit2,
                           milit2, milit1)) %>%
  # create CINC proportion (lower over higher)
  mutate(cincprop = ifelse(cinc1 > cinc2,
                           cinc2/cinc1, cinc1/cinc2)) %>%
  # create weak-link specification using Quick UDS data
  mutate(mindemest = ifelse(xm_qudsest1 > xm_qudsest2,
                            xm_qudsest2, xm_qudsest1)) %>%
  # Create "weak-link" measure of jointly advanced economies
  mutate(minwbgdppc = ifelse(wbgdppc2011est1 > wbgdppc2011est2,
                             wbgdppc2011est2, wbgdppc2011est1)) -> Data

```

Table \@ref(tab:tab-dd) is a formatted version of the results of a logistic regression model of conflict onset using these "dangerous dyads" indicators. Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but `{peacesciencer}` greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.

```{r tab-dd, eval=T}

table_format <- ifelse(is_docx, "huxtable", 'default')

modelsummary(modDD, output=table_format,
             title = 'A "Dangerous Dyads" Analysis of Non-Directed Dyad-Years from \\{peacesciencer\\}',
             stars = c('*' = .05, '+' = .1), gof_omit = "IC|F|Log.|R2$",
             coef_map = c("landcontig" = "Land Contiguity", 
                          "cincprop" = "Dyadic CINC Proportion (Lower/Higher)", 
                          "cowmajdyad" = "CoW Major Power in Dyad",
                          "cow_defense" = "Defense Pact", 
                          "mindemest" = "Dyadic Democracy (Weak-Link)", 
                          "minwbgdppc" = "Dyadic GDP per Capita (Weak-Link)", 
                          "minmilit" = "Dyadic Militarization (Minimum)",
                          "gmlmidspell" = "t",
                          "I(gmlmidspell^2)" = "t^2",
                          "I(gmlmidspell^3)" = "t^3",
                          "(Intercept)" = "Intercept"),
             align=c("lc")) -> tab_modDD

if (!is_docx) {
  tab_modDD %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(1, width = "8cm")
} else {
  tab_modDD %>%
  set_font_size(10) %>%
  set_left_padding(1) %>%
  set_top_padding(1) %>%
  set_bottom_padding(1) %>%
  set_right_padding(1) %>%
  set_bold(., 1, 1:2, TRUE) %>%
  as_flextable()
} 
```

`{peacesciencer}` is also capable of creating data for replications of standard civil conflict analyses. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's [-@fearonlaitin2003eicw] well-cited analysis of civil conflict onset using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with `create_stateyears(system = 'gw')`, creating the full universe of Gleditsch-Ward state years from 1816 to 2020. Whereas the UCDP data included in `{peacesciencer}` are from 1946 to 2019, we can subset ("filter") the data to just those years with `filter(year %in% c(1946, 2019))`. Next, we can use the `add_ucdp_acd()` function to return information about ongoing UCDP conflicts and onsets for these states. `add_ucdp_acd()` takes three arguments: `type`, `issue`, and `only_wars`. `type` is an optional argument for the type of armed conflicts for which the researcher wants information. Options include "extrasystemic", "interstate", "intrastate", and "II" (short for "internationalized intrastate"). If no `type` is specified, the function returns information about ongoing disputes and onsets for all states for all tyeps of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g. `type = c("intrastate", "II")`). `issue` is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict. Options include "territory", "government", and "both". If no `issue` is specified, the function returns information for all conflicts irregarding the particular issue. `only_wars` is an argument that subsets the data to just those with the intensity levels of "war" when `only_wars = TRUE`. The argument defaults to `FALSE`, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, `add_ucdp_acd(type = "intrastate", only_wars = FALSE))` returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.[^adducdpacdalso]

[^adducdpacdalso]: The function also returns some background information of interest to the researcher, including the maximum intensity observed and the IDs associated with all ongoing conflicts in the state that year.

```{r, echo=T, eval=F}
create_stateyears(system = 'gw') %>%
  filter(year %in% c(1946:2019)) %>%
  add_ucdp_acd(type="intrastate", only_wars = FALSE) %>%
  add_peace_years() %>%
  add_democracy() %>%
  add_creg_fractionalization() %>%
  add_sdp_gdp() %>%
  add_rugged_terrain() -> Data
```

Finally, we can add some covariates of interest to these data. `add_peace_years()` calculates peace spells between ongoing conflicts in the data generated by `add_ucdp_acd()`. `add_democracy()` adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' [-@marquez2016qme] extension of Pemstein et al's [-@pemsteinetal2010dc] Unified Democracy Scores). `add_creg_fractionalization()` adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois. `add_sdp_gdp()` will add information about a state's estimated GDP, population, and GDP per capita from the @andersetal2020bbgb simulations. Finally, `add_rugged_terrain()` provide two estimates of the ruggedness of a state's terrain. The first is the terrain ruggedness index calculated by @nunnpuga2012r and the second is the @giblermiller2014etts extension of the natural logged percentage of the state that is mountainous (originally calculated by @fearonlaitin2003eicw). At the end of the pipe, the data returned by `{peacesciencer}` is assigned to an object minimally called `Data`.

`{peacesciencer}`'s tight integration with the `{tidyverse}` permits wide flexibility for the researcher. For example, assume the researcher wants to discern the estimated effect of the same set of covariates on intra-state conflicts at the threshold of war and those intra-state conflicts at or below the threshold of war. The first call included all conflicts with at least 25 deaths, per the UCDP's inclusion rules, and the peace years were calculated for those as well. If the researcher wants a new set of conflicts with a new set of peace years, it would be a simple matter of repeating the pipe-based workflow, but altering the argument in `add_ucdp_acd()` to be `only_wars = TRUE`. `{peacesciencer}` would then calculate the peace years for those (`add_peace_years()`). To avoid confusion with the overlapping column names, the researcher can use some `{tidyverse}` verbs to rename all those conflict variables to have a distinct prefix of `war_` (i.e. `rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .))`) before finally joining these data into the master data frame (i.e. `left_join(Data, .) -> Data`).  Table \@ref(tab:tab-cw) shows the fruits of the data `{peacesciencer}` generated after some post-processing for lagging important variables.

```{r, echo=T, eval=F}
create_stateyears(system = 'gw') %>%
  filter(year %in% c(1946:2019)) %>%
  add_ucdp_acd(type="intrastate", only_wars = TRUE) %>%
  add_peace_years() %>%
  rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .)) %>%
  left_join(Data, .) -> Data
```

```{r tab-cw, eval=T}

table_format <- ifelse(is_docx, "huxtable", 'default')

modelsummary(modCW, output=table_format,
             title = 'A Civil Conflict Analysis of Gleditsch-Ward State-Years in \\{peacesciencer\\}',
             stars = c('*' = .05, '+' = .1), gof_omit = "IC|F|Log.|R2$",
             coef_map = c("l1_wbgdppc2011est	" = "GDP per Capita (Lagged)", 
                          "l1_wbpopest" = "Population Size (Lagged)", 
                          "l1_xm_qudsest" = "Extended UDS Democracy Score (Lagged)",
                          "I(l1_xm_qudsest^2)" = "Extended UDS Democracy Score^2 (Lagged)", 
                          "newlmtnest" = "% Mountainous Terrain (Logged)", 
                          "ethfrac" = "Ethnic Fractionalization", 
                          "relfrac" = "Religious Fractionalization",
                          "ucdpspell" = "t",
                          "I(ucdpspell^2)" = "t^2",
                          "I(ucdpspell^3)" = "t^3",
                          "war_ucdpspell" = "t",
                          "I(war_ucdpspell^2)" = "t^2",
                          "I(war_ucdpspell^3)" = "t^3",
                          "(Intercept)" = "Intercept"),
             align=c("lcc")) -> tab_modCW

if (!is_docx) {
  tab_modCW %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(1, width = "8cm")
} else {
  tab_modCW %>%
  set_font_size(10) %>%
  set_left_padding(1) %>%
  set_top_padding(1) %>%
  set_bottom_padding(1) %>%
  set_right_padding(1) %>%
  set_bold(., 1, 1:2, TRUE) %>%
  as_flextable()
} 
```

# Conclusion

`{peacesciencer}` is already more than capable of creating the kind of data in high demand in peace science. It can create dyad-year and state-year data. It is also generalizable to the dispute data included in the package, allowing for merging into *dispute-year* data as well. This feature showed how it can effectively approximate two types of analyses in wide use in the peace science community. Surely researchers can and will add more information to these simple analyses after using `{peacesciencer}`, but the package already does a lot of the tedious work for researchers. It also does this in a maximally transparent way that conforms well to the DA-RT initiative across all political science.

This is not to say `{peacesciencer}` does everything, but `{peacesciencer}` can only evolve and expand on what it already does well. For example, `{peacesciencer}` does not right now handle *k*-adic data [c.f. @poast2010mdd]. However, the nature of *k*-ads do not make for necessarily problematic data merges and should not be difficult to add for researchers who want it in future updates. Likewise, `{peacesciencer}` contains the Archigos leader data [@goemansetal2009ia] and is able to transform that into interesting data at the dyad-level and state-level, but doing a leader-level analysis with information about leader willingness to use force [e.g. @cartersmith2020fml] is not right now available. However, these are possible expansions on what `{peacesciencer}` already does well. Thus, this package can only evolve to meet new analytical demands for the peace science community.

Finally, a skeptical reader should not think that making the process as simple as possible necessarily facilitates poor decision-making by the user. In cases where it is evident what the user wants (e.g. an estimate of the level of democracy in the state-year), `{peacesciencer}` does the necessary work to provide the user that information. However, the package makes sure to leave important decision-making to the researcher. For example, `add_cow_alliance()` returns information about various types of alliance pledges in the dyad-year---should one exist---but leaves it to the researcher to say whether they want to define the presence of an alliance to be just a defense pledge or any type of alliance pledge. `add_contiguity()` returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity threshold. The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g. the contiguity data are not ordinal and should not be treated as such), how and where data differ (e.g. how CoW state system data differ from Gleditsch-Ward data), and how researchers should consider optimally using its functionality (e.g. `add_ucdp_acd()` probably should not lump all forms of conflict together). `{peacesciencer}` does not endeavor to make researchers lazy or sloppy, and it does not ultimately do this. Instead, `{peacesciencer}` reduces the tedium associated with starting quantitative peace science research, achieving this in a quick, robust, and transparent way.

\newpage

# References

<!-- This option seems to be more pleasing for rmarkdown::pdf_document() with my templates -->
\setlength{\parskip}{6pt}

<!-- If you're using bookdown, this seems to look better with bookdown::pdf_document2() -->
<!-- \setlength{\parskip}{8pt} -->
<!-- \setlength{\parindent}{-0.15in} -->
<!-- \setlength{\leftskip}{0.15in} -->
<!-- \vspace*{-0.35in} -->
<!-- \noindent -->
<!-- \singlespacing -->

### 2021-06-10 user-guide/R/1-prep.R

@@ -1,0 +1,28 @@
+# load libraries ----
+# load the functions I need to process the data
+library(stevedata) # has the raw data
+
+# "load" the data ----
+# Note: in your case, you can load raw data from wherever, even outside the project directory
+# For example, if you're analyzing ANES data, you're not ask to stick it in your project directory
+# If you can share the raw data, create a directory called "data-raw" in the main project directory
+# Then, load it from there.
+data(ESS9GB, package="stevedata")
+
+# "prep" the data ----
+# I can do anything I want here. I can recode things, transform variables, or whatever
+# Here, let's just create a stupid noise variable
+
+Data <- ESS9GB
+
+set.seed(8675309)
+Data$noise <- rnorm(nrow(Data))
+
+# in {tidyverse}
+# set.seed(8675309)
+# ESS9GB %>%
+#   mutate(noise = rnorm(nrow(.))) -> Data
+
+# ^ Notice that I "finished" my data prep into a new object, titled "Data"
+# Now: I save it to data/Data.rds
+saveRDS(Data, "user-guide/data/Data.rds")

### 2021-06-10 user-guide/R/2-analysis.R

@@ -1,0 +1,8 @@
+Data <- readRDS("user-guide/data/Data.rds")
+
+Mods <- list()
+
+Mods$"Default" <- lm(immigsent ~ agea + female + eduyrs + uempla + hinctnta + lrscale, data=Data)
+Mods$"Default + Noise" <- lm(immigsent ~ agea + female + eduyrs + uempla + hinctnta + lrscale + noise, data=Data)
+
+saveRDS(Mods, "user-guide/data/Mods.rds")

### 2021-06-10 user-guide/R/3-sims.R

@@ -1,0 +1,32 @@
+# Libraries first
+library(stevemisc)
+# library(tidyverse)
+library(modelr)
+
+# Load stuff
+Mods <- readRDS("user-guide/data/Mods.rds")
+Data <- readRDS("user-guide/data/Data.rds")
+
+# in {tidyverse}
+# Data %>%
+#   data_grid(lrscale = unique(lrscale), .model = Mods[[1]],
+#             immigsent = 0) %>% na.omit -> newdat
+
+
+Sims <- list()
+
+na.omit(data_grid(Data, lrscale = unique(lrscale), .model = Mods[[1]], immigsent = 0)) -> newdat
+
+Sims$"SQI (Ideology)" <- get_sims(Mods[[1]], newdat = newdat, 1000, 8675309)
+
+Sims[[1]]$lrscale <- rep(seq(0:10), 1000)
+
+
+
+# in {tidyverse}
+# newdat %>%
+#   # repeat this data frame how many times we did simulations
+#   dplyr::slice(rep(row_number(), 1000)) %>%
+#   bind_cols(get_sims(Mods[[1]], newdata = newdat, 1000, 8675309), .) -> Sims$"SQI (Ideology)"
+
+saveRDS(Sims, "user-guide/data/Sims.rds")

### 2021-06-10 user-guide/src/render_docx.R

@@ -1,0 +1,8 @@
+# rmarkdown::render("ms.Rmd", output_file="doc/ms.docx", rmarkdown::word_document(reference_docx = stevetemplates::templ_word()))
+# ^ old, legacy, here if you want to revert to it.
+
+rmarkdown::render("ms.Rmd", output_file="doc/ms.docx",
+                  bookdown::word_document2(reference_docx = stevetemplates::templ_word(),
+                                           toc = FALSE, number_sections = FALSE))
+
+# ^ {bookdown} has table/figure cross-ref ability.

### 2021-06-10 user-guide/doc/ms.tex
add that little bit... little bits...

@@
 
-Finally, the creation and presentation of data in peace science should be 100\% transparent, which \texttt{\{peacesciencer\}} takes seriously in the following ways.
+Finally, the creation and presentation of data in peace science should be 100\% robust and transparent, which \texttt{\{peacesciencer\}} takes seriously in the following ways.
 The website for \texttt{\{peacesciencer\}} has several vignettes that describe its processes in some detail.
 These include how it provides reasonable estimates of democracy that may not be available in the Polity data or the Varieties of Democracy data and how a researcher can whittle dyadic dispute-years into true dyad-years through reasonable case exclusions.
+\texttt{\{peacesciencer\}} subjects itself to a battery of tests before publishing updates, making sure new features do not create duplicate entries in the original data (which is the surest sign of a botched merge).
 The project's Github contains a publicly available \texttt{data-raw} directory that shows how every data set included (and processed) in \texttt{\{peacesciencer\}} was created.
@@
 Thus, this package can only evolve to meet new analytical demands for the peace science community.
+Users are free to request new features as ``issues'' on the project's Github.
 

### 2021-06-10 user-guide/ms.Rmd
add that little bit... little bits...

@@
 
-Finally, the creation and presentation of data in peace science should be 100% transparent, which `{peacesciencer}` takes seriously in the following ways.
+Finally, the creation and presentation of data in peace science should be 100% robust and transparent, which `{peacesciencer}` takes seriously in the following ways.
 The website for `{peacesciencer}` has several vignettes that describe its processes in some detail.
 These include how it provides reasonable estimates of democracy that may not be available in the Polity data or the Varieties of Democracy data and how a researcher can whittle dyadic dispute-years into true dyad-years through reasonable case exclusions.
+`{peacesciencer}` subjects itself to a battery of tests before publishing updates, making sure new features do not create duplicate entries in the original data (which is the surest sign of a botched merge).
 The project's Github contains a publicly available `data-raw` directory that shows how every data set included (and processed) in `{peacesciencer}` was created.
@@
 Thus, this package can only evolve to meet new analytical demands for the peace science community.
+Users are free to request new features as "issues" on the project's Github.
 

### 2021-06-10 user-guide/doc/ms.tex
meh

@@
 Thereafter, it provides a tutorial on how to install and best use \texttt{\{peacesciencer\}} in the R programming language.
-A more comprehensive tutorial follows, showing how \texttt{\{peacesciencer\}} already has a suite of data and functions that can allow for effective replications of a ``dangerous dyads'' type analysis (Bremer 1992) and standard state-year analyses of civil conflict onset (e.g.
-Fearon and Laitin 2003).
+A more comprehensive tutorial follows, showing how \texttt{\{peacesciencer\}} already has a suite of data and functions that can allow for effective replications of a ``dangerous dyads'' type analysis (Bremer, 1992) and standard state-year analyses of civil conflict onset (e.g.
+Fearon \& Laitin, 2003).
 This feature concludes with a discussion of what else \texttt{\{peacesciencer\}} is capable of doing in future updates and how this R package can inform more reasoned design decisions for researchers in peace science.
@@
 \texttt{\{peacesciencer\}} is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch.
-NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars (Bennett, Poast, and Stam 2019).
-NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade (Bennett and Stam 2000).
+NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars (Bennett, Poast \& Stam, 2019).
+NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade (Bennett \& Stam, 2000).
 No matter, \texttt{\{peacesciencer\}} is motivated by the following ideals that led to its creation.
@@
 In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer (c.f.
-Bowers and Voors 2016).
+Bowers \& Voors, 2016).
 This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that \texttt{\{peacesciencer\}} uses.
-There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet (Arel-Bundock 2021b), scrape \emph{any} type of data from the internet (Boehmke 2016), and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero (Arel-Bundock 2021a).
+There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet (Arel-Bundock, 2021a), scrape \emph{any} type of data from the internet (Boehmke, 2016), and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero (Arel-Bundock, 2021b).
 \texttt{\{peacesciencer\}} embraces this.
@@
 Graduate students just beginning their career in peace science scholarship are tasked to learn the substantive material, the characteristics of the underlying data in wide use in the community, and the quantitative methods and statistical software available to them, effectively all at the same time.
-A student interested in civil conflict scholarship will need to learn the UCDP-PRIO armed conflict data in some detail at the same time they learn about the theoretical arguments that informed the creation of the data, likely at the same time they take quantitative methods courses that show them how to load and explore data in a choseen statistical software.
+A student interested in civil conflict scholarship will need to learn the UCDP-PRIO armed conflict data in some detail at the same time they learn about the theoretical arguments that informed the creation of the data, likely at the same time they take quantitative methods courses that show them how to load and explore data in a chosen statistical software.
 At the same time that creating data is increasingly blurred with coding and processing data, students are simultaneously learning to explore data for themselves as they read about the analyses of the data.
@@
 For example, \texttt{create\_dyadyears()} transforms the raw state system membership data for either the Correlates of War or Gleditsch-Ward system into the full universe of post-1816 dyad-years.
-\texttt{add\_gml\_mids()} uses a dyadic dispute-year version of the MID data offered by Gibler, Miller, and Little (2016), which, as the \texttt{data-raw} directory on the project's Github shows, is itself derived from the dispute-level and participant-level versions of the data.
+\texttt{add\_gml\_mids()} uses a dyadic dispute-year version of the MID data offered by Gibler, Miller \& Little (2016), which, as the \texttt{data-raw} directory on the project's Github shows, is itself derived from the dispute-level and participant-level versions of the data.
 \texttt{\{peacesciencer\}} also has some data innovations.
 It has an original data set on capital-to-capital distances, calculating for distance in kilometers between state capitals using the ``Vincenty'' method (i.e.~``as the crow flies'') that also accounts for instances when capitals moved (e.g.~Brazil in 1960, Burundi in 2018).
-It also features an innovation of democracy data, providing reasonable estimates of democracy using the Marquez (2016) of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve, and Melton 2010).\footnote{A vignette on the \texttt{\{peacesciencer\}} website explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.} As of writing, \texttt{\{peacesciencer\}} comes with the following functionality.
+It also features an innovation of democracy data, providing reasonable estimates of democracy using the Marquez (2016) of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve \& Melton, 2010).\footnote{A vignette on the \texttt{\{peacesciencer\}} website explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.} As of writing, \texttt{\{peacesciencer\}} comes with the following functionality.
 
@@
 Data that can be obtained from a stable advanced programming interface---like the World Bank, for example---can be obtained through those other means (e.g.
-Arel-Bundock 2021b).
+Arel-Bundock, 2021a).
 Its coverage will assuredly expand with new additions of interest to the peace science community.
@@
 \texttt{\{peacesciencer\}}'s functions work out of the box, though it works best with one additional package.
-This is \texttt{\{tidyverse\}}, itself a set of packages that share a common form and design (Wickham and Grolemund 2017).
+This is \texttt{\{tidyverse\}}, itself a set of packages that share a common form and design (Wickham \& Grolemund, 2017).
 \texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}} but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality.
@@
 Researchers who want to understand, say, the correlates of inter-state conflict onset almost always do the following.
-They will add the dependent variables from an inter-state conflict data set, like the Gibler-Miller-Little (GML) dispute data set (Gibler, Miller, and Little 2016).
+They will add the dependent variables from an inter-state conflict data set, like the Gibler-Miller-Little (GML) dispute data set (Gibler, Miller \& Little, 2016).
 They will add several indicators to the dyad-year data that are well-traveled in conflict scholarship, including indicators of things like joint alliance contracts and the contiguity relationship of both states in the dyad.
@@
 Put in other words, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)\ \%\textgreater{}\%\ filter\_prd()} is a command in \texttt{\{tidyverse\}} and \texttt{\{peacesciencer\}} to create all non-directed dyad-years from 1816 to 2016 and then subset those data to only the politically relevant dyad-years.
-Because politically relevant dyads depend on information about the major power status for both states as well as their contiguity relationship (Lemke and Reed 2001), the function adds in those data (i.e.~quietly executing \texttt{add\_contiguity()} and \texttt{add\_cow\_majors()}) before subsetting the data to just the dyads that have some kind of contiguity relationship and/or have a major power in it.
+Because politically relevant dyads depend on information about the major power status for both states as well as their contiguity relationship (Lemke \& Reed, 2001), the function adds in those data (i.e.~quietly executing \texttt{add\_contiguity()} and \texttt{add\_cow\_majors()}) before subsetting the data to just the dyads that have some kind of contiguity relationship and/or have a major power in it.
 
@@
 Importantly, I do not recommend manually creating splines or square/cube terms because it creates more problems in adjusting for temporal dependence in model predictions.
-In a regression formula in R, you can specify the D.
-B.
-Carter and Signorino (2010) approach as \texttt{...\ +\ gmlmidspell\ +\ I(gmlmidspell\^{}2)\ +\ I(gmlmidspell\^{}3)}.
-The Beck, Katz, and Tucker (1998) cubic splines approach is \texttt{...\ +\ splines::bs(gmlmidspell,\ 4)}.
+In a regression formula in R, you can specify the Carter \& Signorino (2010) approach as \texttt{...\ +\ gmlmidspell\ +\ I(gmlmidspell\^{}2)\ +\ I(gmlmidspell\^{}3)}.
+The Beck, Katz \& Tucker (1998) cubic splines approach is \texttt{...\ +\ splines::bs(gmlmidspell,\ 4)}.
 This function includes the spell and three splines (hence the 4 in the command).} \texttt{add\_nmc()} will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity.
@@
 \texttt{add\_cow\_alliance()} will add information about whether there was an alliance in the dyad-year by way of whether ther was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge.
-Finally, \texttt{add\_sdp\_gdp()} will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by Anders, Fariss, and Markowitz (2020).
+Finally, \texttt{add\_sdp\_gdp()} will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by Anders, Fariss \& Markowitz (2020).
 
@@
 \texttt{type} is an optional argument for the type of armed conflicts for which the researcher wants information.
-Options include ``extrasystemic,'' ``interstate,'' ``intrastate,'' and ``II'' (short for ``internationalized intrastate'').
+Options include ``extrasystemic'', ``interstate'', ``intrastate'', and ``II'' (short for ``internationalized intrastate'').
 If no \texttt{type} is specified, the function returns information about ongoing disputes and onsets for all states for all tyeps of conflict.
@@
 \texttt{issue} is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict.
-Options include ``territory,'' ``government,'' and ``both.'' If no \texttt{issue} is specified, the function returns information for all conflicts irregarding the particular issue.
+Options include ``territory'', ``government'', and ``both''.
+If no \texttt{issue} is specified, the function returns information for all conflicts irregarding the particular issue.
 \texttt{only\_wars} is an argument that subsets the data to just those with the intensity levels of ``war'' when \texttt{only\_wars\ =\ TRUE}.
@@
 \texttt{add\_creg\_fractionalization()} adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois.
-\texttt{add\_sdp\_gdp()} will add information about a state's estimated GDP, population, and GDP per capita from the Anders, Fariss, and Markowitz (2020) simulations.
+\texttt{add\_sdp\_gdp()} will add information about a state's estimated GDP, population, and GDP per capita from the Anders, Fariss \& Markowitz (2020) simulations.
 Finally, \texttt{add\_rugged\_terrain()} provide two estimates of the ruggedness of a state's terrain.
-The first is the terrain ruggedness index calculated by Nunn and Puga (2012) and the second is the Gibler and Miller (2014) extension of the natural logged percentage of the state that is mountainous (originally calculated by Fearon and Laitin (2003)).
+The first is the terrain ruggedness index calculated by Nunn \& Puga (2012) and the second is the Gibler \& Miller (2014) extension of the natural logged percentage of the state that is mountainous (originally calculated by Fearon \& Laitin (2003)).
 At the end of the pipe, the data returned by \texttt{\{peacesciencer\}} is assigned to an object minimally called \texttt{Data}.
@@
 For example, \texttt{\{peacesciencer\}} does not right now handle \emph{k}-adic data (c.f.
-Poast 2010).
+Poast, 2010).
 However, the nature of \emph{k}-ads do not make for necessarily problematic data merges and should not be difficult to add for researchers who want it in future updates.
-Likewise, \texttt{\{peacesciencer\}} contains the Archigos leader data (Goemans, Gleditsch, and Chiozza 2009) and is able to transform that into interesting data at the dyad-level and state-level, but doing a leader-level analysis with information about leader willingness to use force (e.g.
-J.
-Carter and Smith 2020) is not right now available.
+Likewise, \texttt{\{peacesciencer\}} contains the Archigos leader data (Goemans, Gleditsch \& Chiozza, 2009) and is able to transform that into interesting data at the dyad-level and state-level, but doing a leader-level analysis with information about leader willingness to use force (e.g.
+Carter \& Smith, 2020) is not right now available.
 However, these are possible expansions on what \texttt{\{peacesciencer\}} already does well.
@@
 \leavevmode\hypertarget{ref-andersetal2020bbgb}{}%
-Anders, Therese, Christopher J.
-Fariss, and Jonathan N.
-Markowitz.
-2020.
-{``Bread Before Guns or Butter: Introducing Surplus Domestic Product (SDP).''} \emph{International Studies Quarterly} 64 (2): 392--405.
+Anders, Therese, Christopher J Fariss \& Jonathan N Markowitz (2020) Bread before guns or butter: Introducing surplus domestic product (SDP). \emph{International Studies Quarterly} 64(2): 392--405.
 
 \leavevmode\hypertarget{ref-arelbundock2021m}{}%
-Arel-Bundock, Vincent.
-2021a.
-\emph{Modelsummary: Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready}.
-\url{https://CRAN.R-project.org/package=modelsummary}.
+Arel-Bundock, Vincent (2021b) \emph{Modelsummary: Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready} (\url{https://CRAN.R-project.org/package=modelsummary}).
 
 \leavevmode\hypertarget{ref-arelbundock2021w}{}%
----------. 2021b. \emph{WDI: World Development Indicators and Other World Bank Data}. \url{https://CRAN.R-project.org/package=WDI}.
+Arel-Bundock, Vincent (2021a) \emph{WDI: World Development Indicators and Other World Bank Data} (\url{https://CRAN.R-project.org/package=WDI}).
 
 \leavevmode\hypertarget{ref-becketal1998tts}{}%
-Beck, Nathaniel, Jonathan N.
-Katz, and Richard Tucker.
-1998.
-{``Taking Time Seriously: Time-Series-Cross-Section Analysis with a Binary Dependent Variable.''} \emph{American Journal of Political Science} 42 (4): 1260--88.
+Beck, Nathaniel, Jonathan N Katz \& Richard Tucker (1998) Taking time seriously: Time-series-cross-section analysis with a binary dependent variable.
+\emph{American Journal of Political Science} 42(4): 1260--1288.
 
 \leavevmode\hypertarget{ref-bennettetal2019n}{}%
-Bennett, D. Scott, Paul Poast, and Allan C. Stam. 2019. {``NewGene: An Introduction for Users.''} \emph{Journal of Conflict Resolution} 63 (6): 1579--92.
+Bennett, DScott, Paul Poast \& Allan C Stam (2019) NewGene: An introduction for users. \emph{Journal of Conflict Resolution} 63(6): 1579--1592.
 
 \leavevmode\hypertarget{ref-bennettstam2000e}{}%
-Bennett, D. Scott, and Allan Stam. 2000. {``EUGene: A Conceptual Manual.''} \emph{International Interactions} 26 (2): 179--204.
+Bennett, DScott \& Allan Stam (2000) EUGene: A conceptual manual. \emph{International Interactions} 26(2): 179--204.
 
 \leavevmode\hypertarget{ref-boehmke2016dwr}{}%
-Boehmke, Bradley C. 2016. \emph{Data Wrangling with {R}}. Springer.
+Boehmke, Bradley C (2016) \emph{Data Wrangling with {R}}. Springer.
 
 \leavevmode\hypertarget{ref-bowersvoors2016hiyr}{}%
-Bowers, Jake, and Maarten Voors. 2016. {``How to Improve Your Relationship with Your Future Self.''} \emph{Revista de Ciencia Politica} 36 (3): 829--48.
+Bowers, Jake \& Maarten Voors (2016) How to improve your relationship with your future self. \emph{Revista de Ciencia Politica} 36(3): 829--848.
 
 \leavevmode\hypertarget{ref-bremer1992ddc}{}%
-Bremer, Stuart A. 1992. {``Dangerous Dyads: Conditions Affecting the Likelihood of Interstate War, 1816-1965.''} \emph{Journal of Conflict Resolution} 36 (2): 309--41.
+Bremer, Stuart A (1992) Dangerous dyads: Conditions affecting the likelihood of interstate war, 1816-1965. \emph{Journal of Conflict Resolution} 36(2): 309--341.
 
 \leavevmode\hypertarget{ref-cartersignorino2010bf}{}%
-Carter, David B., and Curtis S. Signorino. 2010. {``Back to the Future: Modeling Time Dependence in Binary Data.''} \emph{Political Analysis} 18 (3): 271--92.
+Carter, David B \& Curtis S Signorino (2010) Back to the future: Modeling time dependence in binary data. \emph{Political Analysis} 18(3): 271--292.
 
 \leavevmode\hypertarget{ref-cartersmith2020fml}{}%
-Carter, Jeff, and Charles E. Smith. 2020. {``A Framework for Measuring Leaders' Willingness to Use Force.''} \emph{American Political Science Review} 114 (4): 1352--58.
+Carter, Jeff \& Charles E Smith (2020) A framework for measuring leaders' willingness to use force. \emph{American Political Science Review} 114(4): 1352--1358.
 
 \leavevmode\hypertarget{ref-fearonlaitin2003eicw}{}%
-Fearon, James D., and David D. Laitin. 2003. {``Ethnicity, Insurgency, and Civil War.''} \emph{American Political Science Review} 97 (1): 75--90.
+Fearon, James D \& David D Laitin (2003) Ethnicity, insurgency, and civil war. \emph{American Political Science Review} 97(1): 75--90.
 
 \leavevmode\hypertarget{ref-giblermiller2014etts}{}%
-Gibler, Douglas M., and Steven V. Miller. 2014. {``External Territorial Threat, State Capacity, and Civil War.''} \emph{Journal of Peace Research} 51 (5): 634--46.
+Gibler, Douglas M \& Steven V Miller (2014) External territorial threat, state capacity, and civil war. \emph{Journal of Peace Research} 51(5): 634--646.
 
 \leavevmode\hypertarget{ref-gibleretal2016amid}{}%
-Gibler, Douglas M., Steven V.
-Miller, and Erin K.
-Little.
-2016.
-{``An Analysis of the {M}ilitarized {I}nterstate {D}ispute {(MID)} Dataset, 1816-2001.''} \emph{International Studies Quarterly} 60 (4): 719--30.
+Gibler, Douglas M, Steven V Miller \& Erin K Little (2016) An analysis of the {M}ilitarized {I}nterstate {D}ispute {(MID)} dataset, 1816-2001. \emph{International Studies Quarterly} 60(4): 719--730.
 
 \leavevmode\hypertarget{ref-goemansetal2009ia}{}%
-Goemans, Henk E., Kristian Skrede Gleditsch, and Giacomo Chiozza. 2009. {``Introducing {A}rchigos: A Dataset on Political Leaders.''} \emph{Journal of Peace Research} 46 (2): 269--83.
+Goemans, Henk E, Kristian Skrede Gleditsch \& Giacomo Chiozza (2009) Introducing {A}rchigos: A dataset on political leaders. \emph{Journal of Peace Research} 46(2): 269--83.
 
 \leavevmode\hypertarget{ref-lemkereed2001rprd}{}%
-Lemke, Douglas, and William Reed. 2001. {``The Relevance of Politically Relevant Dyads.''} \emph{Journal of Conflict Resolution} 45 (1): 126--44.
+Lemke, Douglas \& William Reed (2001) The relevance of politically relevant dyads. \emph{Journal of Conflict Resolution} 45(1): 126--144.
 
 \leavevmode\hypertarget{ref-marquez2016qme}{}%
-Marquez, Xavier. 2016. {``A Quick Method for Extending the {U}nified {D}emocracy {S}cores.''}
+Marquez, Xavier (2016) A quick method for extending the {U}nified {D}emocracy {S}cores.
 
 \leavevmode\hypertarget{ref-nunnpuga2012r}{}%
-Nunn, Nathan, and Diego Puga. 2012. {``Ruggedness: The Blessing of Bad Geography in {A}frica.''} \emph{Review of Economics and Statistics} 94 (1): 20--36.
+Nunn, Nathan \& Diego Puga (2012) Ruggedness: The blessing of bad geography in {A}frica. \emph{Review of Economics and Statistics} 94(1): 20--36.
 
 \leavevmode\hypertarget{ref-pemsteinetal2010dc}{}%
-Pemstein, Daniel, Stephen A. Meserve, and James Melton. 2010. {``Democratic Compromise: A Latent Variable Analysis of Ten Measures of Regime Type.''} \emph{Political Analysis} 18 (4): 426--49.
+Pemstein, Daniel, Stephen A Meserve \& James Melton (2010) Democratic compromise: A latent variable analysis of ten measures of regime type. \emph{Political Analysis} 18(4): 426--449.
 
 \leavevmode\hypertarget{ref-poast2010mdd}{}%
-Poast, Paul. 2010. {``(Mis)using Dyadic Data to Analyze Multilateral Events.''} \emph{Political Analysis} 18 (4): 403--25.
+Poast, Paul (2010) (Mis)using dyadic data to analyze multilateral events. \emph{Political Analysis} 18(4): 403--425.
 
 \leavevmode\hypertarget{ref-wickhamgrolemund2017rds}{}%
-Wickham, Hadley, and Garrett Grolemund. 2017. \emph{{R} for Data Science}. Sebastopol, CA: O'Reilly Media, Inc.
+Wickham, Hadley \& Garrett Grolemund (2017) \emph{{R} for Data Science}. Sebastopol, CA: O'Reilly Media, Inc.
 

### 2021-06-10 user-guide/ms.Rmd
meh

@@
 geometry: margin=1in
+csl: inst/jpr.csl
 mainfont: cochineal

### 2021-06-10 user-guide/doc/ms.tex
okay, I think I anonymized the manuscript well enough. Someone please hire me.

@@
 \noindent \small{This article introduces \texttt{\{peacesciencer\}}, an R package that contains a litany of tools for creating data of widespread interest to the peace science community.
-The package is cross-platform, assuming only a recent installation of the R programming language with some of the enhanced functionality of the broadly popular \texttt{\{tidvyerse\}} packages.
+The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular \texttt{\{tidvyerse\}} packages.
 Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year or state-year analysis (e.g.~democracy data, contiguity data).
@@
 Graduate students, in particular, face unique challenges associated with these developments.
-Graduate students beginning their careers in peace science to learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.
+Graduate students beginning their careers in peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.
 
@@
 This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want.
-\texttt{\{peacesciencer\}} comes with some data innovations as well, including a comprehensive data set on democracy, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts.
+\texttt{\{peacesciencer\}} comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts.
 All are done with the maximum possible transparency.
@@
 The \texttt{data-raw} directory on the project's Github contains information and comments about how every data set was created and the \texttt{R} directory informs the user how each function works.
-The function manuals (\url{http://svmiller.com/peacesciencer/reference/}) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing.
+The function manuals (\url{http://svmiller.com/peacesciencer/reference}) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing.
 Thus, \texttt{\{peacesciencer\}} not only assists a peace scientist with their research, but it does so in a manner that best conforms to the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.
@@
 
-For one, researchers may invest too much time in the construction of a data set that faithfully captures the unit of analysis.
-Assume, for example, a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict.
+For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis.
+Assume a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict.
 How might one do that?
@@
 Second, researchers also invest too much time in retracing steps for peace science analyses for new projects.
-Assume, for example, a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data).
+Assume a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data).
 Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project.
@@
 \texttt{\{peacesciencer\}} embraces this.
-This R programming language package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.
+This R package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.
 
@@
 The function manuals included \texttt{\{peacesciencer\}} contain ample documentation that clarifies what the function is doing, what it returns to the user, and why it is doing it this way.
-Users can also use the project's Github to point out bugs, ask for further clarification, and propose additions.
+Researchers can also use the project's Github to point out bugs, ask for further clarification, and propose additions.
 \texttt{\{peacesciencer\}} takes seriously the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science and endeavors for maximum transparency, leveraging open source and version control software to inform users of what data it uses and how it uses the data.
@@
 \hline
-UCDP Armed Conflicts & Add UCDP armed conflict data by year & S & \ttfamily{add\_ucdp\_acd()}\\
+UCDP Armed Conflicts (v. 20.1) & Add UCDP armed conflict data by year & S & \ttfamily{add\_ucdp\_acd()}\\
 \arrayrulecolor{black}
@@
 \texttt{\{peacesciencer\}}'s functions work out of the box, though it works best with one additional package.
-This is \texttt{\{tidyverse\}}, itself a set of packages that share a common form and design (Wickham \& Grolemund, 2017).
+This is \texttt{\{tidyverse\}}, itself a suit of packages that share a common form and design (Wickham \& Grolemund, 2017).
 \texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}} but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality.
@@
 \texttt{mry}---short for ``most recent year''---determines whether or not the function extends the membership data to the most recently concluded calendar year.
-The function defaults to \texttt{TRUE} in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017).
+The function defaults to \texttt{TRUE} in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to (as of writing) 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017).
 \texttt{create\_dyadyears()} has a third argument---\texttt{directed}---that determines whether the user wants directed dyad-year data or non-directed dyad-year data.
@@
 
-This example would create a full non-directed dyad-year data frame of Correlates of War states from 1816 to 2016, the last year of record in the Correlates of War state system membership data.
+The below example would create a full non-directed dyad-year data frame of Correlates of War states from 1816 to 2016, the last year of record in the Correlates of War state system membership data.
 
@@
 First, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)} will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016.
-This is effectively the same function used above, but with two user-specified overrides.
 
@@
 Researchers can select which one they prefer from these indicators.
-\texttt{add\_cow\_alliance()} will add information about whether there was an alliance in the dyad-year by way of whether ther was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge.
+\texttt{add\_cow\_alliance()} will add information about whether there was an alliance in the dyad-year by way of whether there was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge.
 Finally, \texttt{add\_sdp\_gdp()} will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by Anders, Fariss \& Markowitz (2020).
@@
 \texttt{\{peacesciencer\}} is also capable of creating data for replications of standard civil conflict analyses.
-Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's (2003) well-cited analysis of civil conflict onset using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years.
+Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's (2003) well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years.
 The pipe-based workflow will start with \texttt{create\_stateyears(system\ =\ \textquotesingle{}gw\textquotesingle{})}, creating the full universe of Gleditsch-Ward state years from 1816 to 2020.
-Whereas the UCDP data included in \texttt{\{peacesciencer\}} are from 1946 to 2019, we can subset (``filter'') the data to just those years with \texttt{filter(year\ \%in\%\ c(1946,\ 2019))}.
+Whereas the UCDP data included in \texttt{\{peacesciencer\}} are from 1946 to 2019 (v.
+20.1), we can subset (``filter'') the data to just those years with \texttt{filter(year\ \%in\%\ c(1946,\ 2019))}.
 Next, we can use the \texttt{add\_ucdp\_acd()} function to return information about ongoing UCDP conflicts and onsets for these states.
@@
 \texttt{add\_contiguity()} returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity threshold.
-The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g.~the contiguity data are not ordinal and should not be treated as such), how and where data differ (e.g.~how CoW state system data differ from Gleditsch-Ward data), and how researchers should consider optimally using its functionality (e.g.~\texttt{add\_ucdp\_acd()} probably should not lump all forms of conflict together).
+The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g.~the contiguity data are not ordinal and should not be treated as such), how and where data issues arise (e.g.~how CoW state system data differ from Gleditsch-Ward data and how one is coerced into the other), and how researchers should consider optimally using its functionality (e.g.~\texttt{add\_ucdp\_acd()} probably should not lump all forms of conflict together).
 \texttt{\{peacesciencer\}} does not endeavor to make researchers lazy or sloppy, and it does not ultimately do this.
@@
 \leavevmode\hypertarget{ref-marquez2016qme}{}%
-Marquez, Xavier (2016) A quick method for extending the {U}nified {D}emocracy {S}cores.
+Marquez, Xavier (2016) A quick method for extending the {U}nified {D}emocracy {S}cores (\url{http://dx.doi.org/10.2139/ssrn.2753830}).
 

### 2021-06-10 user-guide/ms.Rmd
okay, I think I anonymized the manuscript well enough. Someone please hire me.

@@
 abstract: "This article introduces `{peacesciencer}`, an R package that contains a litany of tools for creating data of widespread interest to the peace science community.
-The package is cross-platform, assuming only a recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages.
+The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages.
 Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year or state-year analysis (e.g.
 democracy data, contiguity data).
-The software is freely available on CRAN and maintains an active website documenting its features at http://svmiller.com/peacesciencer."
+The software is freely available on CRAN and maintains an active website documenting its features at `r params$ps_website`."
 keywords: "software, statistical analysis, peace science, reproducibility"
@@
   removetitleabstract: ""
+  ps_website: ""
+  ps_manuals: ""
+  ps_github: ""
 anonymous: "`r params$anonymous`"
@@
 Graduate students, in particular, face unique challenges associated with these developments.
-Graduate students beginning their careers in peace science to learn how scholarship informs data in peace science and  how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.
+Graduate students beginning their careers in peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.
 
@@
 This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want.
-`{peacesciencer}` comes with some data innovations as well, including a comprehensive data set on democracy, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts.
+`{peacesciencer}` comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts.
 All are done with the maximum possible transparency.
@@
 The `data-raw` directory on the project's Github contains information and comments about how every data set was created and the `R` directory informs the user how each function works.
-The function manuals (http://svmiller.com/peacesciencer/reference/) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing.
+The function manuals (`r params$ps_manual`) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing.
 Thus, `{peacesciencer}` not only assists a peace scientist with their research, but it does so in a manner that best conforms to the  Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.
@@
 Second, researchers also invest too much time in retracing steps for peace science analyses for new projects.
-Assume, for example, a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data).
+Assume  a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data).
 Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project.
@@
 `{peacesciencer}` embraces this.
-This R programming language package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.
+This R package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.
 
@@
 The function manuals included `{peacesciencer}` contain ample documentation that clarifies what the function is doing, what it returns to the user, and why it is doing it this way.
-Users can also use the project's Github to point out bugs, ask for further clarification, and propose additions.
+Researchers can also use the project's Github to point out bugs, ask for further clarification, and propose additions.
 `{peacesciencer}` takes seriously the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science and endeavors for maximum transparency, leveraging open source and version control software to inform users of what data it uses and how it uses the data.
@@
         "Strategic Rivalries", "Add Thompson and Dreyer (2012) strategic rivalry data by year", "D, S", "add_strategic_rivalries()",
-        "UCDP Armed Conflicts", "Add UCDP armed conflict data by year", "S", "add_ucdp_acd()"
+        "UCDP Armed Conflicts (v. 20.1)", "Add UCDP armed conflict data by year", "S", "add_ucdp_acd()"
         ) -> included
@@
 `{peacesciencer}`'s functions work out of the box, though it works best with one additional package.
-This is `{tidyverse}`, itself a set of packages that share a common form and design [@wickhamgrolemund2017rds].
+This is `{tidyverse}`, itself a suit of packages that share a common form and design [@wickhamgrolemund2017rds].
 `{peacesciencer}` functions make considerable use of the component packages of `{tidyverse}` but installing and loading `{tidyverse}` will allow the researcher to make quicker use of `{peacesciencer}`'s functionality.
@@
 `mry`---short for "most recent year"---determines whether or not the function extends the membership data to the most recently concluded calendar year.
-The function defaults to `TRUE` in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017).
+The function defaults to `TRUE` in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to (as of writing) 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017).
 `create_dyadyears()` has a third argument---`directed`---that determines whether the user wants directed dyad-year data or non-directed dyad-year data.
@@
 First, `create_dyadyears(directed = FALSE, mry = FALSE)` will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016.
-This is effectively the same function used above, but with two user-specified overrides.
 
@@
 Researchers can select which one they prefer from these indicators.
-`add_cow_alliance()` will add information about whether there was an alliance in the dyad-year by way of whether ther was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge.
+`add_cow_alliance()` will add information about whether there was an alliance in the dyad-year by way of whether there was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge.
 Finally, `add_sdp_gdp()` will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by @andersetal2020bbgb.
@@
 `{peacesciencer}` is also capable of creating data for replications of standard civil conflict analyses.
-Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's [-@fearonlaitin2003eicw] well-cited analysis of civil conflict onset using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years.
+Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's [-@fearonlaitin2003eicw] well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years.
 The pipe-based workflow will start with `create_stateyears(system = 'gw')`, creating the full universe of Gleditsch-Ward state years from 1816 to 2020.
-Whereas the UCDP data included in `{peacesciencer}` are from 1946 to 2019, we can subset ("filter") the data to just those years with `filter(year %in% c(1946, 2019))`.
+Whereas the UCDP data included in `{peacesciencer}` are from 1946 to 2019 (v.
+20.1), we can subset ("filter") the data to just those years with `filter(year %in% c(1946, 2019))`.
 Next, we can use the `add_ucdp_acd()` function to return information about ongoing UCDP conflicts and onsets for these states.
@@
 The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g.
-the contiguity data are not ordinal and should not be treated as such), how and where data differ (e.g.
-how CoW state system data differ from Gleditsch-Ward data), and how researchers should consider optimally using its functionality (e.g.
+the contiguity data are not ordinal and should not be treated as such), how and where data issues arise (e.g.
+how CoW state system data differ from Gleditsch-Ward data and how one is coerced into the other), and how researchers should consider optimally using its functionality (e.g.
 `add_ucdp_acd()` probably should not lump all forms of conflict together).

### 2021-06-10 user-guide/src/render_pdf-anon.R

@@ -6,7 +6,10 @@
 
 
 rmarkdown::render("ms.Rmd", output_file="doc/ms-anon.pdf",
-                  params=list(anonymous=TRUE,doublespacing=TRUE,removetitleabstract=TRUE),
+                  params=list(anonymous=TRUE,doublespacing=TRUE,removetitleabstract=TRUE,
+                              ps_website = "[LINK REDACTED FOR PEER REVIEW]",
+                              ps_manuals = "[LINK REDACTED FOR PEER REVIEW]",
+                              ps_github = "[LINK REDACTED FOR PEER REVIEW]"),
                   bookdown::pdf_document2(template = stevetemplates::templ_article2(),
                                           latex_engine = "xelatex", dev= "cairo_pdf",
                                           toc = FALSE, number_sections = FALSE))

### 2021-06-10 README.Rmd
update README

--- 2021-06-04 README.Rmd
@@
 output: github_document
+always_allow_html: TRUE
 ---
@@
 library("badger")
+library("tidyverse")
+library(kableExtra)
+library(peacesciencer)
+library(stevemisc)
 ```
@@
 
-# Usage
+# What's Included in `{peacesciencer}`
 
-The package is very much a work in progress. Right now, it has the following functions:
+The package is already well developed and its functionality continues to expand. The current development version has the following functions.
 
-- `add_archigos()`: adds some summary variables from Archigos about political leaders for state-year or dyad-year data.
-- `add_atop_alliance()`: adds ATOP alliance information to dyad-year data.
-- `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year or state-year data.
-- `add_ccode_to_gw()`: 	Add Correlates of War state system codes to dyad-year or state-year data with Gleditsch-Ward state codes.
-- `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year or state-year data.
-- `add_cow_alliance()`: adds Correlates of War major alliance information to dyad-year data.
-- `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data.
-- `add_cow_mids()`: adds Correlates of War (CoW) Militarized Interstate Dispute (MID) data to dyad-year data frame
-- `add_cow_trade()`: adds Correlates of War trade data to dyad-year or state-year data.
-- `add_democracy()`: adds estimates of democracy/levels of democracy to dyad-year or state-year data.
-- `add_gwcode_to_cow():` adds Gleditsch-Ward state codes to dyad-year or state-year data with Correlates of War state codes.
-- `add_gml_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data.
-- `add_igos()`: adds Correlates of War International Governmental Organizations (IGOs) data to dyad-year or state-year data.
-- `add_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data. (DEPRECATED)
-- `add_minimum_distance()`: adds minimum distance (in kilometers) to dyad-year or state-year data.
-- `add_nmc()`: adds estimates of national material capabilities (from Correlates of War) to dyad-year or state-year data.
-- `add_peace_years()`: Add Peace Years to Your Conflict Data
-- `add_sdp_gdp()`: adds estimates of (gross and surplus) domestic product and population size to dyad-year or state-year data
-- `add_strategic_rivalries()`: Add Thompson and Dreyer's (2012) strategic rivalry data to dyad-year data frame
-- `add_ucdp_acd()`: Add UCDP Armed Conflict Data to state-year data frame
-- `add_ucdp_onsets()`: adds UCDP onsets to state-year data
-- `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format
-- `create_statedays()`: converts Correlates of War or Gleditsch-Ward state membership data into state-day format..
-- `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format.
-- `filter_prd()`: filters dyad-year data frame to just "politically relevant" dyads.
+```{r, echo=F, message=F, warning=F}
+peacesciencer_functions <- tibble(x = as.vector(lsf.str("package:peacesciencer")))
 
-It also has the following data sets:
+db <- tools::Rd_db("peacesciencer")
+db_names <- names(db)
+db_descript <- lapply(db,function(x) {
+    tags <- tools:::RdTags(x)
+    if("\\description" %in% tags){
+      out <- paste(unlist(x[which(tags=="\\description")]),collapse="")
+    }
+    else
+      out <- NULL
+    invisible(out)
+  })
 
-- `archigos`: an abbreviated version of the Archigos data, used internally
-- `atop_alliance:` directed dyad-year alliance data from ATOP
-- `capitals`: a list of capitals and capital transitions for Correlates of War state system members
-- `ccode_democracy`: Correlates of War state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores").
-- `cow_alliance`: directed dyad-year alliance formation data from the Correlates of War
-- `cow_contdir`: Correlates of War Direct Contiguity Data (v. 3.2)
-- `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members
-- `cow_gw_years`: a yearly data frame including information about Correlates of War and Gleditsch-Ward states.
-- `cow_igo_ndy`: non-directed dyad-year data for Correlates of War intergovernmental organizations data.
-- `cow_igo_sy`: state-year data for Correlates of War intergovernmental organizations data.
-- `cow_majors`: Correlates of War major powers data (version: 2016)
-- `cow_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (CoW-MID, v. 5.0)
-- `cow_mid_dirdisps`: Correlates of War Militarized Interstate Dispute data (directed dyadic dispute-year, v. 5.0) 
-- `cow_mid_disps`: Correlates of War Militarized Interstate Dispute data (abbreviated dispute-level, v. 5.0) 
-- `cow_mindist`: the minimum distance between Correlates of War states, in kilometers, in non-directed dyad-year format (1946-2015)
-- `cow_nmc`: Correlates of War National Material Capabilities data (version 5.0)
-- `cow_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States
-- `cow_states`: Correlates of War state system membership data (version: 2016)
-- `cow_trade_sy`: Correlates of War state-year trade data (version 4.0)
-- `creg`: Composition of Religious and Ethnic Groups (CREG) Fractionalization/Polarization Estimates
-- `gml_dirdisp`: directed dispute-year data from version 2.1.1 of the Gibler-Miller-Little inter-state dispute data.
-- `gml_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (Gibler-Miller-Little, v. 2.1.1)
-- `gw_ddy`: a full directed dyad-year data frame of Gleditsch-Ward state system members
-- `gw_mindist`: the minimum distance between  Gleditsch-Ward states, in kilometers, in non-directed dyad-year format (1946-2015)
-- `gw_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States
-- `gw_states`: Gleditsch-Ward independent state system data (version: 2017)
-- `hief`: Historical Index of Ethnic Fractionalization data
-- `maoz_powers`: Zeev Maoz' global/regional power data
-- `rugged`: Rugged/Mountainous Terrain Data
-- `td_rivalries`: Thompson and Dreyer's (2012) Strategic Rivalries, 1494-2010
-- `ucdp_acd`: a (not quite) dyad-year and (not quite) state-year data set on armed conflict episodes
-- `ucdp_onsets`: a state-year data set of UCDP armed conflict onsets/episodes.
+tibble(x = db_names, y = gsub("\n","",unlist(db_descript))) %>%
+  mutate(x = str_sub(x, 1, nchar(x)-3)) %>%
+  left_join(peacesciencer_functions, .) %>%
+  mutate(func = paste0(x,"()")) -> peacesciencer_functions
 
-The workflow is going to look something like this. This is a "tidy"-friendly approach to a data-generating process in quantitative peace science.
+peacesciencer_functions %>%
+  mutate(link = paste0("[",x,"()](http://svmiller.com/peacesciencer/reference/",x,".html)")) -> peacesciencer_functions
+
+tibble(`**Function**` = peacesciencer_functions$link,
+       `**Description**` = peacesciencer_functions$y) %>%
+  arrange(`**Function**`) %>%
+  kable(., format="markdown")
+
+```
+
+The current development version also includes the following data.
+
+```{r, echo=F, message=F, warning=F}
+# library(tidyverse)
+# library(kableExtra)
+peacesciencer_data <- data(package = "peacesciencer")
+
+object_name <- paste0("`",peacesciencer_data$results[, "Item"],"`")
+object_link <- paste0("[",object_name,"](http://svmiller.com/peacesciencer/reference/",peacesciencer_data$results[, "Item"],".html)")
+
+tibble(`**Object Name**` = object_link,
+       `**Description**` = peacesciencer_data$results[, "Title"]) %>%
+  arrange(`**Object Name**`) %>%
+  kable(., format="markdown")
+
+```
+
+<!-- The package is very much a work in progress. Right now, it has the following functions: -->
+
+<!-- - `add_archigos()`: adds some summary variables from Archigos about political leaders for state-year or dyad-year data. -->
+<!-- - `add_atop_alliance()`: adds ATOP alliance information to dyad-year data. -->
+<!-- - `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year or state-year data. -->
+<!-- - `add_ccode_to_gw()`: 	Add Correlates of War state system codes to dyad-year or state-year data with Gleditsch-Ward state codes. -->
+<!-- - `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year or state-year data. -->
+<!-- - `add_cow_alliance()`: adds Correlates of War major alliance information to dyad-year data. -->
+<!-- - `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data. -->
+<!-- - `add_cow_mids()`: adds Correlates of War (CoW) Militarized Interstate Dispute (MID) data to dyad-year data frame -->
+<!-- - `add_cow_trade()`: adds Correlates of War trade data to dyad-year or state-year data. -->
+<!-- - `add_creg_fractionalization()`: add fractionalization/polarization estimates from CREG to your dyad-year or state-year data -->
+<!-- - `add_democracy()`: adds estimates of democracy/levels of democracy to dyad-year or state-year data. -->
+<!-- - `add_gwcode_to_cow():` adds Gleditsch-Ward state codes to dyad-year or state-year data with Correlates of War state codes. -->
+<!-- - `add_gml_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data. -->
+<!-- - `add_igos()`: adds Correlates of War International Governmental Organizations (IGOs) data to dyad-year or state-year data. -->
+<!-- - `add_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data. (DEPRECATED) -->
+<!-- - `add_minimum_distance()`: adds minimum distance (in kilometers) to dyad-year or state-year data. -->
+<!-- - `add_nmc()`: adds estimates of national material capabilities (from Correlates of War) to dyad-year or state-year data. -->
+<!-- - `add_peace_years()`: Add Peace Years to Your Conflict Data -->
+<!-- - `add_sdp_gdp()`: adds estimates of (gross and surplus) domestic product and population size to dyad-year or state-year data -->
+<!-- - `add_strategic_rivalries()`: Add Thompson and Dreyer's (2012) strategic rivalry data to dyad-year data frame -->
+<!-- - `add_ucdp_acd()`: Add UCDP Armed Conflict Data to state-year data frame -->
+<!-- - `add_ucdp_onsets()`: adds UCDP onsets to state-year data -->
+<!-- - `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format -->
+<!-- - `create_statedays()`: converts Correlates of War or Gleditsch-Ward state membership data into state-day format.. -->
+<!-- - `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format. -->
+<!-- - `filter_prd()`: filters dyad-year data frame to just "politically relevant" dyads. -->
+
+<!-- It also has the following data sets: -->
+
+<!-- - `archigos`: an abbreviated version of the Archigos data, used internally -->
+<!-- - `atop_alliance:` directed dyad-year alliance data from ATOP -->
+<!-- - `capitals`: a list of capitals and capital transitions for Correlates of War state system members -->
+<!-- - `ccode_democracy`: Correlates of War state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores").
+-->
+<!-- - `cow_alliance`: directed dyad-year alliance formation data from the Correlates of War -->
+<!-- - `cow_contdir`: Correlates of War Direct Contiguity Data (v. 3.2) -->
+<!-- - `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members -->
+<!-- - `cow_gw_years`: a yearly data frame including information about Correlates of War and Gleditsch-Ward states. -->
+<!-- - `cow_igo_ndy`: non-directed dyad-year data for Correlates of War intergovernmental organizations data. -->
+<!-- - `cow_igo_sy`: state-year data for Correlates of War intergovernmental organizations data. -->
+<!-- - `cow_majors`: Correlates of War major powers data (version: 2016) -->
+<!-- - `cow_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (CoW-MID, v. 5.0) -->
+<!-- - `cow_mid_dirdisps`: Correlates of War Militarized Interstate Dispute data (directed dyadic dispute-year, v. 5.0)  -->
+<!-- - `cow_mid_disps`: Correlates of War Militarized Interstate Dispute data (abbreviated dispute-level, v. 5.0)  -->
+<!-- - `cow_mindist`: the minimum distance between Correlates of War states, in kilometers, in non-directed dyad-year format (1946-2015) -->
+<!-- - `cow_nmc`: Correlates of War National Material Capabilities data (version 5.0) -->
+<!-- - `cow_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States -->
+<!-- - `cow_states`: Correlates of War state system membership data (version: 2016) -->
+<!-- - `cow_trade_sy`: Correlates of War state-year trade data (version 4.0) -->
+<!-- - `creg`: Composition of Religious and Ethnic Groups (CREG) Fractionalization/Polarization Estimates -->
+<!-- - `gml_dirdisp`: directed dispute-year data from version 2.1.1 of the Gibler-Miller-Little inter-state dispute data. -->
+<!-- - `gml_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (Gibler-Miller-Little, v. 2.1.1) -->
+<!-- - `gw_ddy`: a full directed dyad-year data frame of Gleditsch-Ward state system members -->
+<!-- - `gw_mindist`: the minimum distance between  Gleditsch-Ward states, in kilometers, in non-directed dyad-year format (1946-2015) -->
+<!-- - `gw_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States -->
+<!-- - `gw_states`: Gleditsch-Ward independent state system data (version: 2017) -->
+<!-- - `gwcode_democracy`: Gleditsch-Ward state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores").
+-->
+<!-- - `hief`: Historical Index of Ethnic Fractionalization data -->
+<!-- - `maoz_powers`: Zeev Maoz' global/regional power data -->
+<!-- - `rugged`: Rugged/Mountainous Terrain Data -->
+<!-- - `td_rivalries`: Thompson and Dreyer's (2012) Strategic Rivalries, 1494-2010 -->
+<!-- - `ucdp_acd`: a (not quite) dyad-year and (not quite) state-year data set on armed conflict episodes -->
+<!-- - `ucdp_onsets`: a state-year data set of UCDP armed conflict onsets/episodes. -->
+
+# How to Use `{peacesciencer}`
+
+[`{peacesciencer}` has a user's guide](https://github.com/svmiller/peacesciencer/blob/master/user-guide/doc/ms.pdf) that is worth reading.
+The workflow is going to look something like this.
+This is a "tidy"-friendly approach to a data-generating process in quantitative peace science.
 
@@
 Correlates of War state system membership data are current as of Dec.
-31, 2016 and the script can extend that to the end of 2019), and whether the user wants directed or non-directed dyad-year data (`directed`).
+31, 2016 and the script can extend that to the end of the most recently concluded calendar year), and whether the user wants directed or non-directed dyad-year data (`directed`).
 
@@
 All told, the process will look something like this.
+Assume you want to create some data for something analogous to a "dangerous dyads" design for all non-directed dyad-years.
+Here's how you'd do it in `{peacesciencer}`, which is going to be lifted from the source R scripts for the user's guide.
+The first part of this code chunk will lean on core `{peacesciencer}` functionality whereas the other stuff is some post-processing and, as a bonus, some modeling.
 
-```{r, cache=F, message=F}
+```{r, cache=F, message=F, eval=T}
 
-library(tidyverse, quietly = TRUE)
-library(peacesciencer)
+#library(tidyverse)
+#library(peacesciencer)
+# library(stevemisc)
 library(tictoc)
@@
 tic()
-create_dyadyears() %>%
-  # Add GML MIDs 
-  add_gml_mids(keep=NULL) %>%
-  # Add Correlates of War MIDs 
-  add_cow_mids(keep=NULL) %>%
-  # add peace years
+create_dyadyears(directed = FALSE, mry = FALSE) %>%
+  filter_prd() %>%
+  add_gml_mids(keep = NULL) %>%
   add_peace_years() %>%
-  # Add Gleditsch-Ward codes, because we can
-  add_gwcode_to_cow() %>%
-  # Add strategic rivalry data
-  add_strategic_rivalries() %>%
-  # Add trade data
-  add_cow_trade() %>%
-  # Add capital-to-capital distance
-  add_capital_distance() %>%
-  # Add contiguity information
-  add_contiguity() %>%
-  # Add major power data
-  add_cow_majors() %>%
-  # Add democracy variables
+  add_nmc() %>%
   add_democracy() %>%
-  # Add IGOs data
-  add_igos() %>%
-  # Add National Material Capabilities
-  add_nmc() %>%
-  # add alliance data from Correlates of War
   add_cow_alliance() %>%
-  # add alliance data from ATOP
-  add_atop_alliance() %>%
-  # add minimum distance. No default, must specify "cow" or "gw"
-  add_minimum_distance(system = "cow") %>%
-  # add Archigos data
-  add_archigos() %>%
-  # add gross and surplus GDP data
-  add_sdp_gdp(system = "cow") %>%
-  # add terrain ruggedness data, if you wanted it
-  add_rugged_terrain() %>%
-  # you should probably filter to politically relevant dyads earlier than later...
-  # Or not, it's your time and computer processor...
-  filter_prd()
-toc()
+  add_sdp_gdp() -> Data
 
-# state-years now...
+Data %>%
+  mutate(landcontig = ifelse(conttype == 1, 1, 0)) %>%
+  mutate(cowmajdyad = ifelse(cowmaj1 == 1 | cowmaj2 == 1, 1, 0)) %>%
+  # Create estimate of militarization as milper/tpop
+  # Then make a weak-link
+  mutate(milit1 = milper1/tpop1,
+         milit2 = milper2/tpop2,
+         minmilit = ifelse(milit1 > milit2,
+                           milit2, milit1)) %>%
+  # create CINC proportion (lower over higher)
+  mutate(cincprop = ifelse(cinc1 > cinc2,
+                           cinc2/cinc1, cinc1/cinc2)) %>%
+  # create weak-link specification using Quick UDS data
+  mutate(mindemest = ifelse(xm_qudsest1 > xm_qudsest2,
+                            xm_qudsest2, xm_qudsest1)) %>%
+  # Create "weak-link" measure of jointly advanced economies
+  mutate(minwbgdppc = ifelse(wbgdppc2011est1 > wbgdppc2011est2,
+                             wbgdppc2011est2, wbgdppc2011est1)) -> Data
 
-tic()
-create_stateyears() %>%
-  add_gwcode_to_cow() %>%
-  add_ucdp_onsets() %>%
-  add_capital_distance() %>%
-  add_contiguity() %>%
-  add_cow_majors() %>%
-  add_cow_trade() %>%
-  add_democracy() %>%
-  add_igos() %>%
-  add_minimum_distance(system = "cow") %>%
-  add_archigos() %>%
-  add_nmc() %>%
-  add_sdp_gdp(system = "cow") 
+# r2sd() is in {stevemisc}, a {peacesciencer} dependency.
+Data %>%
+  mutate_at(vars("cincprop", "mindemest", "minwbgdppc", "minmilit"),
+            ~r2sd(.)) -> Data
+
+summary(modDD <- glm(gmlmidonset ~ landcontig + cincprop + cowmajdyad + cow_defense +
+               mindemest + minwbgdppc + minmilit +
+               gmlmidspell + I(gmlmidspell^2) + I(gmlmidspell^3), data= Data,
+             family=binomial(link="logit")))
 toc()
@@
 ```
+
+Here is how you might do a standard civil conflict analysis using Gleditsch-Ward states and UCDP conflict data.
+
+```{r, cache=F, message=F, eval=T}
+
+tic()
+create_stateyears(system = 'gw') %>%
+  filter(year %in% c(1946:2019)) %>%
+  add_ucdp_acd(type=c("intrastate"), only_wars = FALSE) %>%
+  add_peace_years() %>%
+  add_democracy() %>%
+  add_creg_fractionalization() %>%
+  add_sdp_gdp() %>%
+  add_rugged_terrain() -> Data
+
+create_stateyears(system = 'gw') %>%
+  filter(year %in% c(1946:2019)) %>%
+  add_ucdp_acd(type=c("intrastate"), only_wars = TRUE) %>%
+  add_peace_years() %>%
+  rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .)) %>%
+  left_join(Data, .) -> Data
+
+Data %>%
+  arrange(gwcode, year) %>%
+  group_by(gwcode) %>%
+  mutate_at(vars("xm_qudsest", "wbgdppc2011est",
+                 "wbpopest"), list(l1 = ~lag(., 1))) %>%
+  rename_at(vars(contains("_l1")),
+            ~paste("l1", gsub("_l1", "", .), sep = "_") ) -> Data
+
+modCW <- list()
+summary(modCW$"All UCDP Conflicts" <- glm(ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
+                    l1_xm_qudsest + I(l1_xm_qudsest^2) +
+                    newlmtnest + ethfrac + relfrac +
+                    ucdpspell + I(ucdpspell^2) + I(ucdpspell^3), data=subset(Data),
+                  family = binomial(link="logit")))
+
+summary(modCW$"Wars Only"  <- glm(war_ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
+                    l1_xm_qudsest + I(l1_xm_qudsest^2) +
+                    newlmtnest + ethfrac + relfrac +
+                    war_ucdpspell + I(war_ucdpspell^2) + I(war_ucdpspell^3), data=subset(Data),
+                  family = binomial(link="logit")))
+
+toc()
+```
+
+# Issues/Requests
+
+`{peacesciencer}` is already more than capable to meet a wide variety of needs in the peace science community.
+Users are free to raise an issue on the project's Github if some feature is not performing as they think it should or if there are additions they would like to see.

### 2021-06-11 docs/ms.tex
move user guide to pkgdown directory

--- 2021-06-10 user-guide/doc/ms.tex
@@
 \title{\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: June 10, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
+\textbf{Current version}: June 11, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
 \author{true}
@@
 \title{\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: June 10, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
+\textbf{Current version}: June 11, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
 

### 2021-06-11 README.Rmd
sfsfsfs

@@
 
-tibble(x = db_names, y = gsub("\n","",unlist(db_descript))) %>%
+tibble(x = db_names, y = gsub("\n"," ",unlist(db_descript))) %>%
   mutate(x = str_sub(x, 1, nchar(x)-3)) %>%
@@
 peacesciencer_functions %>%
-  mutate(link = paste0("[",x,"()](http://svmiller.com/peacesciencer/reference/",x,".html)")) -> peacesciencer_functions
+  mutate(link = paste0("[`",x,"()`](http://svmiller.com/peacesciencer/reference/",x,".html)")) -> peacesciencer_functions
 

### 2021-06-13 data-raw/times/create_times.R
safsdfdsaf

@@ -1,0 +1,63 @@
+library(tidyverse)
+library(peacesciencer)
+library(microbenchmark)
+
+# https://stackoverflow.com/questions/10893611/storing-tic-toc-values-in-r
+# Create some data ----
+
+
+create_bench <- microbenchmark::microbenchmark(
+  `create_statedays()` = create_statedays(),
+  `create_stateyears()` = create_stateyears(),
+  `create_dyadyears()` = create_dyadyears(),
+  unit = "s",
+  times = 100)
+
+saveRDS(create_bench, "~/Dropbox/projects/peacesciencer/data-raw/times/create_bench.rds")
+
+# https://stackoverflow.com/questions/10893611/storing-tic-toc-values-in-r
+# library(tictoc)
+# tic.clearlog()
+#
+# # Create Times list ----
+# create_times <- list()
+#
+#
+#
+# # * create_statedays() -----
+# for (x in 1:100) {
+#   # passing x to tic() makes it a label at time of the matching toc() call.
+#   tic(x)
+#   create_statedays()
+#   gc()
+#   toc(log = TRUE, quiet = TRUE)
+# }
+#
+# create_times$"create_statedays()" <- tibble(x = unlist(lapply(tic.log(format = FALSE), function(x) x$toc - x$tic)))
+# tic.clearlog()
+#
+# # * create_stateyears() -----
+#
+# for (x in 1:100) {
+#   tic(x)
+#   create_stateyears()
+#   gc()
+#   toc(log = TRUE, quiet = TRUE)
+# }
+#
+# create_times$"create_stateyears()" <- tibble(x = unlist(lapply(tic.log(format = FALSE), function(x) x$toc - x$tic)))
+# tic.clearlog()
+#
+# # * create_dyadyears() -----
+#
+# for (x in 1:100) {
+#   tic(x)
+#   create_dyadyears()
+#   gc()
+#   toc(log = TRUE, quiet = TRUE)
+# }
+#
+# create_times$"create_dyadyears()" <- tibble(x = unlist(lapply(tic.log(format = FALSE), function(x) x$toc - x$tic)))
+# tic.clearlog()
+#
+# saveRDS(create_times, "~/Dropbox/projects/peacesciencer/data-raw/times/create_times.rds")

### 2021-06-18 README.Rmd
change/update

@@
 
-```{r echo=FALSE, results="hide", message=FALSE}
+```{r setup, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+```
+
+```{r loadstuff, echo=FALSE, results="hide", message=FALSE}
 library("badger")
@@
 
-#library(tidyverse)
-#library(peacesciencer)
-# library(stevemisc)
+# library(tidyverse) # load this first for most/all things
+# library(peacesciencer) # the package of interest
+# library(stevemisc) # a dependency, but also used for standardizing variables for better interpretation
 library(tictoc)
@@
 # r2sd() is in {stevemisc}, a {peacesciencer} dependency.
+# This is just for a more readable regression output.
 Data %>%
@@
 
-summary(modDD <- glm(gmlmidonset ~ landcontig + cincprop + cowmajdyad + cow_defense +
+broom::tidy(modDD <- glm(gmlmidonset ~ landcontig + cincprop + cowmajdyad + cow_defense +
                mindemest + minwbgdppc + minmilit +
@@
 modCW <- list()
-summary(modCW$"All UCDP Conflicts" <- glm(ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
+broom::tidy(modCW$"All UCDP Conflicts" <- glm(ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
                     l1_xm_qudsest + I(l1_xm_qudsest^2) +
@@
 
-summary(modCW$"Wars Only"  <- glm(war_ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
+broom::tidy(modCW$"Wars Only"  <- glm(war_ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
                     l1_xm_qudsest + I(l1_xm_qudsest^2) +

### 2021-06-18 README.Rmd
let 'er rip

@@
 
+# Citing What You Do in `{peacesciencer}`
+
+You can (and should) cite what you do in `{peacesciencer}`.
+The package includes a data frame of a `BibTeX` file (`ps_bib`) and a function for finding and returning `BibTeX` entries that you can include in your projects.
+This is the `ps_cite()` function.
+The `ps_cite()` function takes a string and does a partial match for relevant keywords (as `KEYWORDS`) associated with entries in the `ps_bib` file.
+For example, you can (and should) cite the package itself.
+
+```{r}
+ps_cite("peacesciencer")
+```
+
+You can see what are the relevant citations to consider using for the data returned by `add_democracy()`
+
+```{r}
+ps_cite("add_democracy()")
+```
+
+You can also return partial matches to see what citations are associated with, say, alliance data in this package.
+
+```{r}
+ps_cite("alliance")
+```
+
+This function might expand in complexity in future releases, but you can use it right now for finding appropriate citations. You an also scan the `ps_bib` data to see what is in there.
+
 # Issues/Requests

### 2021-06-18 R/data.R

@@ -329,7 +329,7 @@ NULL
 #' @details Data originally provided by Gleditsch with no column names. Column names were added before some light re-cleaning in order
 #' to generate these data.
 #'
-#' @references Gleditsch, Kristian S. and Michael D. Ward. 1999. "A Revised List of Independent States since the Congress of Vienna" 25(4): 393--413.
+#' @references Gleditsch, Kristian S. and Michael D. Ward. 1999. "A Revised List of Independent States since the Congress of Vienna" International Interactions 25(4): 393--413.
 #'
 "gw_states"
 
@@ -399,18 +399,6 @@ NULL
 "cow_nmc"
 
 
-#' Citations for Data/Functions Used in \code{peacesciencer}
-#'
-#' This is a master list of references for data/functions used in this package. Do check it out to make sure you're faithfully citing what you're using.
-#'
-#' @format A data frame with two variables:
-#' \describe{
-#' \item{\code{data_function}}{the data or function used or referenced in the package}
-#' \item{\code{citation}}{an appropriate (text) citation you should include in your manuscript}
-#' }
-#'
-"citations"
-
 #' Correlates of War Non-Directed Dyad-Year International Governmental Organizations (IGOs) Data
 #'
 #' This is a non-directed dyad-year version of the Correlates of War IGOs data. I use it internally for merging IGOs data into dyad-year data.
@@ -430,7 +418,7 @@ NULL
 #'
 #' @references
 #'
-#' Pevehouse, Jon C.W., Timothy Nordstron, Roseanne W McManus, Anne Spencer Jamison, “Tracking Organizations in the World: The Correlates of War IGO Version 3.0 datasets”, Journal of Peace Research 57(3): 492-503.
+#' Pevehouse, Jon C.W., Timothy Nordstrom, Roseanne W McManus, Anne Spencer Jamison, 2020. “Tracking Organizations in the World: The Correlates of War IGO Version 3.0 datasets”, Journal of Peace Research 57(3): 492-503.
 #'
 #' Wallace, Michael, and J. David Singer. 1970. "International Governmental Organization in the Global System, 1815-1964." International Organization 24: 239-87.
 #'
@@ -457,7 +445,7 @@ NULL
 #'
 #' @references
 #'
[…]

### 2021-06-18 R/ps_cite.R

@@ -1,0 +1,37 @@
+#' Get \code{BibTeX} Entries Associated with `{peacesciencer}` Data and Functions
+#'
+#' @description \code{ps_cite()} allows the user to get citations to scholarship that they
+#' should include in their papers that incorporate the functions and data in this package.
+#'
+#' @return \code{ps_cite()} takes a character vector and scans the \code{ps_bib} data in
+#' this package to return a \code{BibTeX} citation (or citations) that they should use
+#' to properly cite the material they are using (as generated in this package). The citations
+#' are returned as a full \code{BibTeX} entry (or entries) that they can copy-paste into their
+#' own \code{BibTeX} file.
+#'
+#' @details The base functionality here is simple pattern-matching on keywords in \code{ps_bib}. This
+#' simple pattern-matching is in base R. I assume the user has some familiarity with \code{BibTeX}.
+#'
+#' @author Steven V. Miller
+#'
+#' @param x a character vector
+#'
+#'
+#' @examples
+#'
+#'
+#' # You can do partial matching
+#' ps_cite("democracy")
+#'
+#' # Or more partial matching
+#' ps_cite("alliance")
+#'
+#' # You can also get all citations for a particular function
+#' ps_cite("add_archigos()")
+#'
+
+ps_cite <- function(x) {
+
+  ps_bib %>% filter(grepl(x, KEYWORDS)) %>% df2bib()
+
+}

### 2021-06-18 data-raw/bib/ps_bib.R

@@ -1,0 +1,17 @@
+library(bib2df)
+
+ps_bib <- bib2df("~/Dropbox/projects/peacesciencer/data-raw/bib/ps_bib.bib") %>% filter(CATEGORY != "COMMENT")
+
+save(ps_bib, file="data/ps_bib.rda")
+
+df2bib(ps_bib)
+
+ps_cite <- function(x) {
+
+  ps_bib %>% filter(grepl(x, KEYWORDS)) %>% df2bib()
+
+}
+
+
+ps_cite("alliance") # do a partial match
+ps_cite("add_archigos()") # do an exact match

### 2021-06-21 docs/ms.tex
update the user guide for resubmission

@@
 \title{\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: June 11, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
+\textbf{Current version}: June 21, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
 \author{true}
@@
 \title{\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: June 11, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
+\textbf{Current version}: June 21, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
 
@@
 This data feature tackles a recurring problem for researchers in the peace science community.
-True research reproducibility is achieved creating original dyad-year or state-year data from scratch, though no published guide exists that informs how to do this on their own.
+True research reproducibility is achieved creating original dyad-year or state-year data from scratch, though no published guide exists that informs researchers how to do this on their own.
 Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice.
-In a related manner, researchers may spend too much time reproducing old code for standard information that goes into any dyad-year or state-year analysis---like contiguity relationships and democracy, among many others---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes.
+Researchers may additionally spend too much time reproducing old code for standard information that goes into any dyad-year or state-year analysis---like contiguity relationships and democracy---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes.
 This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well.
@@
 All are done with the maximum possible transparency.
-The project is available for public view of Github (\url{https://github.com/svmiller/peacesciencer}).
+The project is available for public view of Github (\url{https://github.com/svmiller/peacesciencer/}).
 The \texttt{data-raw} directory on the project's Github contains information and comments about how every data set was created and the \texttt{R} directory informs the user how each function works.
@@
 \texttt{\{peacesciencer\}} is primarily born from this question about how to create dyad-year and state-year data from scratch.
-The underlying code that produces these data types is available online and are converted into simple functions for the ease of the researcher.
+The underlying code that produces these data types is available online and \texttt{\{peacesciencer\}} converts these lines of code into simple functions for the ease of the researcher.
 
@@
 
-Third, the creation of the data and the analysis of the data is increasingly becoming one continuous process.
+Third, the creation of the data and the analysis of the data are increasingly becoming one continuous process.
 Not too long ago, it used to be the case that researchers had to download a data set, or create one from scratch (possibly in a spreadsheet or through a program like EUGene).
@@
 
-Fifth, statistical software should be kind to graduate students in peace science with the aforementioned point in mind.
+Fifth, statistical software should be kind to graduate students in peace science with the aforementioned points in mind.
 Graduate students just beginning their career in peace science scholarship are tasked to learn the substantive material, the characteristics of the underlying data in wide use in the community, and the quantitative methods and statistical software available to them, effectively all at the same time.
@@
 
-\texttt{\{peacesciencer\}} comes with a battery of built-in functions for generating the most widespread forms of peace science data.
+\texttt{\{peacesciencer\}} comes with a fully developed suite of built-in functions for generating the most widespread forms of peace science data.
 Each function uses raw or pre-processed data included in \texttt{\{peacesciencer\}}.
@@
 It has an original data set on capital-to-capital distances, calculating for distance in kilometers between state capitals using the ``Vincenty'' method (i.e.~``as the crow flies'') that also accounts for instances when capitals moved (e.g.~Brazil in 1960, Burundi in 2018).
-It also features an innovation of democracy data, providing reasonable estimates of democracy using the Marquez (2016) of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve \& Melton, 2010).\footnote{A vignette on the \texttt{\{peacesciencer\}} website explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.} As of writing, \texttt{\{peacesciencer\}} comes with the following functionality.
+It also features an innovation of democracy data, providing reasonable estimates of democracy using the Marquez (2016) method of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve \& Melton, 2010).\footnote{\texttt{\{peacesciencer\}} comes with a vignette that explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.} As of writing, \texttt{\{peacesciencer\}} comes with the following functionality described in Table \ref{tab:tab-functionality}.
 
-\begingroup\fontsize{7}{9}\selectfont
+\begin{table}
 
-\begin{longtable}[t]{>{}l|>{}l|>{}c|>{}c}
-\caption{\label{tab:unnamed-chunk-1}Features of \{peacesciencer\}}\\
+\caption{\label{tab:tab-functionality}Features of \{peacesciencer\}}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}[t]{>{}l|>{}l|>{}c|>{}c}
 \hline
@@
 \hline
-\end{longtable}
-\endgroup{}
+\end{tabular}
+\end{table}
 
@@
 \texttt{\{peacesciencer\}} can do something like this in the following pipe-based workflow.
-First, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)} will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016.
+First, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)} will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016.\footnote{The conflict data in Gibler, Miller \& Little (2016) do not extend past 2010, but asking for dyad-years from 2010-2016 will just create harmless \texttt{NA}s that will not appear in any analysis.}
 
@@
 The Beck, Katz \& Tucker (1998) cubic splines approach is \texttt{...\ +\ splines::bs(gmlmidspell,\ 4)}.
-This function includes the spell and three splines (hence the 4 in the command).} \texttt{add\_nmc()} will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity.
+This function includes the spell and three splines (hence the 4 in the command).
+Either approach makes for easier model predictions, given R's functionality.} \texttt{add\_nmc()} will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity.
 \texttt{add\_democracy()} will add information about the level of democracy for both states in the dyad using three different indicators.

### 2021-06-21 user-guide/ms.Rmd
update the user guide for resubmission

@@
 All are done with the maximum possible transparency.
-The project is available for public view of Github (https://github.com/svmiller/peacesciencer).
+The project is available for public view of Github (`r params$ps_github`).
 The `data-raw` directory on the project's Github contains information and comments about how every data set was created and the `R` directory informs the user how each function works.
@@
 `{peacesciencer}` is primarily born from this question about how to create dyad-year and state-year data from scratch.
-The underlying code that produces these data types is available online and are converted into simple functions for the ease of the researcher.
+The underlying code that produces these data types is available online and `{peacesciencer}` converts these lines of code into simple functions for the ease of the researcher.
 
@@
 
-`{peacesciencer}` comes with a battery of built-in functions for generating the most widespread forms of peace science data.
+`{peacesciencer}` comes with a fully developed suite of built-in functions for generating the most widespread forms of peace science data.
 Each function uses raw or pre-processed data included in `{peacesciencer}`.
@@
 Brazil in 1960, Burundi in 2018).
-It also features an innovation of democracy data, providing reasonable estimates of democracy using the @marquez2016qme of extending the Unified Democracy Scores (UDS) data [@pemsteinetal2010dc].[^demvig] As of writing, `{peacesciencer}` comes with the following functionality.
+It also features an innovation of democracy data, providing reasonable estimates of democracy using the @marquez2016qme method of extending the Unified Democracy Scores (UDS) data [@pemsteinetal2010dc].[^demvig] As of writing, `{peacesciencer}` comes with the following functionality described in Table \@ref(tab:tab-functionality).
 
-[^demvig]: A vignette on the `{peacesciencer}` website explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.
+[^demvig]: `{peacesciencer}` comes with a vignette that explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.
 
-```{r, echo=F, eval=T}
+```{r tab-functionality, echo=F, eval=T}
 tribble(~data, ~description, ~type, ~func,
@@
 if (is_latex) {
-kable(included, booktabs = FALSE, longtable = TRUE,
+kable(included, #booktabs = FALSE, longtable = TRUE,
     caption = "Features of \\{peacesciencer\\}",
@@
     col.names = c("Data", "Description", "Type", "Function")) %>%
-  kable_styling(font_size = 7) %>%
+  kable_styling(font_size = 8) %>%
   row_spec(0, bold=TRUE) %>%
@@
 `{peacesciencer}` can do something like this in the following pipe-based workflow.
-First, `create_dyadyears(directed = FALSE, mry = FALSE)` will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016.
+First, `create_dyadyears(directed = FALSE, mry = FALSE)` will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016.[^after2010]
+
+[^after2010]: The conflict data in @gibleretal2016amid do not extend past 2010, but asking for dyad-years from 2010-2016 will just create harmless `NA`s that will not appear in any analysis.
 
@@
 This function includes the spell and three splines (hence the 4 in the command).
+Either approach makes for easier model predictions, given R's functionality.
 
@@
                              wbgdppc2011est2, wbgdppc2011est1)) -> Data
-
 ```

### 2021-10-08 README.Rmd
oops
another one

@@
   left_join(peacesciencer_functions, .) %>%
+  # wc_ is a shortcut
+   filter(!str_starts(x, "wc_")) %>%
   mutate(func = paste0(x,"()")) -> peacesciencer_functions
@@
 
-[`{peacesciencer}` has a user's guide](https://github.com/svmiller/peacesciencer/blob/master/user-guide/doc/ms.pdf) that is worth reading.
+[`{peacesciencer}` has a user's guide](http://svmiller.com/peacesciencer/ms.pdf) that is worth reading.
 The workflow is going to look something like this.

### 2021-10-10 README.Rmd
mostly website-related updates

@@
 
-# What's Included in `{peacesciencer}`
-
-The package is already well developed and its functionality continues to expand. The current development version has the following functions.
-
-```{r, echo=F, message=F, warning=F}
-peacesciencer_functions <- tibble(x = as.vector(lsf.str("package:peacesciencer")))
-
-db <- tools::Rd_db("peacesciencer")
-db_names <- names(db)
-db_descript <- lapply(db,function(x) {
-    tags <- tools:::RdTags(x)
-    if("\\description" %in% tags){
-      out <- paste(unlist(x[which(tags=="\\description")]),collapse="")
-    }
-    else
-      out <- NULL
-    invisible(out)
-  })
-
-tibble(x = db_names, y = gsub("\n"," ",unlist(db_descript))) %>%
-  mutate(x = str_sub(x, 1, nchar(x)-3)) %>%
-  left_join(peacesciencer_functions, .) %>%
-  # wc_ is a shortcut
-   filter(!str_starts(x, "wc_")) %>%
-  mutate(func = paste0(x,"()")) -> peacesciencer_functions
-
-peacesciencer_functions %>%
-  mutate(link = paste0("[`",x,"()`](http://svmiller.com/peacesciencer/reference/",x,".html)")) -> peacesciencer_functions
-
-tibble(`**Function**` = peacesciencer_functions$link,
-       `**Description**` = peacesciencer_functions$y) %>%
-  arrange(`**Function**`) %>%
-  kable(., format="markdown")
-
-```
-
-The current development version also includes the following data.
-
-```{r, echo=F, message=F, warning=F}
-# library(tidyverse)
-# library(kableExtra)
-peacesciencer_data <- data(package = "peacesciencer")
-
-object_name <- paste0("`",peacesciencer_data$results[, "Item"],"`")
-object_link <- paste0("[",object_name,"](http://svmiller.com/peacesciencer/reference/",peacesciencer_data$results[, "Item"],".html)")
-
-tibble(`**Object Name**` = object_link,
-       `**Description**` = peacesciencer_data$results[, "Title"]) %>%
-  arrange(`**Object Name**`) %>%
-  kable(., format="markdown")
-
-```
-
-<!-- The package is very much a work in progress. Right now, it has the following functions: -->
-
-<!-- - `add_archigos()`: adds some summary variables from Archigos about political leaders for state-year or dyad-year data. -->
-<!-- - `add_atop_alliance()`: adds ATOP alliance information to dyad-year data. -->
-<!-- - `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year or state-year data. -->
-<!-- - `add_ccode_to_gw()`: 	Add Correlates of War state system codes to dyad-year or state-year data with Gleditsch-Ward state codes. -->
-<!-- - `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year or state-year data. -->
-<!-- - `add_cow_alliance()`: adds Correlates of War major alliance information to dyad-year data. -->
-<!-- - `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data. -->
-<!-- - `add_cow_mids()`: adds Correlates of War (CoW) Militarized Interstate Dispute (MID) data to dyad-year data frame -->
-<!-- - `add_cow_trade()`: adds Correlates of War trade data to dyad-year or state-year data. -->
-<!-- - `add_creg_fractionalization()`: add fractionalization/polarization estimates from CREG to your dyad-year or state-year data -->
-<!-- - `add_democracy()`: adds estimates of democracy/levels of democracy to dyad-year or state-year data. -->
-<!-- - `add_gwcode_to_cow():` adds Gleditsch-Ward state codes to dyad-year or state-year data with Correlates of War state codes. -->
-<!-- - `add_gml_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data. -->
-<!-- - `add_igos()`: adds Correlates of War International Governmental Organizations (IGOs) data to dyad-year or state-year data. -->
-<!-- - `add_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data. (DEPRECATED) -->
-<!-- - `add_minimum_distance()`: adds minimum distance (in kilometers) to dyad-year or state-year data. -->
-<!-- - `add_nmc()`: adds estimates of national material capabilities (from Correlates of War) to dyad-year or state-year data. -->
-<!-- - `add_peace_years()`: Add Peace Years to Your Conflict Data -->
-<!-- - `add_sdp_gdp()`: adds estimates of (gross and surplus) domestic product and population size to dyad-year or state-year data -->
-<!-- - `add_strategic_rivalries()`: Add Thompson and Dreyer's (2012) strategic rivalry data to dyad-year data frame -->
-<!-- - `add_ucdp_acd()`: Add UCDP Armed Conflict Data to state-year data frame -->
-<!-- - `add_ucdp_onsets()`: adds UCDP onsets to state-year data -->
-<!-- - `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format -->
-<!-- - `create_statedays()`: converts Correlates of War or Gleditsch-Ward state membership data into state-day format.. -->
-<!-- - `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format. -->
-<!-- - `filter_prd()`: filters dyad-year data frame to just "politically relevant" dyads. -->
-
-<!-- It also has the following data sets: -->
-
-<!-- - `archigos`: an abbreviated version of the Archigos data, used internally -->
-<!-- - `atop_alliance:` directed dyad-year alliance data from ATOP -->
-<!-- - `capitals`: a list of capitals and capital transitions for Correlates of War state system members -->
-<!-- - `ccode_democracy`: Correlates of War state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores").
--->
-<!-- - `cow_alliance`: directed dyad-year alliance formation data from the Correlates of War -->
-<!-- - `cow_contdir`: Correlates of War Direct Contiguity Data (v. 3.2) -->
-<!-- - `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members -->
-<!-- - `cow_gw_years`: a yearly data frame including information about Correlates of War and Gleditsch-Ward states. -->
-<!-- - `cow_igo_ndy`: non-directed dyad-year data for Correlates of War intergovernmental organizations data. -->
-<!-- - `cow_igo_sy`: state-year data for Correlates of War intergovernmental organizations data. -->
-<!-- - `cow_majors`: Correlates of War major powers data (version: 2016) -->
-<!-- - `cow_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (CoW-MID, v. 5.0) -->
-<!-- - `cow_mid_dirdisps`: Correlates of War Militarized Interstate Dispute data (directed dyadic dispute-year, v. 5.0)  -->
-<!-- - `cow_mid_disps`: Correlates of War Militarized Interstate Dispute data (abbreviated dispute-level, v. 5.0)  -->
-<!-- - `cow_mindist`: the minimum distance between Correlates of War states, in kilometers, in non-directed dyad-year format (1946-2015) -->
-<!-- - `cow_nmc`: Correlates of War National Material Capabilities data (version 5.0) -->
-<!-- - `cow_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States -->
-<!-- - `cow_states`: Correlates of War state system membership data (version: 2016) -->
-<!-- - `cow_trade_sy`: Correlates of War state-year trade data (version 4.0) -->
-<!-- - `creg`: Composition of Religious and Ethnic Groups (CREG) Fractionalization/Polarization Estimates -->
-<!-- - `gml_dirdisp`: directed dispute-year data from version 2.1.1 of the Gibler-Miller-Little inter-state dispute data. -->
-<!-- - `gml_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (Gibler-Miller-Little, v. 2.1.1) -->
-<!-- - `gw_ddy`: a full directed dyad-year data frame of Gleditsch-Ward state system members -->
-<!-- - `gw_mindist`: the minimum distance between  Gleditsch-Ward states, in kilometers, in non-directed dyad-year format (1946-2015) -->
-<!-- - `gw_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States -->
-<!-- - `gw_states`: Gleditsch-Ward independent state system data (version: 2017) -->
-<!-- - `gwcode_democracy`: Gleditsch-Ward state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores").
--->
-<!-- - `hief`: Historical Index of Ethnic Fractionalization data -->
-<!-- - `maoz_powers`: Zeev Maoz' global/regional power data -->
-<!-- - `rugged`: Rugged/Mountainous Terrain Data -->
-<!-- - `td_rivalries`: Thompson and Dreyer's (2012) Strategic Rivalries, 1494-2010 -->
-<!-- - `ucdp_acd`: a (not quite) dyad-year and (not quite) state-year data set on armed conflict episodes -->
-<!-- - `ucdp_onsets`: a state-year data set of UCDP armed conflict onsets/episodes. -->
-
 # How to Use `{peacesciencer}`
 
+New users should read two things to get started.
+[The package's website](http://svmiller.com/peacesciencer/index.html) has an exhaustive list and description of [all the functions and data included](http://svmiller.com/peacesciencer/reference/index.html) in the package.
 [`{peacesciencer}` has a user's guide](http://svmiller.com/peacesciencer/ms.pdf) that is worth reading.
+The user's guide points to its potential uses and benefits while also offering some encouragement for those completely new to the R programming language.
+The package is designed to be accessible to those with no prior experience with R, though completely new users who feel lost or overwhelmed should learn about [the "tidy" approach to R](https://www.tidyverse.org/) to help them get started.
+
 The workflow is going to look something like this.
-This is a "tidy"-friendly approach to a data-generating process in quantitative peace science.
-
 First, start with one of two processes to create either dyad-year or state-year data.
@@
 
-All told, the process will look something like this.
+Here's what this will look like in operation.
 Assume you want to create some data for something analogous to a "dangerous dyads" design for all non-directed dyad-years.

### 2021-10-13 R/data.R
gonna try and push this to CRAN, and shoot for 1.0.0 perhaps with the R&R.

@@ -629,14 +629,16 @@ NULL
 #' Archigos: A (Subset of a) Dataset on Political Leaders
 #'
 #' These are leader-level data drawn from the Archigos data. Space considerations mean I offer here just a few columns
-#' based on these data.
+#' based on these data.  Data are version 4.1.
 #'
 #'
-#' @format A data frame with 3409 observations on the following seven variables.
+#' @format A data frame with 3409 observations on the following 10 variables.
 #' \describe{
 #' \item{\code{ccode}}{a numeric vector for the Correlates of War state code}
+#' \item{\code{obsid}}{a character vector for observation ID}
 #' \item{\code{leadid}}{the unique leader identifier}
 #'  \item{\code{leader}}{the leader name}
+#'   \item{\code{yrborn}}{the year the leader was born}
 #'   \item{\code{gender}}{a categorical variable for leader gender ("M" for men, "W" for women)}
 #' \item{\code{startdate}}{a date for the leader start date}
 #' \item{\code{enddate}}{a date for the leader end date}
@@ -645,12 +647,22 @@ NULL
 #' \item{\code{exitcode}}{a character vector for more information about the leader's exit type}
 #' }
 #' @details Space considerations mean I can only offer a few columns from the overall data. Archigos data are rich with information. Consult
-#' the raw data available on Hein Goeman's website for more. Data are version 4.1.
+#' the raw data available on Hein Goeman's website for more.
+#'
+#' To best conform with data requirements on CRAN, a few leader names were renamed if they included irregular
+#' characters (e.g. umlauts or accents). These leaders are "(Juan Orlando) Hernández" (`HON-2014`),
+#' "(Antonio) Saca González" (`SAL-2004`), "Julián Trujillo Largacha" (`COL-1878`), "César Gaviria Trujillo"
+#' (`COL-1990`), "Gabriel García Moreno" (`ECU-1869`), "Marcos A. Morínigo" (`PAR-1894-1`), "Higinio Morínigo" (`PAR-1940`),
+#' "Sebastián Piñera" (`CHL-2010`), "Sauli Niinistö" (`FIN-2012`), "Louis Gerhard De Geer" (`SWD-1876`),
+#' "Stefan Löfven" (`SWD-2014`), "Lars Løkke Rasmussen" (`DEN-2009`, `DEN-2015`), and "Fernando de Araújo"
[…]

### 2021-10-13 data-raw/bib/ps_bib.R

@@ -3,6 +3,14 @@ library(tidyverse)
 
 ps_bib <- bib2df("~/Dropbox/projects/peacesciencer/data-raw/bib/ps_bib.bib") %>% filter(CATEGORY != "COMMENT")
 
+# I'm going to have to do this iteratively, methinks...
+
+ps_bib %>% mutate_if(is.character,
+                     list(enc = ~stringi::stri_enc_isascii(.))) %>%
+  select(CATEGORY_enc:ncol(.)) %>%
+  summary
+# ^ I think I got it...
+
 save(ps_bib, file="data/ps_bib.rda")
 
 #df2bib(ps_bib)

### 2021-11-01 data-raw/false_cow_dyads.R
small updates

@@ -1,0 +1,67 @@
+library(peacesciencer)
+library(lubridate)
+library(tidyverse)
+
+state_days <- create_statedays()
+
+
+half_cores  <- parallel::detectCores()/2
+library(foreach)
+my.cluster <- parallel::makeCluster(
+  half_cores,
+  type = "PSOCK"
+)
+
+doParallel::registerDoParallel(cl = half_cores)
+foreach::getDoParRegistered()
+library(tictoc)
+
+tic()
+allofthem <- foreach(
+  i = 1816:2016
+) %dopar% {
+
+  state_days %>%
+    filter(year(date) == i) %>%
+    expand(ccode1 = ccode, ccode2 = ccode, date = seq(make_date(i,01,01), make_date(i,12,31), by='1 day')) %>%
+    filter(ccode1 != ccode2) %>%
+    semi_join(., state_days, by=c("ccode1"="ccode", "date"="date")) %>%
+    semi_join(., state_days, by=c("ccode2"="ccode", "date"="date")) %>%
+    mutate(year = year(date)) %>%
+    distinct(ccode1, ccode2, year) -> hold_this
+
+}
+toc()
+parallel::stopCluster(cl = my.cluster)
+
+allofthem %>%
+  bind_rows(.) -> actualdata
+
+actualdata %>% mutate(actual = 1) -> actualdata
+
+cow_states %>%
+  # Select just the stuff we need
+  select(ccode, styear, endyear) %>%
+  # Expand the data, create two ccodes as well
+  expand(ccode1=ccode, ccode2=ccode, year=seq(1816,2016)) %>%
+  # Filter out where ccode1 == ccode2
+  filter(ccode1!=ccode2) %>%
+  # When you're merging into dyad-year data, prepare to do it twice.
+  # Basically: merge in data for ccode1
+  left_join(., cow_states, by=c("ccode1"="ccode")) %>%
+  # ...and filter out cases where the years don't align.
+  filter(year >= styear & year <= endyear) %>%
+  # Get rid of styear and endyear to do it again.
+  select(-styear,-endyear) %>%
+  # And do it again, this time for ccode2
+  left_join(., cow_states, by=c("ccode2"="ccode")) %>%
+  # Again, filter out cases where years don't align.
+  filter(year >= styear & year <= endyear) %>%
+  # And select just what we need.
+  select(ccode1, ccode2, year) -> DDY
+
[…]

### 2021-11-01 data-raw/false_gw_dyads.R

@@ -1,0 +1,73 @@
+library(peacesciencer)
+library(lubridate)
+library(tidyverse)
+
+state_days <- create_statedays(system = 'gw')
+
+
+half_cores  <- parallel::detectCores()/2
+library(foreach)
+my.cluster <- parallel::makeCluster(
+  half_cores,
+  type = "PSOCK"
+)
+
+doParallel::registerDoParallel(cl = half_cores)
+foreach::getDoParRegistered()
+library(tictoc)
+
+
+tic()
+allofthem <- foreach(
+  i = 1816:2017
+) %dopar% {
+
+  state_days %>%
+    filter(year(date) == i) %>%
+    expand(gwcode1 = gwcode, gwcode2 = gwcode, date = seq(make_date(i,01,01), make_date(i,12,31), by='1 day')) %>%
+    filter(gwcode1 != gwcode2) %>%
+    semi_join(., state_days, by=c("gwcode1"="gwcode", "date"="date")) %>%
+    semi_join(., state_days, by=c("gwcode2"="gwcode", "date"="date")) %>%
+    mutate(year = year(date)) %>%
+    distinct(gwcode1, gwcode2, year) -> hold_this
+
+}
+toc()
+
+parallel::stopCluster(cl = my.cluster)
+
+allofthem %>%
+  bind_rows(.) -> actualdata
+
+actualdata %>% mutate(actual = 1) -> actualdata
+
+gw_states %>%
+  mutate(styear = lubridate::year(.data$startdate),
+         endyear = lubridate::year(.data$enddate))  -> gw_states
+
+gw_states %>%
+  # Select just the stuff we need
+  select(.data$gwcode, .data$styear, .data$endyear) %>%
+  # Expand the data, create two ccodes as well
+  expand(gwcode1=.data$gwcode, gwcode2=.data$gwcode, year=seq(1816,2017)) %>%
+  # Filter out where ccode1 == ccode2
+  filter(.data$gwcode1 != .data$gwcode2) %>%
+  left_join(., gw_states, by=c("gwcode1"="gwcode")) %>%
+  # ...and filter out cases where the years don't align.
+  filter(.data$year >= .data$styear & .data$year <= .data$endyear) %>%
+  # Get rid of styear and endyear to do it again.
+  select(-.data$styear,-.data$endyear) %>%
+  # And do it again, this time for ccode2
+  left_join(., gw_states, by=c("gwcode2"="gwcode")) %>%
[…]

### 2021-11-02 data-raw/informal-tests/informal-tests.Rmd
update. I think the leader contributions are sufficiently tested.

---
title: "More Tests"
author: "steve"
date: "`r Sys.Date() `"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Check if Stuff is Already Downloaded

```{r}
library(tidyverse)
library(peacesciencer)
library(testthat)
download_extdata()
```

## Create Some Data

```{r}
cow_leaderyears <- create_leaderyears(standardize_cow = TRUE)
gw_leaderyears <- create_leaderyears(standardize_cow = FALSE)

cow_leaderdyadyears <- create_leaderdyadyears(system = 'cow')
gw_leaderdyadyears <- create_leaderdyadyears(system = 'gw')

cow_leadermids <- cow_leaderyears %>% add_gml_mids() %>% filter(year %in% c(1875:2010))
cow_leaderdyadmids <- cow_leaderdyadyears %>% add_gml_mids() %>% filter(year %in% c(1875:2010))
```

## Run Some Tests

These are tests that add data based on what's stored remotely. I don't trust CRAN to do these tests well---and already so much is tested---so I'll communicate the results of those tests here. The leader-year tests are here too. While these could be done as part of the formal testing process, they would also take forever and the time from `devtools::check()` to `devtools::release()` is already like half an hour or more. In due time, these leader-year tests could---and probably should---go into the formal testing procedure.

```{r}
test_that("Dyad-year additions do not create duplicates", {
  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_trade()))
})
```

```{r}
test_that("Leader-year additions do not create duplicates", {
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_capital_distance()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_capital_distance()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_contiguity()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_cow_majors()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_cow_trade()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_creg_fractionalization()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_creg_fractionalization()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_democracy()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_democracy()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_gml_mids()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_minimum_distance()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_minimum_distance()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_nmc()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_rugged_terrain()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_rugged_terrain()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_sdp_gdp()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_sdp_gdp()))
  expect_equal(nrow(cow_leadermids), nrow(cow_leadermids %>% add_spells()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_gwcode_to_cow()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_ccode_to_gw()))
})
```

```{r}
test_that("Leader-dyad-year additions do not create duplicates", {
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_capital_distance()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_capital_distance()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_contiguity()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_cow_majors()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_cow_trade()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_creg_fractionalization()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_creg_fractionalization()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_democracy()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_democracy()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_gml_mids()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_minimum_distance()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_minimum_distance()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_nmc()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_rugged_terrain()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_rugged_terrain()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_sdp_gdp()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_sdp_gdp()))
  expect_equal(nrow(cow_leaderdyadmids), nrow(cow_leaderdyadmids %>% add_spells()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_gwcode_to_cow()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_ccode_to_gw()))
})
```

### 2021-11-02 R/add_capital_distance.R

@@ -1,15 +1,15 @@
 globalVariables(c('.', 'capitals', 'cow_ddy'))
 
-#' Add capital-to-capital distance to a dyad-year or state-year data frame
+#' Add capital-to-capital distance to a data frame
 #'
 #' @description \code{add_capital_distance()} allows you to add capital-to-capital
-#' distance to a dyad-year or state-year data frame. The capitals are coded in the
+#' distance to a (dyad-year, leader-year, leader-dyad-year, state-year) data frame. The capitals are coded in the
 #' \code{cow_capitals} and \code{gw_capitals}
 #' data frames, along with their latitudes and longitudes. The distance variable that
 #' emerges \code{capdist} is calculated using the "Vincenty" method (i.e. "as the crow
 #' flies") and is expressed in kilometers.
 #'
-#' @return \code{add_capital_distance()} takes a dyad-year or state-year data frame and adds the
+#' @return \code{add_capital_distance()} takes a (dyad-year, leader-year, leader-dyad-year, state-year) data frame and adds the
 #' capital-to-capital distance between the first state and the second state (in dyad-year data) or the minimum
 #' capital-to-capital distance for a given state in a given year. A minor
 #' note about this function: cases of capital transition are recorded in the
@@ -17,13 +17,12 @@ globalVariables(c('.', 'capitals', 'cow_ddy'))
 #' merging into a dyad-year data frame), the Jan. 1 capital is used for calculating
 #' distances.
 #'
-#' @details The function leans on attributes of the data that are provided by the \code{create_dyadyear()} or
-#' \code{create_stateyear()} function. Make sure that function (or data created by that function) appear at the top
-#' of the proverbial pipe.
+#' @details The function leans on attributes of the data that are provided by one of the "create" functions
+#' in this package (e.g. \code{create_dyadyear()} or \code{create_stateyear()}).
 #'
 #' @author Steven V. Miller
 #'
-#' @param data a dyad-year data frame (either "directed" or "non-directed") or state-year data frame
+#' @param data a data frame with appropriate \pkg{peacesciencer} attributes
 #'
 #' @examples
 #'
@@ -116,7 +115,7 @@ globalVariables(c('.', 'capitals', 'cow_ddy'))
 
 add_capital_distance <- function(data) {
 
[…]

### 2021-11-02 R/add_ccode_to_gw.R

@@ -1,10 +1,10 @@
-#' Add Correlates of War state system codes to dyad-year or state-year data with Gleditsch-Ward state codes.
+#' Add Correlates of War state system codes to your data with Gleditsch-Ward state codes.
 #'
 #' @description \code{add_ccode_to_gw()} allows you to match, as well as one can, Correlates of War system membership data
 #' with Gleditsch-Ward system data.
 #'
 #'
-#' @return \code{add_ccode_to_gw()} takes a dyad-year data frame or state-year data frame that already has Gleditsch-Ward
+#' @return \code{add_ccode_to_gw()} takes a (dyad-year, leader-year, leader-dyad-year, state-year) data frame that already has Gleditsch-Ward
 #' state system codes and adds their corollary Correlates of War codes.
 #'
 #' @details The \code{data-raw} directory on the project's Github contains more information about the underlying data that assists
@@ -22,7 +22,7 @@
 #'
 #' @author Steven V. Miller
 #'
-#' @param data a dyad-year data frame (either "directed" or "non-directed") or a state-year data frame.
+#' @param data a data frame with appropriate \pkg{peacesciencer} attributes
 #'
 #'
 #' @examples
@@ -37,9 +37,20 @@
 #'
 add_ccode_to_gw <- function(data) {
 
+  if (any(i <- c("ccode1", "ccode2", "ccode") %in% colnames(data))) {
+
+    stop("Your data already appear to have Correlates of War state codes in it.")
+
+  }
+
+  if (length(attributes(data)$ps_system) == 0 | attributes(data)$ps_system != "gw") {
+    warning("The state system data here do not appear to be Gleditsch-Ward, or at least not declared as such. The function will still run, but you may want to inspect the output.")
+
+  }
+
   gw_cow_years %>% select(.data$gwcode, .data$ccode, .data$year) -> hold_this
 
-  if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type == "dyad_year") {
+  if (length(attributes(data)$ps_data_type) > 0 && attributes(data)$ps_data_type %in% c("dyad_year", "leader_dyad_year")) {
 
     if (!all(i <- c("gwcode1", "gwcode2") %in% colnames(data))) {
 
@@ -60,7 +71,7 @@ add_ccode_to_gw <- function(data) {
 
 
 
[…]

### 2021-11-02 R/add_cow_alliance.R

@@ -1,4 +1,4 @@
-#' Add Correlates of War alliance data to a dyad-year data frame
+#' Add Correlates of War alliance data to a data frame
 #'
 #' @description \code{add_cow_alliance()} allows you to add Correlates of War alliance
 #' data to a dyad-year data frame
@@ -19,7 +19,7 @@
 #'
 #' @author Steven V. Miller
 #'
-#' @param data a dyad-year data frame (either "directed" or "non-directed")
+#' @param data a dyad-year or leader-dyad-year data frame (either "directed" or "non-directed")
 #'
 #' @references
 #'

### 2021-11-05 data-raw/informal-tests/informal-tests.Rmd
update tests
update informal tests, and documentation

@@
 ```{r}
-cow_leaderyears <- create_leaderyears(standardize_cow = TRUE)
-gw_leaderyears <- create_leaderyears(standardize_cow = FALSE)
+cow_leaderyears <- create_leaderyears(standardize = "cow")
+gw_leaderyears <- create_leaderyears(standardize = "gw")
 
@@
   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_trade()))
+  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_fpsim()))
 })

### 2021-11-05 user-guide/appendix.Rmd
prepare appendix and memo for revisions

---
title: "APPENDIX: `{peacesciencer}`: An R Package for Quantitative Peace Science Research"
# author:
# - name: Steven V. Miller
#   affiliation: Clemson University
#   email: svmille@clemson.edu
# thanks: "This is a work in progress. Feedback welcome: svmille@clemson.edu."
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`"
biblio-style: apsr
geometry: margin=1in
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
keywords: "software, statistical analysis, peace science, reproducibility"
endnotes: FALSE
csl: inst/jpr.csl
mainfont: cochineal
sansfont: Linux Biolinum O
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
appendix: TRUE
appendixletter: "A"
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
linkcolor: black
header-includes:
  - \usepackage{longtable}
  - \usepackage{csquotes}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \renewcommand{\thepage}{A-\arabic{page}}
---

```{r setup, include=FALSE}
is_docx <- knitr::pandoc_to("docx")
is_latex <- knitr::pandoc_to("latex")

table_format <- ifelse(is_docx, "huxtable", 'default')

# I don't know how Texas Instruments smart this is, but p-sure default DPI is 96.
# That's not a problem for LaTeX, but it looks not-so-great for Word.
# For Word, let's up that to 600. This should allow for cross-referencing 
# in {bookdown} while allowing for conditional DPI

if (is_latex) {
  conditional_dpi <- 96
  return(conditional_dpi)
} else {
  conditional_dpi <- 600
  return(conditional_dpi)
}

knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE, warning=FALSE,
                      fig.path='doc/figs/',
                      cache.path = 'doc/_cache/',
                      fig.width = 8.5, dpi = conditional_dpi,
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      x2 = paste0(stringr::str_extract(x2, '.*/'), 
                                  "appendix-", 
                                  stringr::str_remove(x2, ".*/"))
                      if (file.rename(x, x2)) x2 else x
                      })

library(tidyverse)

options(knitr.kable.NA = '')
```

\tableofcontents
\newpage

# A Section with a Citation to Get Started

@miller2017etjc

\newpage

# References

\setlength{\parskip}{6pt}

### 2021-11-05 user-guide/memo.Rmd
prepare appendix and memo for revisions

---
# output: 
#   stevetemplates::cover_letter:
#     latex_engine: xelatex 
geometry: margin=1in
header-includes:
  - \linespread{1.05}

author: "&nbsp;"
address: ""
email: ""
url: ""
phone: ""
fontsize: 11pt
mainfont: cochineal
monofont: Fira Code

fontawesome: TRUE
date: "`r format(Sys.time(), '%e %B %Y')`"
greetings: "Dear Editor and Reviewers,"
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`"
biblio-style: apsr
---

```{r setup, include=F}

```

I thank you all for the opportunity to submit revisions of my manuscript "Asfsfssdfsafdsafs3fs" for possible inclusion in *Conflict Management and Peace Science* (*CMPS*). I thank the two reviewers and the editorial team for the helpful comments as well.

# Comments to the Editorial Team

The editorial team noted

Enclosed are the revised manuscript and appendix. I thank the editorial team and the reviewers for their helpful comments and for further consideration of my revisions.

Best regards,

```{r eval=F, echo=FALSE, out.width='30%'}
# knitr::include_graphics('john-hancock.png')
# ^ put your signature here as a PNG or something.
```

The Author

\newpage

# Reviewer #1's Comments

Reviewer #1 (R1) had the following comments on the original version of the manuscript.

\newpage

# Reviewer #2's Comments

I thank Reviewer #2 (R2) for

\newpage

### 2021-11-05 user-guide/ms.Rmd
prepare appendix and memo for revisions

@@
 ---
-title: "`{peacesciencer}`:  A User’s  Guide  for  Quantitative  Peace Science in R"
+title: "`{peacesciencer}`: An R Package for Quantitative Peace Science Research"
 thanks: "Replication files are available on the author's Github account (https://github.com/svmiller/peacesciencer).

### 2021-11-05 user-guide/appendix.Rmd
enh, worth a shot, maybe worth thinking about, we'll see

@@
 appendixletter: "A"
+pandocparas: TRUE
 anonymous: "`r params$anonymous`"
@@
 
+# Vignette: A Discussion of Various Joins in `{peacesciencer}`
+
+```{r child = '../vignettes/joins.Rmd', echo=TRUE}
+```
+
+# Vignette: `{peacesciencer}` Data Versions
+
+```{r child = '../vignettes/versions.Rmd', echo=TRUE}
+```
+
 \newpage

### 2021-11-05 user-guide/src/render_appendix-anon.R

@@ -1,0 +1,8 @@
+the_output <- "latex"
+
+rmarkdown::render("appendix.Rmd", output_file="doc/appendix-anon.pdf",
+                  params=list(anonymous=TRUE,doublespacing=TRUE),
+                  bookdown::pdf_document2(template = stevetemplates::templ_article2(),
+                                          latex_engine = "xelatex", dev= "cairo_pdf",
+                                          toc = FALSE,
+                                          number_sections = FALSE, keep_tex = FALSE))

### 2021-11-06 user-guide/appendix.Rmd
update

@@
 linkcolor: black
+nocite: |
+  @cartersmith2020fml, @andersetal2020bbgb
 header-includes:
@@
 
-# A Section with a Citation to Get Started
+# Workflow Citations
 
-@miller2017etjc
+Let's start with @wickhametal2019wt and @arelbundock2021m. I know I've used others (e.g. bookdown). I'll just have to get on those when the time comes.
 

### 2021-11-07 user-guide/appendix.Rmd
update

@@
 
-Let's start with @wickhametal2019wt and @arelbundock2021m. I know I've used others (e.g. bookdown). I'll just have to get on those when the time comes.
+The following tools and libraries were integral to the creation of the manuscript and I want to at least acknowledge them here.
+Wickham et al.'s [-@wickhametal2019wt] `{tidyverse}` was the primary toolkit for most data transformation techniques outlined here, and indeed forms the implicit foundation of the package itself.[^aboutotherpackages] Regression tables were formatted in `{modelsummary}` [@arelbundock2021m], which itself uses and suggests `{kableExtra}` as a back-end for presentation [@zhu2021k].
+The document itself is a dynamic document using familiar R Markdown syntax [@xieetal2020rmc], knitted to various outputs [@xie2015ddrk], using `{bookdown}` for additional functionality [@xie2016b].
+
+I offer these citations first as an acknowledgement of these important contributions to our collective research productivity.
+Space restrictions in our journal preclude us from citing these important tools, even if they are truly ancillary to the product they help produce.
+No matter, I intend to put this appendix on my website to give Google Scholar the opportunity to find these citations and count them toward the *h*-indexes of the researchers responsible for these important tools.
+
+[^aboutotherpackages]: The appendix will reference component packages of `{tidyverse}`, but follows the encouragement of @wickhametal2019wt to cite the overall suite of packages outlined in @wickhametal2019wt rather than reference the packages individually.
 

### 2021-11-08 R/download_extdata.R
update

@@ -99,7 +99,7 @@
 #' | `srsvaa` | unweighted S (absolute distances, valued alliance data) |
 #' | `srswvaa` | weighted S (absolute distances, valued alliance data) |
 #' | `kappava` | Kappa (squared distances, valued alliance data) |
-#' | `piva` | Pi (suared distances, valued alliance data) |
+#' | `piva` | Pi (squared distances, valued alliance data) |
 #' | `srsba` | Unweighted S (binary alliance data) |
 #' | `srswba` | Weighted S (binary alliance data) |
 #' | `kappaba` | Kappa (binary alliance data) |

### 2021-11-08 docs/ms.tex
no clue why this is happening, and I think it's a UTF-8 thing

@@
 
-\title{\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: June 21, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
+\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
+\textbf{Current version}: November 08, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
 \author{true}
@@
 
-\title{\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: June 21, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
+\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
+\textbf{Current version}: November 08, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
 
@@
              pdfkeywords = {software, statistical analysis, peace science, reproducibility},
-            pdftitle={\texttt{\{peacesciencer\}}: A User's Guide for Quantitative Peace Science in R},
+            pdftitle={\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research},
             colorlinks=true,
@@
 \usepackage{xcolor}
+\usepackage{siunitx}
+\newcolumntype{d}{S[input-symbols = ()]}
 
@@
 
+% 
 \begin{document}
@@
 \midrule
-Land Contiguity & 1.044*\\
- & (0.057)\\
-Dyadic CINC Proportion (Lower/Higher) & 0.444*\\
- & (0.036)\\
-CoW Major Power in Dyad & 0.145*\\
- & \vphantom{1} (0.058)\\
-Defense Pact & -0.102+\\
- & (0.058)\\
-Dyadic Democracy (Weak-Link) & -0.509*\\
- & (0.052)\\
-Dyadic GDP per Capita (Weak-Link) & 0.262*\\
- & (0.050)\\
-Dyadic Militarization (Minimum) & 0.267*\\
- & (0.023)\\
-t & -0.148*\\
- & (0.005)\\
-t\textasciicircum{}2 & 0.002*\\
- & \vphantom{1} (0.000)\\
-t\textasciicircum{}3 & 0.000*\\
- & (0.000)\\
-Intercept & -3.055*\\
- & (0.064)\\
+Land Contiguity & \num{1.044}*\\
+ & (\num{0.057})\\
+Dyadic CINC Proportion (Lower/Higher) & \num{0.444}*\\
+ & (\num{0.036})\\
+CoW Major Power in Dyad & \num{0.145}*\\
+ & \vphantom{1} (\num{0.058})\\
+Defense Pact & \num{-0.102}+\\
+ & (\num{0.058})\\
+Dyadic Democracy (Weak-Link) & \num{-0.509}*\\
+ & (\num{0.052})\\
+Dyadic GDP per Capita (Weak-Link) & \num{0.262}*\\
+ & (\num{0.050})\\
+Dyadic Militarization (Minimum) & \num{0.267}*\\
+ & (\num{0.023})\\
+t & \num{-0.148}*\\
+ & (\num{0.005})\\
+t\textasciicircum2 & \num{0.002}*\\
+ & \vphantom{1} (\num{0.000})\\
+t\textasciicircum3 & \num{0.000}*\\
+ & (\num{0.000})\\
+Intercept & \num{-3.055}*\\
+ & (\num{0.064})\\
 \midrule
-Num.Obs. & 106937\\
+Num.Obs. & \num{106937}\\
 \bottomrule
-\multicolumn{2}{l}{\textsuperscript{} + p $<$ 0.1, * p $<$ 0.05}\\
+\multicolumn{2}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05}\\
 \end{tabular}
@@
 \midrule
-Population Size (Lagged) & 0.229* & 0.272*\\
- & (0.067) & (0.106)\\
-Extended UDS Democracy Score (Lagged) & 0.257 & -0.085\\
- & (0.181) & (0.270)\\
-Extended UDS Democracy Score\textasciicircum{}2 (Lagged) & -0.726* & -0.761*\\
- & (0.211) & (0.352)\\
-\% Mountainous Terrain (Logged) & 0.055 & 0.342*\\
- & (0.067) & (0.112)\\
-Ethnic Fractionalization & 0.442 & 0.333\\
- & (0.358) & (0.554)\\
-Religious Fractionalization & -0.389 & -0.281\\
- & (0.402) & (0.593)\\
-t & -0.074+ & -0.111*\\
- & (0.039) & (0.056)\\
-t\textasciicircum{}2 & 0.004* & 0.005+\\
- & (0.002) & (0.003)\\
-t\textasciicircum{}3 & 0.000* & 0.000+\\
- & (0.000) & (0.000)\\
-Intercept & -5.097* & -6.590*\\
- & (1.351) & (2.084)\\
+Population Size (Lagged) & \num{0.229}* & \num{0.272}*\\
+ & (\num{0.067}) & (\num{0.106})\\
+Extended UDS Democracy Score (Lagged) & \num{0.257} & \num{-0.085}\\
+ & (\num{0.181}) & (\num{0.270})\\
+Extended UDS Democracy Score\textasciicircum2 (Lagged) & \num{-0.726}* & \num{-0.761}*\\
+ & (\num{0.211}) & (\num{0.352})\\
+\% Mountainous Terrain (Logged) & \num{0.055} & \num{0.342}*\\
+ & (\num{0.067}) & (\num{0.112})\\
+Ethnic Fractionalization & \num{0.442} & \num{0.333}\\
+ & (\num{0.358}) & (\num{0.554})\\
+Religious Fractionalization & \num{-0.389} & \num{-0.281}\\
+ & (\num{0.402}) & (\num{0.593})\\
+t & \num{-0.074}+ & \num{-0.111}*\\
+ & (\num{0.039}) & (\num{0.056})\\
+t\textasciicircum2 & \num{0.004}* & \num{0.005}+\\
+ & (\num{0.002}) & (\num{0.003})\\
+t\textasciicircum3 & \num{0.000}* & \num{0.000}+\\
+ & (\num{0.000}) & (\num{0.000})\\
+Intercept & \num{-5.097}* & \num{-6.590}*\\
+ & (\num{1.351}) & (\num{2.084})\\
 \midrule
-Num.Obs. & 8192 & 8192\\
+Num.Obs. & \num{8192} & \num{8192}\\
 \bottomrule
-\multicolumn{3}{l}{\textsuperscript{} + p $<$ 0.1, * p $<$ 0.05}\\
+\multicolumn{3}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05}\\
 \end{tabular}

### 2021-11-09 user-guide/doc/appendix.tex
these are all works in progress.

% ARTICLE 2 ----
% This is just here so I know exactly what I'm looking at in Rstudio when messing with stuff.
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{cochineal}
  \setsansfont[]{Linux Biolinum O}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}
    }
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{APPENDIX: \texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research}
\author{}
\date{November 08, 2021}

% Jesus, okay, everything above this comment is default Pandoc LaTeX template. -----
% ----------------------------------------------------------------------------------
% I think I had assumed beamer and LaTex were somehow different templates.

\usepackage{kantlipsum}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
      {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}
    }
%\fi
\makeatother

\title{APPENDIX: \texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research }

\date{}

\usepackage{titlesec}

% 
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\itshape} % \small\uppercase
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
             pdfkeywords = {software, statistical analysis, peace science, reproducibility},
            pdftitle={APPENDIX: \texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

\renewcommand*\thetable{A.\arabic{table}}
\renewcommand*\thefigure{A.\arabic{figure}}

% This will better treat References as a section when using natbib
% https://tex.stackexchange.com/questions/49962/bibliography-title-fontsize-problem-with-bibtex-and-the-natbib-package

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{longtable}
\usepackage{csquotes}
\LTcapwidth=.95\textwidth
\linespread{1.05}
\usepackage{hyperref}
\usepackage{setspace}
\renewcommand{\thepage}{A-\arabic{page}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\newtheorem{hypothesis}{Hypothesis}

% trick for moving figures to back of document
% really wish we'd knock this shit off with moving tables/figures to back of document
% but, alas...

% 
\begin{document}

% \textsf{\textbf{This is sans-serif bold text.}}
% \textbf{\textsf{This is bold sans-serif text.}}

% \maketitle

{% \usefont{T1}{pnc}{m}{n}
\setlength{\parindent}{0pt}
\thispagestyle{plain}
{%\fontsize{18}{20}\selectfont\raggedright
\maketitle  % title \par

}

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12}

}

}

\vskip -8.5pt

 % removetitleabstract

% We'll put doublespacing here
% Remember to cut it out later before bib
\tableofcontents
\newpage

\hypertarget{workflow-citations}{%
\section{Workflow Citations}\label{workflow-citations}}

The following tools and libraries were integral to the creation of the manuscript and I want to at least acknowledge them here. Wickham et al.'s (2019) \texttt{\{tidyverse\}} was the primary toolkit for most data transformation techniques outlined here, and indeed forms the implicit foundation of the package itself.\footnote{The appendix will reference component packages of \texttt{\{tidyverse\}}, but follows the encouragement of Wickham et al. (2019) to cite the overall suite of packages outlined in Wickham et al. (2019) rather than reference the packages individually.} Regression tables were formatted in \texttt{\{modelsummary\}} (Arel-Bundock, 2021), which itself uses and suggests \texttt{\{kableExtra\}} as a back-end for presentation (Zhu, 2021). The document itself is a dynamic document using familiar R Markdown syntax (Xie, Dervieux \& Riederer, 2020), knitted to various outputs (Xie, 2015), using \texttt{\{bookdown\}} for additional functionality (Xie, 2016).

I offer these citations first as an acknowledgment of these important contributions to our collective research productivity. Space restrictions in our journal preclude us from citing these important tools, even if they are truly ancillary to the product they help produce. No matter, I intend to put this appendix on my website to give Google Scholar the opportunity to find these citations and count them toward the \emph{h}-indexes of the researchers responsible for these important tools.

\hypertarget{vignette-a-discussion-of-various-joins-in-peacesciencer}{%
\section{\texorpdfstring{Vignette: A Discussion of Various Joins in \texttt{\{peacesciencer\}}}{Vignette: A Discussion of Various Joins in \{peacesciencer\}}}\label{vignette-a-discussion-of-various-joins-in-peacesciencer}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{library}\NormalTok{(lubridate)}
\FunctionTok{library}\NormalTok{(peacesciencer)}
\end{Highlighting}
\end{Shaded}

Users who may wish to improve their own data management skills in R by looking how \texttt{\{peacesciencer\}} functions are written will see that basic foundation of \texttt{\{peacesciencer\}}'s functions are so-called ``join'' functions. The ``join'' functions themselves come in \texttt{\{dplyr\}}, a critical dependency for \texttt{\{peacesciencer\}} and the effective engine of \href{https://www.tidyverse.org}{\texttt{\{tidyverse\}}} (which I suggest for a basic workflow tool, and which the user may already be using). Users who have absolutely no idea what these functions are welcome to find more thorough texts about these different types of joins. Their functionality and terminology have \href{https://www.w3schools.com/sql/sql_join.asp}{a clear basis in SQL}, a relational database management system that first appeared in 1974 for data management and data manipulation. My goal here is not to offer a crash course on all these potential ``join'' functions, though helpful \href{https://github.com/gadenbuie/tidyexplain}{visual primers are available in R} and \href{https://blog.codinghorror.com/a-visual-explanation-of-sql-joins/}{SQL}. Instead, I will offer the basic principles these visual primers are communicating as they apply to \texttt{\{peacesciencer\}}.

\hypertarget{left-outer-join}{%
\subsection{Left (Outer) Join}\label{left-outer-join}}

The first type of join is the most important type of join function in \texttt{\{peacesciencer\}}. Indeed, almost every function in this package that deals with adding variables to a type of data created in \texttt{\{peacesciencer\}} includes it. This is the ``left join'' (\texttt{left\_join()} in \texttt{\{dplyr\}}), alternatively known as \href{https://www.w3schools.com/sql/sql_join_left.asp}{the ``outer join'' or ``left outer join'' in the SQL context}, and is a type of ``mutating join'' in the \texttt{\{tidyverse\}} context. In plain English, the \texttt{left\_join()} assumes two data objects---a ``left'' object (\texttt{x}) and a ``right'' object (\texttt{y})---and returns all rows from the left object (\texttt{x}) with matching information in the right object (\texttt{y}) by a set of common matching keys (or columns in both \texttt{x} and \texttt{y}).

Here is a simple example of how this works in the \texttt{\{peacesciencer\}} context. Assume a simple state-year data set of the United States (\texttt{ccode:\ 2}), Canada (\texttt{ccode:\ 20}), and the United Kingdom (\texttt{ccode:\ 200}) for all years from 2016 to 2020. Recreating this simple kind of data is no problem in R and will serve as our ``left object'' (\texttt{x}) for this simple example.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{tibble}\NormalTok{(}\AttributeTok{ccode =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{200}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# rowwise() is a great trick for nesting sequences in tibbles}
  \CommentTok{\# This parlor trick, for example, generates state{-}year data out of raw state }
  \CommentTok{\# data in create\_stateyears()}
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# create a sequence as a nested column}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{year =} \FunctionTok{list}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{2016}\NormalTok{, }\DecValTok{2020}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# unnest the column}
  \FunctionTok{unnest}\NormalTok{(year) }\OtherTok{{-}\textgreater{}}\NormalTok{ x}

\NormalTok{x}
\CommentTok{\#\textgreater{} \# A tibble: 15 x 2}
\CommentTok{\#\textgreater{}    ccode  year}
\CommentTok{\#\textgreater{}    \textless{}dbl\textgreater{} \textless{}int\textgreater{}}
\CommentTok{\#\textgreater{}  1     2  2016}
\CommentTok{\#\textgreater{}  2     2  2017}
\CommentTok{\#\textgreater{}  3     2  2018}
\CommentTok{\#\textgreater{}  4     2  2019}
\CommentTok{\#\textgreater{}  5     2  2020}
\CommentTok{\#\textgreater{}  6    20  2016}
\CommentTok{\#\textgreater{}  7    20  2017}
\CommentTok{\#\textgreater{}  8    20  2018}
\CommentTok{\#\textgreater{}  9    20  2019}
\CommentTok{\#\textgreater{} 10    20  2020}
\CommentTok{\#\textgreater{} 11   200  2016}
\CommentTok{\#\textgreater{} 12   200  2017}
\CommentTok{\#\textgreater{} 13   200  2018}
\CommentTok{\#\textgreater{} 14   200  2019}
\CommentTok{\#\textgreater{} 15   200  2020}
\end{Highlighting}
\end{Shaded}

Let's assume we're building toward the kind of state-year analysis I describe in \href{http://svmiller.com/peacesciencer/ms.pdf}{the manuscript accompanying this package}. For example, the canonical civil conflict analysis by \href{https://www.cambridge.org/core/journals/american-political-science-review/article/abs/ethnicity-insurgency-and-civil-war/B1D5D0E7C782483C5D7E102A61AD6605}{Fearon and Laitin (2003)} has an outcome that varies by year, but several independent variables that are time-invariant and serve as variables for making state-to-state comparisons in their model of civil war onset (e.g ethnic fractionalization, religious fractionalization, terrain ruggedness). In a similar manner, we have a basic ranking of the United States, Canada, and the United Kingdom in our case. Minimally, the United States scores ``low'', Canada scores ``medium'', and the United Kingdom scores ``high'' on some metric. There is no variation by time in this simple example.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{tibble}\NormalTok{(}\AttributeTok{ccode =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{200}\NormalTok{),}
       \AttributeTok{ranking =} \FunctionTok{c}\NormalTok{(}\StringTok{"low"}\NormalTok{, }\StringTok{"medium"}\NormalTok{, }\StringTok{"high"}\NormalTok{)) }\OtherTok{{-}\textgreater{}}\NormalTok{ y}

\NormalTok{y}
\CommentTok{\#\textgreater{} \# A tibble: 3 x 2}
\CommentTok{\#\textgreater{}   ccode ranking}
\CommentTok{\#\textgreater{}   \textless{}dbl\textgreater{} \textless{}chr\textgreater{}  }
\CommentTok{\#\textgreater{} 1     2 low    }
\CommentTok{\#\textgreater{} 2    20 medium }
\CommentTok{\#\textgreater{} 3   200 high}
\end{Highlighting}
\end{Shaded}

This is the ``right object'' (\texttt{y}) that we want to add to the ``left object'' that serves as our main data frame. Notice that \texttt{x} has no variable for the ranking information we want. It does, however, have matching observations for the state identifiers corresponding with the Correlates of War state codes for the U.S., Canada, and the United Kingdom. The left join (as \texttt{left\_join()}) merges \texttt{y} into \texttt{x}, returning all rows of \texttt{x} with matching information in \texttt{y} based on columns they share in common (here: \texttt{ccode}).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# alternatively, as I tend to do it: x \%\textgreater{}\% left\_join(., y)}
\FunctionTok{left\_join}\NormalTok{(x, y)}
\CommentTok{\#\textgreater{} \# A tibble: 15 x 3}
\CommentTok{\#\textgreater{}    ccode  year ranking}
\CommentTok{\#\textgreater{}    \textless{}dbl\textgreater{} \textless{}int\textgreater{} \textless{}chr\textgreater{}  }
\CommentTok{\#\textgreater{}  1     2  2016 low    }
\CommentTok{\#\textgreater{}  2     2  2017 low    }
\CommentTok{\#\textgreater{}  3     2  2018 low    }
\CommentTok{\#\textgreater{}  4     2  2019 low    }
\CommentTok{\#\textgreater{}  5     2  2020 low    }
\CommentTok{\#\textgreater{}  6    20  2016 medium }
\CommentTok{\#\textgreater{}  7    20  2017 medium }
\CommentTok{\#\textgreater{}  8    20  2018 medium }
\CommentTok{\#\textgreater{}  9    20  2019 medium }
\CommentTok{\#\textgreater{} 10    20  2020 medium }
\CommentTok{\#\textgreater{} 11   200  2016 high   }
\CommentTok{\#\textgreater{} 12   200  2017 high   }
\CommentTok{\#\textgreater{} 13   200  2018 high   }
\CommentTok{\#\textgreater{} 14   200  2019 high   }
\CommentTok{\#\textgreater{} 15   200  2020 high}
\end{Highlighting}
\end{Shaded}

This is obviously a very simple example, but it scales well even if there is some additional complexity. For example, let's assume we added a simple five-year panel of Australia (\texttt{ccode:\ 900}) to the ``left object'' (\texttt{x}). However, we have no corresponding information about Australia in the ``right object'' (\texttt{y}). Here is what the left join would produce under these circumstances.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{tibble}\NormalTok{(}\AttributeTok{ccode =} \DecValTok{900}\NormalTok{,}
       \AttributeTok{year =} \FunctionTok{c}\NormalTok{(}\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{bind\_rows}\NormalTok{(x, .) }\OtherTok{{-}\textgreater{}}\NormalTok{ x}

\NormalTok{x}
\CommentTok{\#\textgreater{} \# A tibble: 20 x 2}
\CommentTok{\#\textgreater{}    ccode  year}
\CommentTok{\#\textgreater{}    \textless{}dbl\textgreater{} \textless{}int\textgreater{}}
\CommentTok{\#\textgreater{}  1     2  2016}
\CommentTok{\#\textgreater{}  2     2  2017}
\CommentTok{\#\textgreater{}  3     2  2018}
\CommentTok{\#\textgreater{}  4     2  2019}
\CommentTok{\#\textgreater{}  5     2  2020}
\CommentTok{\#\textgreater{}  6    20  2016}
\CommentTok{\#\textgreater{}  7    20  2017}
\CommentTok{\#\textgreater{}  8    20  2018}
\CommentTok{\#\textgreater{}  9    20  2019}
\CommentTok{\#\textgreater{} 10    20  2020}
\CommentTok{\#\textgreater{} 11   200  2016}
\CommentTok{\#\textgreater{} 12   200  2017}
\CommentTok{\#\textgreater{} 13   200  2018}
\CommentTok{\#\textgreater{} 14   200  2019}
\CommentTok{\#\textgreater{} 15   200  2020}
\CommentTok{\#\textgreater{} 16   900  2016}
\CommentTok{\#\textgreater{} 17   900  2017}
\CommentTok{\#\textgreater{} 18   900  2018}
\CommentTok{\#\textgreater{} 19   900  2019}
\CommentTok{\#\textgreater{} 20   900  2020}

\FunctionTok{left\_join}\NormalTok{(x, y)}
\CommentTok{\#\textgreater{} \# A tibble: 20 x 3}
\CommentTok{\#\textgreater{}    ccode  year ranking}
\CommentTok{\#\textgreater{}    \textless{}dbl\textgreater{} \textless{}int\textgreater{} \textless{}chr\textgreater{}  }
\CommentTok{\#\textgreater{}  1     2  2016 low    }
\CommentTok{\#\textgreater{}  2     2  2017 low    }
\CommentTok{\#\textgreater{}  3     2  2018 low    }
\CommentTok{\#\textgreater{}  4     2  2019 low    }
\CommentTok{\#\textgreater{}  5     2  2020 low    }
\CommentTok{\#\textgreater{}  6    20  2016 medium }
\CommentTok{\#\textgreater{}  7    20  2017 medium }
\CommentTok{\#\textgreater{}  8    20  2018 medium }
\CommentTok{\#\textgreater{}  9    20  2019 medium }
\CommentTok{\#\textgreater{} 10    20  2020 medium }
\CommentTok{\#\textgreater{} 11   200  2016 high   }
\CommentTok{\#\textgreater{} 12   200  2017 high   }
\CommentTok{\#\textgreater{} 13   200  2018 high   }
\CommentTok{\#\textgreater{} 14   200  2019 high   }
\CommentTok{\#\textgreater{} 15   200  2020 high   }
\CommentTok{\#\textgreater{} 16   900  2016 \textless{}NA\textgreater{}   }
\CommentTok{\#\textgreater{} 17   900  2017 \textless{}NA\textgreater{}   }
\CommentTok{\#\textgreater{} 18   900  2018 \textless{}NA\textgreater{}   }
\CommentTok{\#\textgreater{} 19   900  2019 \textless{}NA\textgreater{}   }
\CommentTok{\#\textgreater{} 20   900  2020 \textless{}NA\textgreater{}}
\end{Highlighting}
\end{Shaded}

Because have no ranking for Australia in this simple example, the left join returns NAs for Australia. The original dimensions of \texttt{x} under these conditions is unaffected.

What would happen if we had an observation in \texttt{y} that has no corresponding match in \texttt{x}? For example, let's assume our ranking data also included a ranking for Denmark (\texttt{ccode:\ 390}), though Denmark does not appear in \texttt{x}. Here is what would happen under these circumstances.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{tibble}\NormalTok{(}\AttributeTok{ccode =} \DecValTok{390}\NormalTok{,}
       \AttributeTok{ranking =} \StringTok{"high"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{bind\_rows}\NormalTok{(y, .) }\OtherTok{{-}\textgreater{}}\NormalTok{ y}

\NormalTok{y}
\CommentTok{\#\textgreater{} \# A tibble: 4 x 2}
\CommentTok{\#\textgreater{}   ccode ranking}
\CommentTok{\#\textgreater{}   \textless{}dbl\textgreater{} \textless{}chr\textgreater{}  }
\CommentTok{\#\textgreater{} 1     2 low    }
\CommentTok{\#\textgreater{} 2    20 medium }
\CommentTok{\#\textgreater{} 3   200 high   }
\CommentTok{\#\textgreater{} 4   390 high}

\FunctionTok{left\_join}\NormalTok{(x, y)}
\CommentTok{\#\textgreater{} \# A tibble: 20 x 3}
\CommentTok{\#\textgreater{}    ccode  year ranking}
\CommentTok{\#\textgreater{}    \textless{}dbl\textgreater{} \textless{}int\textgreater{} \textless{}chr\textgreater{}  }
\CommentTok{\#\textgreater{}  1     2  2016 low    }
\CommentTok{\#\textgreater{}  2     2  2017 low    }
\CommentTok{\#\textgreater{}  3     2  2018 low    }
\CommentTok{\#\textgreater{}  4     2  2019 low    }
\CommentTok{\#\textgreater{}  5     2  2020 low    }
\CommentTok{\#\textgreater{}  6    20  2016 medium }
\CommentTok{\#\textgreater{}  7    20  2017 medium }
\CommentTok{\#\textgreater{}  8    20  2018 medium }
\CommentTok{\#\textgreater{}  9    20  2019 medium }
\CommentTok{\#\textgreater{} 10    20  2020 medium }
\CommentTok{\#\textgreater{} 11   200  2016 high   }
\CommentTok{\#\textgreater{} 12   200  2017 high   }
\CommentTok{\#\textgreater{} 13   200  2018 high   }
\CommentTok{\#\textgreater{} 14   200  2019 high   }
\CommentTok{\#\textgreater{} 15   200  2020 high   }
\CommentTok{\#\textgreater{} 16   900  2016 \textless{}NA\textgreater{}   }
\CommentTok{\#\textgreater{} 17   900  2017 \textless{}NA\textgreater{}   }
\CommentTok{\#\textgreater{} 18   900  2018 \textless{}NA\textgreater{}   }
\CommentTok{\#\textgreater{} 19   900  2019 \textless{}NA\textgreater{}   }
\CommentTok{\#\textgreater{} 20   900  2020 \textless{}NA\textgreater{}}
\end{Highlighting}
\end{Shaded}

Notice the output of this left join is identical to the output above. Australia is in \texttt{x}, but not in \texttt{y}. Thus, the rows for Australia are returned but the absence of ranking information for Australia in \texttt{y} means the variable is \texttt{NA} for Australia after the merge. Denmark is in \texttt{y}, but not \texttt{x}. Because the left join returns all rows in \texttt{x} with matching information in \texttt{y}, the absence of observations for Denmark in \texttt{x} means there is nowhere for the ranking information to go in the merge. Thus, Denmark's ranking is effectively ignored.

\hypertarget{why-the-left-join-in-particular}{%
\subsubsection{Why the Left Join, in Particular?}\label{why-the-left-join-in-particular}}

An interested user may ask what's so special about this kind of join that it appears everywhere in \texttt{\{peacesciencer\}}? One is a matter of taste. I could just as well be doing this vignette by reference to the ``right join'', the mirror join to ``left join.'' The right join in \texttt{\{dplyr\}}'s \texttt{right\_join(x,y)} returns all records from \texttt{y} with matching rows in \texttt{x} by common columns, though the equivalency would depend on reversing the order of \texttt{x} and \texttt{y} (i.e.~\texttt{left\_join(x,\ y)} produces the same information as \texttt{right\_join(y,\ x)}). The arrangement of columns would differ in the \texttt{left\_join()} and \texttt{right\_join()} in this simple application even if the same underlying information is there. Ultimately, I tend to \href{http://svmiller.com/blog/2021/01/a-tutorial-on-the-join-family-in-r/\#myadvice}{think ``left-handed''} when it comes to data management and instruct my students to do the same when I introduce them to data transformation in R. I like the intuition, especially in the pipe-based workflow, to start with a master data object at the top of the pipe and keep it ``left'' as I add information to it. It has the benefit of keeping the units of analysis (e.g.~state-years in this simple setup) as the first columns the user sees as well. This is my preferred approach to data transformation and it recurs in \texttt{\{peacesciencer\}} as a result.

Beyond that matter of taste, the left join is everywhere in \texttt{\{peacesciencer\}} because the project endeavors hard to recreate the appropriate universe of cases of interest to the user and allow the user to add stuff to it as they see fit. \texttt{create\_stateyears()} will create the entire universe of state-years from 1816 to the present for a state-year analysis. \texttt{create\_dyadyears()} will create the entire universe of dyad-years from 1816 to the present for a dyad-year analysis. The logic here, as it is implemented in \texttt{\{peacesciencer\}}'s multiple functions, is the type of data the user wants to create has been created for them. The user does not want to expand the data any further than that, though the user may want to do something like reduce the full universe of 1816-2020 state-years to just 1946-2010. However, this is a universe discarded, not a universe that has been augmented or expanded.

With that in mind, every function's use of the left join assumes the data object it receives represents the full universe of cases of interest to the researcher. The left join is just adding information to it, based on matching information in one of its many data sets. When done carefully, the left join is a dutiful way of adding information to a data set without changing the number of rows of the original data set. The number of columns will obviously expand, but the number of rows is unaffected.

\hypertarget{potential-problems-of-the-left-join}{%
\subsubsection{Potential Problems of the Left Join}\label{potential-problems-of-the-left-join}}

``When done carefully'' is doing some heavy-lifting in that last sentence. So, let me explain some situations where the left join will produce problems for the researcher (even if the join itself is doing what it is supposed to do from an operational standpoint).

The first is less of a problem, at least as I have implemented in \texttt{\{peacesciencer\}}, but more of a caution. In the above example, our panel consists of just the U.S., Canada, the United Kingdom, and Australia. We happen to have a ranking for Denmark, but Denmark wasn't in our panel of (effectively, exclusively) Anglophone states. Therefore, no row is created for Denmark. If it were that important that the left join create those rows for Denmark, we should have had it in the first place. In this case, the left join is behaving as it should. We should have had Denmark in the panel before trying to match information to it.

\texttt{\{peacesciencer\}} circumvents this issue by creating universal data (e.g.~all state-years, all dyad-years, all available leader-years) that the user is free to subset as they see fit. Users should run one of the ``create'' functions (e.g.~\texttt{create\_stateyears()}, \texttt{create\_dyadyears()}) at the top of their script before adding information to it because the left join, as implemented everywhere in this package, is building in an assumption that the universe of cases of interest to the user is represented in the ``left object'' for a left outer join. Basically, do not expect the left join to create new rows in \texttt{x} in a situation where there is a state represented in \texttt{y} but not in \texttt{x}. It will not. This type of join assumes the universe of cases of interest to the researcher already appear in the ``left object.''

The second situation is a bigger problem. Sometimes, often when \href{http://svmiller.com/peacesciencer/articles/state-systems.html}{bouncing between information denominated in Correlates of War states and Gleditsch-Ward states}, there is an unwanted duplicate observation in the data frame to be merged into the primary data of interest to the user. Let's go back to our simple example of \texttt{x} and \texttt{y} here. Everything here performs nicely, though Australia (in \texttt{x}) has no ranking and Denmark (in \texttt{y}) is not in our panel of state-years because it wasn't part of the original universe of cases of interest to us.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x}
\CommentTok{\#\textgreater{} \# A tibble: 20 x 2}
\CommentTok{\#\textgreater{}    ccode  year}
\CommentTok{\#\textgreater{}    \textless{}dbl\textgreater{} \textless{}int\textgreater{}}
\CommentTok{\#\textgreater{}  1     2  2016}
\CommentTok{\#\textgreater{}  2     2  2017}
\CommentTok{\#\textgreater{}  3     2  2018}
\CommentTok{\#\textgreater{}  4     2  2019}
\CommentTok{\#\textgreater{}  5     2  2020}
\CommentTok{\#\textgreater{}  6    20  2016}
\CommentTok{\#\textgreater{}  7    20  2017}
\CommentTok{\#\textgreater{}  8    20  2018}
\CommentTok{\#\textgreater{}  9    20  2019}
\CommentTok{\#\textgreater{} 10    20  2020}
\CommentTok{\#\textgreater{} 11   200  2016}
\CommentTok{\#\textgreater{} 12   200  2017}
\CommentTok{\#\textgreater{} 13   200  2018}
\CommentTok{\#\textgreater{} 14   200  2019}
\CommentTok{\#\textgreater{} 15   200  2020}
\CommentTok{\#\textgreater{} 16   900  2016}
\CommentTok{\#\textgreater{} 17   900  2017}
\CommentTok{\#\textgreater{} 18   900  2018}
\CommentTok{\#\textgreater{} 19   900  2019}
\CommentTok{\#\textgreater{} 20   900  2020}
\NormalTok{y}
\CommentTok{\#\textgreater{} \# A tibble: 4 x 2}
\CommentTok{\#\textgreater{}   ccode ranking}
\CommentTok{\#\textgreater{}   \textless{}dbl\textgreater{} \textless{}chr\textgreater{}  }
\CommentTok{\#\textgreater{} 1     2 low    }
\CommentTok{\#\textgreater{} 2    20 medium }
\CommentTok{\#\textgreater{} 3   200 high   }
\CommentTok{\#\textgreater{} 4   390 high}
\end{Highlighting}
\end{Shaded}

Let's assume, however, we mistakenly entered the United Kingdom twice into \texttt{y}. We know these data are supposed to be simple state-level rankings. Each state is supposed to be in there just once. The United Kingdom appears in there twice.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{tibble}\NormalTok{(}\AttributeTok{ccode =} \DecValTok{200}\NormalTok{,}
       \AttributeTok{ranking =} \StringTok{"high"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{bind\_rows}\NormalTok{(y, .) }\OtherTok{{-}\textgreater{}}\NormalTok{ y2}
\end{Highlighting}
\end{Shaded}

If we were to left join \texttt{y2} into \texttt{x}, we get an unwelcome result. The United Kingdom is duplicated for all yearly observations.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{left\_join}\NormalTok{(x, y2) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ data.frame}
\CommentTok{\#\textgreater{}    ccode year ranking}
\CommentTok{\#\textgreater{} 1      2 2016     low}
\CommentTok{\#\textgreater{} 2      2 2017     low}
\CommentTok{\#\textgreater{} 3      2 2018     low}
\CommentTok{\#\textgreater{} 4      2 2019     low}
\CommentTok{\#\textgreater{} 5      2 2020     low}
\CommentTok{\#\textgreater{} 6     20 2016  medium}
\CommentTok{\#\textgreater{} 7     20 2017  medium}
\CommentTok{\#\textgreater{} 8     20 2018  medium}
\CommentTok{\#\textgreater{} 9     20 2019  medium}
\CommentTok{\#\textgreater{} 10    20 2020  medium}
\CommentTok{\#\textgreater{} 11   200 2016    high}
\CommentTok{\#\textgreater{} 12   200 2016    high}
\CommentTok{\#\textgreater{} 13   200 2017    high}
\CommentTok{\#\textgreater{} 14   200 2017    high}
\CommentTok{\#\textgreater{} 15   200 2018    high}
\CommentTok{\#\textgreater{} 16   200 2018    high}
\CommentTok{\#\textgreater{} 17   200 2019    high}
\CommentTok{\#\textgreater{} 18   200 2019    high}
\CommentTok{\#\textgreater{} 19   200 2020    high}
\CommentTok{\#\textgreater{} 20   200 2020    high}
\CommentTok{\#\textgreater{} 21   900 2016    \textless{}NA\textgreater{}}
\CommentTok{\#\textgreater{} 22   900 2017    \textless{}NA\textgreater{}}
\CommentTok{\#\textgreater{} 23   900 2018    \textless{}NA\textgreater{}}
\CommentTok{\#\textgreater{} 24   900 2019    \textless{}NA\textgreater{}}
\CommentTok{\#\textgreater{} 25   900 2020    \textless{}NA\textgreater{}}
\end{Highlighting}
\end{Shaded}

It doesn't matter that the duplicate ranking in \texttt{y2} for the UK was the same. It would be messier, sure, if the ranking was different for the duplicate observation, but it matters more here that it was duplicated. In a panel like this, a user who is not careful will have the effect of overweighting those observations that duplicate. In a simple example like this, subsetting to just complete cases (i.e.~Australia has no ranking), the UK is 50\% of all observations despite the fact it should just be a third of observations. That's not ideal for a researcher.

\texttt{\{peacesciencer\}} goes above and beyond to make sure this doesn't happen in the data it creates. Functions are \href{https://github.com/svmiller/peacesciencer/blob/master/tests/testthat/test_nodamnduplicates.R}{aggressively tested to make sure nothing duplicates}, and various parlor tricks (prominently \href{https://dplyr.tidyverse.org/reference/group_by.html}{group-by} \href{https://dplyr.tidyverse.org/reference/slice.html}{slices}) are used internally to cull those duplicate observations. The release of a function that makes prominent use of the left join is done with the assurance it doesn't create a duplicate. No matter, this is the biggest peril of the left join for a researcher who may want to duplicate what \texttt{\{peacesciencer\}} does on their own. Always inspect the data you merge, and the output.

\hypertarget{semi-join}{%
\subsection{Semi-Join}\label{semi-join}}

The ``semi-join'' (\texttt{semi\_join()} in \texttt{\{dplyr\}}) returns all rows from the left object (\texttt{x}) that have matching values in the right object (\texttt{y}). It is a type of ``filtering join'', which affects the observations and not the variables. It appears just twice in \texttt{\{peacesciencer\}}, serving as a final join in \texttt{create\_leaderdays()} and \texttt{create\_leaderyears()}. In both cases, it serves as a means of standardizing leader data (denominated in the Gleditsch-Ward system, if not necessarily Gleditsch-Ward system dates) to the Correlates of War or Gleditsch-Ward system.

Here is a basic example of what a semi-join is doing in this context, with an illustration of the kind of difficulties that manifest in standardizing Archigos' leader data to the Correlates of War state system. Assume this simple state system that has just two states---``A'' and ``B''---over a two-week period at the start of 1975 (Jan.~1, 1975 to Jan.~14, 1975). In this simple system, ``A'' is a state for the full two week period (Jan.~1-Jan.14) whereas ``B'' is a state for just the first seven days (Jan.~1-Jan.~7) because, let's say, A occupied B and ended its statehood. We also happened to have some leader data for these two states. Over this two week period, our leader data suggests A had just one continuous leader---\href{https://en.wikipedia.org/wiki/Archie_Griffin}{``Archie''}---whereas B had three. \href{https://en.wikipedia.org/wiki/Brian_Baschnagel}{``Brian''} was the leader of B before he retired from office and was replaced by \href{https://en.wikipedia.org/wiki/Cornelius_Greene}{``Cornelius.''} However, he was deposed when A invaded B and was replaced by a puppet head of state, \href{https://en.wikipedia.org/wiki/Pete_Johnson_(American_football)}{``Pete.''} Our data look like this.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{tibble}\NormalTok{(}\AttributeTok{code =} \FunctionTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"B"}\NormalTok{),}
       \AttributeTok{stdate =} \FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{01}\NormalTok{),}
       \AttributeTok{enddate =} \FunctionTok{c}\NormalTok{(}\FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{14}\NormalTok{), }
                   \FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{07}\NormalTok{))) }\OtherTok{{-}\textgreater{}}\NormalTok{ state\_system}

\NormalTok{state\_system}
\CommentTok{\#\textgreater{} \# A tibble: 2 x 3}
\CommentTok{\#\textgreater{}   code  stdate     enddate   }
\CommentTok{\#\textgreater{}   \textless{}chr\textgreater{} \textless{}date\textgreater{}     \textless{}date\textgreater{}    }
\CommentTok{\#\textgreater{} 1 A     1975{-}01{-}01 1975{-}01{-}14}
\CommentTok{\#\textgreater{} 2 B     1975{-}01{-}01 1975{-}01{-}07}

\FunctionTok{tibble}\NormalTok{(}\AttributeTok{code =} \FunctionTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"B"}\NormalTok{, }\StringTok{"B"}\NormalTok{, }\StringTok{"B"}\NormalTok{),}
       \AttributeTok{leader =} \FunctionTok{c}\NormalTok{(}\StringTok{"Archie"}\NormalTok{, }\StringTok{"Brian"}\NormalTok{, }\StringTok{"Cornelius"}\NormalTok{, }\StringTok{"Pete"}\NormalTok{),}
       \AttributeTok{stdate =} \FunctionTok{c}\NormalTok{(}\FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{01}\NormalTok{), }\FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{01}\NormalTok{), }
                  \FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{04}\NormalTok{), }\FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{08}\NormalTok{)),}
       \AttributeTok{enddate =} \FunctionTok{c}\NormalTok{(}\FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{14}\NormalTok{), }\FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{04}\NormalTok{), }
                   \FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{08}\NormalTok{), }\FunctionTok{make\_date}\NormalTok{(}\DecValTok{1975}\NormalTok{, }\DecValTok{01}\NormalTok{, }\DecValTok{14}\NormalTok{))) }\OtherTok{{-}\textgreater{}}\NormalTok{ leaders}

\NormalTok{leaders}
\CommentTok{\#\textgreater{} \# A tibble: 4 x 4}
\CommentTok{\#\textgreater{}   code  leader    stdate     enddate   }
\CommentTok{\#\textgreater{}   \textless{}chr\textgreater{} \textless{}chr\textgreater{}     \textless{}date\textgreater{}     \textless{}date\textgreater{}    }
\CommentTok{\#\textgreater{} 1 A     Archie    1975{-}01{-}01 1975{-}01{-}14}
\CommentTok{\#\textgreater{} 2 B     Brian     1975{-}01{-}01 1975{-}01{-}04}
\CommentTok{\#\textgreater{} 3 B     Cornelius 1975{-}01{-}04 1975{-}01{-}08}
\CommentTok{\#\textgreater{} 4 B     Pete      1975{-}01{-}08 1975{-}01{-}14}
\end{Highlighting}
\end{Shaded}

We can use some basic \texttt{rowwise()} transformation to recast these data as daily, resulting in state-day data and leader-day data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{state\_system }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date =} \FunctionTok{list}\NormalTok{(}\FunctionTok{seq}\NormalTok{(stdate, enddate, }\AttributeTok{by =} \StringTok{\textquotesingle{}1 day\textquotesingle{}}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{unnest}\NormalTok{(date) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(code, date) }\OtherTok{{-}\textgreater{}}\NormalTok{ state\_days}

\NormalTok{state\_days }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ data.frame}
\CommentTok{\#\textgreater{}    code       date}
\CommentTok{\#\textgreater{} 1     A 1975{-}01{-}01}
\CommentTok{\#\textgreater{} 2     A 1975{-}01{-}02}
\CommentTok{\#\textgreater{} 3     A 1975{-}01{-}03}
\CommentTok{\#\textgreater{} 4     A 1975{-}01{-}04}
\CommentTok{\#\textgreater{} 5     A 1975{-}01{-}05}
\CommentTok{\#\textgreater{} 6     A 1975{-}01{-}06}
\CommentTok{\#\textgreater{} 7     A 1975{-}01{-}07}
\CommentTok{\#\textgreater{} 8     A 1975{-}01{-}08}
\CommentTok{\#\textgreater{} 9     A 1975{-}01{-}09}
\CommentTok{\#\textgreater{} 10    A 1975{-}01{-}10}
\CommentTok{\#\textgreater{} 11    A 1975{-}01{-}11}
\CommentTok{\#\textgreater{} 12    A 1975{-}01{-}12}
\CommentTok{\#\textgreater{} 13    A 1975{-}01{-}13}
\CommentTok{\#\textgreater{} 14    A 1975{-}01{-}14}
\CommentTok{\#\textgreater{} 15    B 1975{-}01{-}01}
\CommentTok{\#\textgreater{} 16    B 1975{-}01{-}02}
\CommentTok{\#\textgreater{} 17    B 1975{-}01{-}03}
\CommentTok{\#\textgreater{} 18    B 1975{-}01{-}04}
\CommentTok{\#\textgreater{} 19    B 1975{-}01{-}05}
\CommentTok{\#\textgreater{} 20    B 1975{-}01{-}06}
\CommentTok{\#\textgreater{} 21    B 1975{-}01{-}07}

\NormalTok{leaders }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date =} \FunctionTok{list}\NormalTok{(}\FunctionTok{seq}\NormalTok{(stdate, enddate, }\AttributeTok{by =} \StringTok{\textquotesingle{}1 day\textquotesingle{}}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{unnest}\NormalTok{(date) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(code, leader, date) }\OtherTok{{-}\textgreater{}}\NormalTok{ leader\_days}

\NormalTok{leader\_days }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ data.frame}
\CommentTok{\#\textgreater{}    code    leader       date}
\CommentTok{\#\textgreater{} 1     A    Archie 1975{-}01{-}01}
\CommentTok{\#\textgreater{} 2     A    Archie 1975{-}01{-}02}
\CommentTok{\#\textgreater{} 3     A    Archie 1975{-}01{-}03}
\CommentTok{\#\textgreater{} 4     A    Archie 1975{-}01{-}04}
\CommentTok{\#\textgreater{} 5     A    Archie 1975{-}01{-}05}
\CommentTok{\#\textgreater{} 6     A    Archie 1975{-}01{-}06}
\CommentTok{\#\textgreater{} 7     A    Archie 1975{-}01{-}07}
\CommentTok{\#\textgreater{} 8     A    Archie 1975{-}01{-}08}
\CommentTok{\#\textgreater{} 9     A    Archie 1975{-}01{-}09}
\CommentTok{\#\textgreater{} 10    A    Archie 1975{-}01{-}10}
\CommentTok{\#\textgreater{} 11    A    Archie 1975{-}01{-}11}
\CommentTok{\#\textgreater{} 12    A    Archie 1975{-}01{-}12}
\CommentTok{\#\textgreater{} 13    A    Archie 1975{-}01{-}13}
\CommentTok{\#\textgreater{} 14    A    Archie 1975{-}01{-}14}
\CommentTok{\#\textgreater{} 15    B     Brian 1975{-}01{-}01}
\CommentTok{\#\textgreater{} 16    B     Brian 1975{-}01{-}02}
\CommentTok{\#\textgreater{} 17    B     Brian 1975{-}01{-}03}
\CommentTok{\#\textgreater{} 18    B     Brian 1975{-}01{-}04}
\CommentTok{\#\textgreater{} 19    B Cornelius 1975{-}01{-}04}
\CommentTok{\#\textgreater{} 20    B Cornelius 1975{-}01{-}05}
\CommentTok{\#\textgreater{} 21    B Cornelius 1975{-}01{-}06}
\CommentTok{\#\textgreater{} 22    B Cornelius 1975{-}01{-}07}
\CommentTok{\#\textgreater{} 23    B Cornelius 1975{-}01{-}08}
\CommentTok{\#\textgreater{} 24    B      Pete 1975{-}01{-}08}
\CommentTok{\#\textgreater{} 25    B      Pete 1975{-}01{-}09}
\CommentTok{\#\textgreater{} 26    B      Pete 1975{-}01{-}10}
\CommentTok{\#\textgreater{} 27    B      Pete 1975{-}01{-}11}
\CommentTok{\#\textgreater{} 28    B      Pete 1975{-}01{-}12}
\CommentTok{\#\textgreater{} 29    B      Pete 1975{-}01{-}13}
\CommentTok{\#\textgreater{} 30    B      Pete 1975{-}01{-}14}
\end{Highlighting}
\end{Shaded}

If we wanted to standardize these leader-day data to the state system data, we would semi-join the leader-day data (the left object) with the state-day object (the right object), returning just the leader-day data with valid days in the state system data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{leader\_days }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{semi\_join}\NormalTok{(., state\_days) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  data.frame}
\CommentTok{\#\textgreater{}    code    leader       date}
\CommentTok{\#\textgreater{} 1     A    Archie 1975{-}01{-}01}
\CommentTok{\#\textgreater{} 2     A    Archie 1975{-}01{-}02}
\CommentTok{\#\textgreater{} 3     A    Archie 1975{-}01{-}03}
\CommentTok{\#\textgreater{} 4     A    Archie 1975{-}01{-}04}
\CommentTok{\#\textgreater{} 5     A    Archie 1975{-}01{-}05}
\CommentTok{\#\textgreater{} 6     A    Archie 1975{-}01{-}06}
\CommentTok{\#\textgreater{} 7     A    Archie 1975{-}01{-}07}
\CommentTok{\#\textgreater{} 8     A    Archie 1975{-}01{-}08}
\CommentTok{\#\textgreater{} 9     A    Archie 1975{-}01{-}09}
\CommentTok{\#\textgreater{} 10    A    Archie 1975{-}01{-}10}
\CommentTok{\#\textgreater{} 11    A    Archie 1975{-}01{-}11}
\CommentTok{\#\textgreater{} 12    A    Archie 1975{-}01{-}12}
\CommentTok{\#\textgreater{} 13    A    Archie 1975{-}01{-}13}
\CommentTok{\#\textgreater{} 14    A    Archie 1975{-}01{-}14}
\CommentTok{\#\textgreater{} 15    B     Brian 1975{-}01{-}01}
\CommentTok{\#\textgreater{} 16    B     Brian 1975{-}01{-}02}
\CommentTok{\#\textgreater{} 17    B     Brian 1975{-}01{-}03}
\CommentTok{\#\textgreater{} 18    B     Brian 1975{-}01{-}04}
\CommentTok{\#\textgreater{} 19    B Cornelius 1975{-}01{-}04}
\CommentTok{\#\textgreater{} 20    B Cornelius 1975{-}01{-}05}
\CommentTok{\#\textgreater{} 21    B Cornelius 1975{-}01{-}06}
\CommentTok{\#\textgreater{} 22    B Cornelius 1975{-}01{-}07}
\end{Highlighting}
\end{Shaded}

Notice that Pete drops from these data because, in this simple example, Pete was a puppet head of state installed by Archie when state A invaded and occupied B. The semi-join here is simply standardizing the leader data to the state system data, which is effectively what's happening with the semi-joins in \texttt{create\_leaderdays()} (and its aggregation function: \texttt{create\_leaderyears()}).

\hypertarget{anti-join}{%
\subsection{Anti-Join}\label{anti-join}}

The anti-join is another type of filtering join, returning all rows from the left object (\texttt{x}) with\emph{out} a match in the right object (\texttt{y}). This type of join appears just once in \texttt{\{peacesciencer\}}. Prominently, \texttt{\{peacesciencer\}} prepares and presents two data sets in this package---\texttt{false\_cow\_dyads} and \texttt{false\_gw\_dyads}---that represent directed dyad-years in the Correlates of War and Gleditsch-Ward systems that were active in the same year, but never at the same time on the same year.

Here are those dyads for context.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{false\_cow\_dyads}
\CommentTok{\#\textgreater{} \# A tibble: 60 x 4}
\CommentTok{\#\textgreater{}    ccode1 ccode2  year in\_ps}
\CommentTok{\#\textgreater{}     \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{} \textless{}int\textgreater{} \textless{}dbl\textgreater{}}
\CommentTok{\#\textgreater{}  1    115    817  1975     1}
\CommentTok{\#\textgreater{}  2    210    255  1945     1}
\CommentTok{\#\textgreater{}  3    211    255  1945     1}
\CommentTok{\#\textgreater{}  4    223    678  1990     1}
\CommentTok{\#\textgreater{}  5    223    680  1990     1}
\CommentTok{\#\textgreater{}  6    255    210  1945     1}
\CommentTok{\#\textgreater{}  7    255    211  1945     1}
\CommentTok{\#\textgreater{}  8    255    260  1990     1}
\CommentTok{\#\textgreater{}  9    255    265  1990     1}
\CommentTok{\#\textgreater{} 10    255    290  1945     1}
\CommentTok{\#\textgreater{} \# ... with 50 more rows}

\NormalTok{false\_gw\_dyads}
\CommentTok{\#\textgreater{} \# A tibble: 38 x 4}
\CommentTok{\#\textgreater{}    gwcode1 gwcode2  year in\_ps}
\CommentTok{\#\textgreater{}      \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{} \textless{}int\textgreater{} \textless{}dbl\textgreater{}}
\CommentTok{\#\textgreater{}  1      99     100  1830     1}
\CommentTok{\#\textgreater{}  2      99     211  1830     1}
\CommentTok{\#\textgreater{}  3     100      99  1830     1}
\CommentTok{\#\textgreater{}  4     100     615  1830     1}
\CommentTok{\#\textgreater{}  5     115     817  1975     1}
\CommentTok{\#\textgreater{}  6     211      99  1830     1}
\CommentTok{\#\textgreater{}  7     211     615  1830     1}
\CommentTok{\#\textgreater{}  8     255     850  1945     1}
\CommentTok{\#\textgreater{}  9     300     305  1918     1}
\CommentTok{\#\textgreater{} 10     300     345  1918     1}
\CommentTok{\#\textgreater{} \# ... with 28 more rows}
\end{Highlighting}
\end{Shaded}

These were created by \href{https://github.com/svmiller/peacesciencer/blob/master/data-raw/false_cow_dyads.R}{two} \href{https://github.com/svmiller/peacesciencer/blob/master/data-raw/false_gw_dyads.R}{scripts} that, for each year in the respective state system data, creates every possible \emph{daily} dyadic pairing and truncates the dyads to just those that had at least one day of overlap. This is a computationally demanding procedure compared to what \texttt{\{peacesciencer\}} does (which creates every possible dyadic pair in a given year, given the state system data supplied to it). However, it creates the possibility of same false dyads in a given year that showed no overlap.

Consider the case of Suriname (115) and the Republic of Vietnam (817) in 1975 as illustrative here.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{check\_both }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x) \{}

\NormalTok{  gw\_states }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{data =} \StringTok{"G{-}W"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{filter}\NormalTok{(gwcode }\SpecialCharTok{\%in\%}\NormalTok{ x) }\OtherTok{{-}\textgreater{}}\NormalTok{ gwrows}

\NormalTok{  cow\_states }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{startdate =} \FunctionTok{ymd}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(styear,}\StringTok{"/"}\NormalTok{,stmonth, }\StringTok{"/"}\NormalTok{, stday)),}
           \AttributeTok{enddate =} \FunctionTok{ymd}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(endyear,}\StringTok{"/"}\NormalTok{,endmonth,}\StringTok{"/"}\NormalTok{,endday))) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{select}\NormalTok{(stateabb}\SpecialCharTok{:}\NormalTok{statenme, startdate, enddate) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{data =} \StringTok{"CoW"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{rename}\NormalTok{(}\AttributeTok{statename =}\NormalTok{ statenme) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{filter}\NormalTok{(ccode }\SpecialCharTok{\%in\%}\NormalTok{ x) }\OtherTok{{-}\textgreater{}}\NormalTok{ cowrows}

\NormalTok{  dat }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(gwrows, cowrows) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{select}\NormalTok{(gwcode, ccode, stateabb, }\FunctionTok{everything}\NormalTok{())}

  \FunctionTok{return}\NormalTok{(dat)}
\NormalTok{\}}

\FunctionTok{check\_both}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{115}\NormalTok{, }\DecValTok{817}\NormalTok{))}
\CommentTok{\#\textgreater{} \# A tibble: 4 x 7}
\CommentTok{\#\textgreater{}   gwcode ccode stateabb statename            startdate  enddate    data }
\CommentTok{\#\textgreater{}    \textless{}dbl\textgreater{} \textless{}dbl\textgreater{} \textless{}chr\textgreater{}    \textless{}chr\textgreater{}                \textless{}date\textgreater{}     \textless{}date\textgreater{}     \textless{}chr\textgreater{}}
\CommentTok{\#\textgreater{} 1    115    NA SUR      Surinam              1975{-}11{-}25 2017{-}12{-}31 G{-}W  }
\CommentTok{\#\textgreater{} 2    817    NA RVN      Vietnam, Republic of 1954{-}05{-}01 1975{-}04{-}30 G{-}W  }
\CommentTok{\#\textgreater{} 3     NA   115 SUR      Suriname             1975{-}11{-}25 2016{-}12{-}31 CoW  }
\CommentTok{\#\textgreater{} 4     NA   817 RVN      Republic of Vietnam  1954{-}06{-}04 1975{-}04{-}30 CoW}
\end{Highlighting}
\end{Shaded}

Notice both Suriname and Republic of Vietnam were both active in 1975. Suriname appears on Nov.~25, 1975 whereas the Republic of Vietnam exits on April 30, 1975. However, there is no daily overlap between the two because they did not exist at any point on the same day in 1975. These are false dyads. \href{https://github.com/svmiller/peacesciencer/blob/master/R/create_dyadyears.R}{\texttt{anti\_join()} is used in the \texttt{create\_dyadyears()}} function to remove these observations before presenting them to the user.

Here is a simple example of what an anti-join is doing with these examples in mind.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{valid\_dyads }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}\AttributeTok{ccode1 =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{200}\NormalTok{),}
                      \AttributeTok{ccode2 =} \FunctionTok{c}\NormalTok{(}\DecValTok{20}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{900}\NormalTok{),}
                      \AttributeTok{year =} \FunctionTok{c}\NormalTok{(}\DecValTok{2016}\NormalTok{, }\DecValTok{2017}\NormalTok{, }\DecValTok{2018}\NormalTok{))}

\NormalTok{valid\_dyads }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{bind\_rows}\NormalTok{(., false\_cow\_dyads }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(ccode1}\SpecialCharTok{:}\NormalTok{year)) }\OtherTok{{-}\textgreater{}}\NormalTok{ valid\_and\_invalid}

\NormalTok{valid\_and\_invalid }
\CommentTok{\#\textgreater{} \# A tibble: 63 x 3}
\CommentTok{\#\textgreater{}    ccode1 ccode2  year}
\CommentTok{\#\textgreater{}     \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}}
\CommentTok{\#\textgreater{}  1      2     20  2016}
\CommentTok{\#\textgreater{}  2     20    200  2017}
\CommentTok{\#\textgreater{}  3    200    900  2018}
\CommentTok{\#\textgreater{}  4    115    817  1975}
\CommentTok{\#\textgreater{}  5    210    255  1945}
\CommentTok{\#\textgreater{}  6    211    255  1945}
\CommentTok{\#\textgreater{}  7    223    678  1990}
\CommentTok{\#\textgreater{}  8    223    680  1990}
\CommentTok{\#\textgreater{}  9    255    210  1945}
\CommentTok{\#\textgreater{} 10    255    211  1945}
\CommentTok{\#\textgreater{} \# ... with 53 more rows}

\NormalTok{valid\_and\_invalid }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# remove those invalid dyads{-}years}
  \FunctionTok{anti\_join}\NormalTok{(., false\_cow\_dyads)}
\CommentTok{\#\textgreater{} \# A tibble: 3 x 3}
\CommentTok{\#\textgreater{}   ccode1 ccode2  year}
\CommentTok{\#\textgreater{}    \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}}
\CommentTok{\#\textgreater{} 1      2     20  2016}
\CommentTok{\#\textgreater{} 2     20    200  2017}
\CommentTok{\#\textgreater{} 3    200    900  2018}
\end{Highlighting}
\end{Shaded}

\hypertarget{vignette-peacesciencer-data-versions}{%
\section{\texorpdfstring{Vignette: \texttt{\{peacesciencer\}} Data Versions}{Vignette: \{peacesciencer\} Data Versions}}\label{vignette-peacesciencer-data-versions}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{library}\NormalTok{(peacesciencer)}
\FunctionTok{library}\NormalTok{(kableExtra)}
\end{Highlighting}
\end{Shaded}

These are the data versions available in \texttt{\{peacesciencer\}}. Do note the user can find specific data versions with more targeted use of the \texttt{ps\_version()} function in this package. Without an argument, \texttt{ps\_version()} produces a data frame of all the data versions in this package.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ps\_version}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kbl}\NormalTok{(., }\AttributeTok{caption =} \StringTok{"Data Versions in \textasciigrave{}\{peacesciencer\}\textasciigrave{}"}\NormalTok{,}
      \AttributeTok{align =}\FunctionTok{c}\NormalTok{(}\StringTok{"c"}\NormalTok{, }\StringTok{"l"}\NormalTok{, }\StringTok{"c"}\NormalTok{, }\StringTok{"c"}\NormalTok{), }
      \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{longtable =} \ConstantTok{TRUE}\NormalTok{)  }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{position =} \StringTok{"center"}\NormalTok{, }\AttributeTok{full\_width =}\NormalTok{ F, }
                \AttributeTok{bootstrap\_options =} \StringTok{"striped"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{row\_spec}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{bold=}\ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[t]{clcc}
\caption{\label{tab:unnamed-chunk-20}Data Versions in `{peacesciencer}`}\\
\toprule
\textbf{category} & \textbf{data} & \textbf{version} & \textbf{bibtexkey}\\
\midrule
states & Correlates of War State System Membership & 2016 & cowstates2016\\
leaders & LEAD & 2015 & ellisetal2015lead\\
leaders & Archigos & 4.1 & goemansetal2009ia\\
alliance & ATOP & 5 & leedsetal2002atop\\
democracy & Polity & 2017 & marshalletal2017p\\
\addlinespace
democracy & \{QuickUDS\} & 0.2.3 & marquez2016qme\\
democracy & V-Dem & 10 & coppedgeetal2020vdem\\
capitals & \{peacesciencer\} & 2020 & peacesciencer-package\\
contiguity & Correlates of War Direct Contiguity & 3.2 & stinnettetal2002cow\\
igo & Correlates of War IGOs & 3 & pevehouseetal2020tow\\
\addlinespace
majors & Correlates of War & 2016 & cowstates2016\\
conflict\_interstate & Correlates of War Militarized Interstate Disputes & 5 & palmeretal2021mid5\\
distance & \{Cshapes\} & 2 & schvitz2021mis\\
capabilities & Correlates of War National Material Capabilities & 5 & singer1987rcwd\\
gdp & SDP & 2020 & andersetal2020bbgb\\
\addlinespace
sdp & SDP & 2020 & andersetal2020bbgb\\
population & SDP & 2020 & andersetal2020bbgb\\
trade & Correlates of War Trade & 4 & barbierietal2009td\\
conflict\_intrastate & Correlates of War Intra-State War & 4.1 & dixonsarkees2016giw\\
conflict\_interstate & Correlates of War Inter-State War & 4 & sarkeeswayman2010rw\\
\addlinespace
fractionalization & CREG & 2012 & nardulli2012creg\\
polarization & CREG & 2012 & nardulli2012creg\\
conflict\_interstate & Gibler-Miller-Little (GML) & 2.2.1 & gibleretal2016amid\\
states & Gleditsch-Ward & 2017 & gledtischward1999rlis\\
leaders & Leader Willingness to Use Force & 2020 & cartersmith2020fml\\
\addlinespace
terrain & Ruggedness & 2012 & nunnpuga2012r\\
terrain & \% Mountainous & 2014 & giblermiller2014etts\\
rivalries & Thompson and Dreyer & 2012 & thompsondreyer2012hir\\
conflict\_intrastate & UCDP Armed Conflicts & 20.1 & gleditschetal2002ac\\
conflict\_intrastate & UCDP Onsets & 19.1 & pettersson2019ov\\
\addlinespace
dyadic\_similarity & FPSIM & 2 & haege2011cc\\
\bottomrule
\end{longtable}

Every data set that is used by a function in this package is included in this table (and with the \texttt{ps\_version()} function). Users can see a category of the type of data (which can be used for more careful searches of the underlying data), a description of the data set, a version number associated with that data set in \texttt{\{peacesciencer\}}, along with a BibTeX key. Users should interpret data versions as years as instances where the data are not formally versioned, per se, and the year corresponds with a year of a last update or a year of publication. For example, \href{https://doi.org/10.1093/isq/sqaa013}{Anders et al.~(2020)} released their state-year simulations of population, surplus domestic product, and gross domestic product in 2020. The data are not formally versioned and the year corresponds with, in this case, the publication.

Users can use the BibTeX key (\texttt{bibtexkey}) to search \href{http://svmiller.com/peacesciencer/reference/ps_bib.html}{the dataframe of citations} in this package. For example, we can use \texttt{ps\_cite()} to get a full citation for Carter and Smith's (2020) estimates of leader willingness to use force.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ps\_cite}\NormalTok{(}\StringTok{"cartersmith2020fml"}\NormalTok{, }\AttributeTok{column =} \StringTok{"bibtexkey"}\NormalTok{)}
\CommentTok{\#\textgreater{} @Article\{cartersmith2020fml,}
\CommentTok{\#\textgreater{}   Author = \{Jeff Carter and Charles E. Smith\},}
\CommentTok{\#\textgreater{}   Journal = \{American Political Science Review\},}
\CommentTok{\#\textgreater{}   Number = \{4\},}
\CommentTok{\#\textgreater{}   Pages = \{1352{-}{-}1358\},}
\CommentTok{\#\textgreater{}   Title = \{A Framework for Measuring Leaders’ Willingness to Use Force\},}
\CommentTok{\#\textgreater{}   Volume = \{114\},}
\CommentTok{\#\textgreater{}   Year = \{2020\},}
\CommentTok{\#\textgreater{}   Keywords = \{lwuf, add\_lwuf()\}}
\CommentTok{\#\textgreater{} \}}
\end{Highlighting}
\end{Shaded}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{)}
\CommentTok{\#\textgreater{} \# A tibble: 2,029,622 x 3}
\CommentTok{\#\textgreater{}    gwcode1 gwcode2  year}
\CommentTok{\#\textgreater{}      \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{} \textless{}int\textgreater{}}
\CommentTok{\#\textgreater{}  1       2      20  1867}
\CommentTok{\#\textgreater{}  2       2      20  1868}
\CommentTok{\#\textgreater{}  3       2      20  1869}
\CommentTok{\#\textgreater{}  4       2      20  1870}
\CommentTok{\#\textgreater{}  5       2      20  1871}
\CommentTok{\#\textgreater{}  6       2      20  1872}
\CommentTok{\#\textgreater{}  7       2      20  1873}
\CommentTok{\#\textgreater{}  8       2      20  1874}
\CommentTok{\#\textgreater{}  9       2      20  1875}
\CommentTok{\#\textgreater{} 10       2      20  1876}
\CommentTok{\#\textgreater{} \# ... with 2,029,612 more rows}
\end{Highlighting}
\end{Shaded}

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parskip}{6pt}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-andersetal2020bbgb}{}%
Anders, Therese, Christopher J Fariss \& Jonathan N Markowitz (2020) Bread before guns or butter: Introducing surplus domestic product (SDP). \emph{International Studies Quarterly} 64(2): 392--405.

\leavevmode\hypertarget{ref-arelbundock2021m}{}%
Arel-Bundock, Vincent (2021) \emph{Modelsummary: Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready} (\url{https://CRAN.R-project.org/package=modelsummary}).

\leavevmode\hypertarget{ref-cartersmith2020fml}{}%
Carter, Jeff \& Charles E Smith (2020) A framework for measuring leaders' willingness to use force. \emph{American Political Science Review} 114(4): 1352--1358.

\leavevmode\hypertarget{ref-wickhametal2019wt}{}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, Alex Hayes, Lionel Henry, Jim Hester, Max Kuhn, Thomas Lin Pedersen, Evan Miller, Stephan Milton Bache, Kirill Müller, Jeroen Ooms, David Robinson, Dana Paige Seidel, Vitalie Spinu, Kohske Takahashi, Davis Vaughan, Claus Wilke, Kara Woo \& Hiroaki Yutani (2019) Welcome to the {tidyverse}. \emph{Journal of Open Source Software} 4(43): 1686.

\leavevmode\hypertarget{ref-xie2015ddrk}{}%
Xie, Yihui (2015) \emph{Dynamic Documents with {R} and Knitr}. Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-xie2016b}{}%
Xie, Yihui (2016) \emph{Bookdown: Authoring Books and Technical Documents with {R} Markdown}. Boca Raton, Florida: Chapman; Hall/CRC (\url{https://bookdown.org/yihui/bookdown}).

\leavevmode\hypertarget{ref-xieetal2020rmc}{}%
Xie, Yihui, Christophe Dervieux \& Emily Riederer (2020) \emph{R Markdown Cookbook}. Boca Raton, Florida: Chapman; Hall/CRC (\url{https://bookdown.org/yihui/rmarkdown-cookbook}).

\leavevmode\hypertarget{ref-zhu2021k}{}%
Zhu, Hao (2021) \emph{kableExtra: Construct Complex Table with {``Kable''} and Pipe Syntax} (\url{https://CRAN.R-project.org/package=kableExtra}).

\end{CSLReferences}

\end{document}

### 2021-11-09 user-guide/src/render_appendix.R

@@ -1,7 +1,7 @@
 the_output <- "latex"
 
-rmarkdown::render("appendix.Rmd", 
-                  output_file="doc/appendix.pdf", 
-                  bookdown::pdf_document2(template = stevetemplates::templ_article2(), 
+rmarkdown::render("appendix.Rmd",
+                  output_file="doc/appendix.pdf",
+                  bookdown::pdf_document2(template = stevetemplates::templ_article2(),
                                           latex_engine = "xelatex", dev="cairo_pdf", toc = FALSE,
                                           number_sections = FALSE, keep_tex = FALSE))

### 2021-11-09 user-guide/src/render_pdf-anon.R

@@ -11,6 +11,7 @@ rmarkdown::render("ms.Rmd", output_file="doc/ms-anon.pdf",
                               ps_manuals = "[LINK REDACTED FOR PEER REVIEW]",
                               ps_github = "[LINK REDACTED FOR PEER REVIEW]"),
                   bookdown::pdf_document2(template = stevetemplates::templ_article2(),
+                                          keep_tex = FALSE,
                                           latex_engine = "xelatex", dev= "cairo_pdf",
                                           toc = FALSE, number_sections = FALSE))
 

### 2021-11-09 user-guide/src/render_pdf.R

@@ -11,7 +11,7 @@ rmarkdown::render("ms.Rmd",
                               ps_manuals = "http://svmiller.com/peacesciencer/reference",
                               ps_github = "https://github.com/svmiller/peacesciencer/"),
                   bookdown::pdf_document2(template = stevetemplates::templ_article2(),
-                                          keep_tex = TRUE,
+                                          keep_tex = FALSE,
                                           latex_engine = "xelatex", dev="cairo_pdf",
                                           toc = FALSE, number_sections = FALSE))
 

### 2021-11-09 user-guide/appendix.Rmd
update
works in progress

@@
 library(tidyverse)
+library(peacesciencer)
 
@@
 
-I offer these citations first as an acknowledgement of these important contributions to our collective research productivity.
+I offer these citations first as an acknowledgment of these important contributions to our collective research productivity.
 Space restrictions in our journal preclude us from citing these important tools, even if they are truly ancillary to the product they help produce.
@@
 
+```{r, echo=TRUE}
+create_dyadyears(system = 'gw')
+```
+
 # References

### 2021-11-09 user-guide/ms.Rmd
works in progress

@@
 removetitleabstract: "`r params$removetitleabstract`"
-alwaays_allow_html: TRUE
+#always_allow_html: TRUE
 header-includes:
@@
   - \usepackage{booktabs}
+  - \usepackage{csquotes}
   - \LTcapwidth=.95\textwidth
@@
   - \usepackage{setspace}
-  - \let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
-  - \usepackage{colortbl}
-  - \arrayrulecolor{black}
+#  - \let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
+#  - \usepackage{colortbl}
+#  - \arrayrulecolor{black}
+#  - \usepackage{threeparttable}
+#  - \usepackage{threeparttablex}
 #  - \midrule\relax
@@
 is_html <- knitr::pandoc_to("html")
+
+table_format <- ifelse(is_docx, "huxtable", 'default')
 
@@
 
-knitr::opts_chunk$set(echo = FALSE, cache = FALSE,
+knitr::opts_chunk$set(cache = FALSE,
                       message=FALSE, warning=FALSE,
+                      collapse = TRUE,
+                      comment = "#>",
                       fig.path='doc/figs/',
@@
 library(tidyverse)
+library(peacesciencer)
 library(stevemisc)
+
 library(modelsummary)
-library(huxtable)
 library(kableExtra)
-library(peacesciencer)
 
@@
     }
-
 ```
@@
 Software options exist that allow the researcher to load data, create data, clean data, analyze data, and present the results of the analysis all within one program.
-`{peacesciencer}`, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and Rstudio (its free-for-use integrated desktop environment [IDE]).
+`{peacesciencer}`, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and Rstudio (its free-for-use integrated development environment [IDE]).
 
@@
 
-`{peacesciencer}` comes with a fully developed suite of built-in functions for generating the most widespread forms of peace science data.
-Each function uses raw or pre-processed data included in `{peacesciencer}`.
+`{peacesciencer}` comes with a fully developed suite of built-in functions for generating some of the most widespread forms of peace science data and populating the data with important variables that recur in many quantitative analyses.
+The core functionality of `{peacesciencer}` reduces to two broad categories of functions.
+These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis.
+Table \@ref(tab:tab-createfunctions) and Table \@ref(tab:tab-addfunctions) list these core functions as of version 0.7, the version of this package slated for release alongside the manuscript.
+
+Table \@ref(tab:tab-createfunctions) are functions that crease base data frames for a researcher starting an original project.
+They serve as functions that communicate the units of analysis supported in this package and that the package is capable of generating for an interested user.
+For example, `create_stateyears()` will generate the full universe of state-years from the Correlates of War [@cowstates2016, v.
+2016] or Gleditsch-Ward [@gledtischward1999rlis, v.2017] system, encompassing all state-years to the most recently concluded calendar year, depending on the arguments supplied to the user in the function.
+`create_leaderyears()` will generate the full universe of leader-years from the Archigos leader data [@goemansetal2009ia, v.
+4.1], optionally standardizing leader-years to the Gleditsch-Ward or Correlates of War state system data.
+As of version 0.7, `{peacesciencer}` is capable of creating full dyad-year data, leader-day data, leader-dyad-year data, leader-year data, state-day data, and state-year data.[^leaderdyadyear] A vignette on the package's website shows how users can create other forms of data from these functions as well (e.g.
+leader-months, state-quarters).
+
+[^leaderdyadyear]: Leader-dyad-year data are too time-consuming to calculate each time and too size-prohibitive to include in the package.
+These data can be downloaded through `download_extdata()` for much easier use.
+
+```{r tab-createfunctions, echo=F}
+
+tribble(~func, ~description,
+        "create_dyadyears()", "Create dyad-years from state system membership data",
+        "create_leaderdays()", "Create leader-day data from Archigos",
+        "create_leaderdyadyears()", "Create leader-dyad-years data from Archigos",
+        "create_leaderyears()", "Create leader-years data from Archigos",
+        "create_statedays()", "Create state-days from state system membership data",
+        "create_stateyears()", "Create state-years from state system membership data",
+) -> create_functions
+
+if (is_latex) {
+kable(create_functions, booktabs = TRUE, longtable = TRUE,
+    caption = "Functions that Create Base Data",
+    align = c("l","l"),
+    linesep = "",
+    col.names = c("Function", "Description"))  %>%
+   row_spec(0, bold=TRUE) %>%
+    column_spec(1, monospace = TRUE)
+} else {
+  kable(create_functions, booktabs = TRUE, longtable = TRUE,
+    caption = "Functions that Create Base Data",
+    align = c("l","l"),
+    linesep = "",
+    col.names = c("Function", "Description"))  %>%
+   row_spec(0, bold=TRUE) %>%
+    column_spec(1, monospace = TRUE)
+}
+
+```
+
+Table \@ref(tab:tab-addfunctions) lists the main functions in `{peacesciencer}` that add information or subset the number of rows of the data to just those of interest to the user, describing these functions and listing whether they are applicable to dyad-year (`D`), leader-year (`L`), leader-dyad-year (`LD`), state-year (`S`) data or specialty functions applicable to just the dyadic conflict data (`C`).[^whittle] All these functions use raw or pre-processed data included in the package.
+For example, `add_gml_mids()` uses a dyadic dispute-year version of the MID data offered by Gibler et al.
+[-@gibleretal2016amid, v.
+2.2.1] and merges in information about whether there was an ongoing MID or MID onset in a dyad-year, leader-year, leader-dyad-year, or state-year.
+`{peacesciencer}` also has some data innovations included in these functions.
+For example, `add_capital_distance()` calculates distance between state capitals in kilometers using the Vincenty method (i.e.
+"as the crow flies") based on an original data set of state capitals that accounts for instances when capitals moved (e.g.
+Brazil in 1960, Burundi in 2018).
+`add_democracy()` does more than just add Polity data to a data set.
+The data underpinning the function feature an innovation in democracy data, providing reasonable estimates of democracy using the @marquez2016qme method of extending the Unified Democracy Scores (UDS) data [@pemsteinetal2010dc] in addition to Polity estimates [@marshalletal2017p, v.
+2017] and V-Dem estimates [@coppedgeetal2020vdem, v.
+10].[^demvig]
+
+[^demvig]: `{peacesciencer}` comes with a vignette that explains just how much unnecessary missingness pervades data on democracy in the absence of an innovation like this, and how this missingness might adversely affect inferences of inter-state or intra-state conflict.
+
+[^whittle]: The "whittle" (`wc_`) class of functions are a suite of functions that whittle dyadic dispute-years to true dyad-year data.
+They are effectively used in `add_cow_mids()` and `add_gml_mids()` for dyad-year data, but are offered here to allow users to employ their own case exclusion rules to create their own dyad-year conflict data.
+A vignette on `{peacesciencer}`'s website explains these case exclusion rules in some detail.
+
+```{r tab-addfunctions, echo=F}
+
+tribble(~func, ~type, ~description,
+        "add_archigos()", "D,S", "Add Archigos political leader information to a data frame",
+        "add_atop_alliance()", "D, LD", "Add Alliance Treaty Obligations and Provisions (ATOP) alliance data to a data frame",
+        "add_capital_distance()", "D, L, LD, S", "Add capital-to-capital distance to a data frame",
+        "add_ccode_to_gw()", "D, L, LD, S", "Match CoW state codes to G-W state codes",
+        "add_contiguity()", "D, L, LD, S", "Add CoW direct contiguity information to a data frame",
+        "add_cow_alliance()", "D, LD", "Add CoW alliance data to a data frame",
+        "add_cow_majors()", "D, L, LD, S", "Add CoW major power information to a data frame",
+        "add_cow_mids()", "D", "Add CoW-MID data to a dyad-year data frame",
+        "add_cow_trade()", "D, L, LD, S", "Add CoW trade data to a data frame",
+        "add_cow_wars()", "D, S", "Add CoW war data to a dyad-year or state-year data frame",
+        "add_creg_fractionalization()", "D, L, LD, S", "Add fractionalization/polarization estimates from CREG to a data frame",
+        "add_democracy()", "D, L, LD, S", "Add democracy information to a data frame",
+        "add_fpsim()", "D, LD", "Add dyadic foreign policy similarity measures to a data frame",
+        "add_gml_mids()", "D, L, LD, S", "Add GML MID data to a data frame",
+        "add_gwcode_to_cow()", "D, L, LD, S", "Match G-W state codes to CoW state codes",
+        "add_igos()", "D, S", "Add CoW IGO data to a data frame",
+        "add_lead()", "L, LD", "Add leader experience/attributes (LEAD) to data",
+        "add_lwuf()", "L, LD", "Add leader willingness to use force estimates to data",
+        "add_minimum_distance()", "D, L, LD, S", "Add minimum distance data to data frame",
+        "add_nmc()", "D, L, LD, S", "Add minimum distance estimates (in km) to data frame",
+        "add_peace_years()", "D, S", "Add 'peace years' to dyad-year/state-year conflict data",
+        "add_rugged_terrain()", "D, L, LD, S", "Add rugged terrain information to a data frame",
+        "add_sdp_gdp()", "D, L, LD, S", "Add (surplus, gross) domestic product data to a data frame",
+        "add_spells()", "D, L, LD, S", "Add duration 'spells' to a data frame",
+        "add_strategic_rivalries()", "D, S", "Add strategic rivalry information to a data frame",
+        "add_ucdp_acd()", "S", "Add UCDP Armed Conflict Data to state-year data frame",
+        "add_ucdp_onsets", "S", "Add UCDP onsets to state-year data",
+        "filter_prd()", "D, LD", "Filter dyad-year data to just politically relevant dyads",
+        "wc_duration()", "C", "Whittle Duplicate Conflict-Years by (Minimum or Maximum) Conflict Duration",
+        "wc_fatality()", "C", "Whittle Duplicate Conflict-Years by Highest Fatality",
+        "wc_hostility()", "C", "Whittle Duplicate Conflict-Years by Highest Hostility",
+        "wc_jds()", "C", "Whittle Duplicate Conflict-Years by Just Dropping Something",
+        "wc_onsets()", "C", "Whittle Uniqute Conflict Onset-Years from Conflict-Year Data",
+        "wc_recip()", "C", "Whittle Duplicate Conflict-Years by Conflict Reciprocation",
+        "wc_stmon()", "C", "Whittle Duplicate Conflict-Years by Lowest Start Month"
+) -> add_functions
+
+if (is_latex) {
+kable(add_functions, booktabs = TRUE, longtable = TRUE,
+    caption = "Functions that Add Data",
+    align = c("l", "c", "l"),
+    linesep = "",
+   col.names = c("Function", "Type", "Description"))  %>%
+    kable_styling(font_size = 9) %>%
+   row_spec(0, bold=TRUE) %>%
+    column_spec(1, monospace = TRUE) %>%
+    column_spec(3, width="30em")
+} else {
+  kable(add_functions, booktabs = TRUE, longtable = TRUE,
+    caption = "Functions that Add Data",
+    align = c("l", "c", "l"),
+    linesep = "",
+    col.names = c("Function",  "Type", "Description"))  %>%
+   row_spec(0, bold=TRUE) %>%
+    column_spec(1, monospace = TRUE)
+}
+
+```
+
+<!-- Each function uses raw or pre-processed data included in `{peacesciencer}`.
 For example, `create_dyadyears()` transforms the raw state system membership data for either the Correlates of War or Gleditsch-Ward system into the full universe of post-1816 dyad-years.
 `add_gml_mids()` uses a dyadic dispute-year version of the MID data offered by @gibleretal2016amid, which, as the `data-raw` directory on the project's Github shows, is itself derived from the dispute-level and participant-level versions of the data.
-`{peacesciencer}` also has some data innovations.
-It has an original data set on capital-to-capital distances, calculating for distance in kilometers between state capitals using the "Vincenty" method (i.e.
-"as the crow flies") that also accounts for instances when capitals moved (e.g.
-Brazil in 1960, Burundi in 2018).
-It also features an innovation of democracy data, providing reasonable estimates of democracy using the @marquez2016qme method of extending the Unified Democracy Scores (UDS) data [@pemsteinetal2010dc].[^demvig] As of writing, `{peacesciencer}` comes with the following functionality described in Table \@ref(tab:tab-functionality).
+-->
 
-[^demvig]: `{peacesciencer}` comes with a vignette that explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.
-
-```{r tab-functionality, echo=F, eval=T}
+```{r tab-functionality, echo=F, eval=F}
 tribble(~data, ~description, ~type, ~func,
-        "(CoW, G-W) Dyad-Year Data", "Create dyad-years from state system membership data", "", "create_dyadyears()",
-        "(CoW, G-W) State-Year Data", "Create state-years from state system membership data", "", "create_stateyears()",
-        "(CoW, G-W) State-Day Data", "Create state-days from state system membership data", "", "create_statedays()",
-        "Archigos (v. 4.1)", "Add Archigos political leader information", "D, S", "add_archigos()",
+        "(CoW, G-W) Dyad-Year Data", "Create dyad-years from state system membership data", " ", "create_dyadyears()",
+        "(CoW, G-W) State-Year Data", "Create state-years from state system membership data", " ", "create_stateyears()",
+        "(CoW, G-W) State-Day Data", "Create state-days from state system membership data", " ", "create_statedays()",
+        "Archigos Leader-Day Data", "Create leader-day data from Archigos", " ", "create_leaderdays()",
+        "Archigos Leader-Year Data", "Create leader-year data from Archigos", " ", "create_leadeyears()",
+        "Archigos Leader-Dyad-Year Data", "Create leader-dyad-year data from Archigos", " ", "create_leaderdyadyears()",
+        "Archigos", "Add Archigos political leader information", "D, S", "add_archigos()",
         "ATOP Alliances (v. 5.0)", "Add ATOP alliance data", "D", "add_atop_alliance()",
@@
         "Convert G-W codes to CoW codes", "Add matching G-W codes for CoW states by year", "D, S", "add_gwcode_to_cow()",
-        "CoW National Capabilities (v. 5.0)", "Adds capabilities of states by year", "D, S", "add_nmc()",
-        "CoW Direct Contiguity (v. 3.2)", "Adds contiguity information by year", "D, S", "add_contiguity()",
-        "CoW Alliances (v. 4.1)", "Adds CoW alliance data by year", "D", "add_cow_alliance()",
-        "CoW IGOs (v. 3.0)", "Adds CoW IGO data by year", "D, S", "add_igos()",
-        "CoW Major Powers (v. 2016)", "Adds CoW major power by year", "D, S", "add_cow_majors()",
-        "CoW MIDs (v. 5.0)", "Add CoW MID information to data", "D", "add_cow_mids()",
-        "CoW Trade Data (v. 4.0)", "Add CoW trade information to data", "D, S", "add_cow_trade()",
+        "CoW National Capabilities", "Adds capabilities of states by year", "D, S", "add_nmc()",
+        "CoW Direct Contiguity", "Adds contiguity information by year", "D, S", "add_contiguity()",
+        "CoW Alliances", "Adds CoW alliance data by year", "D", "add_cow_alliance()",
+        "CoW IGOs", "Adds CoW IGO data by year", "D, S", "add_igos()",
+        "CoW Major Powers", "Adds CoW major power by year", "D, S", "add_cow_majors()",
+        "CoW MIDs", "Add CoW MID information to data", "D", "add_cow_mids()",
+        "CoW Trade Data", "Add CoW trade information to data", "D, S", "add_cow_trade()",
         "CoW Wars", "Add CoW inter-state (v. 4.0) and intra-state (v. 4.1) war data", "D", "add_cow_wars()",
-        "CREG (v. 1.02)", "Add ethnic/religious fractionalization/polarization by year", "D, S", "add_creg_fractionalization()",
+        "CREG", "Add ethnic/religious fractionalization/polarization by year", "D, S", "add_creg_fractionalization()",
         "Democracy", "Add Polity (v. 2017), V-Dem (v. 10), and Marquez' (2016) UDS extensions", "D,S", "add_democracy()",
-        "GML MIDs (v. 2.1.1)", "Add Gibler-Miller-Little MIDs to data", "D", "add_gml_mids()",
-        "Minimum Distance (v. 2016)", "Adds minimum distance data by year", "D, S", "add_minimum_distance()",
+        "GML MIDs", "Add Gibler-Miller-Little MIDs to data", "D", "add_gml_mids()",
+        "Minimum Distance", "Adds minimum distance data by year", "D, S", "add_minimum_distance()",
         "Peace Years", "Calculate peace years for various disputes", "D, S", "add_peace_years()",
@@
 
+# if (is_latex) {
+# kable(included, #booktabs = FALSE, longtable = TRUE,
+#     caption = "Features of \\{peacesciencer\\}",
+#     align = c("l","l","c", "c"),
+#     col.names = c("Data", "Description", "Type", "Function")) %>%
+#   kable_styling(font_size = 8) %>%
+#   row_spec(0, bold=TRUE) %>%
+#   column_spec(4, monospace = TRUE) %>%
+#   column_spec(c(1,2,3), border_left = FALSE) %>%
+#   row_spec((head(c(1:nrow(included)), -1)), extra_latex_after = "\\arrayrulecolor{white}") %>%
+#   row_spec((included %>% 
+#              nrow()), extra_latex_after = "\\arrayrulecolor{black}")
+# } else {
+#   kable(included,
+#     caption = "Features of \\{peacesciencer\\}",
+#     align = c("l","l","c", "c"),
+#     col.names = c("Data", "Description", "Type", "Function")) %>%
+#   # kable_styling(font_size = 7) %>%
+#   row_spec(0, bold=TRUE) %>%
+#   column_spec(4, monospace = TRUE) 
+# }
+
 if (is_latex) {
-kable(included, #booktabs = FALSE, longtable = TRUE,
-    caption = "Features of \\{peacesciencer\\}",
+kable(included, booktabs = TRUE, longtable = FALSE,
+    caption = "Features of This Package",
     align = c("l","l","c", "c"),
+    linesep = "",
     col.names = c("Data", "Description", "Type", "Function"))  %>%
@@
    row_spec(0, bold=TRUE) %>%
-  column_spec(4, monospace = TRUE) %>%
-  column_spec(c(1,2,3), border_left = FALSE) %>%
-  row_spec((head(c(1:nrow(included)), -1)), extra_latex_after = "\\arrayrulecolor{white}") %>%
-  row_spec((included %>% 
-             nrow()), extra_latex_after = "\\arrayrulecolor{black}")
+    column_spec(4, monospace = TRUE)
 } else {
@@
 @arelbundock2021w].
-Its coverage will assuredly expand with new additions of interest to the peace science community.
+Its coverage will assuredly expand with new additions of interest to the peace science community, though the package already offers a lot to meet researcher needs.
 
@@
 This assumes at least some familiarity with the R programming language.
-Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.0.
+Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.2.
 `{peacesciencer}` is designed to be as user-friendly as possible.
@@
 
-`{peacesciencer}`'s functions work out of the box, though it works best with one additional package.
-This is `{tidyverse}`, itself a suit of packages that share a common form and design [@wickhamgrolemund2017rds].
+`{peacesciencer}`'s functions work out of the box, though users should find their experience augmented by two additional downloads.
+First, RStudio offers an IDE that serves as a user-friendly graphical user interface (GUI) over what is, at its core, a programming language with a command-line interface.
+Rstudio's design will make it much easier for users to experiment with `{peacesciencer}`'s functionality and read its documentation to assist them with the user of these functions.
+The second additional download is `{tidyverse}`, itself a suit of packages that share a common form and design [@wickhametal2019wt].
 `{peacesciencer}` functions make considerable use of the component packages of `{tidyverse}` but installing and loading `{tidyverse}` will allow the researcher to make quicker use of `{peacesciencer}`'s functionality.
-A user can open an R session and install both packages as follows.
+A user can open an R session by way of Rstudio and install both packages as follows.
 
@@
 
-R packages once installed need to be loaded with every R session. The user can load both with the `library()` function in R.
+R packages once installed need to be loaded with every R session (i.e. every time the user opens RStudio). The user can load both with the `library()` function in R.
 
@@
 # A Tutorial on How to Use `{peacesciencer}`
+
+I encourage users who are using `{peacesciencer}` for the first time, especially those who are learning R for the first time because of their interest in this package, to approach the package with an idea of the kind of data they want to create for the sake of a project.
+The core functionality of `{peacesciencer}` begins with the creation of a data frame, after which can be populated with various indicators of interest.
+No matter, the suggested use of `{peacesciencer}` begins with the creation of a data frame.
+To start, assume a new user without much familiarity with the R programming language installed Rstudio, `{tidyverse}`, and `{peacesciencer}` with the idea of using `{peacesciencer}` to help them start a new research project that seeks to explain civil conflict onset across state-years.
+Toward that end, they have identified their unit of analysis is state-year and can be created with the `create_stateyears()`.
+To get started, I encourage entering the following command in the console in Rstudio.
+
+```{r, eval=F, echo=T}
+?create_stateyears()
+```
+
+This will open a documentation file for this function, which is itself quite verbose and informative for the user about what the function is doing.
+In this case, the documentation file will show there are three arguments in this function, each with built-in defaults.
+This function will allow the user to choose a state system for which they want state-years (`system`, which accepts either "cow" or "gw" for CoW state system data or Gleditsch-Ward state system data and defaults to CoW), whether they want to extend the state system to the most recently concluded calendar year (`mry`, which defaults to TRUE), and whether they may want to additionally subset the years to a more narrow temporal domain they may already have in mind (`subset_years`, which defaults to no subset of the data, returning all possible state-years).
+If the user simply ran the function with no overrides, the function would return all Correlates of War state-years from 1816 to the most recent year (2020, as of writing).
+
+```{r}
+create_stateyears()
+```
+
+A user who approaches `{peacesciencer}` with a project in mind will see they can better tailor this function to what they want.
+Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the UCDP armed conflict data.
+They will also know they have no use for pre-1946 observations since civil conflict data, certainly UCDP's suite of data, has coverage only from 1946 forward.
+Thus, the user can supply some additional arguments to tailor the creation of data to just what they want (here: all Gleditsch-Ward state-years from 1946 to 2019).
+
+```{r}
+create_stateyears(system = 'gw', subset_years = c(1946:2019))
+```
+
+Another reader may be interested in using `{peacesciencer}` for a research project using a leader-year unit of analysis, which can be created with `create_leaderyears()`.
+Users can read more about what this function is doing by consulting the documentation file in R with the following command.
+
+```{r, eval=F, echo=T}
+?create_leaderyears()
+```
+
+This function has three arguments denoting the leader system to inform the creation of the leader-year data (i.e.
+Archigos) and two other arguments: `standardize` and `subset_years`.
+The `standardize` arguments, which when it defaults to "none", returns all leader-years as presented in the raw Archigos data.
+Archigos' leader data are nominally denominated in the Gleditsch-Ward state system data, if not necessarily Gleditsch-Ward state system dates (e.g.
+Archigos often has leader entries prior to state system entry).
+Thus, the user can standardize leader-year data to Gleditsch-Ward state system dates of CoW state system dates.
+Finally, the user can subset the leader-year data to a more narrow temporal domain that may interest them with the `subset_years` function.
+If a user is interested in creating a base data of leader-years for an analysis of inter-state conflict initiation with the GML MID data, they can create the data that interests them with the following function.
+Notice how `create_leaderyears()` also returns information about the leader too, like the leader's approximate age that year, their gender, and information about their tenure.
+
+```{r}
+create_leaderyears(standardize = "cow", subset_years = c(1870:2010))
+```
 
@@
 The following example creates a full state-year data frame of Gleditsch-Ward states from 1816 to 2020.
-
-```{r, echo=T, eval=F}
-create_stateyears(system = 'gw')
-```
 
@@
 
-```{r tab-dd, eval=T}
+```{r tab-dd, eval=T, echo=F}
 
@@
 
-```{r tab-cw, eval=T}
+```{r tab-cw, eval=T, echo=F}
 

### 2021-11-09 user-guide/ms.Rmd
upload/update, ms will be a work in progress.

@@
 The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages.
-Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year or state-year analysis (e.g.
+Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year, leader-year, leader-dyad-year, or state-year analysis (e.g.
 democracy data, contiguity data).
@@
 This data feature tackles a recurring problem for researchers in the peace science community.
-True research reproducibility is achieved creating original dyad-year or state-year data from scratch, though no published guide exists that informs researchers how to do this on their own.
+True research reproducibility is best achieved creating data from scratch, though no published guide exists that informs researchers how to do this on their own.
 Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice.
-Researchers may additionally spend too much time reproducing old code for standard information that goes into any dyad-year or state-year analysis---like contiguity relationships and democracy---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes.
+Researchers may additionally spend too much time reproducing old code for standard information that goes into any dyadic or monadic analysis---like contiguity relationships and democracy---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes.
 This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well.
@@
 Built around the free and open source R programming language, `{peacesciencer}` contains a suite of data and functions for creating data of interest to researchers.
-Researchers can use `{peacesciencer}` to create dyad-year, state-year, and even state-day data from scratch.
+Researchers can use `{peacesciencer}` to create dyad-year, leader-year, leader-dyad-year, and state-year data (among some others) from scratch.
 Afterward, they can add a variety of standard information (e.g.
@@
 This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want.
-`{peacesciencer}` comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts.
+`{peacesciencer}` comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing conflicts.
 All are done with the maximum possible transparency.
@@
 No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question.
-After all, EUGene's primary impetus was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g.
+After all, EUGene's main value---if not its primary impetus---was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g.
 dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit.
-`{peacesciencer}` is primarily born from this question about how to create dyad-year and state-year data from scratch.
+`{peacesciencer}` is primarily born from this question about how to create these data from scratch.
 The underlying code that produces these data types is available online and `{peacesciencer}` converts these lines of code into simple functions for the ease of the researcher.
@@
 Second, researchers also invest too much time in retracing steps for peace science analyses for new projects.
-Assume  a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data).
+Assume  a researcher finished a state-year analysis on the correlates of civil conflict onset a few years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data).
 Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project.
@@
 
-Fifth, statistical software should be kind to graduate students in peace science with the aforementioned points in mind.
+<!-- Fifth, statistical software should be kind to graduate students in peace science with the aforementioned points in mind.
 Graduate students just beginning their career in peace science scholarship are tasked to learn the substantive material, the characteristics of the underlying data in wide use in the community, and the quantitative methods and statistical software available to them, effectively all at the same time.
@@
 Graduate students can use `{peacesciencer}` to learn about peace science data and computer programming in R at the same time.
+-->
 
@@
 
-`{peacesciencer}` has a simple "pipe"-based workflow.
-The top of the "pipe" is the first function to create the data.
-These options include `create_dyadyears()`, `create_stateyears()`, and `create_statedays()`.
-All options include two arguments: `system` and `mry`.
-`system` takes one of two values---"cow" or "gw"---for what state system membership data the user wants.
-If `system` is not specified, the function defaults to Correlates of War system membership data given the prominence of the Correlates of War project in the conflict data ecosystem.
-`mry`---short for "most recent year"---determines whether or not the function extends the membership data to the most recently concluded calendar year.
-The function defaults to `TRUE` in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to (as of writing) 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017).
-`create_dyadyears()` has a third argument---`directed`---that determines whether the user wants directed dyad-year data or non-directed dyad-year data.
-The function defaults to `TRUE`, returning directed dyad-year data.
+Researchers using `{peacesciencer}` to create data for their research project should start with one of these "create" functions.
+Whether state-year, dyad-year, or a leader-level analysis, these functions will create the full universe of cases of interest to a researcher for the unit of analysis that serves as the core of their project.
 
-The following example creates a full state-year data frame of Gleditsch-Ward states from 1816 to 2020.
+## Adding Information to Data Created in `{peacesciencer}`
 
-The below example would create a full non-directed dyad-year data frame of Correlates of War states from 1816 to 2016, the last year of record in the Correlates of War state system membership data.
+After creating the base data of interest to their project, researchers can begin to add information they want with the suite of functions outlined in Table \@ref(tab:tab-addfunctions).
+For example, researcher interested in a dyad-year analysis of inter-state disputes can create a non-directed dyad-year data set from 1816 to 2010 in `{peacesciencer}` with the `create_dyadyears()` function, one of the aforementioned "create" function.
+The following would create the base data of interest to the user (i.e.
+all non-directed dyad-years from 1816 to 2010).
 
-```{r, echo=T, eval=F}
-create_dyadyears(directed = FALSE, mry = FALSE)
+```{r}
+create_dyadyears(directed = FALSE, subset_years = c(1816:2010))
 ```
 
-This example would create a full state-day data frame from the Gleditsch-Ward system for all states from Jan. 1, 1816 to Dec. 31, 2020. 
-```{r, echo=T, eval=F}
-create_statedays(system = 'gw')
+Adding information to this data frame is a simple matter of joining a series of functions together in a "pipe." The "pipe"---represented as `%>%` in the code below---is an operator built into `{tidyverse}` that allows (`R`) users to pass forward expressions or functions.
+These pipes are common in the programming world and, as the code below will show, have the benefit of changing code in a way that is more intuitive and easier to both read and write for the user.
+While these functions can be modified to work without `{tidyverse}` installed and loaded into the session, the user should find their experience with `{peacesciencer}` improved by this important package.
+
+For example, assume the researcher wants just all politically relevant, non-directed dyad-years, where political relevance is traditionally understood as a dyadic relationship involving some form of a contiguity relationship or a major power [@lemkereed2001rprd].
+`create_dyadyears(directed = FALSE, subset_years = c(1816:2010))` created the full universe of non-directed dyad-years from 1816 to 2010, though this full universe includes "irrelevant" dyads like Nigeria-Mongolia and Canada-Estonia.
+Reducing the data to just politically relevand dyads is simple in `{peacesciencer}` and `{tidyverse}`.
+Users first create the data they want (here: `create_dyadyears(directed = FALSE, subset_years = c(1816:2010))`), follow it with the pipe operator (`%>%`), and then add another function from `{peacesciencer}` (here: `filter_prd()`, which also quietly executes `add_contiguity()` and `add_cow_majors()`).
+
+```{r}
+# create base data, and pipe (%>%) to next function
+create_dyadyears(directed = FALSE, subset_years = c(1816:2010)) %>%
+  # subset data to politically relevant dyads (PRDs)
+  filter_prd()
 ```
 
-There are not too many data sets in the peace science community with daily data, certainly dating to the 19th century, but `{peacesciencer}` can greatly reduce the costs of creating the kind of data the user may want.
+Here, the user has created all non-directed dyad-years from 1816 to 2010 and then subset the data to just those with a major power or with some kind of contiguity relationship.
 
-## Add Standard Information to Dyad-Year or State-Year Data
+Users will find that the ease of the "pipe" will allow them greater agency in creating the full data set they may want for an analysis.
+Indeed, the pipe has the effect of forming something analogous to a drop-down menu, in which the user can "select" additional data they may want simply by specifying the function in `{peacesciencer}` that creates the data for them.
+For example, a researcher can follow `filter_prd()` with another pipe and communicate they want information about ongoing conflicts and conflict onsets from the Gibler-Miller-Little (GML) dispute data set [@gibleretal2016amid].
+Following `filter_prd()` with `add_gml_mids(keep = NULL)` will add information about ongoing conflicts and onsets in a given dyad-year.[^keepnull]
 
-The next step in the `{peacesciencer}` "pipe"-based workflow will also be among the biggest time-savers for researchers using it.
-Researchers who want to understand, say, the correlates of inter-state conflict onset almost always do the following.
-They will add the dependent variables from an inter-state conflict data set, like the Gibler-Miller-Little (GML) dispute data set [@gibleretal2016amid].
-They will add several indicators to the dyad-year data that are well-traveled in conflict scholarship, including indicators of things like joint alliance contracts and the contiguity relationship of both states in the dyad.
-`{peacesciencer}` has great coverage for a lot of these data and can even allow almost full-scale replications with the data it offers.
+[^keepnull]: `keep = NULL` in this context will focus the information returned to just the information about ongoing conflicts and onsets, discarding potentially unwanted columns about things like Side A, hostility levels, and other information included in the data set.
+Type `?add_gml_mids()` for more information.
 
-For example, assume a researcher wanted to replicate something analogous to Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis about the causes of conflict onset.
-`{peacesciencer}` can do something like this in the following pipe-based workflow.
-First, `create_dyadyears(directed = FALSE, mry = FALSE)` will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016.[^after2010]
-
-[^after2010]: The conflict data in @gibleretal2016amid do not extend past 2010, but asking for dyad-years from 2010-2016 will just create harmless `NA`s that will not appear in any analysis.
-
-After creating the base data they want, the user can add data to it with the "pipe" operator.
-The `{tidyverse}` represents this pipe operator as `%>%`.
-Ending each line with this pipe passes forward an object (here: data) into the next function on the next line.
-Put in other words, `create_dyadyears(directed = FALSE, mry = FALSE) %>% filter_prd()` is a command in `{tidyverse}` and `{peacesciencer}` to create all non-directed dyad-years from 1816 to 2016 and then subset those data to only the politically relevant dyad-years.
-Because politically relevant dyads depend on information about the major power status for both states as well as their contiguity relationship [@lemkereed2001rprd], the function adds in those data (i.e.
-quietly executing `add_contiguity()` and `add_cow_majors()`) before subsetting the data to just the dyads that have some kind of contiguity relationship and/or have a major power in it.
-
-```{r, echo=T, eval=F}
-create_dyadyears(directed = FALSE, mry = FALSE) %>%
+```{r}
+# create base data, and pipe (%>%) to next function
+create_dyadyears(directed = FALSE, subset_years = c(1816:2010)) %>%
+  # subset data to politically relevant dyads (PRDs), pipe to next function
   filter_prd() %>%
-  add_gml_mids(keep = NULL) %>%
-  add_peace_years() %>%
-  add_nmc() %>%
-  add_democracy() %>%
-  add_cow_alliance() %>%
-  add_sdp_gdp() -> Data
+  # add conflict information from GML-MID data,
+  add_gml_mids(keep = NULL) 
 ```
 
-Next, the pipe-based workflow continues with `add_gml_mids(keep = NULL)`, which will add information about whether there was an ongoing dispute in the dyad (and whether it was a dispute onset) from the GML dispute data.[^keepnull] `add_peace_years()` will calculate peace years between ongoing disputes in these politically relevant non-directed dyads.[^onpy] `add_nmc()` will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity.
-`add_democracy()` will add information about the level of democracy for both states in the dyad using three different indicators.
-Researchers can select which one they prefer from these indicators.
-`add_cow_alliance()` will add information about whether there was an alliance in the dyad-year by way of whether there was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge.
-Finally, `add_sdp_gdp()` will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by @andersetal2020bbgb.
-
-[^keepnull]: `keep = NULL` is optional and removes supporting information about dyadic dispute-participants (e.g.
-dispute-level hostility, dispute-level outcome) in order to reduce the size of the data set.
+Users can also calculate peace-years for these conflicts with `add_peace_years()` by using the pipe to pass forward the data set and applying the `add_peace_years()` function to it.[^onpy]
 
@@
 
+```{r}
+# create base data, and pipe (%>%) to next function
+create_dyadyears(directed = FALSE, subset_years = c(1816:2010)) %>%
+  # subset data to politically relevant dyads (PRDs), pipe to next function
+  filter_prd() %>%
+  # add conflict information from GML-MID data, pipe to next function
+  add_gml_mids(keep = NULL) %>%
+  # add peace years
+  add_peace_years()
+```
+
+Researchers should see that `{peacesciencer}`'s functionality can scale up nicely from there.
+For example, the following would round out the kind of information necessary to replicate Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis by adding information about national material capabilities for both states in the dyad (`add_nmc()`), estimates of democracy for both states in the dyad (`add_democracy()`), information about alliance commitments in the dyad-year (`add_cow_alliance()`), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by @andersetal2020bbgb.
 Whereas `add_sdp_gdp()` is the last command in the pipe-based workflow, the `{peacesciencer}` call ends by assigning to an object called `Data`.
@@
 `->`).
+
+```{r}
+# create base data, and pipe (%>%) to next function
+create_dyadyears(directed = FALSE, subset_years = c(1816:2010)) %>%
+  # subset data to politically relevant dyads (PRDs), pipe to next function
+  filter_prd() %>%
+  # add conflict information from GML-MID data, pipe to next function
+  add_gml_mids(keep = NULL) %>%
+  # add peace years, pipe to next function
+  add_peace_years() %>%
+  # add capabilities data, pipe to next function
+  add_nmc() %>%
+  # add some estimates about democracy for each state, pipe to next function
+  add_democracy() %>%
+  # add information about alliance commitments in dyad-year
+  add_cow_alliance() %>%
+  # finish with information about population and GDP/SDP
+  # and then assign to object, called, minimally, 'Data'
+  add_sdp_gdp() -> Data
+
+Data
+```
+
 If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like `haven::write_dta(Data, "my-data.dta")` and import it into Stata when they are done.
+No matter, `{peacesciencer}` has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding `-9s` in these data to be `NA` in the National Material Capabilities data.
+In this particular application, it has already created the main data required for a replication of @bremer1992ddc.
+There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some "weak-link" indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the @marquez2016qme for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.
+Table \@ref(tab:tab-dd) is a formatted version of the results of a logistic regression model of conflict onset using these "dangerous dyads" indicators.
+Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but `{peacesciencer}` greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.
 
-`{peacesciencer}` has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding `-9s` in these data to be `NA` in the National Material Capabilities data.
-There is only some slight data work to create the desired indicators for a statistical model of conflict onset.
-Here, briefly, these functions lean on `{tidyverse}` "verbs" and base R terminology to do things like creating a dummy variable for whether the dyad was land-contiguous, whether there was a major power on either side of the non-directed dyad, creating some "weak-link" indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the @marquez2016qme for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.
-
-```{r, echo=T, eval=F}
+```{r, echo=F, eval=F}
 Data %>%
@@
 ```
-
-Table \@ref(tab:tab-dd) is a formatted version of the results of a logistic regression model of conflict onset using these "dangerous dyads" indicators.
-Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but `{peacesciencer}` greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.
 
@@
 
-`{peacesciencer}` is also capable of creating data for replications of standard civil conflict analyses.
+## A Basic Replication of Fearon and Laitin (2003) in `{peacesciencer}`
+
+`{peacesciencer}` is capable of generating data for replications of analyses at multiple levels beyond dyad-year, including leader-year, leader-dyad-year, and state-year.
 Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's [-@fearonlaitin2003eicw] well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years.
-The pipe-based workflow will start with `create_stateyears(system = 'gw')`, creating the full universe of Gleditsch-Ward state years from 1816 to 2020.
-Whereas the UCDP data included in `{peacesciencer}` are from 1946 to 2019 (v.
-20.1), we can subset ("filter") the data to just those years with `filter(year %in% c(1946, 2019))`.
+The pipe-based workflow will start with `create_stateyears(system = 'gw', subset_years = c(1946:2019))`, creating the full universe of Gleditsch-Ward state years and subsetting them to just 1946-2019 (because the UCDP data included in `{peacesciencer}` include just the observations in that time frame).
 Next, we can use the `add_ucdp_acd()` function to return information about ongoing UCDP conflicts and onsets for these states.
@@
 ```{r, echo=T, eval=F}
-create_stateyears(system = 'gw') %>%
+create_stateyears(system = 'gw', subset_years = c(1946:2019)) %>%
   filter(year %in% c(1946:2019)) %>%
@@
 ```{r, echo=T, eval=F}
-create_stateyears(system = 'gw') %>%
-  filter(year %in% c(1946:2019)) %>%
+create_stateyears(system = 'gw', subset_years = c(1946:2019)) %>%
   add_ucdp_acd(type="intrastate", only_wars = TRUE) %>%
@@
              stars = c('*' = .05, '+' = .1), gof_omit = "IC|F|Log.|R2$",
-             coef_map = c("l1_wbgdppc2011est	" = "GDP per Capita (Lagged)", 
+             coef_map = c("l1_wbgdppc2011est" = "GDP per Capita (Lagged)", 
                           "l1_wbpopest" = "Population Size (Lagged)", 
@@
 `{peacesciencer}` is already more than capable of creating the kind of data in high demand in peace science.
-It can create dyad-year and state-year data.
+It can create dyad-year, leader-year, leader-dyad-year, and state-year data (among others).
 It is also generalizable to the dispute data included in the package, allowing for merging into *dispute-year* data as well.
@@
 It also does this in a maximally transparent way that conforms well to the DA-RT initiative across all political science.
+This is not to say `{peacesciencer}` does everything, but `{peacesciencer}` can only evolve and expand on what it already does well.
+This package can only evolve to meet new analytical demands for the peace science community.
+Users are free to request new features as "issues" on the project's Github.
 
-This is not to say `{peacesciencer}` does everything, but `{peacesciencer}` can only evolve and expand on what it already does well.
+<!-- This is not to say `{peacesciencer}` does everything, but `{peacesciencer}` can only evolve and expand on what it already does well.
 For example, `{peacesciencer}` does not right now handle *k*-adic data [c.f.
@@
 Users are free to request new features as "issues" on the project's Github.
+-->
 

### 2021-11-09 R/create_stateyears.R

@@ -87,6 +87,12 @@ create_stateyears <- function(system = "cow", mry = TRUE, subset_years) {
   }
 
 
+  if (!missing(subset_years)) {
+    data <- subset(data, data$year %in% subset_years)
+  } else {
+    data <- data
+  }
+
 
   return(data)
 }

### 2021-11-09 user-guide/R/1-dangerous-dyads-prd.R

@@ -3,13 +3,21 @@ library(peacesciencer)
 library(stevemisc)
 
 
-create_dyadyears(directed = FALSE, mry = FALSE) %>%
+create_dyadyears(directed = FALSE, subset_years = c(1816:2010)) %>%
+  # subset data to politically relevant dyads (PRDs), pipe to next function
   filter_prd() %>%
+  # add conflict information from GML-MID data, pipe to next function
   add_gml_mids(keep = NULL) %>%
+  # add peace years, pipe to next function
   add_peace_years() %>%
+  # add capabilities data, pipe to next function
   add_nmc() %>%
+  # add some estimates about democracy for each state, pipe to next function
   add_democracy() %>%
+  # add information about alliance commitments in dyad-year
   add_cow_alliance() %>%
+  # finish with information about population and GDP/SDP
+  # and then assign to object, called, minimally, 'Data'
   add_sdp_gdp() -> Data
 
 

### 2021-11-10 user-guide/ms.Rmd
update

@@
     column_spec(1, monospace = TRUE) %>%
-    column_spec(3, width="30em")
+    column_spec(3, width="20em")
 } else {
@@
 
-```r
+```{r, echo=T, eval=FALSE}
 install.packages(c("tidyverse", "peacesciencer"))
@@
 
-```r
+```{r, echo=T, eval=FALSE}
 library(tidyverse)
@@
 
-I encourage users who are using `{peacesciencer}` for the first time, especially those who are learning R for the first time because of their interest in this package, to approach the package with an idea of the kind of data they want to create for the sake of a project.
+I encourage users who are using `{peacesciencer}` for the first time, especially those who are learning R for the first time because of their interest in this package, to approach `{peacesciencer}` with an idea of the kind of data they want to create for the sake of a project.
 The core functionality of `{peacesciencer}` begins with the creation of a data frame, after which can be populated with various indicators of interest.
@@
 
-## Adding Information to Data Created in `{peacesciencer}`
+## Creating Dyad-Year Data and Adding Dyad-Year Data in `{peacesciencer}`
 
@@
 
-Users can also calculate peace-years for these conflicts with `add_peace_years()` by using the pipe to pass forward the data set and applying the `add_peace_years()` function to it.[^onpy]
+Users can also calculate peace-years for these conflicts with `add_spells()` by using the pipe to pass forward the data set and applying the `add_spells()` function to it.[^onpy]
 
-[^onpy]: `add_peace_years()` will only calculate the peace years---titled, in this case, `gmlmidspell`---and will leave the temporal dependence adjustment to the taste of the researcher.
+[^onpy]: `add_spells()` will only calculate the peace years---titled, in this case, `gmlmidspell`---and will leave the temporal dependence adjustment to the taste of the researcher.
 Importantly, I do not recommend manually creating splines or square/cube terms because it creates more problems in adjusting for temporal dependence in model predictions.
@@
   add_gml_mids(keep = NULL) %>%
-  # add peace years
-  add_peace_years()
+  # add peace years ("spells")
+  add_spells()
 ```
@@
   add_gml_mids(keep = NULL) %>%
-  # add peace years, pipe to next function
-  add_peace_years() %>%
+  # add peace years ("spells"), pipe to next function
+  add_spells() %>%
   # add capabilities data, pipe to next function
@@
 
-## A Basic Replication of Fearon and Laitin (2003) in `{peacesciencer}`
+## Creating State-Year Data and Adding State-Year Data in `{peacesciencer}`
 
-`{peacesciencer}` is capable of generating data for replications of analyses at multiple levels beyond dyad-year, including leader-year, leader-dyad-year, and state-year.
+`{peacesciencer}` is capable of generating data for replications of analyses at multiple levels beyond dyad-year.
 Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's [-@fearonlaitin2003eicw] well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years.
@@
   add_ucdp_acd(type="intrastate", only_wars = FALSE) %>%
-  add_peace_years() %>%
+  add_spells() %>%
   add_democracy() %>%
@@
   add_ucdp_acd(type="intrastate", only_wars = TRUE) %>%
-  add_peace_years() %>%
+  add_spells() %>%
   rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .)) %>%
@@
   tab_modCW %>%
+  set_font_size(10) %>%
+  set_left_padding(1) %>%
+  set_top_padding(1) %>%
+  set_bottom_padding(1) %>%
+  set_right_padding(1) %>%
+  set_bold(., 1, 1:2, TRUE) %>%
+  as_flextable()
+} 
+```
+
+## Creating Leader-Year Data and Adding Leader-Year Data in `{peacesciencer}`
+
+`{peacesciencer}` also has support for newer levels of analysis in the peace science community, prominently leader-levels of analysis.
+There has been considerable emphasis in peace science research to emphasize that state leaders, not "states", make foreign policy decisions that may lead to war.
+Understanding the attributes of leaders themselves are critical to the core research questions of the community [@ellisetal2015lead; @goemansetal2009ia; @horowitzstam2014hpme] and `{peacesciencer}` wants to help researchers toward that end.
+
+Table \@ref(tab:tab-addfunctions) shows there are some dedicated functions for populating leader-level data with leader-specific information, in addition to adding state-year-level information (e.g.
+democracy, capabilities) to leader-level data.
+Table \@ref(tab:tab-createfunctions) shows support for creating leader-level data that are standardized to either the CoW or G-W state system data.
+Suppose a researcher wanted to create leader-year data, standardized to CoW system dates, for an analysis of inter-state dispute initiation analogous to @horowitzstam2014hpme.
+The user would first start with creating the base data (`create_leaderyears(standardize = "cow", subset_years=c(1875:2010))`), which also generates some leader attributes (e.g.
+estimated leader age, year in office, and leader gender).
+From there, they would add information about leader conflict behavior in the year with `add_gml_mids()` and calculate peace-years with the `add_spells()` function.[^oninitiation] `add_lead()` will add a battery of leader experience and attributes variables from data created by @ellisetal2015lead, including whether the leader had military experience, combat experience, was a rebel fighter, and many more.
+Finally, `add_nmc()` and `add_democracy()` will add state-year estimates of national capabilities and democracy for assisting in making state-to-state comparisons, even for leader-level analyses.
+
+```{r, eval=F}
+create_leaderyears(standardize = "cow", subset_years=c(1875:2010)) %>%
+  add_gml_mids() %>%
+  add_spells() %>%
+  add_lead() %>%
+  mutate(milservice_nocombat = ifelse(milservice == 1 & combat == 0, 1, 0)) %>%
+  add_nmc() %>%
+  add_democracy() -> Data
+
+```
+
+[^oninitiation]: The documentation for `add_gml_mids()`, along with the `data-raw` directory on Github, provides ample explanation for how inter-state conflicts in the GML conflict data are connected to specific leaders.
+Since inter-state dispute data can have missing days, connecting leaders to dispute onsets involves some important archival research that is only possible because of the information collected by @gibleretal2016amid over 10 years of scholarship.
+The documentation also considers the multiple ways in which users can more flexibly code the concept of "initiation" based on participant-level summaries.
+The default coding of initiation, based on this exhaustive archival work, is to code a leader initiating a dispute if they were on Side A or joined the conflict at any point after the onset of the dispute.
+Users can change this to focus on just Side A, or just originators on Side A.
+This is the `init` argument in `add_gml_mids()`.
+
+```{r tab-ly, eval=T, echo=F}
+
+table_format <- ifelse(is_docx, "huxtable", 'default')
+
+modelsummary(modLY, output=table_format,
+             title = 'A Leader-Year Analysis of Inter-State Dispute Initiation in \\{peacesciencer\\}',
+             stars = c('*' = .05, '+' = .1), gof_omit = "IC|F|Log.|R2$",
+             coef_map = c("genderM" = "Male Leader", 
+                          "leaderage" = "Leader Age", 
+                          "yrinoffice" = "Leader Tenure (in Years)",
+                          "milservice_nocombat" = "Military Service (No Combat)", 
+                          "combat" = "Military Serivce (with Combat)", 
+                          "rebel" = "Previous Rebel Experience", 
+                          "warwin" = "Prior War Win",
+                          "warloss" = "Prior War Loss",
+                          "rebelwin" = "Prior Rebel Win",
+                          "rebelloss" = "Prior Rebel Loss",
+                          "xm_qudsest" = "Extended UDS Democracy Score",
+                          "cinc" = "CINC",
+                          "gmlmidinitspell" = "t",
+                          "I(gmlmidinitspell^2)" = "t^2",
+                          "I(gmlmidinitspell^3)" = "t^3",
+                          "(Intercept)" = "Intercept"),
+             align=c("lc")) -> tab_modLY
+
+if (!is_docx) {
+  tab_modLY %>%
+  row_spec(0, bold=TRUE) %>%
+  column_spec(1, width = "8cm")
+} else {
+  tab_modLY %>%
   set_font_size(10) %>%

### 2021-11-10 R/add_creg_fractionalization.R

@@ -1,36 +1,50 @@
 #' Add fractionalization/polarization estimates from CREG to a data frame
 #'
-#' @description \code{add_creg_fractionalization()} allows you to add information
-#' about the fractionalization/polarization of a state's ethnic and religious groups to your dyad-year
-#' or state-year data.
+#' @description
 #'
-#' @return \code{add_creg_fractionalization()} takes a dyad-year, leader-year, leader-dyad-year, or state-data frame,
-#' whether the primary state identifiers are from the Correlates of War system or the Gleditsch-Ward system, and
-#' returns information about the
-#' fractionalization and polarization of the state(s) in a given year. The function returns four additional
-#' columns when the data are state-year and returns eight additional columns when the data are state-year (or leader-year).
-#' The columns returned are the fractionalization of ethnic groups, the polarization of ethnic groups, the
-#' fractionalization of religious groups, and the polarization of religious groups. When the data are dyad-year (or leader-dyad-year),
-#' the return doubles because it provides information for both states in the dyad.
+#' \code{add_creg_fractionalization()} allows you to add information about the
+#' fractionalization/polarization of a state's ethnic and religious groups to
+#' your data.
 #'
-#' @details Please see the information for the underlying data \code{creg}, and the
-#' associated R script in the \code{data-raw} directory, to see how these data are generated.
+#' @return \code{add_creg_fractionalization()} takes a dyad-year, leader-year,
+#' leader-dyad-year, or state-data frame, whether the primary state
+#' identifiers are from the Correlates of War system or the Gleditsch-Ward
+#' system, and  returns information about the fractionalization and
+#' polarization of the state(s) in a given year. The function returns four
+#' additional columns when the data are state-year and returns eight
+#' additional columns when the data are state-year (or leader-year).
+#' The columns returned are the fractionalization of ethnic groups, the
+#' polarization of ethnic groups, the fractionalization of religious groups,
+#' and the polarization of religious groups. When the data are dyad-year
+#' (or leader-dyad-year), the return doubles because it provides information
+#' for both states in the dyad.
 #'
-#' The \code{creg} data have a few duplicates. When standardizing to true CoW codes, the duplicates
-#' concern Serbia/Yugoslavia in 1991 and 1992 as well as Russia/the Soviet Union in 1991. When standardizing
[…]

### 2021-11-10 user-guide/ms.Rmd
update

@@
 The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages.
-Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year, leader-year, leader-dyad-year, or state-year analysis (e.g.
+Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate common types of data from scratch and to merge in ubiquitous indicators included in almost every analysis (e.g.
 democracy data, contiguity data).
@@
 Thereafter, it provides a tutorial on how to install and best use `{peacesciencer}` in the R programming language.
-A more comprehensive tutorial follows, showing how `{peacesciencer}` already has a suite of data and functions that can allow for effective replications of a "dangerous dyads" type analysis [@bremer1992ddc] and standard state-year analyses of civil conflict onset [e.g.
-@fearonlaitin2003eicw].
-This feature concludes with a discussion of what else `{peacesciencer}` is capable of doing in future updates and how this R package can inform more reasoned design decisions for researchers in peace science.
+A more comprehensive tutorial follows, showing how `{peacesciencer}` already has a suite of data and functions that can allow for effective replications of a "dangerous dyads" type analysis [@bremer1992ddc], standard state-year analyses of civil conflict onset [e.g.
+@fearonlaitin2003eicw], and even leader-year analyses of inter-state dispute initiation [@horowitzstam2014hpme].
+This feature concludes with a comparison of `{peacesciencer}` with other alternatives and a discussion of how `{peacesciencer}` can inform more reasoned design decisions for researchers in peace science.
 
@@
 
-`{peacesciencer}` is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch.
-NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars [@bennettetal2019n].
-NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade [@bennettstam2000e].
-No matter, `{peacesciencer}` is motivated by the following ideals that led to its creation.
-
+`{peacesciencer}` is motivated by the following ideals that led to its creation.
 For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis.
@@
 This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that `{peacesciencer}` uses.
-There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet [@arelbundock2021w], scrape *any* type of data from the internet [@boehmke2016dwr], and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero [@arelbundock2021m].
+There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet [@arelbundock2021w] and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero [@arelbundock2021m].
 `{peacesciencer}` embraces this.
@@
 These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis.
-Table \@ref(tab:tab-createfunctions) and Table \@ref(tab:tab-addfunctions) list these core functions as of version 0.7, the version of this package slated for release alongside the manuscript.
+Table \@ref(tab:tab-createfunctions) and Table \@ref(tab:tab-addfunctions) list these core functions as of version 0.7, the version of this package slated for release alongside the manuscript.[^helperfunctions]
 
-Table \@ref(tab:tab-createfunctions) are functions that crease base data frames for a researcher starting an original project.
+[^helperfunctions]: `{peacesciencer}` also has miscellaneous "helper" functions that are not belabored in these manuscript.
+These include functions that allow the user to use `{peacesciencer}` functions with data created outside the package (`declare_attributes()`), download external data for quicker use of other functions (`download_extdata()`), and functions that communicate suggested citations (`ps_cite()`), among a few other functions.
+See the package's website for more information (`r params$ps_website`).
+
+Table \@ref(tab:tab-createfunctions) are functions that create base data frames for a researcher starting an original project.
 They serve as functions that communicate the units of analysis supported in this package and that the package is capable of generating for an interested user.
@@
 ```{r, echo=T, eval=FALSE}
+# Install the packages for use
 install.packages(c("tidyverse", "peacesciencer"))
@@
 ```{r, echo=T, eval=FALSE}
+# Load the packages for use
 library(tidyverse)
@@
 ```{r, eval=F, echo=T}
+# See documentation for create_stateyears() function.
 ?create_stateyears()
@@
 In this case, the documentation file will show there are three arguments in this function, each with built-in defaults.
-This function will allow the user to choose a state system for which they want state-years (`system`, which accepts either "cow" or "gw" for CoW state system data or Gleditsch-Ward state system data and defaults to CoW), whether they want to extend the state system to the most recently concluded calendar year (`mry`, which defaults to TRUE), and whether they may want to additionally subset the years to a more narrow temporal domain they may already have in mind (`subset_years`, which defaults to no subset of the data, returning all possible state-years).
+This function will allow the user to choose a state system for which they want state-years (`system`, which accepts either "cow" or "gw" for Correlates of War [CoW] state system data or Gleditsch-Ward [G-W] state system data and defaults to CoW), whether they want to extend the state system to the most recently concluded calendar year (`mry`, which defaults to TRUE), and whether they may want to additionally subset the years to a more narrow temporal domain they may already have in mind (`subset_years`, which defaults to no subset of the data, returning all possible state-years).
 If the user simply ran the function with no overrides, the function would return all Correlates of War state-years from 1816 to the most recent year (2020, as of writing).
@@
 ```{r}
+# Create all CoW state-years from 1816 to most recently concluded calendar year.
 create_stateyears()
@@
 ```{r}
+# Create all G-W state years from 1946 to 2019.
 create_stateyears(system = 'gw', subset_years = c(1946:2019))
@@
 ```{r, eval=F, echo=T}
+# See documentation for create_leaderyears() function.
 ?create_leaderyears()
@@
 ```{r}
+# Create all leader-years for CoW states from 1870 to 2010.
 create_leaderyears(standardize = "cow", subset_years = c(1870:2010))
@@
 
-## Creating Dyad-Year Data and Adding Dyad-Year Data in `{peacesciencer}`
+## Creating Dyad-Year Data and Adding to Dyad-Year Data in `{peacesciencer}`
 
@@
 ```{r}
+# Create all non-directed dyad-years from 1816 to 2010.
 create_dyadyears(directed = FALSE, subset_years = c(1816:2010))
@@
 
-## Creating State-Year Data and Adding State-Year Data in `{peacesciencer}`
+## Creating State-Year Data and Adding to State-Year Data in `{peacesciencer}`
 
@@
 ```{r, echo=T, eval=F}
+# Create a set of state-year data for a civil conflict analysis.
 create_stateyears(system = 'gw', subset_years = c(1946:2019)) %>%
@@
 ```{r, echo=T, eval=F}
+# Repeat the process, but for a new DV of just wars
 create_stateyears(system = 'gw', subset_years = c(1946:2019)) %>%
@@
 
-## Creating Leader-Year Data and Adding Leader-Year Data in `{peacesciencer}`
+## Creating Leader-Year Data and Adding to Leader-Year Data in `{peacesciencer}`
 
@@
 Finally, `add_nmc()` and `add_democracy()` will add state-year estimates of national capabilities and democracy for assisting in making state-to-state comparisons, even for leader-level analyses.
+Some light recoding of the data created in `{peacesciencer}` and some regression modeling reproduces the results presented in Table \@ref(tab:tab-ly), itself an approximation of the kind of leader-year analysis exemplified in @horowitzstam2014hpme.
 
 ```{r, eval=F}
+# Create a set of leader-year data for an analysis of conflict initiation.
 create_leaderyears(standardize = "cow", subset_years=c(1875:2010)) %>%
@@
   add_lead() %>%
-  mutate(milservice_nocombat = ifelse(milservice == 1 & combat == 0, 1, 0)) %>%
   add_nmc() %>%
@@
 
+# A Comparison with Other Alternative Approaches
+
+`{peacesciencer}` is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch.
+Alternatives exist, some more inaccessible than others.
+NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars [@bennettetal2019n].
+NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade [@bennettstam2000e].
+Finally, peace scientists well-versed in Structured Query Language (SQL) could use one of several "join" transformations---e.g.
+"cross join", "inner join", "left join"---to create dyad-year data from state system data, even if this might amount to a detour in the peace scientist's research agenda for this particular task.[^joins] A comparison of `{peacesciencer}` with EUGene and NewGene suggests the following benefits of the package while emphasizing areas where other options may have some advantages.
+
+[^joins]: That said, `{peacesciencer}` makes ample use of SQL-like "joins" for the construction of base data and functions that merge data into base data created in the package.
+A discussion of these joins appears as a vignette on the package's website.
+
+EUGene is the clear inspiration for this package.
+Though its original impetus was the generation of expected utility data for evaluating @bdmlalman1992wr (i.e.
+the "EU" in "EUGene"), the software became quite popular for peace science scholars in the early 2000s for helping them start new projects from scratch with important data already provided.
+`{peacesciencer}` covers all the same units of analysis that EUGene covers, as of version 3.2.[^v3212] EUGene has more explicit support for dyadic dispute-year data whereas `{peacesciencer}` treats dyadic dispute-year data as a derivation of dyad-year data, with respect to functions that populate base data with additional information.[^whittledeclare] `{peacesciencer}` may reflect more current research interests in the peace science community.
+This is why there are leader-level data and support for the G-W (and UCDP) ecosystem of data that EUGene does not have, but there is no function yet for things like calculating Tau-b or expected utility values.
+Importantly, though, `{peacesciencer}` mimics the verbosity of EUGene's user manual.
+EUGene's user manual was amply informative about what it was calculating and why it was doing what it was doing for some defaults.
+Likewise, `{peacesciencer}`'s documentation strives to be as informative as possible as to what it is doing and why it is doing what it is doing for some defaults.
+A scholar who remembers EUGene well will ideally think of `{peacesciencer}` as the most faithful approximation of what that software did for the community at the time.
+It has the added benefit of being agnostic to the researcher's operating system, having greater flexibility of data types supported, and  better reflecting more current frontiers in the community.
+It does have the drawback of requiring at least some level of comfort with the R programming language.
+
+[^whittledeclare]: For example, the "whittle" class of functions (i.e.
+`wc_` in Table \@ref(tab:tab-addfunctions)) are designed for dyadic dispute-year data while a helper function included in the package---`declare_attributes()`---can allow a user to treat dyadic dispute-year data as dyad-year data for the sake of using any of the other supported functions in Table \@ref(tab:tab-addfunctions).
+
+[^v3212]: The latest version, as of writing, is v.
+3.212 released in 2017.
+However, this is apparently a bug fix on string variables and that the underlying functionality was still last updated in 2007 with the release of version 3.2.
+
+NewGene is the latest evolution of EUGene, at least as a stand-alone executable program for creating the kind of data of interest to the peace science community.
+NewGene's greatest strength, relative to EUGene and even `{peacesciencer}`, is its support for *k*-adic data [c.f.
+@poast2010mdd].
+*k*-adic levels of analysis are not yet supported in `{peacesciencer}` and users interested in generating *k*-adic data should consider downloading NewGene.
+No matter, `{peacesciencer}` has several superlatives in relation to NewGene.
+It supports leader-year and leader-dyad-year data where NewGene does not.
+It offers support for the G-W (and UCDP) ecosystem of data whereas NewGene does not and is mostly aimed for researchers interested in inter-state conflict.
+NewGene deviates a bit from EUGene by only indirectly asking users what kind of data they want (e.g.
+state-year, dyad-year) and without providing too much detail about what it is doing and why it is doing what it is doing.
+For example, NewGene only indirectly states the unit of analysis of the data to be generated near the top of its interface by asking the user how many country columns they want.
+This is an indirect way of the user to get data that are state-year, dyad-year, triad-year (etc.), which are then expanded and populated with data at various levels.
+`{peacesciencer}`, much like EUGene, is more explicit, encouraging the user to be upfront about what their unit of analysis is and what are the units of analysis that can be plausibly plugged into the data the user is creating.
+`{peacesciencer}`, again, does expect at least some level of comfort with the R programming language, but even this comes with greater ease of interpreting what is the unit of analysis and what are the primary spatial and temporal units that serve as the basis of the data.
+
 # Conclusion
@@
 It is also generalizable to the dispute data included in the package, allowing for merging into *dispute-year* data as well.
-This feature showed how it can effectively approximate two types of analyses in wide use in the peace science community.
+This feature showed how it can effectively approximate three types of analyses in wide use in the peace science community.
 Surely researchers can and will add more information to these simple analyses after using `{peacesciencer}`, but the package already does a lot of the tedious work for researchers.

### 2021-11-11 user-guide/ms.Rmd
another day, another update

@@
 create_stateyears(system = 'gw', subset_years = c(1946:2019)) %>%
-  filter(year %in% c(1946:2019)) %>%
   add_ucdp_acd(type="intrastate", only_wars = FALSE) %>%
@@
 
-# A Comparison with Other Alternative Approaches
+# A Comparison with Other Approaches
 
@@
 NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade [@bennettstam2000e].
-Finally, peace scientists well-versed in Structured Query Language (SQL) could use one of several "join" transformations---e.g.
-"cross join", "inner join", "left join"---to create dyad-year data from state system data, even if this might amount to a detour in the peace scientist's research agenda for this particular task.[^joins] A comparison of `{peacesciencer}` with EUGene and NewGene suggests the following benefits of the package while emphasizing areas where other options may have some advantages.
+Finally, peace scientists well-versed in Structured Query Language (SQL) could use one of several "join" transformations to create dyad-year data from state system data, even if this might amount to a detour in the peace scientist's research agenda for this particular task.[^joins] A comparison of `{peacesciencer}` with EUGene and NewGene suggests the following benefits of the package while emphasizing areas where other options may have some advantages.
 

### 2021-11-11 user-guide/R/1-create-gw_statedays.R

@@ -1,6 +1,0 @@
-library(tidyverse)
-library(peacesciencer)
-
-create_statedays(system = 'gw') -> gw_statedays
-
-saveRDS(gw_statedays, "data/gw_statedays.rds")

### 2021-11-11 user-guide/R/2-analysis.R

@@ -1,8 +1,0 @@
-Data <- readRDS("user-guide/data/Data.rds")
-
-Mods <- list()
-
-Mods$"Default" <- lm(immigsent ~ agea + female + eduyrs + uempla + hinctnta + lrscale, data=Data)
-Mods$"Default + Noise" <- lm(immigsent ~ agea + female + eduyrs + uempla + hinctnta + lrscale + noise, data=Data)
-
-saveRDS(Mods, "user-guide/data/Mods.rds")

### 2021-11-11 user-guide/ms.Rmd
update site

@@
 This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well.
-Graduate students, in particular, face unique challenges associated with these developments.
-Graduate students beginning their careers in peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.
+Graduate students and other beginners in the field face unique challenges associated with these developments.
+Students just learning peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.
 
@@
 All are done with the maximum possible transparency.
-The project is available for public view of Github (`r params$ps_github`).
-The `data-raw` directory on the project's Github contains information and comments about how every data set was created and the `R` directory informs the user how each function works.
+The project is available for public view on Github (`r params$ps_github`).
+The `data-raw` directory on the project's Github contains information and comments about how every data set was created.
 The function manuals (`r params$ps_manual`) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing.
@@
 This data feature proceeds in the following fashion.
-The next section expands what need this package fills for peace scientists in the present.
+The next section expands what need this package fills for peace scientists.
 Afterward, it provides an overview of what is included in `{peacesciencer}` to help researchers more quickly conduct the kind of quantitative research they want.
@@
 A more comprehensive tutorial follows, showing how `{peacesciencer}` already has a suite of data and functions that can allow for effective replications of a "dangerous dyads" type analysis [@bremer1992ddc], standard state-year analyses of civil conflict onset [e.g.
-@fearonlaitin2003eicw], and even leader-year analyses of inter-state dispute initiation [@horowitzstam2014hpme].
+@fearonlaitin2003eicw], and even leader-year analyses of inter-state dispute initiation [e.g.
+@horowitzstam2014hpme].
 This feature concludes with a comparison of `{peacesciencer}` with other alternatives and a discussion of how `{peacesciencer}` can inform more reasoned design decisions for researchers in peace science.
@@
 
-`{peacesciencer}` is motivated by the following ideals that led to its creation.
+`{peacesciencer}` is motivated by the following observations and ideals that led to its creation.
 For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis.
@@
 Software options exist that allow the researcher to load data, create data, clean data, analyze data, and present the results of the analysis all within one program.
-`{peacesciencer}`, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and Rstudio (its free-for-use integrated development environment [IDE]).
+`{peacesciencer}`, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and RStudio (its free-for-use integrated development environment [IDE]).
 
@@
 As of version 0.7, `{peacesciencer}` is capable of creating full dyad-year data, leader-day data, leader-dyad-year data, leader-year data, state-day data, and state-year data.[^leaderdyadyear] A vignette on the package's website shows how users can create other forms of data from these functions as well (e.g.
-leader-months, state-quarters).
+dyadic-dispute-year, leader-months, state-quarters).
 
@@
 First, RStudio offers an IDE that serves as a user-friendly graphical user interface (GUI) over what is, at its core, a programming language with a command-line interface.
-Rstudio's design will make it much easier for users to experiment with `{peacesciencer}`'s functionality and read its documentation to assist them with the user of these functions.
+RStudio's design will make it much easier for users to experiment with `{peacesciencer}`'s functionality and read its documentation to assist them with the use of these functions.
 The second additional download is `{tidyverse}`, itself a suit of packages that share a common form and design [@wickhametal2019wt].
 `{peacesciencer}` functions make considerable use of the component packages of `{tidyverse}` but installing and loading `{tidyverse}` will allow the researcher to make quicker use of `{peacesciencer}`'s functionality.
-A user can open an R session by way of Rstudio and install both packages as follows.
+A user can open an R session by way of RStudio and install both packages as follows.
 
@@
 No matter, the suggested use of `{peacesciencer}` begins with the creation of a data frame.
-To start, assume a new user without much familiarity with the R programming language installed Rstudio, `{tidyverse}`, and `{peacesciencer}` with the idea of using `{peacesciencer}` to help them start a new research project that seeks to explain civil conflict onset across state-years.
+To start, assume a new user without much familiarity with the R programming language installed RStudio, `{tidyverse}`, and `{peacesciencer}` with the idea of using `{peacesciencer}` to help them start a new research project that seeks to explain civil conflict onset across state-years.
 Toward that end, they have identified their unit of analysis is state-year and can be created with the `create_stateyears()`.
-To get started, I encourage entering the following command in the console in Rstudio.
+To get started, I encourage entering the following command in the console in RStudio.
 

### 2021-11-11 user-guide/ms.Rmd
update

@@
 RStudio's design will make it much easier for users to experiment with `{peacesciencer}`'s functionality and read its documentation to assist them with the use of these functions.
-The second additional download is `{tidyverse}`, itself a suit of packages that share a common form and design [@wickhametal2019wt].
-`{peacesciencer}` functions make considerable use of the component packages of `{tidyverse}` but installing and loading `{tidyverse}` will allow the researcher to make quicker use of `{peacesciencer}`'s functionality.
-A user can open an R session by way of RStudio and install both packages as follows.
+The second additional download is `{tidyverse}`, itself a suite of packages that share a common form and design [@wickhametal2019wt].
+`{peacesciencer}` functions make considerable use of the component packages of `{tidyverse}`, and `{peacesciencer}` can work without it, but installing and loading `{tidyverse}` will allow the researcher to make quicker use of `{peacesciencer}`'s functionality.
+A user can open an R session by way of RStudio and install both packages as follows.[^poundsign]
+
+[^poundsign]: R users will recognize that pound signs (`#`) allow for commenting code.
+Users new to R should understand its use here is primarily for exposition of what the code snippets in this manuscript are doing.
 
@@
 In this case, the documentation file will show there are three arguments in this function, each with built-in defaults.
-This function will allow the user to choose a state system for which they want state-years (`system`, which accepts either "cow" or "gw" for Correlates of War [CoW] state system data or Gleditsch-Ward [G-W] state system data and defaults to CoW), whether they want to extend the state system to the most recently concluded calendar year (`mry`, which defaults to TRUE), and whether they may want to additionally subset the years to a more narrow temporal domain they may already have in mind (`subset_years`, which defaults to no subset of the data, returning all possible state-years).
+This function will allow the user to choose a state system for which they want state-years (`system`, which accepts either "cow" or "gw" for Correlates of War [CoW] state system data or Gleditsch-Ward [G-W] state system data and defaults to "cow"), whether they want to extend the state system to the most recently concluded calendar year (`mry`, which defaults to TRUE), and whether they may want to additionally subset the years to a more narrow temporal domain they may already have in mind (`subset_years`, which defaults to no subset of the data, returning all possible state-years).
 If the user simply ran the function with no overrides, the function would return all Correlates of War state-years from 1816 to the most recent year (2020, as of writing).
@@
 Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the UCDP armed conflict data.
-They will also know they have no use for pre-1946 observations since civil conflict data, certainly UCDP's suite of data, has coverage only from 1946 forward.
+They will also know they have no use for pre-1946 observations since civil conflict data, certainly UCDP's suite of data, have coverage only from 1946 forward.
 Thus, the user can supply some additional arguments to tailor the creation of data to just what they want (here: all Gleditsch-Ward state-years from 1946 to 2019).
@@
 Archigos' leader data are nominally denominated in the Gleditsch-Ward state system data, if not necessarily Gleditsch-Ward state system dates (e.g.
-Archigos often has leader entries prior to state system entry).
+Archigos often has leader entries prior to state system entry in a few cases).
 Thus, the user can standardize leader-year data to Gleditsch-Ward state system dates of CoW state system dates.
 Finally, the user can subset the leader-year data to a more narrow temporal domain that may interest them with the `subset_years` function.
-If a user is interested in creating a base data of leader-years for an analysis of inter-state conflict initiation with the GML MID data, they can create the data that interests them with the following function.
+If a user is interested in creating a data set of leader-years for an analysis of inter-state conflict initiation with the GML MID data, they can create the data that interests them with the following function.
 Notice how `create_leaderyears()` also returns information about the leader too, like the leader's approximate age that year, their gender, and information about their tenure.
@@
 These pipes are common in the programming world and, as the code below will show, have the benefit of changing code in a way that is more intuitive and easier to both read and write for the user.
-While these functions can be modified to work without `{tidyverse}` installed and loaded into the session, the user should find their experience with `{peacesciencer}` improved by this important package.
+While these functions can be modified to work without `{tidyverse}` installed and loaded into the session, the user will find their experience with `{peacesciencer}` is only improved by this important package.
 
@@
 `create_dyadyears(directed = FALSE, subset_years = c(1816:2010))` created the full universe of non-directed dyad-years from 1816 to 2010, though this full universe includes "irrelevant" dyads like Nigeria-Mongolia and Canada-Estonia.
-Reducing the data to just politically relevand dyads is simple in `{peacesciencer}` and `{tidyverse}`.
+Reducing the data to just politically relevant dyads is simple in `{peacesciencer}` and `{tidyverse}`.
 Users first create the data they want (here: `create_dyadyears(directed = FALSE, subset_years = c(1816:2010))`), follow it with the pipe operator (`%>%`), and then add another function from `{peacesciencer}` (here: `filter_prd()`, which also quietly executes `add_contiguity()` and `add_cow_majors()`).
@@
 Users will find that the ease of the "pipe" will allow them greater agency in creating the full data set they may want for an analysis.
-Indeed, the pipe has the effect of forming something analogous to a drop-down menu, in which the user can "select" additional data they may want simply by specifying the function in `{peacesciencer}` that creates the data for them.
+Indeed, the pipe has the effect of forming something analogous to a drop-down menu, in which the user can "select" additional data/commands they may want simply by specifying the function in `{peacesciencer}` that does what they want.
 For example, a researcher can follow `filter_prd()` with another pipe and communicate they want information about ongoing conflicts and conflict onsets from the Gibler-Miller-Little (GML) dispute data set [@gibleretal2016amid].
@@
 If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like `haven::write_dta(Data, "my-data.dta")` and import it into Stata when they are done.
-No matter, `{peacesciencer}` has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding `-9s` in these data to be `NA` in the National Material Capabilities data.
+No matter, `{peacesciencer}` has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding `-9s` in these data to be `NA` in the capabilities data.
 In this particular application, it has already created the main data required for a replication of @bremer1992ddc.
@@
 Options include "territory", "government", and "both".
-If no `issue` is specified, the function returns information for all conflicts irregarding the particular issue.
+If no `issue` is specified, the function returns information for all conflicts regardless of the particular issue.
 `only_wars` is an argument that subsets the data to just those with the intensity levels of "war" when `only_wars = TRUE`.
@@
 `left_join(Data, .) -> Data`).
-Table \@ref(tab:tab-cw) shows the fruits of the data `{peacesciencer}` generated after some post-processing for lagging important variables.
+Table \@ref(tab:tab-cw) shows the fruits of the data `{peacesciencer}` generated after some post-processing and lagging important variables.
 
@@
 Table \@ref(tab:tab-createfunctions) shows support for creating leader-level data that are standardized to either the CoW or G-W state system data.
-Suppose a researcher wanted to create leader-year data, standardized to CoW system dates, for an analysis of inter-state dispute initiation analogous to @horowitzstam2014hpme.
+Suppose a researcher wanted to create leader-year data, standardized to CoW system dates, for an analysis of inter-state dispute initiation analogous to what @horowitzstam2014hpme do in their analysis.
 The user would first start with creating the base data (`create_leaderyears(standardize = "cow", subset_years=c(1875:2010))`), which also generates some leader attributes (e.g.
@@
 [^whittledeclare]: For example, the "whittle" class of functions (i.e.
-`wc_` in Table \@ref(tab:tab-addfunctions)) are designed for dyadic dispute-year data while a helper function included in the package---`declare_attributes()`---can allow a user to treat dyadic dispute-year data as dyad-year data for the sake of using any of the other supported functions in Table \@ref(tab:tab-addfunctions).
+`wc_` in Table \@ref(tab:tab-addfunctions)) are designed for dyadic dispute-year data also included in the package while a helper function---`declare_attributes()`---can allow a user to treat other dyadic dispute-year data as dyad-year data for the sake of using any of the other supported functions in Table \@ref(tab:tab-addfunctions).
 
@@
 No matter, `{peacesciencer}` has several superlatives in relation to NewGene.
-It supports leader-year and leader-dyad-year data where NewGene does not.
+It supports leader-year and leader-dyad-year data while NewGene does not.
 It offers support for the G-W (and UCDP) ecosystem of data whereas NewGene does not and is mostly aimed for researchers interested in inter-state conflict.
@@
 `{peacesciencer}` does not endeavor to make researchers lazy or sloppy, and it does not ultimately do this.
-Instead, `{peacesciencer}` reduces the tedium associated with starting quantitative peace science research, achieving this in a quick, robust, and transparent way.
+Instead, `{peacesciencer}` encourages reasoned design decisions by the user up front and reduces the tedium associated with starting quantitative peace science research.
+It achieves this in a quick, robust, and transparent way.
 

### 2021-11-16 user-guide/ms.Rmd
update

@@
 The core functionality of `{peacesciencer}` reduces to two broad categories of functions.
-These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis.
-Table \@ref(tab:tab-createfunctions) and Table \@ref(tab:tab-addfunctions) list these core functions as of version 0.7, the version of this package slated for release alongside the manuscript.[^helperfunctions]
+These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis.[^leftjoin] Table \@ref(tab:tab-createfunctions) and Table \@ref(tab:tab-addfunctions) list these core functions as of version 0.7, the version of this package slated for release alongside the manuscript.[^helperfunctions]
+
+[^leftjoin]: In both cases, SQL-like "joins"---especially the left (outer) join---are the primary workhorse in these functions.
+A vignette on the package website, included in the appendix, explains these various joins, why they are used, and what issues may arise from them.
 
@@
 
-Table \@ref(tab:tab-createfunctions) are functions that create base data frames for a researcher starting an original project.
+Table \@ref(tab:tab-createfunctions) lists functions that create base data frames for a researcher starting an original project.
 They serve as functions that communicate the units of analysis supported in this package and that the package is capable of generating for an interested user.
@@
 
-[^whittle]: The "whittle" (`wc_`) class of functions are a suite of functions that whittle dyadic dispute-years to true dyad-year data.
+[^whittle]: The "whittle" (`wc_`) class of functions is a suite of functions that whittle dyadic dispute-years to true dyad-year data.
 They are effectively used in `add_cow_mids()` and `add_gml_mids()` for dyad-year data, but are offered here to allow users to employ their own case exclusion rules to create their own dyad-year conflict data.
@@
 Users new to R should understand its use here is primarily for exposition of what the code snippets in this manuscript are doing.
+In subsequent code snippets, `#>` will indicate console output the user can expect from these functions.
 
@@
 I encourage users who are using `{peacesciencer}` for the first time, especially those who are learning R for the first time because of their interest in this package, to approach `{peacesciencer}` with an idea of the kind of data they want to create for the sake of a project.
-The core functionality of `{peacesciencer}` begins with the creation of a data frame, after which can be populated with various indicators of interest.
+The core functionality of `{peacesciencer}` begins with the creation of a data frame, which can then be populated with various indicators of interest.
 No matter, the suggested use of `{peacesciencer}` begins with the creation of a data frame.
@@
 Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the UCDP armed conflict data.
-They will also know they have no use for pre-1946 observations since civil conflict data, certainly UCDP's suite of data, have coverage only from 1946 forward.
+They will also know they have no use for pre-1946 observations since civil conflict data typically have coverage only from 1946 forward.
 Thus, the user can supply some additional arguments to tailor the creation of data to just what they want (here: all Gleditsch-Ward state-years from 1946 to 2019).
@@
 
-Another reader may be interested in using `{peacesciencer}` for a research project using a leader-year unit of analysis, which can be created with `create_leaderyears()`.
+Another reader may be interested in using `{peacesciencer}` for a research project using a leader-year unit of analysis.
+These data can be created with `create_leaderyears()`.
 Users can read more about what this function is doing by consulting the documentation file in R with the following command.
@@
 
-Adding information to this data frame is a simple matter of joining a series of functions together in a "pipe." The "pipe"---represented as `%>%` in the code below---is an operator built into `{tidyverse}` that allows (`R`) users to pass forward expressions or functions.
+Adding information to this data frame is a simple matter of joining a series of functions together in a "pipe." The "pipe"---represented as `%>%` in the code below---is an operator built into `{tidyverse}` that allows users to pass forward expressions or functions.
 These pipes are common in the programming world and, as the code below will show, have the benefit of changing code in a way that is more intuitive and easier to both read and write for the user.
@@
 Finally, we can add some covariates of interest to these data.
-`add_peace_years()` calculates peace spells between ongoing conflicts in the data generated by `add_ucdp_acd()`.
+`add_spells()` calculates peace spells between ongoing conflicts in the data generated by `add_ucdp_acd()`.
 `add_democracy()` adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' [-@marquez2016qme] extension of Pemstein et al's [-@pemsteinetal2010dc] Unified Democracy Scores).
@@
 Table \@ref(tab:tab-createfunctions) shows support for creating leader-level data that are standardized to either the CoW or G-W state system data.
-Suppose a researcher wanted to create leader-year data, standardized to CoW system dates, for an analysis of inter-state dispute initiation analogous to what @horowitzstam2014hpme do in their analysis.
+Suppose a researcher wanted to create leader-year data, standardized to CoW system dates, for an analysis of inter-state dispute initiation analogous to what @horowitzstam2014hpme do.
 The user would first start with creating the base data (`create_leaderyears(standardize = "cow", subset_years=c(1875:2010))`), which also generates some leader attributes (e.g.
@@
 This is an indirect way of the user to get data that are state-year, dyad-year, triad-year (etc.), which are then expanded and populated with data at various levels.
-`{peacesciencer}`, much like EUGene, is more explicit, encouraging the user to be upfront about what their unit of analysis is and what are the units of analysis that can be plausibly plugged into the data the user is creating.
+`{peacesciencer}`, much like EUGene, is more explicit, encouraging the user to be upfront about what their unit of analysis is and what are the data that can be plausibly plugged into the data the user is creating.
 `{peacesciencer}`, again, does expect at least some level of comfort with the R programming language, but even this comes with greater ease of interpreting what is the unit of analysis and what are the primary spatial and temporal units that serve as the basis of the data.

### 2021-12-22 user-guide/ms.Rmd
batch update of stuff

@@
 removetitleabstract: "`r params$removetitleabstract`"
+nocite: |
+  @cartersmith2020fml, @marshalletal2017p, @coppedgeetal2020vdem, @gleditschetal2002ac, @thompsondreyer2012hir, @cowstates2016, @stinnettetal2002cow, @pevehouseetal2020tow, @schvitz2021mis, @palmeretal2021mid5, @singer1987rcwd, @nardulli2012creg, @gledtischward1999rlis, @nunnpuga2012r, @haege2011cc, @pettersson2019ov, @leedsetal2002atop, @barbierietal2009td
 #always_allow_html: TRUE
@@
         "add_minimum_distance()", "D, L, LD, S", "Add minimum distance data to data frame",
-        "add_nmc()", "D, L, LD, S", "Add minimum distance estimates (in km) to data frame",
+        "add_nmc()", "D, L, LD, S", "Add Correlates of War National Military Capabilities data",
         "add_peace_years()", "D, S", "Add 'peace years' to dyad-year/state-year conflict data",
@@
         "Archigos", "Add Archigos political leader information", "D, S", "add_archigos()",
-        "ATOP Alliances (v. 5.0)", "Add ATOP alliance data", "D", "add_atop_alliance()",
+        "ATOP Alliances", "Add ATOP alliance data", "D", "add_atop_alliance()",
         "Capital-to-Capital Distances", "Add capital-to-capital distance", "D, S", "add_capital_distance()",
@@
         "CoW Trade Data", "Add CoW trade information to data", "D, S", "add_cow_trade()",
-        "CoW Wars", "Add CoW inter-state (v. 4.0) and intra-state (v. 4.1) war data", "D", "add_cow_wars()",
+        "CoW Wars", "Add CoW inter-state and intra-state war data", "D", "add_cow_wars()",
         "CREG", "Add ethnic/religious fractionalization/polarization by year", "D, S", "add_creg_fractionalization()",
-        "Democracy", "Add Polity (v. 2017), V-Dem (v. 10), and Marquez' (2016) UDS extensions", "D,S", "add_democracy()",
+        "Democracy", "Add Polity, V-Dem, and Marquez' (2016) UDS extensions", "D,S", "add_democracy()",
         "GML MIDs", "Add Gibler-Miller-Little MIDs to data", "D", "add_gml_mids()",
@@
         "Strategic Rivalries", "Add Thompson and Dreyer (2012) strategic rivalry data by year", "D, S", "add_strategic_rivalries()",
-        "UCDP Armed Conflicts (v. 20.1)", "Add UCDP armed conflict data by year", "S", "add_ucdp_acd()"
+        "UCDP Armed Conflicts", "Add UCDP armed conflict data by year", "S", "add_ucdp_acd()"
         ) -> included
@@
 `{peacesciencer}` covers all the same units of analysis that EUGene covers, as of version 3.2.[^v3212] EUGene has more explicit support for dyadic dispute-year data whereas `{peacesciencer}` treats dyadic dispute-year data as a derivation of dyad-year data, with respect to functions that populate base data with additional information.[^whittledeclare] `{peacesciencer}` may reflect more current research interests in the peace science community.
-This is why there are leader-level data and support for the G-W (and UCDP) ecosystem of data that EUGene does not have, but there is no function yet for things like calculating Tau-b or expected utility values.
+This is why there are leader-level data and support for the G-W (and UCDP) ecosystem of data that EUGene does not have, but there is no function yet for things like calculating expected utility values.
 Importantly, though, `{peacesciencer}` mimics the verbosity of EUGene's user manual.
@@
 how CoW state system data differ from Gleditsch-Ward data and how one is coerced into the other), and how researchers should consider optimally using its functionality (e.g.
-`add_ucdp_acd()` probably should not lump all forms of conflict together).
+`add_ucdp_acd()` probably should not lump all forms of conflict together [c.f.
+@giblermiller2021apm]).
 `{peacesciencer}` does not endeavor to make researchers lazy or sloppy, and it does not ultimately do this.
-Instead, `{peacesciencer}` encourages reasoned design decisions by the user up front and reduces the tedium associated with starting quantitative peace science research.
+Instead, `{peacesciencer}` encourages well-reasoned design decisions by the user up front and reduces the tedium associated with starting quantitative peace science research.
 It achieves this in a quick, robust, and transparent way.

### 2021-12-22 R/add_fpsim.R

@@ -46,7 +46,7 @@
 #' An evaluation of the data, the article, and an email exchange
 #' with the author leads to the following points the user should
 #' consider. What follows is a rationale for why users should think of
-#' kappa a default measure for dyadic foreign policy similarity, though
+#' kappa as a default measure for dyadic foreign policy similarity, though
 #' why the "valued" equivalent for the alliance data is an inadvisable
 #' default. The example at the end of the document offers the operational
 #' "nudge" for what the user should want from this function.

### 2021-12-22 R/data.R

@@ -434,7 +434,7 @@ NULL
 
 #'  Correlates of War National Military Capabilities Data
 #'
-#' These are version 5.0 of the Correlates of War National Military Capabilities data. Data omit
+#' These are version 6.0 of the Correlates of War National Military Capabilities data. Data omit
 #' the state abbreviation and version identifier for consideration.
 #'
 #' @format A data frame with 15171 observations on the following 9 variables.

### 2021-12-22 data-raw/cow_nmc.R

@@ -1,6 +1,6 @@
 library(tidyverse)
 
-cow_nmc <- read_csv("/home/steve/Dropbox/data/cow/cinc/NMC_5_0.csv") %>%
+cow_nmc <- read_csv("/home/steve/Dropbox/data/cow/nmc/6.0/NMC-60-abridged.csv") %>%
   select(-version, -stateabb) %>%
   mutate_at(vars(-ccode, -year), ~ifelse(. == -9, NA, .)) %>% print()
 

### 2021-12-23 user-guide/ms.Rmd
update

@@
 The project's Github contains a publicly available `data-raw` directory that shows how every data set included (and processed) in `{peacesciencer}` was created.
-The function manuals included `{peacesciencer}` contain ample documentation that clarifies what the function is doing, what it returns to the user, and why it is doing it this way.
+The function manuals included `{peacesciencer}` contain ample documentation that clarify what each function is doing, what it returns to the user, and why it is doing it this way.
 Researchers can also use the project's Github to point out bugs, ask for further clarification, and propose additions.
@@
 The core functionality of `{peacesciencer}` reduces to two broad categories of functions.
-These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis.[^leftjoin] Table \@ref(tab:tab-createfunctions) and Table \@ref(tab:tab-addfunctions) list these core functions as of version 0.7, the version of this package slated for release alongside the manuscript.[^helperfunctions]
+These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis.[^leftjoin] Table \@ref(tab:tab-createfunctions) and Table \@ref(tab:tab-addfunctions) list these core functions as of version 1.0, the version of this package slated for release alongside the manuscript.[^helperfunctions]
 
@@
 4.1], optionally standardizing leader-years to the Gleditsch-Ward or Correlates of War state system data.
-As of version 0.7, `{peacesciencer}` is capable of creating full dyad-year data, leader-day data, leader-dyad-year data, leader-year data, state-day data, and state-year data.[^leaderdyadyear] A vignette on the package's website shows how users can create other forms of data from these functions as well (e.g.
+As of version 1.0, `{peacesciencer}` is capable of creating full dyad-year data, leader-day data, leader-dyad-year data, leader-year data, state-day data, and state-year data.[^leaderdyadyear] A vignette on the package's website shows how users can create other forms of data from these functions as well (e.g.
 dyadic-dispute-year, leader-months, state-quarters).
@@
 @arelbundock2021w].
-Its coverage will assuredly expand with new additions of interest to the peace science community, though the package already offers a lot to meet researcher needs.
+Its coverage will assuredly expand with new additions of interest to the peace science community, though the package already offers a lot to meet researcher needs.[^ps_version]
+
+[^ps_version]: The version of `{peacesciencer}` slated for release includes the most up-to-date-possible versions of data sets referenced in this manuscript.
+One helper function, `ps_version()`, will inform the user about the versions of the data included in the installed package.
+`{peacesciencer}` promises prompt package updates as new versions become available.
 
@@
 For example, assume the researcher wants just all politically relevant, non-directed dyad-years, where political relevance is traditionally understood as a dyadic relationship involving some form of a contiguity relationship or a major power [@lemkereed2001rprd].
-`create_dyadyears(directed = FALSE, subset_years = c(1816:2010))` created the full universe of non-directed dyad-years from 1816 to 2010, though this full universe includes "irrelevant" dyads like Nigeria-Mongolia and Canada-Estonia.
+`create_dyadyears(directed = FALSE, subset_years = c(1816:2010))` created the full universe of non-directed dyad-years from 1816 to 2010, though this full universe includes "irrelevant" dyads like Nigeria-Mongolia and Estonia-Rwanda.
 Reducing the data to just politically relevant dyads is simple in `{peacesciencer}` and `{tidyverse}`.
@@
 In this particular application, it has already created the main data required for a replication of @bremer1992ddc.
-There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some "weak-link" indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the @marquez2016qme for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.
+There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some "weak-link" indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the @marquez2016qme method for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.
 Table \@ref(tab:tab-dd) is a formatted version of the results of a logistic regression model of conflict onset using these "dangerous dyads" indicators.
@@
 Options include "extrasystemic", "interstate", "intrastate", and "II" (short for "internationalized intrastate").
-If no `type` is specified, the function returns information about ongoing disputes and onsets for all states for all tyeps of conflict.
+If no `type` is specified, the function returns information about ongoing disputes and onsets for all states for all types of conflict.
 If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g.

### 2021-12-23 user-guide/memo.Rmd
updates

@@
 ```{r setup, include=F}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+
+library(tidyverse)
+library(peacesciencer)
+library(kableExtra)
+
+is_docx <- knitr::pandoc_to("docx")
+is_latex <- knitr::pandoc_to("latex")
+is_html <- knitr::pandoc_to("html")
 
@@
 
-I thank you all for the opportunity to submit revisions of my manuscript "Asfsfssdfsafdsafs3fs" for possible inclusion in *Conflict Management and Peace Science* (*CMPS*).
-I thank the two reviewers and the editorial team for the helpful comments as well.
+I thank the editor and the two reviewers to their reception of my manuscript "`{peacesciencer}`: An R Package for Quantitative Peace Science Research" for inclusion in *Conflict Management and Peace Science* (*CMPS*).
+A conditional acceptance is always a welcome gift from the profession during the holidays.
+In what follows, I address the editor about the minor changes to the manuscript the editor suggested in order to transition the manuscript to an outright acceptance.
 
-# Comments to the Editorial Team
+First, I included the citations to *all* the data referenced in the manuscript (i.e.
+all the data included in `{peacesciencer}`), even if the manuscript itself did not make explicit use of the data.
+R2 (rightly) noted that a project that is concerned with proper attribution should have a manuscript that is comparably as concerned with proper attribution.
+R2's application here was the ATOP data.
+In this case, the manuscript notes coverage of both the ATOP and Correlates of War alliance data, uses (and cites) the Correlates of War alliance data in one application, and incidentally cited neither source outright.
+It was an oversight to not reference @gibler2009ima, though R2 felt  @leedsetal2002atop should be cited as well.
+This version of the manuscript now cites all data sets included in `{peacesciencer}`.
 
-The editorial team noted
+R2 also advised me of some data updates that should be included in the package, encouraging me to keep everything as up-to-date as possible.
+R2's case here was version 6.0 of the Correlates of War National Material Capabilities data, suggesting I include these updates in the package.
+The version of `{peacesciencer}` slated for release alongside the manuscript (v.
+1.0.0) now has these updates.
+I did want to use this particular case as a plea to the editor about why I structured the manuscript the way I did.
+Notice the manuscript is light on listing version numbers for the data included in the manuscript.
+In the case of the capabilities data referenced by R2, the data were released on July 22, 2021 and the manuscript was initially submitted earlier that same month.
+Thus, the peer review process almost concluded by missing this update entirely.
+I mention this to note that it stands to reason that something like this will almost invariably happen again at the copy-editing stage.
+Indeed, the manuscript notes that the `create_dyadyears()` and `create_stateyears()` functions will generate data to the most recently concluded calendar year.
+Right now, that is 2020.
+In a week, it will be 2021.
 
-Enclosed are the revised manuscript and appendix. I thank the editorial team and the reviewers for their helpful comments and for further consideration of my revisions.
+I mention this to note that what I offer to the editorial team, the reviewers, and the peace science community is effectively a sincere "scout's honor" that the package will remain as up-to-date as possible, even if interested users may have to bring to my attention updates that I may miss.
+Features will invariably fall out of date and need to be updated.
+If anything, I think this is a unique advantage, of a kind, to `{peacesciencer}`.
+There is nothing in this package that will break upon the release of some new data set, since the data are effectively stored in-house.
+It just means there is a more current version of the data that needs to be properly cleaned and included in a future update.
+CRAN---the central repository for R packages---is generally very quick to update packages.
+All CRAN requires is about a month between updates (so they are not spammed with busy work done on a volunteer basis).
+The timeline from submission to publication of an R package update is almost always less than week (holidays notwithstanding), and routinely (from my experience) as quick as a day.
+
+In other words, these updates (and more) are going to be part of a new version (1.0.0) slated for release alongside the manuscript.
+In most cases, I avoid deliberate reference to version numbers of data included in the package in case they fall out of date prior to the appearance of the manuscript in print.
+However, I do believe updating `{peacesciencer}` will be much more streamlined, and much quicker, than alternatives like NewGene and EUGene.
+
+Finally, I thank the editor for encouraging me to at least acknowledge, if not outright incorporate, R1's suggestion to show how much time is saved in `{peacesciencer}` relative to doing things manually.
+I plan to make this another vignette around the time the paper comes to print.
+The application here will be the `create_dyadyears()` function.
+If R1 is interested, I encourage them to time `create_dyadyears()` in `{peacesciencer}` with a procedure similar to what I outline here at the link in this footnote.[^falsedyadslink] This procedure uses multi-core processing to identify every single day overlap of Correlates of War dyads for all years from 1816 to 2016.
+I use this mainly to identify what I term in the memo and appendix to be "false" dyads like Suriname and South Vietnam (i.e.
+states that exist in the same year, but never at the same time in the year).
+A user can see how much time is required for this procedure that ultimately ensures that these "false" dyads do not appear in their dyad-year data, and how much time is ultimately spared them by `{peacesciencer}` and the `create_dyadyears()` function.
+
+[^falsedyadslink]: https://github.com/svmiller/peacesciencer/blob/master/data-raw/false_cow_dyads.R
+
+All told, I think the manuscript is ready for outright acceptance and I humbly thank again the editor and the reviewers for their interest.
+This has been one of the more enjoyable review experiences in my over 10 years in the profession.
+Enclosed are the revised manuscript and appendix.
 
@@
 
-# Reviewer #1's Comments
+# Previous Memo to the Editorial Team
 
-Reviewer #1 (R1) had the following comments on the original version of the manuscript.
+*This was the memo to the editorial team and two reviewers for the first round of revisions.
+I include it here for posterity, though some text was slightly changed for presentation (i.e.
+I saw my own typos).*
+
+I thank you all for the opportunity to submit revisions of my manuscript "`{peacesciencer}`: An R Package for Quantitative Peace Science Research" for possible inclusion in *Conflict Management and Peace Science* (*CMPS*).
+I thank the two reviewers and the editorial team for the helpful comments as well.
+In what follows, I address the editor about the changes to the manuscript that came in light of reviewer feedback and continued development of the package.
+Thereafter, I address the reviewers' comments individually.
+
+##  Comments to the Editorial Team
+
+The package and paper evolved considerably since the first version I submitted for review at *CMPS*. In addition to the slight name change, I want to drive home the following major changes.
+
+First, the manuscript is much longer now.
+An email inquiry to the editorial team informed me that there is no different word limit for data features relative to other article types that *CMPS* would accept.
+The manuscript I submitted for review was under 6,000 words.
+This version is just under 9,500 words.
+
+Second, `{peacesciencer}` is *much* more feature-rich.
+It has robust support for leader-level analyses.
+The manuscript even shows how you could use built-in functions and data to approximate the kind of analysis that @horowitzstam2014hpme did.
+
+Third, there are more vignettes for the package that discuss some important topics about this package, some raised during the peer-review process.
+These are also included in the appendix submitted alongside the memo and the manuscript.
+
+All told, I think the manuscript and the package are much stronger now for these changes, and I think (and certianly hope) the editorial team and reviewers would agree.
+Enclosed are the revised manuscript and appendix.
+I thank the editorial team and the reviewers for their helpful comments and for further consideration of my revisions.
 
@@
 
-# Reviewer #2's Comments
+## Reviewer #1's Comments
 
-I thank Reviewer #2 (R2) for
+Reviewer #1 (R1) had the following comments on the original version of the manuscript.
+First, I want to thank R1 for this feedback, and for the editorial team for selecting R1 to review this manuscript.
+R1 had exactly the kind of expertise I wanted to get from this process: someone who has good insight into writing software.
+
+I have identified the following major issues raised by R1 and write here how I have revised the manuscript and the package in light of these issues.
+
+## The documentation lacks detail about how these merges work
+
+R1 mentioned that the documentation lacks some detail about how the merges work and, in particular, draws attention to how dyadic dispute-year data are coerced into true dyad-year data.
+I want to mention three things in response.
+
+First, this has always been communicated to the user if they looked at the underlying code, but I understand the reviewer's point that this should not be necessary.
+Toward that end, `add_cow_mids()` and `add_gml_mids()` will now communicate what are the case exclusion rules it employs in coercing dyadic dispute-year data to true dyad-year data.
+This communication appears as a message when the respective function finishes.
+
+Second, this has long been a vignette on the package's website.
+While more care could be given to point the user in this direction, there is a dedicated vignette that belabors these case exclusion rules.
+I include this in the appendix as well.
+
+Third, the release of v.
+0.6.0 of `{peacesciencer}` came with a class of what I term "whittle" functions.
+These are specialty functions that coerce dyadic dispute-year data into dyad-year in a given sequence that the researcher deems best fits their particular needs.
+Table \@ref(tab:tab-whittlefunctions) lists these functions, partly abbreviated from Table 2 of the manuscript.
+For example, `wc_onsets()`---itself a shortcut for the more verbose `whittle_conflicts_onsets()`---will take dyadic dispute-year data (either Correlates of War [CoW] or Gibler-Miller-Little [GML]) and, where duplicates exist in a dyad-year, keep only those observations that are unique dispute onsets.
+`wc_fatality()`---itself a shortcut for the more verbose `whittle_conflicts_fatality()`---will take dyadic dispute-year data (either CoW or GML) and select those with the highest observed dispute-level fatality.
+
+```{r tab-whittlefunctions, echo=F}
+
+tribble(~func, ~type, ~description,
+        "wc_fatality()", "C", "Whittle Duplicate Conflict-Years by Highest Fatality",
+        "wc_hostility()", "C", "Whittle Duplicate Conflict-Years by Highest Hostility",
+        "wc_jds()", "C", "Whittle Duplicate Conflict-Years by Just Dropping Something",
+        "wc_onsets()", "C", "Whittle Uniqute Conflict Onset-Years from Conflict-Year Data",
+        "wc_recip()", "C", "Whittle Duplicate Conflict-Years by Conflict Reciprocation",
+        "wc_stmon()", "C", "Whittle Duplicate Conflict-Years by Lowest Start Month"
+) %>% select(-type) -> whittle_functions
+
+if (is_latex) {
+kable(whittle_functions, booktabs = TRUE, longtable = TRUE,
+    caption = "The 'whittle' Class of Funtions",
+    align = c("l",  "l"),
+    linesep = "",
+   col.names = c("Function",  "Description"))  %>%
+   row_spec(0, bold=TRUE) %>%
+    column_spec(1, monospace = TRUE) %>%
+    column_spec(2, width="25em")
+} else {
+kable(whittle_functions, booktabs = TRUE, longtable = TRUE,
+    caption = "The 'whittle' Class of Funtions",
+    align = c("l",  "l"),
+    linesep = "",
+   col.names = c("Function",  "Description"))  %>%
+    kable_styling(font_size = 9) %>%
+   row_spec(0, bold=TRUE) %>%
+    column_spec(1, monospace = TRUE) 
+}
+
+```
+
+The vignette on the package's website, combined with the message from, for example, the `add_gml_mids()` function, means these two functions will produce the equivalent dyad-year data for an analysis of dispute onsets (even if the column names will differ).
+
+```{r, eval=F}
+# Approach 1
+cow_ddy %>% add_gml_mids()
+
+# Approach 2
+gml_dirdisp %>%
+  whittle_conflicts_onsets() %>%
+  whittle_conflicts_fatality() %>%
+  whittle_conflicts_hostility() %>%
+  whittle_conflicts_duration(durtype = "mindur") %>%
+  whittle_conflicts_duration(durtype = "maxdur") %>%
+  whittle_conflicts_reciprocation() %>%
+  whittle_conflicts_startmonth() %>%
+  whittle_conflicts_jds() %>%
+  left_join(cow_ddy, .)
+```
+
+I thank R1 for encouraging me to do this---it really was no chore since I had already effectively written these functions---and I think the functionality of the package improved for it.
+
+### Better explain the left join
+
+R1 also deduced from the underlying code that the left (outer) join is everywhere in this package and is the primary workhorse for the functions that populate a base data set created in this package.
+R1 encouraged me to acknowledge this and explain it in some detail.
+
+In light of some other feedback, I felt an entire section on the left join would have been a detour from the flow of the manuscript.
+Instead, I do acknowledge this type of join as the primary workhorse of the functions included in the package and reference a vignette on the website, included as an appendix in this submission, about what this type of join is doing, why I use it the way I do, and what issues may arise from them.
+I summarize this as follows.
+
+In short, the left join is a type of "mutating" join that adds information from one data set into another based on some common matching keys.
+Assuming two data objects, `x` and `y`, then `left_join()` returns all rows from `x` (i.e.
+a "left" object) with matching information in `y` (i.e.
+a "right" object) based on those matching keys across `x` and `y`.
+Its use in this package has the following assumptions.
+First, the universe of cases of interest to the user---as generated by one of the "create" functions in the package---is in the "left" object.
+A user may, for example, have a state-year analysis of European countries and want to merge in some, say, GDP information that has observations as well for African countries.
+Information from the African countries will *not* be merged into the left object with this type of join.
+Second, there are no unwanted duplicate observations in the "right" object, which would create unwanted duplicate observations in the "left" object.
+If, for example, the left object has just one observation for a unified Germany in 1990, but the right object has two observations for Germany in 1990 (e.g..
+pre- and post-unification), this type of join will duplicate those observations in the left object.
+Thus, the package invests considerable care in making sure the functions contained in this package do not create additional rows in the output beyond the dimensions of the data supplied to it.
+
+I think this discussion makes more sense in the appendix or as a vignette on the package's website, but I thank R1 for encouraging me to acknowledge this.
+
+### Communicate version numbers of data
+
+R1 liked that I included a `ps_cite()` function for appropriately citing the material I use, but felt that the package needed a corollary for citing the version of the data included in the package.
+Thus, I have included a function---`ps_version()`---that produces a data set of version numbers associated with the data included in this package.
+Here is a preview of what this function will do.
+
+```{r}
+# What are all the version numbers?
+ps_version()
+
+# What are the version numbers of data where category == "democracy"?
+ps_version("democracy")
+```
+
+A vignette on the package website, also included in the appendix, explains how to interpret some of these version numbers.
+For example, a version number that is a year includes data that are not formally "versioned", per se, [c.f.
+@andersetal2020bbgb; @cartersmith2020fml] and that the data included reflect the last update or date of publication.
+
+### Include a license file
+
+R1 drew my attention to the *Journal of Open Source Software* and encouraged me to follow their lead in adequately licensing this software.
+Thus, the software bundle now includes an actual GPL-2 license file and not just a reference to GPL-2 licensing.
+
+### Do a better job engaging with other software packages
+
+R1 also felt the version of the manuscript I submitted for peer review could do a better job engaging with alternatives to what I propose in the paper.
+Namely, R1 wanted to know what is the current state of this practice and how `{peacesciencer}` is an improvement (or at least different).
+To accommodate this, I dedicated an entire section of the revised manuscript to how `{peacesciencer}` compares with some alternatives.
+I will also add to that discussion here.
+
+For one, I want to communicate to R1 and the editorial team that I searched high and low for alternatives to what I propose here and found only NewGene and, before it, EUGene.
+I will confess up front that the initial impetus for what became `{peacesciencer}` was constantly having to remind myself how to create dyad-year data based on state system membership.
+I did a Google search for that---"how to create dyad-year data" (without quotes)---and the first search result was to a blog post of mine that itself became a part of the code included in this package.
+The fifth search result was even to a Stack Overflow post on this exact same question over eight years ago, that I incidentally had asked.
+Someone on the Stata forum had actually asked this question in 2014 as well.[^stataforum] The bulk of the search results on top of that point to this particular package, in addition to references to NewGene and EUGene.
+I mention this to say that I do not think many alternatives to `{peacesciencer}` exist, certainly as practitioners in our field would think to search for them.
+
+[^stataforum]: https://www.statalist.org/forums/forum/general-stata-discussion/general/511471-creating-many-dyads
+
+My hunch is that the current state of the practice, at least for the core functionality of creating raw data of interest to IR researchers, is one that relies on data transformations that owe to some functionality built into Structured Query Language (SQL).
+R1 deduced from the raw code that the left (outer) join is everywhere in this package.
+The implementation may be in R, but the procedure is very much SQL at heart (e.g.
+left joins, inner joins).
+The revised manuscript acknowledges this, and does say that a researcher who wants to create state-year or dyad-year data from state system membership data, or leader-year and leader-dyad-year data from Archigos, could just as well invest time in learning SQL to accomplish this.
+The process of creating `{peacesciencer}` has made me appreciate SQL more---and I am convinced there is important value for students in learning it---but learning SQL just to create these types of data amounts to a detour in the research process.[^c++]
+
+[^c++]: I am unaware exactly how EUGene and NewGene are programmed.
+EUGene's documentation for version 3.2 says the application itself was written in the "Borland Delphi language", which I think pertains to the integrated development environment and not the programming language in the software itself.
+Delphi has long had SQL support, at least from its first release, and I think Delphi itself is derivative of Pascal.
+@bennettetal2019n thank what appears to be a C++ programmer in the acknowledgment section of their manuscript and greeting to NewGene announces its C++ basis in the bottom right corner, but some debugging output given by NewGene suggests that important SQL transformations are happening underneath the proverbial hood.
+I will confess to having no real operational knowledge of C++, Delphi, or Pascal.
+
+EUGene is the clear inspiration for this package.
+I acknowledge R1's minor point that using EUGene to create raw data from scratch was not the main point of EUGene---it just happened to be how most of us used it over 10 years ago---but the point remains EUGene is effectively retired as a software package.
+It was last updated in 2017 with some minor bug fixes, though these bug fixes do not come with any change in functionality from a version (3.2) that was last released in 2007.
+This means `{peacesciencer}` has a major advantage over EUGene in better meeting current research interests and needs.
+It has support for leader-level data and leader-level analyses where EUGene does not.
+It has support for the ecosystem of data denominated in Gleditsch-Ward state system data where EUGene does not.
+Functionally, this means `{peacesciencer}` can support the needs of civil conflict scholars working with UCDP-PRIO data where EUGene cannot.
+A researcher who may still want to calculate older variables like Tau-b and expected utility values will not find that in `{peacesciencer}` right now, but `{peacesciencer}` is more than capable of meeting current needs in the peace science community.
+
+NewGene is the closest thing that approximates what `{peacesciencer}` does, among software options still actively supported.
+The manuscript mentions where the two differ.
+NewGene has a current advantage over `{peacesciencer}` in the construction of *k*-adic data, and it right now has some variables that `{peacesciencer}` could easily incorporate in future updates (e.g.
+ICOW territorial claim data).
+`{peacesciencer}` has important support for civil conflict data where NewGene's focus is almost entirely aimed for researchers interested in inter-state conflict.
+I also want to dedicate some additional space here to clarifying some important differences between the two that I do not think the lay reader will understand or appreciate, but I am sure the reviewer will because of their demonstrated insight into how `{peacesciencer}` is working and how to write good software for this community.
+
+For one, the manuscript does make a reference to the fact that `{peacesciencer}` is more "explicit" than NewGene, which is not as "explicit" as EUGene.
+I want to clarify what I mean here because I think it touches on multiple issues.
+For one, as I do reference in the manuscript, NewGene is vague on what it understands to be the unit of analysis.
+NewGene never formally tells the user to declare their unit of analysis in the interface it provides and the closest it does come to this is near the top where it prompts the user to state how many (Correlates of War) country code columns they want.
+If the user selects 1, NewGene generates monadic data over a temporal domain the user specifies in the next tab.
+If the user selects 2, NewGene generates dyadic data over a temporal domain the user specifies in the next tab.
+With three onward, the user is getting some *k*-adic data over a temporal domain the user specifies in the next tab.
+This would otherwise be no real issue if not for the fact that NewGene offers leader-level data.
+A researcher can think of leader-level analyses as derivations of state-level, dyadic-level, or *k*-adic-level analyses the extent to which leaders are nested in states, but leader-level units of analysis are still distinct from these other units of analysis.
+Thus, if I am a naive researcher using NewGene for the first time asking for just one state-year column over a given period of time and checking boxes for Archigos information in return, do I know if I am getting true leader-year data or just state-year data?[^faq] `{peacesciencer}` forces this upfront by telling the user to state what is their unit of analysis before anything else is generated alongside it.
+NewGene only gets the user to state their unit of analysis after they indicate the information they want, which, in a sense, conflates the operationalization of the research design with the conceptualization of the research design.
+
+[^faq]: The FAQ on NewGene's website advertises how you would do this when it explains why states are the main decision-making units, though the documentation (`FAQ 14`) [describes a NewGene interface that is no longer current as version 2.0](http://www.newgenesoftware.org/download/i/mark_dl/u/4012918385/4629218818/Why%20are%20countries%20the%20main%20DMU.pdf).
+This leaves the user---this particular user---in the dark as to how one could get something like leader-year data with some state-year information (e.g.
+capabilities) using NewGene.
+
+On a related note, NewGene is similarly vague about merging in data between different data ecosystems, or even among the same data ecosystem.
+For example, assume a researcher is using NewGene to create a state-year data set that has some national material capabilities data (denominated in the Correlates of War state system) along with some IPE data (which, at least in part, is denominated in the Gleditsch-Ward ecosystem).
+Which system is the master system here?
+After selecting where the user wants the data saved, they must then pick what is the "top-level variable group" (either the data containing the capabilities or the data containing some IPE variables).
+I think this suboptimal because it does not force the user to state upfront what is the universe of their cases (i.e.
+in the `{peacesciencer}` case, the Correlates of War state system over a period of time or the Gleditsch-Ward system over a period of time).
+This may seem trivial, but it is not.
+It again encourages the researcher to conflate the conceptualization of the research design with the operationalization of the research design.
+Stating upfront what is the universe of cases is an important design decision.
+`{peacesciencer}` does this and NewGene does not appear to do this.
+Base data approximating the universe of cases should be created *first* before data are appended to them and users should declare what is the population to which they are trying to infer (i.e.
+the state system data in the quantitative peace science context) before collecting information.
+
+Think of what `{countrycode}` does for us as analogous to what is happening here, and why this is important.
+R users in our field know about this package well since it is the most useful package for matching one state system classification to another, including support for Correlates of War system, Gleditsch-Ward system, Federal Information Processing Standards, UN M49, and much more [@arelbundocketal2018c].
+However, anyone who uses this package for, say, converting Correlates of War state codes to ISO 3166-1 alpha-3 codes for the sake of merging in World Bank data [c.f.
+@arelbundock2021w], are acknowledging---if implicitly---that their master system is the Correlates of War system.
+Their unit of analysis is Correlates of War states over a certain time period.
+It just happens to be the case that Correlates of War state codes can mostly be matched to ISO 3166-1 alpha-3 codes for the cause of merging in information about some economic indicator of interest.
+However, this matching and merging procedure does not change the fact that there is a master system, which remains the master system, even after merging.
+That should be declared *before* anything else is added to it, which is what `{peacesciencer}` does with the `create_` family of functions.
+In terms of current and best practices, I think teaching researchers to acknowledge this is one of those subtly important things that we should be doing better since---as `{countrycode}`'s development shows and what `{peacesciencer}`'s raw data and comments emphasize on a more focused scale---this can be quite challenging.[^statesystems]
+
+[^statesystems]: A vignette on the package website talks in greater detail how `{peacesciencer}` handles the integration of Correlates of War state system data with Gleditsch-Ward state system data while also reiterating this same point.
+
+R1 raised an issue, addressed above, that `{peacesciencer}` could better inform the user to how dyadic dispute-year data are coerced into dyad-year data.
+This information had always been in `{peacesciencer}` and was a vignette on the package's website, though a recent release (v.
+0.6.0) gave users the option to employ their own case exclusion rules to eliminate duplicate dyad-year observations.
+Toward that same end, NewGene is maximally non-committal about what would be an optimal approach.
+Indeed, one of NewGene's YouTube videos, published March 6, 2020, tells the user to eliminate duplicate observations through Excel.[^newgeneduplicate] The video describing the procedure has only music overlaying it (no commentary) and describes a procedure that is brute-force, with no questions asked, and with no explanation given to what it is being eliminated.
+`{peacesciencer}`'s transparency and explicitness are major improvements here.
+
+[^newgeneduplicate]: https://www.youtube.com/watch?v=ZBlW77lD4iQ
+
+Finally, I am quite confident of the quality control that goes into `{peacesciencer}` whereas I am less confident in some of the quality control that goes into what NewGene is doing.
+The following are just two things I noted in comparing v.
+0.7.0 of `{peacesciencer}` with version 2.0 of NewGene.
+I encourage the reviewer to see these things for themselves.
+
+First, create a simple state-year data set in NewGene that asks for the Correlates of War state code, year, and iron and steel production estimate from the country capabilities tab.
+Add to it a request to get a Gleditsch-Ward state code and the number of islands from the IPE data resource tab at the bottom.
+Subset the years to 1985 to 1995 in the next tab and declare the capabilities data to have the top-level variable group.
+Save the output to some spreadsheet and open that.
+Find where `ccode` is 255 and the year is greater than or equal to 1990.
+Notice that there is no number of islands information for Germany those years because NewGene did not internalize that what is Gleditsch-Ward's `260` in the state system data is Correlates of War's `255` after the unification of Germany.
+A similar problem occurs for Yemen, whose integration is treated identically.
+The application here is minimal, and perhaps no one is interested in the number of islands Germany and Yemen have, but this points to a bigger problem in how NewGene tries to integrate Correlates of War data with Gleditsch-Ward data.
+
+Second, create a simple dyad-year data set in NewGene (i.e.
+ask for two Correlates of War state columns near the top of the interface) and ask for the Correlates of War state code, year, and iron and steel production estimate from the country capabilities tab.
+In the next tab, ask for a full temporal domain of 1816 to 2020 and save to a spreadsheet.
+Open the spreadsheet and find where `ccode_1` is 115 and `ccode_2` is 817, with the operative year in question being 1975.
+This dyad will be there; it should not be.
+The Republic of Vietnam exits the Correlates of War state system on April 30, 1975 whereas Suriname enters the state system on November 25, 1975.
+Both appear in the same year, but not at the same time.
+These are false dyads that happen when expanding the state system data based on unique observations of state-years, but not state-days.[^falsedyads] `{peacesciencer}` identifies these and removes them in the `create_dyadyears()` function.
+
+[^falsedyads]: After loading `{peacesciencer}` into the R session, the reader can see what these "false" dyads are in the Correlates of War state system with `false_cow_dyads`.
+There is a Gleditsch-Ward equivalent as well (`false_gw_dyads`).
+
+All told, I think the reader will agree that `{peacesciencer}` is a useful solution to some important research design problems, albeit one that requires at least some familiarity with the R programming language.
+It has several innovations that are not available in alternative programs, along with more clarity as to what it is doing.
+The maximal transparency of `{peacesciencer}` allows for greater confidence in quality control as well.
+
+### Include some simpler code snippets for specific use cases
+
+R1 felt that the code snippets I included got too complicated too quickly and asked for simpler code snippets to highlight some more specific cases. I have tried to do this in a few ways.
+
+First, the code snippets I include in the manuscript are amply annotated with comments about what each function is doing.
+This is a made possible and easy by using R Markdown as front-end for writing the manuscript, allowing me to seamlessly execute code and use comments to explain what each code snippet is doing.
+
+Second, the section on creating and amending a dyad-year data set is done somewhat piecemeal.
+I start with the creation of the dyad-year data.
+I then talk about how the pipe operator allows the user to create dyad-year data and and subset the data to just politically relevant dyads.
+I then talk about how the user can extend this process forward another step by creating dyad-year data, subsetting the data to just politically relevant dyads, and then populating the data with information about ongoing inter-state disputes and dispute onsets from the Gibler-Miller-Little (GML) data set.
+This entire section shows the reader how to build the data they want, piece by piece, using `{peacesciencer}` functions and basic features of the `{tidyverse}`.
+The point is well-taken that the code snippets got too complicated without much explanation.
+I believe this version of the manuscript strikes a balance between both.
+There are simple use cases along with a guide for more thorough data-building.
+Everything is amply annotated with an explanation to what is happening.
+
+Third, and this is more a comment about what I cannot do, but I interpreted the comment from R1 here as more about the manuscript and not the documentation files included in the package (which R1 indicates they have seen).
+Thus, I worked hard to make clearer what the code is doing in the manuscript, but R documentation files do not have a lot of opportunity for highlighting specific use cases because of CRAN testing requirements.
+CRAN requires every test of a function in the documentation file be quick enough to execute in under 5 seconds on their amalgam of testing computers.
+Some of these testing computers can be quite slow, and indeed some of the functions in this package can be a little slow as well even on fast computers.
+I circumvent some of this with the package's website (which does execute every piece of code in the documentation file) as well as a vignette that showcases various parlor tricks (included in the appendix).
+However, I am unable to be too creative in the documentation files of the R package itself.
 
 \newpage
+
+## Reviewer #2's Comments
+
+I thank Reviewer #2 (R2) for their comments.
+On the balance, I read R2's feedback as supportive of what I had offered.
+However, they suggested more functionality should be included and the manuscript could be better streamlined for the reader.
+I address these main points below.
+
+### Introduce leader-level functionality
+
+R2 felt leader-level functions and leader-level data would be a major improvement over the version of the package I had submitted for peer review.
+As of v.
+0.7.0, `{peacesciencer}` now has multiple functions and built-in data for researchers interested in doing leader-level analyses.
+I will summarize these below.
+
+`create_leaderdays()`, `create_leaderyears()`, and `create_leaderdyadyears()` will create leader-day, leader-year, and leader-dyad-year data.
+Depending on the `standardize` argument specified in the function, the function will optionally standardize these data to the Correlates of War and Gleditsch-Ward state system.
+Though Archigos (the underlying leader data) is nominally denominated in Gleditsch-Ward state codes, the leader data have observations for leaders outside state system dates.
+Think of Lynden Pindling (`BHM-1967`) as at least one illustration here.
+Archigos has his tenure starting in 1967 though statehood for Bahamas begins in 1973.
+Thus, the user will still need to standardize the leader data to Gleditsch-Ward system dates, if they want this.
+
+`{peacesciencer}` also has some specialty functions designed for leader-level analyses.
+`add_lead()` and `add_lwuf()` will add information to leader-level analyses based on information from the LEAD data set [@ellisetal2015lead] and estimates about leader willingness to use force as generated by @cartersmith2020fml.
+Both depend on the `LEAD` and `lwuf` data also supplied in the package.
+`add_gml_mids()` also painstakingly matches leaders to dispute onset for leader-level analyses (both monadic and dyadic) on dispute initiation.
+`add_spells()` is a more flexible peace-spell function designed for the unbalanced nature of leader panels.
+
+Most other functions primarily written around state-year or dyad-year data will also work with leader-level data, For example, `add_democracy()` will return democracy estimates in the state-year, even for leader-year data.
+Where appropriate, users are reminded in the documentation files that data like these (e.g.
+democracy estimates in `add_democracy()`) are fundamentally state-year and that extensions to leader-level data should be understood as approximations for leaders in a given state-year.
+
+In addition, there is a helper data set provided in the package---`leader_codes`---that is used internally for helping sync leader observations across varying leader codes across Archigos 4.1, Archigos 2.9, and the LEAD data set.
+These data treat version 4.1 of the Archigos data as the gospel leader data (if you will) for which the observation ID (`obsid`) is the master code indicating a leader tenure period.
+It also builds in an assumption that various observations that duplicate in the LEAD data should not have duplicated.
+This concerns Francisco Aguilar Barquer (who appears twice), Emile Reuter (who appears twice), and Gunnar Thoroddsen (who appears three times) in the LEAD data despite having uninterrupted tenures in office.
+None of the covariates associated with these leaders change in the LEAD data, which is why I assume they were duplicates.
+
+### Better explain the pipe operator
+
+R2 felt the discussion around the pipe operator (`%>%`) was a bit distracting and asked to make it easier for the reader to follow along with what it is doing and why it is important.
+R2 alternatively felt it might be better to minimize the discussion of the pipe operator altogether.
+
+This is challenging, because I do see R2's point.
+The manuscript is primarily about the R package, its benefits, and why people should use it.
+It also dovetails with, effectively, an R primer that also elevates the `{tidyverse}`.
+The extent to which I have primarily written this as almost an advocacy piece for those who still otherwise do their statistical analysis in Stata to use this package (and R) to at least generate their data, R2 is right that I risk losing the reader.
+Thus, from that perspective, it may make some sense to minimize the discussion of the pipe operator since it is not a core function native to the package.
+
+However, I also strongly believe that the package becomes that much harder to use in the absence of the pipe operator.
+Compare the two code snippets below for creating Correlates of War state-years from 1816 to 2020 and populating it with information about major power status, democracy, and suprplus/gross domestic product information.
+The first approach uses the pipe operator as I advocate it.
+The second approach builds the same data, but without the pipe operator.
+
+```{r, message=F}
+# Approach 1, with the pipe operator
+create_stateyears() %>%
+  add_cow_majors() %>%
+  add_democracy() %>%
+  add_sdp_gdp() -> A
+
+# Approach 2, without the pipe operator
+add_sdp_gdp(add_democracy(add_cow_majors(create_stateyears()))) -> B
+
+# Are these two things identical?
+identical(A, B)
+```
+
+The data sets are identical, but that is also less the point.
+The first approach is much more readable than the second approach.
+I think the pipe operator is worth mentioning, and worth advocating, because it makes what is still---at its core---a programming language with a command-line interface seem *much* more intuitive.
+I think this is true even for beginners without much experience in R.
+I mention in the manuscript that the pipe operator has the effect of forming something analogous to a drop-down menu, in which the user can "select" additional data/commands they may want simply by specifying the function built into the package.
+The pipe operator is clearly not necessary, as the second approach in the code snippet shows, but it makes `{peacesciencer}` much more user-friendly.
+
+That said, I understand R2's point that I cannot lose the reader here, and I risked doing so in the first draft.
+I believe this version has much more explicit description of what the pipe operator is and how to use it.
+This, in particular, concerns the guide on creating dyad-year data for analysis, also addressing R1's point that the manuscript could benefit from simpler code snippets.
+
+### The passage about the graduate students adds no value to the manuscript
+
+R2 felt I could drop the passage about the graduate students---not because they disagreed with it, but because they felt it added no real value to the paper.
+I think this is fair, and I removed that material from this version of the mansucript.
+
+\newpage

### 2021-12-23 user-guide/ms.Rmd
updates

@@
 Researchers should see that `{peacesciencer}`'s functionality can scale up nicely from there.
-For example, the following would round out the kind of information necessary to replicate Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis by adding information about national material capabilities for both states in the dyad (`add_nmc()`), estimates of democracy for both states in the dyad (`add_democracy()`), information about alliance commitments in the dyad-year (`add_cow_alliance()`), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by @andersetal2020bbgb.
+For example, the following would round out the kind of information necessary to replicate Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis by adding information about national material capabilities for both states in the dyad (`add_nmc()`), estimates of democracy for both states in the dyad (`add_democracy()`), information about alliance commitments in the dyad-year (`add_cow_alliance()`, by way of @gibler2009ima), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by @andersetal2020bbgb.
 Whereas `add_sdp_gdp()` is the last command in the pipe-based workflow, the `{peacesciencer}` call ends by assigning to an object called `Data`.

### 2022-01-12 manuscript/appendix.Rmd
rename directory

--- 2021-11-09 user-guide/appendix.Rmd
@@
 
+# Vignette: Create Different Kinds of Data in `{peacesciencer}`
+
+```{r child = '../vignettes/different-data-types.Rmd', echo=TRUE}
+```
+
 # Vignette: A Discussion of Various Joins in `{peacesciencer}`
@@
 ```{r child = '../vignettes/joins.Rmd', echo=TRUE}
+```
+
+# Vignette: How `{peacesciencer}` Coerces Dispute-Year Data into Dyad-Year Data
+
+```{r child = '../vignettes/coerce-dispute-year-dyad-year.Rmd', echo=TRUE}
+```
+
+# Vignette: Various Parlor Tricks in `{peacesciencer}`
+
+```{r child = '../vignettes/parlor-tricks.Rmd', echo=TRUE}
+```
+
+# Vignette: A Discussion of Correlates of War and Gleditsch-Ward Systems and `{peacesciencer}`
+
+```{r child = '../vignettes/state-systems.Rmd', echo=TRUE}
 ```

### 2022-01-12 manuscript/ms.Rmd
rename directory

--- 2021-12-23 user-guide/ms.Rmd
@@
 
-[^helperfunctions]: `{peacesciencer}` also has miscellaneous "helper" functions that are not belabored in these manuscript.
+[^helperfunctions]: `{peacesciencer}` also has miscellaneous "helper" functions that are not belabored in this manuscript.
 These include functions that allow the user to use `{peacesciencer}` functions with data created outside the package (`declare_attributes()`), download external data for quicker use of other functions (`download_extdata()`), and functions that communicate suggested citations (`ps_cite()`), among a few other functions.
@@
 This function will allow the user to choose a state system for which they want state-years (`system`, which accepts either "cow" or "gw" for Correlates of War [CoW] state system data or Gleditsch-Ward [G-W] state system data and defaults to "cow"), whether they want to extend the state system to the most recently concluded calendar year (`mry`, which defaults to TRUE), and whether they may want to additionally subset the years to a more narrow temporal domain they may already have in mind (`subset_years`, which defaults to no subset of the data, returning all possible state-years).
-If the user simply ran the function with no overrides, the function would return all Correlates of War state-years from 1816 to the most recent year (2020, as of writing).
+If the user simply ran the function with no overrides, the function would return all Correlates of War state-years from 1816 to the most recent year (2021, as of writing).
 

### 2022-01-14 manuscript/ms.Rmd
fira code

@@
 mainfont: cochineal
+monofont: Fira Code
 #sansfont: Linux Biolinum O 

### 2022-01-15 manuscript/appendix.Rmd
update, mostly Fira Code

@@
 title: "APPENDIX: `{peacesciencer}`: An R Package for Quantitative Peace Science Research"
-# author:
-# - name: Steven V. Miller
-#   affiliation: Clemson University
-#   email: svmille@clemson.edu
+author:
+- name: Steven V. Miller
+  affiliation: Clemson University
+  email: svmille@clemson.edu
 # thanks: "This is a work in progress. Feedback welcome: svmille@clemson.edu."
@@
 sansfont: Linux Biolinum O
+monofont: Fira Code
 sansitup: FALSE

### 2022-01-29 docs/ms.tex
gonna eventually send this to Sage.

@@
   \setmainfont[]{cochineal}
+  \setmonofont[]{Fira Code}
 \fi
@@
 \title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: November 08, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
+\textbf{Current version}: January 29, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
 \author{true}
@@
 \title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: November 08, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
+\textbf{Current version}: January 29, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
 
@@
 \usepackage{booktabs}
+\usepackage{csquotes}
 \LTcapwidth=.95\textwidth
@@
 \usepackage{setspace}
-\let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
-\usepackage{colortbl}
-\arrayrulecolor{black}
 \usepackage{booktabs}
@@
 % 
+% Optional code chunks ------
+% SOURCE: https://stackoverflow.com/questions/50702942/does-rmarkdown-allow-captions-and-references-for-code-chunks
+
 \begin{document}
@@
 The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular \texttt{\{tidvyerse\}} packages.
-Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year or state-year analysis (e.g.~democracy data, contiguity data).
+Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate common types of data from scratch and to merge in ubiquitous indicators included in almost every analysis (e.g.~democracy data, contiguity data).
 The software is freely available on CRAN and maintains an active website documenting its features at \url{http://svmiller.com/peacesciencer}.}
@@
 
-This data feature tackles a recurring problem for researchers in the peace science community.
-True research reproducibility is achieved creating original dyad-year or state-year data from scratch, though no published guide exists that informs researchers how to do this on their own.
+This manuscript tackles a recurring problem for researchers in the peace science community.
+True research reproducibility is best achieved creating data from scratch, though no published guide exists that informs researchers how to do this on their own.
 Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice.
-Researchers may additionally spend too much time reproducing old code for standard information that goes into any dyad-year or state-year analysis---like contiguity relationships and democracy---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes.
+Researchers may additionally spend too much time reproducing old code for standard information that goes into any dyadic or monadic analysis---like contiguity relationships and democracy---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes.
 This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well.
-Graduate students, in particular, face unique challenges associated with these developments.
-Graduate students beginning their careers in peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.
+Graduate students and other beginners in the field face unique challenges associated with these developments.
+Students just learning peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.
 
@@
 Built around the free and open source R programming language, \texttt{\{peacesciencer\}} contains a suite of data and functions for creating data of interest to researchers.
-Researchers can use \texttt{\{peacesciencer\}} to create dyad-year, state-year, and even state-day data from scratch.
+Researchers can use \texttt{\{peacesciencer\}} to create dyad-year, leader-year, leader-dyad-year, and state-year data (among some others) from scratch.
 Afterward, they can add a variety of standard information (e.g.~contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command.
 This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want.
-\texttt{\{peacesciencer\}} comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts.
+\texttt{\{peacesciencer\}} comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing conflicts.
 All are done with the maximum possible transparency.
-The project is available for public view of Github (\url{https://github.com/svmiller/peacesciencer/}).
-The \texttt{data-raw} directory on the project's Github contains information and comments about how every data set was created and the \texttt{R} directory informs the user how each function works.
+The project is available for public view on Github (\url{https://github.com/svmiller/peacesciencer/}).
+The \texttt{data-raw} directory on the project's Github contains information and comments about how every data set was created.
 The function manuals (\url{http://svmiller.com/peacesciencer/reference}) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing.
@@
 This data feature proceeds in the following fashion.
-The next section expands what need this package fills for peace scientists in the present.
+The next section expands what need this package fills for peace scientists.
 Afterward, it provides an overview of what is included in \texttt{\{peacesciencer\}} to help researchers more quickly conduct the kind of quantitative research they want.
 Thereafter, it provides a tutorial on how to install and best use \texttt{\{peacesciencer\}} in the R programming language.
-A more comprehensive tutorial follows, showing how \texttt{\{peacesciencer\}} already has a suite of data and functions that can allow for effective replications of a ``dangerous dyads'' type analysis (Bremer, 1992) and standard state-year analyses of civil conflict onset (e.g.
-Fearon \& Laitin, 2003).
-This feature concludes with a discussion of what else \texttt{\{peacesciencer\}} is capable of doing in future updates and how this R package can inform more reasoned design decisions for researchers in peace science.
+A more comprehensive tutorial follows, showing how \texttt{\{peacesciencer\}} already has a suite of data and functions that can allow for effective replications of a ``dangerous dyads'' type analysis (Bremer, 1992), standard state-year analyses of civil conflict onset (e.g.
+Fearon \& Laitin, 2003), and even leader-year analyses of inter-state dispute initiation (e.g.
+Horowitz \& Stam, 2014).
+This feature concludes with a comparison of \texttt{\{peacesciencer\}} with other alternatives and a discussion of how \texttt{\{peacesciencer\}} can inform more reasoned design decisions for researchers in peace science.
 
@@
 
-\texttt{\{peacesciencer\}} is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch.
-NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars (Bennett, Poast \& Stam, 2019).
-NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade (Bennett \& Stam, 2000).
-No matter, \texttt{\{peacesciencer\}} is motivated by the following ideals that led to its creation.
-
+\texttt{\{peacesciencer\}} is motivated by the following observations and ideals that led to its creation.
 For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis.
@@
 No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question.
-After all, EUGene's primary impetus was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g.~dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit.
-\texttt{\{peacesciencer\}} is primarily born from this question about how to create dyad-year and state-year data from scratch.
+After all, EUGene's main value---if not its primary impetus---was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g.~dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit.
+\texttt{\{peacesciencer\}} is primarily born from this question about how to create these data from scratch.
 The underlying code that produces these data types is available online and \texttt{\{peacesciencer\}} converts these lines of code into simple functions for the ease of the researcher.
@@
 Second, researchers also invest too much time in retracing steps for peace science analyses for new projects.
-Assume a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data).
+Assume a researcher finished a state-year analysis on the correlates of civil conflict onset a few years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data).
 Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project.
@@
 Software options exist that allow the researcher to load data, create data, clean data, analyze data, and present the results of the analysis all within one program.
-\texttt{\{peacesciencer\}}, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and Rstudio (its free-for-use integrated desktop environment {[}IDE{]}).
+\texttt{\{peacesciencer\}}, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and RStudio (its free-for-use integrated development environment {[}IDE{]}).
 
@@
 This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that \texttt{\{peacesciencer\}} uses.
-There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet (Arel-Bundock, 2021a), scrape \emph{any} type of data from the internet (Boehmke, 2016), and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero (Arel-Bundock, 2021b).
+There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet (Arel-Bundock, 2021a) and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero (Arel-Bundock, 2021b).
 \texttt{\{peacesciencer\}} embraces this.
 This R package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.
-
-Fifth, statistical software should be kind to graduate students in peace science with the aforementioned points in mind.
-Graduate students just beginning their career in peace science scholarship are tasked to learn the substantive material, the characteristics of the underlying data in wide use in the community, and the quantitative methods and statistical software available to them, effectively all at the same time.
-A student interested in civil conflict scholarship will need to learn the UCDP-PRIO armed conflict data in some detail at the same time they learn about the theoretical arguments that informed the creation of the data, likely at the same time they take quantitative methods courses that show them how to load and explore data in a chosen statistical software.
-At the same time that creating data is increasingly blurred with coding and processing data, students are simultaneously learning to explore data for themselves as they read about the analyses of the data.
-\texttt{\{peacesciencer\}} aims to make this process easier for graduate students using the features of the R programming language.
-Graduate students can use \texttt{\{peacesciencer\}} to learn about peace science data and computer programming in R at the same time.
 
@@
 The project's Github contains a publicly available \texttt{data-raw} directory that shows how every data set included (and processed) in \texttt{\{peacesciencer\}} was created.
-The function manuals included \texttt{\{peacesciencer\}} contain ample documentation that clarifies what the function is doing, what it returns to the user, and why it is doing it this way.
+The function manuals included \texttt{\{peacesciencer\}} contain ample documentation that clarify what each function is doing, what it returns to the user, and why it is doing it this way.
 Researchers can also use the project's Github to point out bugs, ask for further clarification, and propose additions.
@@
 
-\texttt{\{peacesciencer\}} comes with a fully developed suite of built-in functions for generating the most widespread forms of peace science data.
-Each function uses raw or pre-processed data included in \texttt{\{peacesciencer\}}.
-For example, \texttt{create\_dyadyears()} transforms the raw state system membership data for either the Correlates of War or Gleditsch-Ward system into the full universe of post-1816 dyad-years.
-\texttt{add\_gml\_mids()} uses a dyadic dispute-year version of the MID data offered by Gibler, Miller \& Little (2016), which, as the \texttt{data-raw} directory on the project's Github shows, is itself derived from the dispute-level and participant-level versions of the data.
-\texttt{\{peacesciencer\}} also has some data innovations.
-It has an original data set on capital-to-capital distances, calculating for distance in kilometers between state capitals using the ``Vincenty'' method (i.e.~``as the crow flies'') that also accounts for instances when capitals moved (e.g.~Brazil in 1960, Burundi in 2018).
-It also features an innovation of democracy data, providing reasonable estimates of democracy using the Marquez (2016) method of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve \& Melton, 2010).\footnote{\texttt{\{peacesciencer\}} comes with a vignette that explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.} As of writing, \texttt{\{peacesciencer\}} comes with the following functionality described in Table \ref{tab:tab-functionality}.
-
-\begin{table}
-
-\caption{\label{tab:tab-functionality}Features of \{peacesciencer\}}
-\centering
-\fontsize{8}{10}\selectfont
-\begin{tabular}[t]{>{}l|>{}l|>{}c|>{}c}
-\hline
-\textbf{Data} & \textbf{Description} & \textbf{Type} & \textbf{Function}\\
-\hline
-(CoW, G-W) Dyad-Year Data & Create dyad-years from state system membership data &  & \ttfamily{create\_dyadyears()}\\
-\arrayrulecolor{white}
-\hline
-(CoW, G-W) State-Year Data & Create state-years from state system membership data &  & \ttfamily{create\_stateyears()}\\
-\arrayrulecolor{white}
-\hline
-(CoW, G-W) State-Day Data & Create state-days from state system membership data &  & \ttfamily{create\_statedays()}\\
-\arrayrulecolor{white}
-\hline
-Archigos (v. 4.1) & Add Archigos political leader information & D, S & \ttfamily{add\_archigos()}\\
-\arrayrulecolor{white}
-\hline
-ATOP Alliances (v. 5.0) & Add ATOP alliance data & D & \ttfamily{add\_atop\_alliance()}\\
-\arrayrulecolor{white}
-\hline
-Capital-to-Capital Distances & Add capital-to-capital distance & D, S & \ttfamily{add\_capital\_distance()}\\
-\arrayrulecolor{white}
-\hline
-Convert CoW codes to G-W codes & Add matching CoW codes for G-W states by year & D, S & \ttfamily{add\_ccode\_to\_gw()}\\
-\arrayrulecolor{white}
-\hline
-Convert G-W codes to CoW codes & Add matching G-W codes for CoW states by year & D, S & \ttfamily{add\_gwcode\_to\_cow()}\\
-\arrayrulecolor{white}
-\hline
-CoW National Capabilities (v. 5.0) & Adds capabilities of states by year & D, S & \ttfamily{add\_nmc()}\\
-\arrayrulecolor{white}
-\hline
-CoW Direct Contiguity (v. 3.2) & Adds contiguity information by year & D, S & \ttfamily{add\_contiguity()}\\
-\arrayrulecolor{white}
-\hline
-CoW Alliances (v. 4.1) & Adds CoW alliance data by year & D & \ttfamily{add\_cow\_alliance()}\\
-\arrayrulecolor{white}
-\hline
-CoW IGOs (v. 3.0) & Adds CoW IGO data by year & D, S & \ttfamily{add\_igos()}\\
-\arrayrulecolor{white}
-\hline
-CoW Major Powers (v. 2016) & Adds CoW major power by year & D, S & \ttfamily{add\_cow\_majors()}\\
-\arrayrulecolor{white}
-\hline
-CoW MIDs (v. 5.0) & Add CoW MID information to data & D & \ttfamily{add\_cow\_mids()}\\
-\arrayrulecolor{white}
-\hline
-CoW Trade Data (v. 4.0) & Add CoW trade information to data & D, S & \ttfamily{add\_cow\_trade()}\\
-\arrayrulecolor{white}
-\hline
-CoW Wars & Add CoW inter-state (v. 4.0) and intra-state (v. 4.1) war data & D & \ttfamily{add\_cow\_wars()}\\
-\arrayrulecolor{white}
-\hline
-CREG (v. 1.02) & Add ethnic/religious fractionalization/polarization by year & D, S & \ttfamily{add\_creg\_fractionalization()}\\
-\arrayrulecolor{white}
-\hline
-Democracy & Add Polity (v. 2017), V-Dem (v. 10), and Marquez' (2016) UDS extensions & D,S & \ttfamily{add\_democracy()}\\
-\arrayrulecolor{white}
-\hline
-GML MIDs (v. 2.1.1) & Add Gibler-Miller-Little MIDs to data & D & \ttfamily{add\_gml\_mids()}\\
-\arrayrulecolor{white}
-\hline
-Minimum Distance (v. 2016) & Adds minimum distance data by year & D, S & \ttfamily{add\_minimum\_distance()}\\
-\arrayrulecolor{white}
-\hline
-Peace Years & Calculate peace years for various disputes & D, S & \ttfamily{add\_peace\_years()}\\
-\arrayrulecolor{white}
-\hline
-Rugged/Mountainous Terrain & Add 'ruggedness'/'mountainous' terrain data by year & D, S & \ttfamily{add\_rugged\_terrain()}\\
-\arrayrulecolor{white}
-\hline
-(Surplus/Gross) Domestic Product & Add Anders et al. (2020) GDP/SDP simulations by year & D, S & \ttfamily{add\_sdp\_gdp()}\\
-\arrayrulecolor{white}
-\hline
-Strategic Rivalries & Add Thompson and Dreyer (2012) strategic rivalry data by year & D, S & \ttfamily{add\_strategic\_rivalries()}\\
-\arrayrulecolor{white}
-\hline
-UCDP Armed Conflicts (v. 20.1) & Add UCDP armed conflict data by year & S & \ttfamily{add\_ucdp\_acd()}\\
-\arrayrulecolor{black}
-\hline
-\end{tabular}
-\end{table}
+\texttt{\{peacesciencer\}} comes with a fully developed suite of built-in functions for generating some of the most widespread forms of peace science data and populating the data with important variables that recur in many quantitative analyses.
+The core functionality of \texttt{\{peacesciencer\}} reduces to two broad categories of functions.
+These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis.\footnote{In both cases, SQL-like ``joins''---especially the left (outer) join---are the primary workhorse in these functions.
+A vignette on the package website, included in the appendix, explains these various joins, why they are used, and what issues may arise from them.} Table \ref{tab:tab-createfunctions} and Table \ref{tab:tab-addfunctions} list these core functions as of version 1.0, the version of this package slated for release alongside the manuscript.\footnote{\texttt{\{peacesciencer\}} also has miscellaneous ``helper'' functions that are not belabored in this manuscript.
+These include functions that allow the user to use \texttt{\{peacesciencer\}} functions with data created outside the package (\texttt{declare\_attributes()}), download external data for quicker use of other functions (\texttt{download\_extdata()}), and functions that communicate suggested citations (\texttt{ps\_cite()}), among a few other functions.
+See the package's website for more information (\url{http://svmiller.com/peacesciencer}).}
+
+Table \ref{tab:tab-createfunctions} lists functions that create base data frames for a researcher starting an original project.
+They serve as functions that communicate the units of analysis supported in this package and that the package is capable of generating for an interested user.
+For example, \texttt{create\_stateyears()} will generate the full universe of state-years from the Correlates of War (Correlates of War, 2011: v.
+2016) or Gleditsch-Ward (Gleditsch \& Ward, 1999: v.
+2017) system, encompassing all state-years to the most recently concluded calendar year, depending on the arguments supplied to the user in the function.
+\texttt{create\_leaderyears()} will generate the full universe of leader-years from the Archigos leader data (Goemans, Gleditsch \& Chiozza, 2009: v.
+4.1), optionally standardizing leader-years to the Gleditsch-Ward or Correlates of War state system data.
+As of version 1.0, \texttt{\{peacesciencer\}} is capable of creating full dyad-year data, leader-day data, leader-dyad-year data, leader-year data, state-day data, and state-year data.\footnote{Leader-dyad-year data are too time-consuming to calculate each time and too size-prohibitive to include in the package.
+These data can be downloaded through \texttt{download\_extdata()} for much easier use.} A vignette on the package's website shows how users can create other forms of data from these functions as well (e.g.~dyadic-dispute-year, leader-months, state-quarters).
+
+\begin{longtable}[t]{>{}ll}
+\caption{\label{tab:tab-createfunctions}Functions that Create Base Data}\\
+\toprule
+\textbf{Function} & \textbf{Description}\\
+\midrule
+\ttfamily{create\_dyadyears()} & Create dyad-years from state system membership data\\
+\ttfamily{create\_leaderdays()} & Create leader-day data from Archigos\\
+\ttfamily{create\_leaderdyadyears()} & Create leader-dyad-years data from Archigos\\
+\ttfamily{create\_leaderyears()} & Create leader-years data from Archigos\\
+\ttfamily{create\_statedays()} & Create state-days from state system membership data\\
+\ttfamily{create\_stateyears()} & Create state-years from state system membership data\\
+\bottomrule
+\end{longtable}
+
+Table \ref{tab:tab-addfunctions} lists the main functions in \texttt{\{peacesciencer\}} that add information or subset the number of rows of the data to just those of interest to the user, describing these functions and listing whether they are applicable to dyad-year (\texttt{D}), leader-year (\texttt{L}), leader-dyad-year (\texttt{LD}), state-year (\texttt{S}) data or specialty functions applicable to just the dyadic conflict data (\texttt{C}).\footnote{The ``whittle'' (\texttt{wc\_}) class of functions is a suite of functions that whittle dyadic dispute-years to true dyad-year data.
+They are effectively used in \texttt{add\_cow\_mids()} and \texttt{add\_gml\_mids()} for dyad-year data, but are offered here to allow users to employ their own case exclusion rules to create their own dyad-year conflict data.
+A vignette on \texttt{\{peacesciencer\}}'s website explains these case exclusion rules in some detail.} All these functions use raw or pre-processed data included in the package.
+For example, \texttt{add\_gml\_mids()} uses a dyadic dispute-year version of the MID data offered by Gibler et al.
+(2016: v.
+2.2.1) and merges in information about whether there was an ongoing MID or MID onset in a dyad-year, leader-year, leader-dyad-year, or state-year.
+\texttt{\{peacesciencer\}} also has some data innovations included in these functions.
+For example, \texttt{add\_capital\_distance()} calculates distance between state capitals in kilometers using the Vincenty method (i.e.~``as the crow flies'') based on an original data set of state capitals that accounts for instances when capitals moved (e.g.~Brazil in 1960, Burundi in 2018).
+\texttt{add\_democracy()} does more than just add Polity data to a data set.
+The data underpinning the function feature an innovation in democracy data, providing reasonable estimates of democracy using the Marquez (2016) method of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve \& Melton, 2010) in addition to Polity estimates (Marshall, Gurr \& Jaggers, 2017: v.
+2017) and V-Dem estimates (Coppedge et al., 2020: v.
+10).\footnote{\texttt{\{peacesciencer\}} comes with a vignette that explains just how much unnecessary missingness pervades data on democracy in the absence of an innovation like this, and how this missingness might adversely affect inferences of inter-state or intra-state conflict.}
+
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{longtable}[t]{>{}lc>{\raggedright\arraybackslash}p{20em}}
+\caption{\label{tab:tab-addfunctions}Functions that Add Data}\\
+\toprule
+\textbf{Function} & \textbf{Type} & \textbf{Description}\\
+\midrule
+\ttfamily{add\_archigos()} & D,S & Add Archigos political leader information to a data frame\\
+\ttfamily{add\_atop\_alliance()} & D, LD & Add Alliance Treaty Obligations and Provisions (ATOP) alliance data to a data frame\\
+\ttfamily{add\_capital\_distance()} & D, L, LD, S & Add capital-to-capital distance to a data frame\\
+\ttfamily{add\_ccode\_to\_gw()} & D, L, LD, S & Match CoW state codes to G-W state codes\\
+\ttfamily{add\_contiguity()} & D, L, LD, S & Add CoW direct contiguity information to a data frame\\
+\ttfamily{add\_cow\_alliance()} & D, LD & Add CoW alliance data to a data frame\\
+\ttfamily{add\_cow\_majors()} & D, L, LD, S & Add CoW major power information to a data frame\\
+\ttfamily{add\_cow\_mids()} & D & Add CoW-MID data to a dyad-year data frame\\
+\ttfamily{add\_cow\_trade()} & D, L, LD, S & Add CoW trade data to a data frame\\
+\ttfamily{add\_cow\_wars()} & D, S & Add CoW war data to a dyad-year or state-year data frame\\
+\ttfamily{add\_creg\_fractionalization()} & D, L, LD, S & Add fractionalization/polarization estimates from CREG to a data frame\\
+\ttfamily{add\_democracy()} & D, L, LD, S & Add democracy information to a data frame\\
+\ttfamily{add\_fpsim()} & D, LD & Add dyadic foreign policy similarity measures to a data frame\\
+\ttfamily{add\_gml\_mids()} & D, L, LD, S & Add GML MID data to a data frame\\
+\ttfamily{add\_gwcode\_to\_cow()} & D, L, LD, S & Match G-W state codes to CoW state codes\\
+\ttfamily{add\_igos()} & D, S & Add CoW IGO data to a data frame\\
+\ttfamily{add\_lead()} & L, LD & Add leader experience/attributes (LEAD) to data\\
+\ttfamily{add\_lwuf()} & L, LD & Add leader willingness to use force estimates to data\\
+\ttfamily{add\_minimum\_distance()} & D, L, LD, S & Add minimum distance data to data frame\\
+\ttfamily{add\_nmc()} & D, L, LD, S & Add Correlates of War National Military Capabilities data\\
+\ttfamily{add\_peace\_years()} & D, S & Add 'peace years' to dyad-year/state-year conflict data\\
+\ttfamily{add\_rugged\_terrain()} & D, L, LD, S & Add rugged terrain information to a data frame\\
+\ttfamily{add\_sdp\_gdp()} & D, L, LD, S & Add (surplus, gross) domestic product data to a data frame\\
+\ttfamily{add\_spells()} & D, L, LD, S & Add duration 'spells' to a data frame\\
+\ttfamily{add\_strategic\_rivalries()} & D, S & Add strategic rivalry information to a data frame\\
+\ttfamily{add\_ucdp\_acd()} & S & Add UCDP Armed Conflict Data to state-year data frame\\
+\ttfamily{add\_ucdp\_onsets} & S & Add UCDP onsets to state-year data\\
+\ttfamily{filter\_prd()} & D, LD & Filter dyad-year data to just politically relevant dyads\\
+\ttfamily{wc\_duration()} & C & Whittle Duplicate Conflict-Years by (Minimum or Maximum) Conflict Duration\\
+\ttfamily{wc\_fatality()} & C & Whittle Duplicate Conflict-Years by Highest Fatality\\
+\ttfamily{wc\_hostility()} & C & Whittle Duplicate Conflict-Years by Highest Hostility\\
+\ttfamily{wc\_jds()} & C & Whittle Duplicate Conflict-Years by Just Dropping Something\\
+\ttfamily{wc\_onsets()} & C & Whittle Uniqute Conflict Onset-Years from Conflict-Year Data\\
+\ttfamily{wc\_recip()} & C & Whittle Duplicate Conflict-Years by Conflict Reciprocation\\
+\ttfamily{wc\_stmon()} & C & Whittle Duplicate Conflict-Years by Lowest Start Month\\
+\bottomrule
+\end{longtable}
+\endgroup{}
 
@@
 Arel-Bundock, 2021a).
-Its coverage will assuredly expand with new additions of interest to the peace science community.
+Its coverage will assuredly expand with new additions of interest to the peace science community, though the package already offers a lot to meet researcher needs.\footnote{The version of \texttt{\{peacesciencer\}} slated for release includes the most up-to-date-possible versions of data sets referenced in this manuscript.
+One helper function, \texttt{ps\_version()}, will inform the user about the versions of the data included in the installed package.
+\texttt{\{peacesciencer\}} promises prompt package updates as new versions become available.}
 
@@
 This assumes at least some familiarity with the R programming language.
-Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.0.
+Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.2.
 \texttt{\{peacesciencer\}} is designed to be as user-friendly as possible.
@@
 
-\texttt{\{peacesciencer\}}'s functions work out of the box, though it works best with one additional package.
-This is \texttt{\{tidyverse\}}, itself a suit of packages that share a common form and design (Wickham \& Grolemund, 2017).
-\texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}} but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality.
-A user can open an R session and install both packages as follows.
-
-\begin{Shaded}
-\begin{Highlighting}[]
+\texttt{\{peacesciencer\}}'s functions work out of the box, though users should find their experience augmented by two additional downloads.
+First, RStudio offers an IDE that serves as a user-friendly graphical user interface (GUI) over what is, at its core, a programming language with a command-line interface.
+RStudio's design will make it much easier for users to experiment with \texttt{\{peacesciencer\}}'s functionality and read its documentation to assist them with the use of these functions.
+The second additional download is \texttt{\{tidyverse\}}, itself a suite of packages that share a common form and design (Wickham et al., 2019).
+\texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}}, and \texttt{\{peacesciencer\}} can work without it, but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality.
+A user can open an R session by way of RStudio and install both packages as follows.\footnote{R users will recognize that pound signs (\texttt{\#}) allow for commenting code.
+Users new to R should understand its use here is primarily for exposition of what the code snippets in this manuscript are doing.
+In subsequent code snippets, \texttt{\#\textgreater{}} will indicate console output the user can expect from these functions.}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Install the packages for use}
 \FunctionTok{install.packages}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"tidyverse"}\NormalTok{, }\StringTok{"peacesciencer"}\NormalTok{))}
@@
 
-R packages once installed need to be loaded with every R session. The user can load both with the \texttt{library()} function in R.
-
-\begin{Shaded}
-\begin{Highlighting}[]
+R packages once installed need to be loaded with every R session (i.e.~every time the user opens RStudio). The user can load both with the \texttt{library()} function in R.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Load the packages for use}
 \FunctionTok{library}\NormalTok{(tidyverse)}
@@
 
-\texttt{\{peacesciencer\}} has a simple ``pipe''-based workflow.
-The top of the ``pipe'' is the first function to create the data.
-These options include \texttt{create\_dyadyears()}, \texttt{create\_stateyears()}, and \texttt{create\_statedays()}.
-All options include two arguments: \texttt{system} and \texttt{mry}.
-\texttt{system} takes one of two values---``cow'' or ``gw''---for what state system membership data the user wants.
-If \texttt{system} is not specified, the function defaults to Correlates of War system membership data given the prominence of the Correlates of War project in the conflict data ecosystem.
-\texttt{mry}---short for ``most recent year''---determines whether or not the function extends the membership data to the most recently concluded calendar year.
-The function defaults to \texttt{TRUE} in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to (as of writing) 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017).
-\texttt{create\_dyadyears()} has a third argument---\texttt{directed}---that determines whether the user wants directed dyad-year data or non-directed dyad-year data.
-The function defaults to \texttt{TRUE}, returning directed dyad-year data.
-
-The following example creates a full state-year data frame of Gleditsch-Ward states from 1816 to 2020.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{create\_stateyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-The below example would create a full non-directed dyad-year data frame of Correlates of War states from 1816 to 2016, the last year of record in the Correlates of War state system membership data.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{directed =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{mry =} \ConstantTok{FALSE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-This example would create a full state-day data frame from the Gleditsch-Ward system for all states from Jan.~1, 1816 to Dec.~31, 2020.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{create\_statedays}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-There are not too many data sets in the peace science community with daily data, certainly dating to the 19th century, but \texttt{\{peacesciencer\}} can greatly reduce the costs of creating the kind of data the user may want.
-
-\hypertarget{add-standard-information-to-dyad-year-or-state-year-data}{%
-\subsection{Add Standard Information to Dyad-Year or State-Year Data}\label{add-standard-information-to-dyad-year-or-state-year-data}}
-
-The next step in the \texttt{\{peacesciencer\}} ``pipe''-based workflow will also be among the biggest time-savers for researchers using it.
-Researchers who want to understand, say, the correlates of inter-state conflict onset almost always do the following.
-They will add the dependent variables from an inter-state conflict data set, like the Gibler-Miller-Little (GML) dispute data set (Gibler, Miller \& Little, 2016).
-They will add several indicators to the dyad-year data that are well-traveled in conflict scholarship, including indicators of things like joint alliance contracts and the contiguity relationship of both states in the dyad.
-\texttt{\{peacesciencer\}} has great coverage for a lot of these data and can even allow almost full-scale replications with the data it offers.
-
-For example, assume a researcher wanted to replicate something analogous to Bremer's (1992) famous ``dangerous dyads'' analysis about the causes of conflict onset.
-\texttt{\{peacesciencer\}} can do something like this in the following pipe-based workflow.
-First, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)} will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016.\footnote{The conflict data in Gibler, Miller \& Little (2016) do not extend past 2010, but asking for dyad-years from 2010-2016 will just create harmless \texttt{NA}s that will not appear in any analysis.}
-
-After creating the base data they want, the user can add data to it with the ``pipe'' operator.
-The \texttt{\{tidyverse\}} represents this pipe operator as \texttt{\%\textgreater{}\%}.
-Ending each line with this pipe passes forward an object (here: data) into the next function on the next line.
-Put in other words, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)\ \%\textgreater{}\%\ filter\_prd()} is a command in \texttt{\{tidyverse\}} and \texttt{\{peacesciencer\}} to create all non-directed dyad-years from 1816 to 2016 and then subset those data to only the politically relevant dyad-years.
-Because politically relevant dyads depend on information about the major power status for both states as well as their contiguity relationship (Lemke \& Reed, 2001), the function adds in those data (i.e.~quietly executing \texttt{add\_contiguity()} and \texttt{add\_cow\_majors()}) before subsetting the data to just the dyads that have some kind of contiguity relationship and/or have a major power in it.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{directed =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{mry =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+I encourage users who are using \texttt{\{peacesciencer\}} for the first time, especially those who are learning R for the first time because of their interest in this package, to approach \texttt{\{peacesciencer\}} with an idea of the kind of data they want to create for the sake of a project.
+The core functionality of \texttt{\{peacesciencer\}} begins with the creation of a data frame, which can then be populated with various indicators of interest.
+No matter, the suggested use of \texttt{\{peacesciencer\}} begins with the creation of a data frame.
+To start, assume a new user without much familiarity with the R programming language installed RStudio, \texttt{\{tidyverse\}}, and \texttt{\{peacesciencer\}} with the idea of using \texttt{\{peacesciencer\}} to help them start a new research project that seeks to explain civil conflict onset across state-years.
+Toward that end, they have identified their unit of analysis is state-year and can be created with the \texttt{create\_stateyears()}.
+To get started, I encourage entering the following command in the console in RStudio.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# See documentation for create\_stateyears() function.}
+\NormalTok{?}\FunctionTok{create\_stateyears}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+This will open a documentation file for this function, which is itself quite verbose and informative for the user about what the function is doing.
+In this case, the documentation file will show there are three arguments in this function, each with built-in defaults.
+This function will allow the user to choose a state system for which they want state-years (\texttt{system}, which accepts either ``cow'' or ``gw'' for Correlates of War {[}CoW{]} state system data or Gleditsch-Ward {[}G-W{]} state system data and defaults to ``cow''), whether they want to extend the state system to the most recently concluded calendar year (\texttt{mry}, which defaults to TRUE), and whether they may want to additionally subset the years to a more narrow temporal domain they may already have in mind (\texttt{subset\_years}, which defaults to no subset of the data, returning all possible state-years).
+If the user simply ran the function with no overrides, the function would return all Correlates of War state-years from 1816 to the most recent year (2021, as of writing).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Create all CoW state{-}years from 1816 to most recently concluded calendar year.}
+\FunctionTok{create\_stateyears}\NormalTok{()}
+\CommentTok{\#\textgreater{} \# A tibble: 16,926 x 3}
+\CommentTok{\#\textgreater{}    ccode statenme                  year}
+\CommentTok{\#\textgreater{}    \textless{}dbl\textgreater{} \textless{}chr\textgreater{}                    \textless{}int\textgreater{}}
+\CommentTok{\#\textgreater{}  1     2 United States of America  1816}
+\CommentTok{\#\textgreater{}  2     2 United States of America  1817}
+\CommentTok{\#\textgreater{}  3     2 United States of America  1818}
+\CommentTok{\#\textgreater{}  4     2 United States of America  1819}
+\CommentTok{\#\textgreater{}  5     2 United States of America  1820}
+\CommentTok{\#\textgreater{}  6     2 United States of America  1821}
+\CommentTok{\#\textgreater{}  7     2 United States of America  1822}
+\CommentTok{\#\textgreater{}  8     2 United States of America  1823}
+\CommentTok{\#\textgreater{}  9     2 United States of America  1824}
+\CommentTok{\#\textgreater{} 10     2 United States of America  1825}
+\CommentTok{\#\textgreater{} \# ... with 16,916 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+A user who approaches \texttt{\{peacesciencer\}} with a project in mind will see they can better tailor this function to what they want.
+Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the UCDP armed conflict data.
+They will also know they have no use for pre-1946 observations since civil conflict data typically have coverage only from 1946 forward.
+Thus, the user can supply some additional arguments to tailor the creation of data to just what they want (here: all Gleditsch-Ward state-years from 1946 to 2019).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Create all G{-}W state years from 1946 to 2019.}
+\FunctionTok{create\_stateyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{, }\AttributeTok{subset\_years =} \FunctionTok{c}\NormalTok{(}\DecValTok{1946}\SpecialCharTok{:}\DecValTok{2019}\NormalTok{))}
+\CommentTok{\#\textgreater{} \# A tibble: 10,490 x 3}
+\CommentTok{\#\textgreater{}    gwcode statename                 year}
+\CommentTok{\#\textgreater{}     \textless{}dbl\textgreater{} \textless{}chr\textgreater{}                    \textless{}int\textgreater{}}
+\CommentTok{\#\textgreater{}  1      2 United States of America  1946}
+\CommentTok{\#\textgreater{}  2      2 United States of America  1947}
+\CommentTok{\#\textgreater{}  3      2 United States of America  1948}
+\CommentTok{\#\textgreater{}  4      2 United States of America  1949}
+\CommentTok{\#\textgreater{}  5      2 United States of America  1950}
+\CommentTok{\#\textgreater{}  6      2 United States of America  1951}
+\CommentTok{\#\textgreater{}  7      2 United States of America  1952}
+\CommentTok{\#\textgreater{}  8      2 United States of America  1953}
+\CommentTok{\#\textgreater{}  9      2 United States of America  1954}
+\CommentTok{\#\textgreater{} 10      2 United States of America  1955}
+\CommentTok{\#\textgreater{} \# ... with 10,480 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+Another reader may be interested in using \texttt{\{peacesciencer\}} for a research project using a leader-year unit of analysis.
+These data can be created with \texttt{create\_leaderyears()}.
+Users can read more about what this function is doing by consulting the documentation file in R with the following command.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# See documentation for create\_leaderyears() function.}
+\NormalTok{?}\FunctionTok{create\_leaderyears}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+This function has three arguments denoting the leader system to inform the creation of the leader-year data (i.e.~Archigos) and two other arguments: \texttt{standardize} and \texttt{subset\_years}.
+The \texttt{standardize} argument, which when it defaults to ``none'', returns all leader-years as presented in the raw Archigos data.
+Archigos' leader data are nominally denominated in the Gleditsch-Ward state system data, if not necessarily Gleditsch-Ward state system dates (e.g.~Archigos often has leader entries prior to state system entry in a few cases).
+Thus, the user can standardize leader-year data to Gleditsch-Ward state system dates of CoW state system dates.
+Finally, the user can subset the leader-year data to a more narrow temporal domain that may interest them with the \texttt{subset\_years} function.
+If a user is interested in creating a data set of leader-years for an analysis of inter-state conflict initiation with the GML MID data, they can create the data that interests them with the following function.
+Notice how \texttt{create\_leaderyears()} also returns information about the leader too, like the leader's approximate age that year, their gender, and information about their tenure.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Create all leader{-}years for CoW states from 1870 to 2010.}
+\FunctionTok{create\_leaderyears}\NormalTok{(}\AttributeTok{standardize =} \StringTok{"cow"}\NormalTok{, }\AttributeTok{subset\_years =} \FunctionTok{c}\NormalTok{(}\DecValTok{1870}\SpecialCharTok{:}\DecValTok{2010}\NormalTok{))}
+\CommentTok{\#\textgreater{} \# A tibble: 15,074 x 7}
+\CommentTok{\#\textgreater{}    obsid    ccode leader gender  year yrinoffice leaderage}
+\CommentTok{\#\textgreater{}    \textless{}chr\textgreater{}    \textless{}dbl\textgreater{} \textless{}chr\textgreater{}  \textless{}chr\textgreater{}  \textless{}dbl\textgreater{}      \textless{}dbl\textgreater{}     \textless{}dbl\textgreater{}}
+\CommentTok{\#\textgreater{}  1 USA{-}1869     2 Grant  M       1870          2        48}
+\CommentTok{\#\textgreater{}  2 USA{-}1869     2 Grant  M       1871          3        49}
+\CommentTok{\#\textgreater{}  3 USA{-}1869     2 Grant  M       1872          4        50}
+\CommentTok{\#\textgreater{}  4 USA{-}1869     2 Grant  M       1873          5        51}
+\CommentTok{\#\textgreater{}  5 USA{-}1869     2 Grant  M       1874          6        52}
+\CommentTok{\#\textgreater{}  6 USA{-}1869     2 Grant  M       1875          7        53}
+\CommentTok{\#\textgreater{}  7 USA{-}1869     2 Grant  M       1876          8        54}
+\CommentTok{\#\textgreater{}  8 USA{-}1869     2 Grant  M       1877          9        55}
+\CommentTok{\#\textgreater{}  9 USA{-}1877     2 Hayes  M       1877          1        55}
+\CommentTok{\#\textgreater{} 10 USA{-}1877     2 Hayes  M       1878          2        56}
+\CommentTok{\#\textgreater{} \# ... with 15,064 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+Researchers using \texttt{\{peacesciencer\}} to create data for their research project should start with one of these ``create'' functions.
+Whether state-year, dyad-year, or a leader-level analysis, these functions will create the full universe of cases of interest to a researcher for the unit of analysis that serves as the core of their project.
+
+\hypertarget{creating-dyad-year-data-and-adding-to-dyad-year-data-in-peacesciencer}{%
+\subsection{\texorpdfstring{Creating Dyad-Year Data and Adding to Dyad-Year Data in \texttt{\{peacesciencer\}}}{Creating Dyad-Year Data and Adding to Dyad-Year Data in \{peacesciencer\}}}\label{creating-dyad-year-data-and-adding-to-dyad-year-data-in-peacesciencer}}
+
+After creating the base data of interest to their project, researchers can begin to add information they want with the suite of functions outlined in Table \ref{tab:tab-addfunctions}.
+For example, a researcher interested in a dyad-year analysis of inter-state disputes can create a non-directed dyad-year data set from 1816 to 2010 in \texttt{\{peacesciencer\}} with the \texttt{create\_dyadyears()} function, one of the aforementioned ``create'' function.
+The following would create the base data of interest to the user (i.e.~all non-directed dyad-years from 1816 to 2010).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Create all non{-}directed dyad{-}years from 1816 to 2010.}
+\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{directed =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{subset\_years =} \FunctionTok{c}\NormalTok{(}\DecValTok{1816}\SpecialCharTok{:}\DecValTok{2010}\NormalTok{))}
+\CommentTok{\#\textgreater{} \# A tibble: 842,655 x 3}
+\CommentTok{\#\textgreater{}    ccode1 ccode2  year}
+\CommentTok{\#\textgreater{}     \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{} \textless{}int\textgreater{}}
+\CommentTok{\#\textgreater{}  1      2     20  1920}
+\CommentTok{\#\textgreater{}  2      2     20  1921}
+\CommentTok{\#\textgreater{}  3      2     20  1922}
+\CommentTok{\#\textgreater{}  4      2     20  1923}
+\CommentTok{\#\textgreater{}  5      2     20  1924}
+\CommentTok{\#\textgreater{}  6      2     20  1925}
+\CommentTok{\#\textgreater{}  7      2     20  1926}
+\CommentTok{\#\textgreater{}  8      2     20  1927}
+\CommentTok{\#\textgreater{}  9      2     20  1928}
+\CommentTok{\#\textgreater{} 10      2     20  1929}
+\CommentTok{\#\textgreater{} \# ... with 842,645 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+Adding information to this data frame is a simple matter of joining a series of functions together in a ``pipe.'' The ``pipe''---represented as \texttt{\%\textgreater{}\%} in the code below---is an operator built into \texttt{\{tidyverse\}} that allows users to pass forward expressions or functions.
+These pipes are common in the programming world and, as the code below will show, have the benefit of changing code in a way that is more intuitive and easier to both read and write for the user.
+While these functions can be modified to work without \texttt{\{tidyverse\}} installed and loaded into the session, the user will find their experience with \texttt{\{peacesciencer\}} is only improved by this important package.
+
+For example, assume the researcher wants just all politically relevant, non-directed dyad-years, where political relevance is traditionally understood as a dyadic relationship involving some form of a contiguity relationship or a major power (Lemke \& Reed, 2001).
+\texttt{create\_dyadyears(directed\ =\ FALSE,\ subset\_years\ =\ c(1816:2010))} created the full universe of non-directed dyad-years from 1816 to 2010, though this full universe includes ``irrelevant'' dyads like Nigeria-Mongolia and Estonia-Rwanda.
+Reducing the data to just politically relevant dyads is simple in \texttt{\{peacesciencer\}} and \texttt{\{tidyverse\}}.
+Users first create the data they want (here: \texttt{create\_dyadyears(directed\ =\ FALSE,\ subset\_years\ =\ c(1816:2010))}), follow it with the pipe operator (\texttt{\%\textgreater{}\%}), and then add another function from \texttt{\{peacesciencer\}} (here: \texttt{filter\_prd()}, which also quietly executes \texttt{add\_contiguity()} and \texttt{add\_cow\_majors()}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# create base data, and pipe (\%\textgreater{}\%) to next function}
+\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{directed =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{subset\_years =} \FunctionTok{c}\NormalTok{(}\DecValTok{1816}\SpecialCharTok{:}\DecValTok{2010}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# subset data to politically relevant dyads (PRDs)}
+  \FunctionTok{filter\_prd}\NormalTok{()}
+\CommentTok{\#\textgreater{} \# A tibble: 112,282 x 7}
+\CommentTok{\#\textgreater{}    ccode1 ccode2  year conttype cowmaj1 cowmaj2   prd}
+\CommentTok{\#\textgreater{}     \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{} \textless{}int\textgreater{}    \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}}
+\CommentTok{\#\textgreater{}  1      2     20  1920        1       1       0     1}
+\CommentTok{\#\textgreater{}  2      2     20  1921        1       1       0     1}
+\CommentTok{\#\textgreater{}  3      2     20  1922        1       1       0     1}
+\CommentTok{\#\textgreater{}  4      2     20  1923        1       1       0     1}
+\CommentTok{\#\textgreater{}  5      2     20  1924        1       1       0     1}
+\CommentTok{\#\textgreater{}  6      2     20  1925        1       1       0     1}
+\CommentTok{\#\textgreater{}  7      2     20  1926        1       1       0     1}
+\CommentTok{\#\textgreater{}  8      2     20  1927        1       1       0     1}
+\CommentTok{\#\textgreater{}  9      2     20  1928        1       1       0     1}
+\CommentTok{\#\textgreater{} 10      2     20  1929        1       1       0     1}
+\CommentTok{\#\textgreater{} \# ... with 112,272 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+Here, the user has created all non-directed dyad-years from 1816 to 2010 and then subset the data to just those with a major power or with some kind of contiguity relationship.
+
+Users will find that the ease of the ``pipe'' will allow them greater agency in creating the full data set they may want for an analysis.
+Indeed, the pipe has the effect of forming something analogous to a drop-down menu, in which the user can ``select'' additional data/commands they may want simply by specifying the function in \texttt{\{peacesciencer\}} that does what they want.
+For example, a researcher can follow \texttt{filter\_prd()} with another pipe and communicate they want information about ongoing conflicts and conflict onsets from the Gibler-Miller-Little (GML) dispute data set (Gibler, Miller \& Little, 2016).
+Following \texttt{filter\_prd()} with \texttt{add\_gml\_mids(keep\ =\ NULL)} will add information about ongoing conflicts and onsets in a given dyad-year.\footnote{\texttt{keep\ =\ NULL} in this context will focus the information returned to just the information about ongoing conflicts and onsets, discarding potentially unwanted columns about things like Side A, hostility levels, and other information included in the data set.
+Type \texttt{?add\_gml\_mids()} for more information.}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# create base data, and pipe (\%\textgreater{}\%) to next function}
+\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{directed =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{subset\_years =} \FunctionTok{c}\NormalTok{(}\DecValTok{1816}\SpecialCharTok{:}\DecValTok{2010}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# subset data to politically relevant dyads (PRDs), pipe to next function}
   \FunctionTok{filter\_prd}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_gml\_mids}\NormalTok{(}\AttributeTok{keep =} \ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_peace\_years}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_nmc}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_democracy}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_cow\_alliance}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_sdp\_gdp}\NormalTok{() }\OtherTok{{-}\textgreater{}}\NormalTok{ Data}
-\end{Highlighting}
-\end{Shaded}
-
-Next, the pipe-based workflow continues with \texttt{add\_gml\_mids(keep\ =\ NULL)}, which will add information about whether there was an ongoing dispute in the dyad (and whether it was a dispute onset) from the GML dispute data.\footnote{\texttt{keep\ =\ NULL} is optional and removes supporting information about dyadic dispute-participants (e.g.~dispute-level hostility, dispute-level outcome) in order to reduce the size of the data set.} \texttt{add\_peace\_years()} will calculate peace years between ongoing disputes in these politically relevant non-directed dyads.\footnote{\texttt{add\_peace\_years()} will only calculate the peace years---titled, in this case, \texttt{gmlmidspell}---and will leave the temporal dependence adjustment to the taste of the researcher.
+  \CommentTok{\# add conflict information from GML{-}MID data,}
+  \FunctionTok{add\_gml\_mids}\NormalTok{(}\AttributeTok{keep =} \ConstantTok{NULL}\NormalTok{) }
+\CommentTok{\#\textgreater{} \# A tibble: 112,282 x 9}
+\CommentTok{\#\textgreater{}    ccode1 ccode2  year conttype cowmaj1 cowmaj2   prd gmlmidonset gmlmidongoing}
+\CommentTok{\#\textgreater{}     \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}    \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}       \textless{}dbl\textgreater{}         \textless{}dbl\textgreater{}}
+\CommentTok{\#\textgreater{}  1      2     20  1920        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  2      2     20  1921        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  3      2     20  1922        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  4      2     20  1923        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  5      2     20  1924        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  6      2     20  1925        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  7      2     20  1926        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  8      2     20  1927        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  9      2     20  1928        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{} 10      2     20  1929        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{} \# ... with 112,272 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+Users can also calculate peace-years for these conflicts with \texttt{add\_spells()} by using the pipe to pass forward the data set and applying the \texttt{add\_spells()} function to it.\footnote{\texttt{add\_spells()} will only calculate the peace years---titled, in this case, \texttt{gmlmidspell}---and will leave the temporal dependence adjustment to the taste of the researcher.
 Importantly, I do not recommend manually creating splines or square/cube terms because it creates more problems in adjusting for temporal dependence in model predictions.
@@
 This function includes the spell and three splines (hence the 4 in the command).
-Either approach makes for easier model predictions, given R's functionality.} \texttt{add\_nmc()} will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity.
-\texttt{add\_democracy()} will add information about the level of democracy for both states in the dyad using three different indicators.
-Researchers can select which one they prefer from these indicators.
-\texttt{add\_cow\_alliance()} will add information about whether there was an alliance in the dyad-year by way of whether there was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge.
-Finally, \texttt{add\_sdp\_gdp()} will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by Anders, Fariss \& Markowitz (2020).
-
+Either approach makes for easier model predictions, given R's functionality.}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# create base data, and pipe (\%\textgreater{}\%) to next function}
+\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{directed =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{subset\_years =} \FunctionTok{c}\NormalTok{(}\DecValTok{1816}\SpecialCharTok{:}\DecValTok{2010}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# subset data to politically relevant dyads (PRDs), pipe to next function}
+  \FunctionTok{filter\_prd}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# add conflict information from GML{-}MID data, pipe to next function}
+  \FunctionTok{add\_gml\_mids}\NormalTok{(}\AttributeTok{keep =} \ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# add peace years ("spells")}
+  \FunctionTok{add\_spells}\NormalTok{()}
+\CommentTok{\#\textgreater{} \# A tibble: 112,282 x 10}
+\CommentTok{\#\textgreater{}    ccode1 ccode2  year conttype cowmaj1 cowmaj2   prd gmlmidonset gmlmidongoing}
+\CommentTok{\#\textgreater{}     \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}    \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}       \textless{}dbl\textgreater{}         \textless{}dbl\textgreater{}}
+\CommentTok{\#\textgreater{}  1      2     20  1920        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  2      2     20  1921        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  3      2     20  1922        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  4      2     20  1923        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  5      2     20  1924        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  6      2     20  1925        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  7      2     20  1926        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  8      2     20  1927        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  9      2     20  1928        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{} 10      2     20  1929        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{} \# ... with 112,272 more rows, and 1 more variable: gmlmidspell \textless{}dbl\textgreater{}}
+\end{Highlighting}
+\end{Shaded}
+
+Researchers should see that \texttt{\{peacesciencer\}}'s functionality can scale up nicely from there.
+For example, the following would round out the kind of information necessary to replicate Bremer's (1992) famous ``dangerous dyads'' analysis by adding information about national material capabilities for both states in the dyad (\texttt{add\_nmc()}), estimates of democracy for both states in the dyad (\texttt{add\_democracy()}), information about alliance commitments in the dyad-year (\texttt{add\_cow\_alliance()}, by way of Gibler (2009)), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by Anders, Fariss \& Markowitz (2020).
 Whereas \texttt{add\_sdp\_gdp()} is the last command in the pipe-based workflow, the \texttt{\{peacesciencer\}} call ends by assigning to an object called \texttt{Data}.
 This type of assignment is done with the ``right hand'' assignment operator (i.e.~\texttt{-\textgreater{}}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# create base data, and pipe (\%\textgreater{}\%) to next function}
+\FunctionTok{create\_dyadyears}\NormalTok{(}\AttributeTok{directed =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{subset\_years =} \FunctionTok{c}\NormalTok{(}\DecValTok{1816}\SpecialCharTok{:}\DecValTok{2010}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# subset data to politically relevant dyads (PRDs), pipe to next function}
+  \FunctionTok{filter\_prd}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# add conflict information from GML{-}MID data, pipe to next function}
+  \FunctionTok{add\_gml\_mids}\NormalTok{(}\AttributeTok{keep =} \ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# add peace years ("spells"), pipe to next function}
+  \FunctionTok{add\_spells}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# add capabilities data, pipe to next function}
+  \FunctionTok{add\_nmc}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# add some estimates about democracy for each state, pipe to next function}
+  \FunctionTok{add\_democracy}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# add information about alliance commitments in dyad{-}year}
+  \FunctionTok{add\_cow\_alliance}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# finish with information about population and GDP/SDP}
+  \CommentTok{\# and then assign to object, called, minimally, \textquotesingle{}Data\textquotesingle{}}
+  \FunctionTok{add\_sdp\_gdp}\NormalTok{() }\OtherTok{{-}\textgreater{}}\NormalTok{ Data}
+
+\NormalTok{Data}
+\CommentTok{\#\textgreater{} \# A tibble: 112,282 x 42}
+\CommentTok{\#\textgreater{}    ccode1 ccode2  year conttype cowmaj1 cowmaj2   prd gmlmidonset gmlmidongoing}
+\CommentTok{\#\textgreater{}     \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}    \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}       \textless{}dbl\textgreater{}         \textless{}dbl\textgreater{}}
+\CommentTok{\#\textgreater{}  1      2     20  1920        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  2      2     20  1921        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  3      2     20  1922        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  4      2     20  1923        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  5      2     20  1924        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  6      2     20  1925        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  7      2     20  1926        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  8      2     20  1927        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{}  9      2     20  1928        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{} 10      2     20  1929        1       1       0     1           0             0}
+\CommentTok{\#\textgreater{} \# ... with 112,272 more rows, and 33 more variables: gmlmidspell \textless{}dbl\textgreater{},}
+\CommentTok{\#\textgreater{} \#   milex1 \textless{}dbl\textgreater{}, milper1 \textless{}dbl\textgreater{}, irst1 \textless{}dbl\textgreater{}, pec1 \textless{}dbl\textgreater{}, tpop1 \textless{}dbl\textgreater{},}
+\CommentTok{\#\textgreater{} \#   upop1 \textless{}dbl\textgreater{}, cinc1 \textless{}dbl\textgreater{}, milex2 \textless{}dbl\textgreater{}, milper2 \textless{}dbl\textgreater{}, irst2 \textless{}dbl\textgreater{},}
+\CommentTok{\#\textgreater{} \#   pec2 \textless{}dbl\textgreater{}, tpop2 \textless{}dbl\textgreater{}, upop2 \textless{}dbl\textgreater{}, cinc2 \textless{}dbl\textgreater{}, v2x\_polyarchy1 \textless{}dbl\textgreater{},}
+\CommentTok{\#\textgreater{} \#   polity21 \textless{}dbl\textgreater{}, xm\_qudsest1 \textless{}dbl\textgreater{}, v2x\_polyarchy2 \textless{}dbl\textgreater{}, polity22 \textless{}dbl\textgreater{},}
+\CommentTok{\#\textgreater{} \#   xm\_qudsest2 \textless{}dbl\textgreater{}, cow\_defense \textless{}dbl\textgreater{}, cow\_neutral \textless{}dbl\textgreater{}, cow\_nonagg \textless{}dbl\textgreater{},}
+\CommentTok{\#\textgreater{} \#   cow\_entente \textless{}dbl\textgreater{}, wbgdp2011est1 \textless{}dbl\textgreater{}, wbpopest1 \textless{}dbl\textgreater{}, sdpest1 \textless{}dbl\textgreater{}, ...}
+\end{Highlighting}
+\end{Shaded}
+
 If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like \texttt{haven::write\_dta(Data,\ "my-data.dta")} and import it into Stata when they are done.
-
-\texttt{\{peacesciencer\}} has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding \texttt{-9s} in these data to be \texttt{NA} in the National Material Capabilities data.
-There is only some slight data work to create the desired indicators for a statistical model of conflict onset.
-Here, briefly, these functions lean on \texttt{\{tidyverse\}} ``verbs'' and base R terminology to do things like creating a dummy variable for whether the dyad was land-contiguous, whether there was a major power on either side of the non-directed dyad, creating some ``weak-link'' indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the Marquez (2016) for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{Data }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{landcontig =} \FunctionTok{ifelse}\NormalTok{(conttype }\SpecialCharTok{==} \DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cowmajdyad =} \FunctionTok{ifelse}\NormalTok{(cowmaj1 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{|}\NormalTok{ cowmaj2 }\SpecialCharTok{==} \DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\# Create estimate of militarization as milper/tpop}
-  \CommentTok{\# Then make a weak{-}link}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{milit1 =}\NormalTok{ milper1}\SpecialCharTok{/}\NormalTok{tpop1,}
-         \AttributeTok{milit2 =}\NormalTok{ milper2}\SpecialCharTok{/}\NormalTok{tpop2,}
-         \AttributeTok{minmilit =} \FunctionTok{ifelse}\NormalTok{(milit1 }\SpecialCharTok{\textgreater{}}\NormalTok{ milit2,}
-\NormalTok{                           milit2, milit1)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\# create CINC proportion (lower over higher)}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cincprop =} \FunctionTok{ifelse}\NormalTok{(cinc1 }\SpecialCharTok{\textgreater{}}\NormalTok{ cinc2,}
-\NormalTok{                           cinc2}\SpecialCharTok{/}\NormalTok{cinc1, cinc1}\SpecialCharTok{/}\NormalTok{cinc2)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\# create weak{-}link specification using Quick UDS data}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{mindemest =} \FunctionTok{ifelse}\NormalTok{(xm\_qudsest1 }\SpecialCharTok{\textgreater{}}\NormalTok{ xm\_qudsest2,}
-\NormalTok{                            xm\_qudsest2, xm\_qudsest1)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\# Create "weak{-}link" measure of jointly advanced economies}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{minwbgdppc =} \FunctionTok{ifelse}\NormalTok{(wbgdppc2011est1 }\SpecialCharTok{\textgreater{}}\NormalTok{ wbgdppc2011est2,}
-\NormalTok{                             wbgdppc2011est2, wbgdppc2011est1)) }\OtherTok{{-}\textgreater{}}\NormalTok{ Data}
-\end{Highlighting}
-\end{Shaded}
-
+No matter, \texttt{\{peacesciencer\}} has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding \texttt{-9s} in these data to be \texttt{NA} in the capabilities data.
+In this particular application, it has already created the main data required for a replication of Bremer (1992).
+There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some ``weak-link'' indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the Marquez (2016) method for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.
 Table \ref{tab:tab-dd} is a formatted version of the results of a logistic regression model of conflict onset using these ``dangerous dyads'' indicators.
@@
 \midrule
-Land Contiguity & \num{1.044}*\\
+Land Contiguity & \num{1.062}*\\
+ & \vphantom{1} (\num{0.057})\\
+Dyadic CINC Proportion (Lower/Higher) & \num{0.453}*\\
+ & (\num{0.036})\\
+CoW Major Power in Dyad & \num{0.144}*\\
  & (\num{0.057})\\
-Dyadic CINC Proportion (Lower/Higher) & \num{0.444}*\\
- & (\num{0.036})\\
-CoW Major Power in Dyad & \num{0.145}*\\
- & \vphantom{1} (\num{0.058})\\
-Defense Pact & \num{-0.102}+\\
+Defense Pact & \num{-0.119}*\\
  & (\num{0.058})\\
-Dyadic Democracy (Weak-Link) & \num{-0.509}*\\
+Dyadic Democracy (Weak-Link) & \num{-0.493}*\\
  & (\num{0.052})\\
-Dyadic GDP per Capita (Weak-Link) & \num{0.262}*\\
- & (\num{0.050})\\
-Dyadic Militarization (Minimum) & \num{0.267}*\\
+Dyadic GDP per Capita (Weak-Link) & \num{0.293}*\\
+ & (\num{0.051})\\
+Dyadic Militarization (Minimum) & \num{0.263}*\\
  & (\num{0.023})\\
-t & \num{-0.148}*\\
+t & \num{-0.146}*\\
  & (\num{0.005})\\
@@
  & (\num{0.000})\\
-Intercept & \num{-3.055}*\\
- & (\num{0.064})\\
+Intercept & \num{-3.045}*\\
+ & (\num{0.063})\\
 \midrule
-Num.Obs. & \num{106937}\\
+Num.Obs. & \num{103919}\\
 \bottomrule
@@
 
-\texttt{\{peacesciencer\}} is also capable of creating data for replications of standard civil conflict analyses.
+\hypertarget{creating-state-year-data-and-adding-to-state-year-data-in-peacesciencer}{%
+\subsection{\texorpdfstring{Creating State-Year Data and Adding to State-Year Data in \texttt{\{peacesciencer\}}}{Creating State-Year Data and Adding to State-Year Data in \{peacesciencer\}}}\label{creating-state-year-data-and-adding-to-state-year-data-in-peacesciencer}}
+
+\texttt{\{peacesciencer\}} is capable of generating data for replications of analyses at multiple levels beyond dyad-year.
 Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's (2003) well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years.
-The pipe-based workflow will start with \texttt{create\_stateyears(system\ =\ \textquotesingle{}gw\textquotesingle{})}, creating the full universe of Gleditsch-Ward state years from 1816 to 2020.
-Whereas the UCDP data included in \texttt{\{peacesciencer\}} are from 1946 to 2019 (v.
-20.1), we can subset (``filter'') the data to just those years with \texttt{filter(year\ \%in\%\ c(1946,\ 2019))}.
+The pipe-based workflow will start with \texttt{create\_stateyears(system\ =\ \textquotesingle{}gw\textquotesingle{},\ subset\_years\ =\ c(1946:2019))}, creating the full universe of Gleditsch-Ward state years and subsetting them to just 1946-2019 (because the UCDP data included in \texttt{\{peacesciencer\}} include just the observations in that time frame).
 Next, we can use the \texttt{add\_ucdp\_acd()} function to return information about ongoing UCDP conflicts and onsets for these states.
@@
 Options include ``extrasystemic'', ``interstate'', ``intrastate'', and ``II'' (short for ``internationalized intrastate'').
-If no \texttt{type} is specified, the function returns information about ongoing disputes and onsets for all states for all tyeps of conflict.
+If no \texttt{type} is specified, the function returns information about ongoing disputes and onsets for all states for all types of conflict.
 If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g.~\texttt{type\ =\ c("intrastate",\ "II")}).
@@
 Options include ``territory'', ``government'', and ``both''.
-If no \texttt{issue} is specified, the function returns information for all conflicts irregarding the particular issue.
+If no \texttt{issue} is specified, the function returns information for all conflicts regardless of the particular issue.
 \texttt{only\_wars} is an argument that subsets the data to just those with the intensity levels of ``war'' when \texttt{only\_wars\ =\ TRUE}.
@@
 \begin{Highlighting}[]
-\FunctionTok{create\_stateyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(year }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1946}\SpecialCharTok{:}\DecValTok{2019}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+\CommentTok{\# Create a set of state{-}year data for a civil conflict analysis.}
+\FunctionTok{create\_stateyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{, }\AttributeTok{subset\_years =} \FunctionTok{c}\NormalTok{(}\DecValTok{1946}\SpecialCharTok{:}\DecValTok{2019}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{add\_ucdp\_acd}\NormalTok{(}\AttributeTok{type=}\StringTok{"intrastate"}\NormalTok{, }\AttributeTok{only\_wars =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_peace\_years}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{add\_spells}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{add\_democracy}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
@@
 Finally, we can add some covariates of interest to these data.
-\texttt{add\_peace\_years()} calculates peace spells between ongoing conflicts in the data generated by \texttt{add\_ucdp\_acd()}.
+\texttt{add\_spells()} calculates peace spells between ongoing conflicts in the data generated by \texttt{add\_ucdp\_acd()}.
 \texttt{add\_democracy()} adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' (2016) extension of Pemstein et al's (2010) Unified Democracy Scores).
@@
 To avoid confusion with the overlapping column names, the researcher can use some \texttt{\{tidyverse\}} verbs to rename all those conflict variables to have a distinct prefix of \texttt{war\_} (i.e.~\texttt{rename\_at(vars(ucdpongoing:ucdpspell),\ \textasciitilde{}paste0("war\_",\ .))}) before finally joining these data into the master data frame (i.e.~\texttt{left\_join(Data,\ .)\ -\textgreater{}\ Data}).
-Table \ref{tab:tab-cw} shows the fruits of the data \texttt{\{peacesciencer\}} generated after some post-processing for lagging important variables.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{create\_stateyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(year }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1946}\SpecialCharTok{:}\DecValTok{2019}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+Table \ref{tab:tab-cw} shows the fruits of the data \texttt{\{peacesciencer\}} generated after some post-processing and lagging important variables.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Repeat the process, but for a new DV of just wars}
+\FunctionTok{create\_stateyears}\NormalTok{(}\AttributeTok{system =} \StringTok{\textquotesingle{}gw\textquotesingle{}}\NormalTok{, }\AttributeTok{subset\_years =} \FunctionTok{c}\NormalTok{(}\DecValTok{1946}\SpecialCharTok{:}\DecValTok{2019}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{add\_ucdp\_acd}\NormalTok{(}\AttributeTok{type=}\StringTok{"intrastate"}\NormalTok{, }\AttributeTok{only\_wars =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_peace\_years}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{add\_spells}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{rename\_at}\NormalTok{(}\FunctionTok{vars}\NormalTok{(ucdpongoing}\SpecialCharTok{:}\NormalTok{ucdpspell), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{paste0}\NormalTok{(}\StringTok{"war\_"}\NormalTok{, .)) }\SpecialCharTok{\%\textgreater{}\%}
@@
 \midrule
+GDP per Capita (Lagged) & \num{-0.285}* & \num{-0.343}*\\
+ & (\num{0.110}) & (\num{0.172})\\
 Population Size (Lagged) & \num{0.229}* & \num{0.272}*\\
@@
  & (\num{0.000}) & (\num{0.000})\\
-Intercept & \num{-5.097}* & \num{-6.590}*\\
+Intercept & \num{-5.098}* & \num{-6.591}*\\
  & (\num{1.351}) & (\num{2.084})\\
@@
 
+\hypertarget{creating-leader-year-data-and-adding-to-leader-year-data-in-peacesciencer}{%
+\subsection{\texorpdfstring{Creating Leader-Year Data and Adding to Leader-Year Data in \texttt{\{peacesciencer\}}}{Creating Leader-Year Data and Adding to Leader-Year Data in \{peacesciencer\}}}\label{creating-leader-year-data-and-adding-to-leader-year-data-in-peacesciencer}}
+
+\texttt{\{peacesciencer\}} also has support for newer levels of analysis in the peace science community, prominently leader-levels of analysis.
+There has been considerable emphasis in peace science research to emphasize that state leaders, not ``states'', make foreign policy decisions that may lead to war.
+Understanding the attributes of leaders themselves are critical to the core research questions of the community (Goemans, Gleditsch \& Chiozza, 2009; Horowitz \& Stam, 2014; Ellis, Horowitz \& Stam, 2015) and \texttt{\{peacesciencer\}} wants to help researchers toward that end.
+
+Table \ref{tab:tab-addfunctions} shows there are some dedicated functions for populating leader-level data with leader-specific information, in addition to adding state-year-level information (e.g.~democracy, capabilities) to leader-level data.
+Table \ref{tab:tab-createfunctions} shows support for creating leader-level data that are standardized to either the CoW or G-W state system data.
+Suppose a researcher wanted to create leader-year data, standardized to CoW system dates, for an analysis of inter-state dispute initiation analogous to what Horowitz \& Stam (2014) do.
+The user would first start with creating the base data (\texttt{create\_leaderyears(standardize\ =\ "cow",\ subset\_years=c(1875:2010))}), which also generates some leader attributes (e.g.~estimated leader age, year in office, and leader gender).
+From there, they would add information about leader conflict behavior in the year with \texttt{add\_gml\_mids()} and calculate peace-years with the \texttt{add\_spells()} function.\footnote{The documentation for \texttt{add\_gml\_mids()}, along with the \texttt{data-raw} directory on Github, provides ample explanation for how inter-state conflicts in the GML conflict data are connected to specific leaders.
+Since inter-state dispute data can have missing days, connecting leaders to dispute onsets involves some important archival research that is only possible because of the information collected by Gibler, Miller \& Little (2016) over 10 years of scholarship.
+The documentation also considers the multiple ways in which users can more flexibly code the concept of ``initiation'' based on participant-level summaries.
+The default coding of initiation, based on this exhaustive archival work, is to code a leader initiating a dispute if they were on Side A or joined the conflict at any point after the onset of the dispute.
+Users can change this to focus on just Side A, or just originators on Side A.
+This is the \texttt{init} argument in \texttt{add\_gml\_mids()}.} \texttt{add\_lead()} will add a battery of leader experience and attributes variables from data created by Ellis, Horowitz \& Stam (2015), including whether the leader had military experience, combat experience, was a rebel fighter, and many more.
+Finally, \texttt{add\_nmc()} and \texttt{add\_democracy()} will add state-year estimates of national capabilities and democracy for assisting in making state-to-state comparisons, even for leader-level analyses.
+Some light recoding of the data created in \texttt{\{peacesciencer\}} and some regression modeling reproduces the results presented in Table \ref{tab:tab-ly}, itself an approximation of the kind of leader-year analysis exemplified in Horowitz \& Stam (2014).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Create a set of leader{-}year data for an analysis of conflict initiation.}
+\FunctionTok{create\_leaderyears}\NormalTok{(}\AttributeTok{standardize =} \StringTok{"cow"}\NormalTok{, }\AttributeTok{subset\_years=}\FunctionTok{c}\NormalTok{(}\DecValTok{1875}\SpecialCharTok{:}\DecValTok{2010}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{add\_gml\_mids}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{add\_spells}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{add\_lead}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{add\_nmc}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{add\_democracy}\NormalTok{() }\OtherTok{{-}\textgreater{}}\NormalTok{ Data}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}
+
+\caption{\label{tab:tab-ly}A Leader-Year Analysis of Inter-State Dispute Initiation in \{peacesciencer\}}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{8cm}c}
+\toprule
+\textbf{ } & \textbf{Model 1}\\
+\midrule
+Male Leader & \num{-0.752}*\\
+ & (\num{0.206})\\
+Leader Age & \num{0.010}*\\
+ & (\num{0.002})\\
+Leader Tenure (in Years) & \num{0.027}*\\
+ & \vphantom{1} (\num{0.004})\\
+Military Service (No Combat) & \num{0.332}*\\
+ & (\num{0.084})\\
+Military Serivce (with Combat) & \num{0.132}\\
+ & (\num{0.090})\\
+Previous Rebel Experience & \num{0.287}*\\
+ & (\num{0.069})\\
+Prior War Win & \num{0.033}\\
+ & (\num{0.107})\\
+Prior War Loss & \num{0.090}\\
+ & (\num{0.106})\\
+Prior Rebel Win & \num{-0.136}\\
+ & (\num{0.102})\\
+Prior Rebel Loss & \num{0.328}*\\
+ & (\num{0.142})\\
+Extended UDS Democracy Score & \num{-0.180}*\\
+ & (\num{0.037})\\
+CINC & \num{11.858}*\\
+ & (\num{0.626})\\
+t & \num{-0.247}*\\
+ & (\num{0.028})\\
+t\textasciicircum2 & \num{0.010}*\\
+ & (\num{0.004})\\
+t\textasciicircum3 & \num{0.000}+\\
+ & (\num{0.000})\\
+Intercept & \num{-1.816}*\\
+ & (\num{0.243})\\
+\midrule
+Num.Obs. & \num{13839}\\
+\bottomrule
+\multicolumn{2}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05}\\
+\end{tabular}
+\end{table}
+
+\hypertarget{a-comparison-with-other-approaches}{%
+\section{A Comparison with Other Approaches}\label{a-comparison-with-other-approaches}}
+
+\texttt{\{peacesciencer\}} is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch.
+Alternatives exist, some more inaccessible than others.
+NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars (Bennett, Poast \& Stam, 2019).
+NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade (Bennett \& Stam, 2000).
+Finally, peace scientists well-versed in Structured Query Language (SQL) could use one of several ``join'' transformations to create dyad-year data from state system data, even if this might amount to a detour in the peace scientist's research agenda for this particular task.\footnote{That said, \texttt{\{peacesciencer\}} makes ample use of SQL-like ``joins'' for the construction of base data and functions that merge data into base data created in the package.
+A discussion of these joins appears as a vignette on the package's website.} A comparison of \texttt{\{peacesciencer\}} with EUGene and NewGene suggests the following benefits of the package while emphasizing areas where other options may have some advantages.
+
+EUGene is the clear inspiration for this package.
+Though its original impetus was the generation of expected utility data for evaluating Bueno de Mesquita \& Lalman (1992) (i.e.~the ``EU'' in ``EUGene''), the software became quite popular for peace science scholars in the early 2000s for helping them start new projects from scratch with important data already provided.
+\texttt{\{peacesciencer\}} covers all the same units of analysis that EUGene covers, as of version 3.2.\footnote{The latest version, as of writing, is v.
+3.212 released in 2017.
+However, this is apparently a bug fix on string variables and that the underlying functionality was still last updated in 2007 with the release of version 3.2.} EUGene has more explicit support for dyadic dispute-year data whereas \texttt{\{peacesciencer\}} treats dyadic dispute-year data as a derivation of dyad-year data, with respect to functions that populate base data with additional information.\footnote{For example, the ``whittle'' class of functions (i.e.~\texttt{wc\_} in Table \ref{tab:tab-addfunctions}) are designed for dyadic dispute-year data also included in the package while a helper function---\texttt{declare\_attributes()}---can allow a user to treat other dyadic dispute-year data as dyad-year data for the sake of using any of the other supported functions in Table \ref{tab:tab-addfunctions}.} \texttt{\{peacesciencer\}} may reflect more current research interests in the peace science community.
+This is why there are leader-level data and support for the G-W (and UCDP) ecosystem of data that EUGene does not have, but there is no function yet for things like calculating expected utility values.
+Importantly, though, \texttt{\{peacesciencer\}} mimics the verbosity of EUGene's user manual.
+EUGene's user manual was amply informative about what it was calculating and why it was doing what it was doing for some defaults.
+Likewise, \texttt{\{peacesciencer\}}'s documentation strives to be as informative as possible as to what it is doing and why it is doing what it is doing for some defaults.
+A scholar who remembers EUGene well will ideally think of \texttt{\{peacesciencer\}} as the most faithful approximation of what that software did for the community at the time.
+It has the added benefit of being agnostic to the researcher's operating system, having greater flexibility of data types supported, and better reflecting more current frontiers in the community.
+It does have the drawback of requiring at least some level of comfort with the R programming language.
+
+NewGene is the latest evolution of EUGene, at least as a stand-alone executable program for creating the kind of data of interest to the peace science community.
+NewGene's greatest strength, relative to EUGene and even \texttt{\{peacesciencer\}}, is its support for \emph{k}-adic data (c.f.
+Poast, 2010).
+\emph{k}-adic levels of analysis are not yet supported in \texttt{\{peacesciencer\}} and users interested in generating \emph{k}-adic data should consider downloading NewGene.
+No matter, \texttt{\{peacesciencer\}} has several superlatives in relation to NewGene.
+It supports leader-year and leader-dyad-year data while NewGene does not.
+It offers support for the G-W (and UCDP) ecosystem of data whereas NewGene does not and is mostly aimed for researchers interested in inter-state conflict.
+NewGene deviates a bit from EUGene by only indirectly asking users what kind of data they want (e.g.~state-year, dyad-year) and without providing too much detail about what it is doing and why it is doing what it is doing.
+For example, NewGene only indirectly states the unit of analysis of the data to be generated near the top of its interface by asking the user how many country columns they want.
+This is an indirect way of the user to get data that are state-year, dyad-year, triad-year (etc.), which are then expanded and populated with data at various levels.
+\texttt{\{peacesciencer\}}, much like EUGene, is more explicit, encouraging the user to be upfront about what their unit of analysis is and what are the data that can be plausibly plugged into the data the user is creating.
+\texttt{\{peacesciencer\}}, again, does expect at least some level of comfort with the R programming language, but even this comes with greater ease of interpreting what is the unit of analysis and what are the primary spatial and temporal units that serve as the basis of the data.
+
 \hypertarget{conclusion}{%
@@
 \texttt{\{peacesciencer\}} is already more than capable of creating the kind of data in high demand in peace science.
-It can create dyad-year and state-year data.
+It can create dyad-year, leader-year, leader-dyad-year, and state-year data (among others).
 It is also generalizable to the dispute data included in the package, allowing for merging into \emph{dispute-year} data as well.
-This feature showed how it can effectively approximate two types of analyses in wide use in the peace science community.
+This feature showed how it can effectively approximate three types of analyses in wide use in the peace science community.
 Surely researchers can and will add more information to these simple analyses after using \texttt{\{peacesciencer\}}, but the package already does a lot of the tedious work for researchers.
 It also does this in a maximally transparent way that conforms well to the DA-RT initiative across all political science.
-
 This is not to say \texttt{\{peacesciencer\}} does everything, but \texttt{\{peacesciencer\}} can only evolve and expand on what it already does well.
-For example, \texttt{\{peacesciencer\}} does not right now handle \emph{k}-adic data (c.f.
-Poast, 2010).
-However, the nature of \emph{k}-ads do not make for necessarily problematic data merges and should not be difficult to add for researchers who want it in future updates.
-Likewise, \texttt{\{peacesciencer\}} contains the Archigos leader data (Goemans, Gleditsch \& Chiozza, 2009) and is able to transform that into interesting data at the dyad-level and state-level, but doing a leader-level analysis with information about leader willingness to use force (e.g.
-Carter \& Smith, 2020) is not right now available.
-However, these are possible expansions on what \texttt{\{peacesciencer\}} already does well.
-Thus, this package can only evolve to meet new analytical demands for the peace science community.
+This package can only evolve to meet new analytical demands for the peace science community.
 Users are free to request new features as ``issues'' on the project's Github.
@@
 \texttt{add\_contiguity()} returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity threshold.
-The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g.~the contiguity data are not ordinal and should not be treated as such), how and where data issues arise (e.g.~how CoW state system data differ from Gleditsch-Ward data and how one is coerced into the other), and how researchers should consider optimally using its functionality (e.g.~\texttt{add\_ucdp\_acd()} probably should not lump all forms of conflict together).
+The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g.~the contiguity data are not ordinal and should not be treated as such), how and where data issues arise (e.g.~how CoW state system data differ from Gleditsch-Ward data and how one is coerced into the other), and how researchers should consider optimally using its functionality (e.g.~\texttt{add\_ucdp\_acd()} probably should not lump all forms of conflict together (c.f.
+Gibler \& Miller, Forthcoming)).
 \texttt{\{peacesciencer\}} does not endeavor to make researchers lazy or sloppy, and it does not ultimately do this.
-Instead, \texttt{\{peacesciencer\}} reduces the tedium associated with starting quantitative peace science research, achieving this in a quick, robust, and transparent way.
+Instead, \texttt{\{peacesciencer\}} encourages well-reasoned design decisions by the user up front and reduces the tedium associated with starting quantitative peace science research.
+It achieves this in a quick, robust, and transparent way.
 
@@
 
+\leavevmode\hypertarget{ref-barbierietal2009td}{}%
+Barbieri, Katherine, Omar MG Keshk \& Brian Pollins (2009) Trading data: Evaluating our assumptions and coding rules. \emph{Conflict Management and Peace Science} 26(5): 471--491.
+
 \leavevmode\hypertarget{ref-becketal1998tts}{}%
@@
 
-\leavevmode\hypertarget{ref-boehmke2016dwr}{}%
-Boehmke, Bradley C (2016) \emph{Data Wrangling with {R}}. Springer.
-
 \leavevmode\hypertarget{ref-bowersvoors2016hiyr}{}%
@@
 
+\leavevmode\hypertarget{ref-bdmlalman1992wr}{}%
+Bueno de Mesquita, Bruce \& David Lalman (1992) \emph{War and Reason: Domestic and International Imperatives}. Yale University Press.
+
 \leavevmode\hypertarget{ref-cartersignorino2010bf}{}%
@@
 
+\leavevmode\hypertarget{ref-coppedgeetal2020vdem}{}%
+Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I Lindberg, Jan Teorell, David Altman, Michael Bernhard, MSteven Fish, Adam Glynn, Allen Hicken, Anna Luhrmann, Kyle L Marquardt, Kelly McMann, Pamela Paxton, Daniel Pemstein, Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Agnes Cornell, Lisa Gastaldi, Haakon Gjerløw, Valeriya Mechkova, Johannes von Römer, Aksel Sundtröm, Eitan Tzelgov, Luca Uberti, Yi-ting Wang, Tore Wig \& Daniel Ziblatt (2020) V-dem codebook v10.
+
+\leavevmode\hypertarget{ref-cowstates2016}{}%
+Correlates of War (2011) State system membership list, v2016 (\url{http://correlatesofwar.org}).
+
+\leavevmode\hypertarget{ref-ellisetal2015lead}{}%
+Ellis, Cali Mortenson, Michael C Horowitz \& Allan C Stam (2015) Introducing the {LEAD} data set. \emph{International Interactions} 41(4): 718--741.
+
 \leavevmode\hypertarget{ref-fearonlaitin2003eicw}{}%
@@
 
+\leavevmode\hypertarget{ref-gibler2009ima}{}%
+Gibler, Douglas M (2009) \emph{International Military Alliances, 1648-2008}. Washington DC: CQ Press.
+
 \leavevmode\hypertarget{ref-giblermiller2014etts}{}%
@@
 
+\leavevmode\hypertarget{ref-giblermiller2021apm}{}%
+Gibler, Douglas M \& Steven V Miller (Forthcoming) An appraisal of {P}roject {M}ars and the {D}ivided {A}rmies argument. \emph{International Studies Quarterly} (Forthcoming).
+
 \leavevmode\hypertarget{ref-gibleretal2016amid}{}%
@@
 
+\leavevmode\hypertarget{ref-gledtischward1999rlis}{}%
+Gleditsch, Kristian S \& Michael D Ward (1999) A revised list of independent states since the {C}ongress of {V}ienna. \emph{International Interactions} 25(4): 393--413.
+
+\leavevmode\hypertarget{ref-gleditschetal2002ac}{}%
+Gleditsch, Nils Petter, Peter Wallensteen, Mikael Eriksson, Margareta Sollenberg \& Havard Strand (2002) Armed conflict 1946-2001: A new dataset. \emph{Journal of Peace Research} 39(5): 615--637.
+
 \leavevmode\hypertarget{ref-goemansetal2009ia}{}%
@@
 
+\leavevmode\hypertarget{ref-haege2011cc}{}%
+Häge, Frank (2011) Choice or circumstance? Adjusting measures of foreign policy similarity for chance agreement. \emph{Political Analysis} 19(3): 287--305.
+
+\leavevmode\hypertarget{ref-horowitzstam2014hpme}{}%
+Horowitz, Michael C \& Allan C Stam (2014) How prior military experience influences the future militarized behavior of leaders. \emph{International Organization} 68(3): 527--559.
+
+\leavevmode\hypertarget{ref-leedsetal2002atop}{}%
+Leeds, Bretty Ashley, Jeffrey M Ritter, Sara McLaughlin Mitchell \& Andrew G Long (2002) Alliance treaty obligations and provisions, 1815-1944. \emph{International Interactions} 28: 237--260.
+
 \leavevmode\hypertarget{ref-lemkereed2001rprd}{}%
@@
 
+\leavevmode\hypertarget{ref-marshalletal2017p}{}%
+Marshall, Monty G, Ted Robert Gurr \& Keith Jaggers (2017) Polity {IV} project: Political regime characteristics and transitions, 1800-2016.
+
 \leavevmode\hypertarget{ref-nunnpuga2012r}{}%
@@
 
-\leavevmode\hypertarget{ref-wickhamgrolemund2017rds}{}%
-Wickham, Hadley \& Garrett Grolemund (2017) \emph{{R} for Data Science}. Sebastopol, CA: O'Reilly Media, Inc.
+\leavevmode\hypertarget{ref-schvitz2021mis}{}%
+Schvitz, Guy, Seraina Rüegger, Luc Girardin, Lars-Erik Cederman, Nils Weidmann \& Kristian Skrede Gleditsch (2021) {Mapping The International System, 1886-2017: The {CS}hapes 2.0 Dataset}.
+\emph{Journal of Conflict Resolution} (\url{https://journals.sagepub.com/doi/full/10.1177/00220027211013563}).
+
+\leavevmode\hypertarget{ref-singer1987rcwd}{}%
+Singer, David J (1987) Reconstructing the {C}orrelates of {W}ar dataset on material capabilities of states, 1816-1985. \emph{International Interactions} 14(1): 115--32.
+
+\leavevmode\hypertarget{ref-stinnettetal2002cow}{}%
+Stinnett, Douglas M, Jaroslav Tir, Philip Schafer, Paul F Diehl \& Charles S Gochman (2002) The {C}orrelates of {W}ar {P}roject {D}irect {C}ontiguity {D}ata, version 3.
+\emph{Conflict Management and Peace Science} 19(2): 58--66.
+
+\leavevmode\hypertarget{ref-thompsondreyer2012hir}{}%
+Thompson, William R \& David R Dreyer (2012) \emph{Handbook of International Rivalries, 1494-2010}. Washington DC: CQ Press.
+
+\leavevmode\hypertarget{ref-wickhametal2019wt}{}%
+Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, Alex Hayes, Lionel Henry, Jim Hester, Max Kuhn, Thomas Lin Pedersen, Evan Miller, Stephan Milton Bache, Kirill Müller, Jeroen Ooms, David Robinson, Dana Paige Seidel, Vitalie Spinu, Kohske Takahashi, Davis Vaughan, Claus Wilke, Kara Woo \& Hiroaki Yutani (2019) Welcome to the {tidyverse}.
+\emph{Journal of Open Source Software} 4(43): 1686.
 

### 2022-01-29 manuscript/ms.Rmd
update, mostly Fira Code
gonna eventually send this to Sage.

@@
 
-This data feature tackles a recurring problem for researchers in the peace science community.
+This manuscript tackles a recurring problem for researchers in the peace science community.
 True research reproducibility is best achieved creating data from scratch, though no published guide exists that informs researchers how to do this on their own.
@@
 Archigos) and two other arguments: `standardize` and `subset_years`.
-The `standardize` arguments, which when it defaults to "none", returns all leader-years as presented in the raw Archigos data.
+The `standardize` argument, which when it defaults to "none", returns all leader-years as presented in the raw Archigos data.
 Archigos' leader data are nominally denominated in the Gleditsch-Ward state system data, if not necessarily Gleditsch-Ward state system dates (e.g.
@@
 After creating the base data of interest to their project, researchers can begin to add information they want with the suite of functions outlined in Table \@ref(tab:tab-addfunctions).
-For example, researcher interested in a dyad-year analysis of inter-state disputes can create a non-directed dyad-year data set from 1816 to 2010 in `{peacesciencer}` with the `create_dyadyears()` function, one of the aforementioned "create" function.
+For example, a researcher interested in a dyad-year analysis of inter-state disputes can create a non-directed dyad-year data set from 1816 to 2010 in `{peacesciencer}` with the `create_dyadyears()` function, one of the aforementioned "create" function.
 The following would create the base data of interest to the user (i.e.

### 2022-01-29 manuscript/src/render_pdf.R

@@ -11,7 +11,7 @@ rmarkdown::render("ms.Rmd",
                               ps_manuals = "http://svmiller.com/peacesciencer/reference",
                               ps_github = "https://github.com/svmiller/peacesciencer/"),
                   bookdown::pdf_document2(template = stevetemplates::templ_article2(),
-                                          keep_tex = FALSE,
+                                          keep_tex = TRUE,
                                           latex_engine = "xelatex", dev="cairo_pdf",
                                           toc = FALSE, number_sections = FALSE))
 

### 2022-03-22 docs/ms.tex
add, with professional easter egg

@@
 \title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: January 29, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
+\textbf{Current version}: March 03, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
 \author{true}
@@
 \title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}).
-\textbf{Current version}: January 29, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
+\textbf{Current version}: March 03, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }
 
@@
             bookmarks=true,
-            pdfauthor={Steven V. Miller (Clemson University)},
+            pdfauthor={Steven V. Miller (Stockholm University)},
              pdfkeywords = {software, statistical analysis, peace science, reproducibility},
@@
    \vskip 13.5pt\relax \normalsize\fontsize{11}{12}
-   \MakeUppercase{Steven V. Miller}, \small{Clemson University}
+   \MakeUppercase{Steven V. Miller}, \small{Stockholm University}
 
@@
 
-\texttt{\{peacesciencer\}} is designed to address these problems.
+\texttt{\{peacesciencer\}} addresses these problems.
 Built around the free and open source R programming language, \texttt{\{peacesciencer\}} contains a suite of data and functions for creating data of interest to researchers.
@@
 \ttfamily{add\_fpsim()} & D, LD & Add dyadic foreign policy similarity measures to a data frame\\
-\ttfamily{add\_gml\_mids()} & D, L, LD, S & Add GML MID data to a data frame\\
+\ttfamily{add\_gml\_mids()} & D, L, LD, S & Add Gibler-Miller-Little (GML) MID data to a data frame\\
 \ttfamily{add\_gwcode\_to\_cow()} & D, L, LD, S & Match G-W state codes to CoW state codes\\
-\ttfamily{add\_igos()} & D, S & Add CoW IGO data to a data frame\\
+\ttfamily{add\_igos()} & D, S & Add CoW intergovernmental organizations (IGO) data to a data frame\\
 \ttfamily{add\_lead()} & L, LD & Add leader experience/attributes (LEAD) to data\\
@@
 \ttfamily{add\_strategic\_rivalries()} & D, S & Add strategic rivalry information to a data frame\\
-\ttfamily{add\_ucdp\_acd()} & S & Add UCDP Armed Conflict Data to state-year data frame\\
+\ttfamily{add\_ucdp\_acd()} & S & Add Uppsala Conflict Data Program (UCDP) Armed Conflict Data to state-year data frame\\
 \ttfamily{add\_ucdp\_onsets} & S & Add UCDP onsets to state-year data\\
@@
 This assumes at least some familiarity with the R programming language.
-Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.2.
+Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.3.
 \texttt{\{peacesciencer\}} is designed to be as user-friendly as possible.
@@
 \texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}}, and \texttt{\{peacesciencer\}} can work without it, but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality.
-A user can open an R session by way of RStudio and install both packages as follows.\footnote{R users will recognize that pound signs (\texttt{\#}) allow for commenting code.
+A user can open an R session by way of RStudio and install both packages as follows.\footnote{R users will recognize that hash signs (\texttt{\#}) allow for commenting code.
 Users new to R should understand its use here is primarily for exposition of what the code snippets in this manuscript are doing.
@@
 A user who approaches \texttt{\{peacesciencer\}} with a project in mind will see they can better tailor this function to what they want.
-Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the UCDP armed conflict data.
+Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the Uppsala Conflict Data Program (UCDP) armed conflict data.
 They will also know they have no use for pre-1946 observations since civil conflict data typically have coverage only from 1946 forward.
@@
 Researchers using \texttt{\{peacesciencer\}} to create data for their research project should start with one of these ``create'' functions.
-Whether state-year, dyad-year, or a leader-level analysis, these functions will create the full universe of cases of interest to a researcher for the unit of analysis that serves as the core of their project.
+Whether state-year, dyad-year, or a leader-level analysis, these functions will create the full universe of cases of interest to a researcher.
 
@@
 Researchers should see that \texttt{\{peacesciencer\}}'s functionality can scale up nicely from there.
-For example, the following would round out the kind of information necessary to replicate Bremer's (1992) famous ``dangerous dyads'' analysis by adding information about national material capabilities for both states in the dyad (\texttt{add\_nmc()}), estimates of democracy for both states in the dyad (\texttt{add\_democracy()}), information about alliance commitments in the dyad-year (\texttt{add\_cow\_alliance()}, by way of Gibler (2009)), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by Anders, Fariss \& Markowitz (2020).
+For example, the following would round out the kind of information necessary to replicate Bremer's (1992) famous ``dangerous dyads'' analysis by adding information about national material capabilities (e.g.~the composite index of national capabilities {[}CINC{]}) for both states in the dyad (\texttt{add\_nmc()}), estimates of democracy for both states in the dyad (\texttt{add\_democracy()}), information about alliance commitments in the dyad-year (\texttt{add\_cow\_alliance()}, by way of Gibler (2009)), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by Anders, Fariss \& Markowitz (2020).
 Whereas \texttt{add\_sdp\_gdp()} is the last command in the pipe-based workflow, the \texttt{\{peacesciencer\}} call ends by assigning to an object called \texttt{Data}.
@@
 If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like \texttt{haven::write\_dta(Data,\ "my-data.dta")} and import it into Stata when they are done.
-No matter, \texttt{\{peacesciencer\}} has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding \texttt{-9s} in these data to be \texttt{NA} in the capabilities data.
+No matter, \texttt{\{peacesciencer\}} has pre-processed, cleaned, recoded, and merged the desired data that have greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding \texttt{-9s} in these data to be \texttt{NA} in the capabilities data.
 In this particular application, it has already created the main data required for a replication of Bremer (1992).
-There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some ``weak-link'' indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the Marquez (2016) method for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.
-Table \ref{tab:tab-dd} is a formatted version of the results of a logistic regression model of conflict onset using these ``dangerous dyads'' indicators.
+There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some ``weak-link'' indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the Marquez (2016) method for extending the Unified Democracy Scores {[}UDS{]} data), and the GDP per capita in the dyad.
+Table \ref{tab:tab-dd} is a formatted version of the results of a logistic regression model of conflict onset using these ``dangerous dyads'' indicators and temporal adjustment variables (\emph{t}).
 Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but \texttt{\{peacesciencer\}} greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.
@@
 The argument defaults to \texttt{FALSE}, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths.
-In this application, \texttt{add\_ucdp\_acd(type\ =\ "intrastate",\ only\_wars\ =\ FALSE))} returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.\footnote{The function also returns some background information of interest to the researcher, including the maximum intensity observed and the IDs associated with all ongoing conflicts in the state that year.}
+In this application, \texttt{add\_ucdp\_acd(type\ =\ "intrastate",\ only\_wars\ =\ FALSE)} returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.\footnote{The function also returns some background information of interest to the researcher, including the maximum intensity observed and the IDs associated with all ongoing conflicts in the state that year.}
 
@@
 \texttt{add\_spells()} calculates peace spells between ongoing conflicts in the data generated by \texttt{add\_ucdp\_acd()}.
-\texttt{add\_democracy()} adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' (2016) extension of Pemstein et al's (2010) Unified Democracy Scores).
+\texttt{add\_democracy()} adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' (2016) extension of Pemstein et al's (2010) UDS data).
 \texttt{add\_creg\_fractionalization()} adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois.
@@
  & (\num{0.067}) & (\num{0.106})\\
-Extended UDS Democracy Score (Lagged) & \num{0.257} & \num{-0.085}\\
+Extended UDS (Lagged) & \num{0.257} & \num{-0.085}\\
  & (\num{0.181}) & (\num{0.270})\\
-Extended UDS Democracy Score\textasciicircum2 (Lagged) & \num{-0.726}* & \num{-0.761}*\\
+Extended UDS\textasciicircum2 (Lagged) & \num{-0.726}* & \num{-0.761}*\\
  & (\num{0.211}) & (\num{0.352})\\
@@
 For example, \texttt{add\_cow\_alliance()} returns information about various types of alliance pledges in the dyad-year---should one exist---but leaves it to the researcher to say whether they want to define the presence of an alliance to be just a defense pledge or any type of alliance pledge.
-\texttt{add\_contiguity()} returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity threshold.
+\texttt{add\_contiguity()} returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity relationship.
 The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g.~the contiguity data are not ordinal and should not be treated as such), how and where data issues arise (e.g.~how CoW state system data differ from Gleditsch-Ward data and how one is coerced into the other), and how researchers should consider optimally using its functionality (e.g.~\texttt{add\_ucdp\_acd()} probably should not lump all forms of conflict together (c.f.
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-andersetal2020bbgb}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-andersetal2020bbgb}{}}%
 Anders, Therese, Christopher J Fariss \& Jonathan N Markowitz (2020) Bread before guns or butter: Introducing surplus domestic product (SDP). \emph{International Studies Quarterly} 64(2): 392--405.
 
-\leavevmode\hypertarget{ref-arelbundock2021m}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-arelbundock2021m}{}}%
 Arel-Bundock, Vincent (2021b) \emph{Modelsummary: Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready} (\url{https://CRAN.R-project.org/package=modelsummary}).
 
-\leavevmode\hypertarget{ref-arelbundock2021w}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-arelbundock2021w}{}}%
 Arel-Bundock, Vincent (2021a) \emph{WDI: World Development Indicators and Other World Bank Data} (\url{https://CRAN.R-project.org/package=WDI}).
 
-\leavevmode\hypertarget{ref-barbierietal2009td}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-barbierietal2009td}{}}%
 Barbieri, Katherine, Omar MG Keshk \& Brian Pollins (2009) Trading data: Evaluating our assumptions and coding rules. \emph{Conflict Management and Peace Science} 26(5): 471--491.
 
-\leavevmode\hypertarget{ref-becketal1998tts}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-becketal1998tts}{}}%
 Beck, Nathaniel, Jonathan N Katz \& Richard Tucker (1998) Taking time seriously: Time-series-cross-section analysis with a binary dependent variable.
@@
 
-\leavevmode\hypertarget{ref-bennettetal2019n}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bennettetal2019n}{}}%
 Bennett, DScott, Paul Poast \& Allan C Stam (2019) NewGene: An introduction for users. \emph{Journal of Conflict Resolution} 63(6): 1579--1592.
 
-\leavevmode\hypertarget{ref-bennettstam2000e}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bennettstam2000e}{}}%
 Bennett, DScott \& Allan Stam (2000) EUGene: A conceptual manual. \emph{International Interactions} 26(2): 179--204.
 
-\leavevmode\hypertarget{ref-bowersvoors2016hiyr}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bowersvoors2016hiyr}{}}%
 Bowers, Jake \& Maarten Voors (2016) How to improve your relationship with your future self. \emph{Revista de Ciencia Politica} 36(3): 829--848.
 
-\leavevmode\hypertarget{ref-bremer1992ddc}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bremer1992ddc}{}}%
 Bremer, Stuart A (1992) Dangerous dyads: Conditions affecting the likelihood of interstate war, 1816-1965. \emph{Journal of Conflict Resolution} 36(2): 309--341.
 
-\leavevmode\hypertarget{ref-bdmlalman1992wr}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bdmlalman1992wr}{}}%
 Bueno de Mesquita, Bruce \& David Lalman (1992) \emph{War and Reason: Domestic and International Imperatives}. Yale University Press.
 
-\leavevmode\hypertarget{ref-cartersignorino2010bf}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-cartersignorino2010bf}{}}%
 Carter, David B \& Curtis S Signorino (2010) Back to the future: Modeling time dependence in binary data. \emph{Political Analysis} 18(3): 271--292.
 
-\leavevmode\hypertarget{ref-cartersmith2020fml}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-cartersmith2020fml}{}}%
 Carter, Jeff \& Charles E Smith (2020) A framework for measuring leaders' willingness to use force. \emph{American Political Science Review} 114(4): 1352--1358.
 
-\leavevmode\hypertarget{ref-coppedgeetal2020vdem}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-coppedgeetal2020vdem}{}}%
 Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I Lindberg, Jan Teorell, David Altman, Michael Bernhard, MSteven Fish, Adam Glynn, Allen Hicken, Anna Luhrmann, Kyle L Marquardt, Kelly McMann, Pamela Paxton, Daniel Pemstein, Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Agnes Cornell, Lisa Gastaldi, Haakon Gjerløw, Valeriya Mechkova, Johannes von Römer, Aksel Sundtröm, Eitan Tzelgov, Luca Uberti, Yi-ting Wang, Tore Wig \& Daniel Ziblatt (2020) V-dem codebook v10.
 
-\leavevmode\hypertarget{ref-cowstates2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-cowstates2016}{}}%
 Correlates of War (2011) State system membership list, v2016 (\url{http://correlatesofwar.org}).
 
-\leavevmode\hypertarget{ref-ellisetal2015lead}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-ellisetal2015lead}{}}%
 Ellis, Cali Mortenson, Michael C Horowitz \& Allan C Stam (2015) Introducing the {LEAD} data set. \emph{International Interactions} 41(4): 718--741.
 
-\leavevmode\hypertarget{ref-fearonlaitin2003eicw}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-fearonlaitin2003eicw}{}}%
 Fearon, James D \& David D Laitin (2003) Ethnicity, insurgency, and civil war. \emph{American Political Science Review} 97(1): 75--90.
 
-\leavevmode\hypertarget{ref-gibler2009ima}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gibler2009ima}{}}%
 Gibler, Douglas M (2009) \emph{International Military Alliances, 1648-2008}. Washington DC: CQ Press.
 
-\leavevmode\hypertarget{ref-giblermiller2014etts}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-giblermiller2014etts}{}}%
 Gibler, Douglas M \& Steven V Miller (2014) External territorial threat, state capacity, and civil war. \emph{Journal of Peace Research} 51(5): 634--646.
 
-\leavevmode\hypertarget{ref-giblermiller2021apm}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-giblermiller2021apm}{}}%
 Gibler, Douglas M \& Steven V Miller (Forthcoming) An appraisal of {P}roject {M}ars and the {D}ivided {A}rmies argument. \emph{International Studies Quarterly} (Forthcoming).
 
-\leavevmode\hypertarget{ref-gibleretal2016amid}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gibleretal2016amid}{}}%
 Gibler, Douglas M, Steven V Miller \& Erin K Little (2016) An analysis of the {M}ilitarized {I}nterstate {D}ispute {(MID)} dataset, 1816-2001. \emph{International Studies Quarterly} 60(4): 719--730.
 
-\leavevmode\hypertarget{ref-gledtischward1999rlis}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gledtischward1999rlis}{}}%
 Gleditsch, Kristian S \& Michael D Ward (1999) A revised list of independent states since the {C}ongress of {V}ienna. \emph{International Interactions} 25(4): 393--413.
 
-\leavevmode\hypertarget{ref-gleditschetal2002ac}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gleditschetal2002ac}{}}%
 Gleditsch, Nils Petter, Peter Wallensteen, Mikael Eriksson, Margareta Sollenberg \& Havard Strand (2002) Armed conflict 1946-2001: A new dataset. \emph{Journal of Peace Research} 39(5): 615--637.
 
-\leavevmode\hypertarget{ref-goemansetal2009ia}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-goemansetal2009ia}{}}%
 Goemans, Henk E, Kristian Skrede Gleditsch \& Giacomo Chiozza (2009) Introducing {A}rchigos: A dataset on political leaders. \emph{Journal of Peace Research} 46(2): 269--83.
 
-\leavevmode\hypertarget{ref-haege2011cc}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-haege2011cc}{}}%
 Häge, Frank (2011) Choice or circumstance? Adjusting measures of foreign policy similarity for chance agreement. \emph{Political Analysis} 19(3): 287--305.
 
-\leavevmode\hypertarget{ref-horowitzstam2014hpme}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-horowitzstam2014hpme}{}}%
 Horowitz, Michael C \& Allan C Stam (2014) How prior military experience influences the future militarized behavior of leaders. \emph{International Organization} 68(3): 527--559.
 
-\leavevmode\hypertarget{ref-leedsetal2002atop}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-leedsetal2002atop}{}}%
 Leeds, Bretty Ashley, Jeffrey M Ritter, Sara McLaughlin Mitchell \& Andrew G Long (2002) Alliance treaty obligations and provisions, 1815-1944. \emph{International Interactions} 28: 237--260.
 
-\leavevmode\hypertarget{ref-lemkereed2001rprd}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-lemkereed2001rprd}{}}%
 Lemke, Douglas \& William Reed (2001) The relevance of politically relevant dyads. \emph{Journal of Conflict Resolution} 45(1): 126--144.
 
-\leavevmode\hypertarget{ref-marquez2016qme}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-marquez2016qme}{}}%
 Marquez, Xavier (2016) A quick method for extending the {U}nified {D}emocracy {S}cores (\url{http://dx.doi.org/10.2139/ssrn.2753830}).
 
-\leavevmode\hypertarget{ref-marshalletal2017p}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-marshalletal2017p}{}}%
 Marshall, Monty G, Ted Robert Gurr \& Keith Jaggers (2017) Polity {IV} project: Political regime characteristics and transitions, 1800-2016.
 
-\leavevmode\hypertarget{ref-nunnpuga2012r}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-nunnpuga2012r}{}}%
 Nunn, Nathan \& Diego Puga (2012) Ruggedness: The blessing of bad geography in {A}frica. \emph{Review of Economics and Statistics} 94(1): 20--36.
 
-\leavevmode\hypertarget{ref-pemsteinetal2010dc}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-pemsteinetal2010dc}{}}%
 Pemstein, Daniel, Stephen A Meserve \& James Melton (2010) Democratic compromise: A latent variable analysis of ten measures of regime type. \emph{Political Analysis} 18(4): 426--449.
 
-\leavevmode\hypertarget{ref-poast2010mdd}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-poast2010mdd}{}}%
 Poast, Paul (2010) (Mis)using dyadic data to analyze multilateral events. \emph{Political Analysis} 18(4): 403--425.
 
-\leavevmode\hypertarget{ref-schvitz2021mis}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-schvitz2021mis}{}}%
 Schvitz, Guy, Seraina Rüegger, Luc Girardin, Lars-Erik Cederman, Nils Weidmann \& Kristian Skrede Gleditsch (2021) {Mapping The International System, 1886-2017: The {CS}hapes 2.0 Dataset}.
@@
 
-\leavevmode\hypertarget{ref-singer1987rcwd}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-singer1987rcwd}{}}%
 Singer, David J (1987) Reconstructing the {C}orrelates of {W}ar dataset on material capabilities of states, 1816-1985. \emph{International Interactions} 14(1): 115--32.
 
-\leavevmode\hypertarget{ref-stinnettetal2002cow}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-stinnettetal2002cow}{}}%
 Stinnett, Douglas M, Jaroslav Tir, Philip Schafer, Paul F Diehl \& Charles S Gochman (2002) The {C}orrelates of {W}ar {P}roject {D}irect {C}ontiguity {D}ata, version 3.
@@
 
-\leavevmode\hypertarget{ref-thompsondreyer2012hir}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-thompsondreyer2012hir}{}}%
 Thompson, William R \& David R Dreyer (2012) \emph{Handbook of International Rivalries, 1494-2010}. Washington DC: CQ Press.
 
-\leavevmode\hypertarget{ref-wickhametal2019wt}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-wickhametal2019wt}{}}%
 Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, Alex Hayes, Lionel Henry, Jim Hester, Max Kuhn, Thomas Lin Pedersen, Evan Miller, Stephan Milton Bache, Kirill Müller, Jeroen Ooms, David Robinson, Dana Paige Seidel, Vitalie Spinu, Kohske Takahashi, Davis Vaughan, Claus Wilke, Kara Woo \& Hiroaki Yutani (2019) Welcome to the {tidyverse}.

### 2022-03-22 manuscript/ms.Rmd
add, with professional easter egg

@@
 - name: Steven V. Miller
-  affiliation: Clemson University
+  affiliation: Stockholm University
 bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`" # This would be more general
@@
 
-`{peacesciencer}` is designed to address these problems.
+`{peacesciencer}` addresses these problems.
 Built around the free and open source R programming language, `{peacesciencer}` contains a suite of data and functions for creating data of interest to researchers.
@@
 After all, EUGene's main value---if not its primary impetus---was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g.
-dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit.
+dyad-year, state-year, dispute-year) along the way that users could amend as they saw fit.
 `{peacesciencer}` is primarily born from this question about how to create these data from scratch.
@@
         "add_fpsim()", "D, LD", "Add dyadic foreign policy similarity measures to a data frame",
-        "add_gml_mids()", "D, L, LD, S", "Add GML MID data to a data frame",
+        "add_gml_mids()", "D, L, LD, S", "Add Gibler-Miller-Little (GML) MID data to a data frame",
         "add_gwcode_to_cow()", "D, L, LD, S", "Match G-W state codes to CoW state codes",
-        "add_igos()", "D, S", "Add CoW IGO data to a data frame",
+        "add_igos()", "D, S", "Add CoW intergovernmental organizations (IGO) data to a data frame",
         "add_lead()", "L, LD", "Add leader experience/attributes (LEAD) to data",
@@
         "add_strategic_rivalries()", "D, S", "Add strategic rivalry information to a data frame",
-        "add_ucdp_acd()", "S", "Add UCDP Armed Conflict Data to state-year data frame",
+        "add_ucdp_acd()", "S", "Add Uppsala Conflict Data Program (UCDP) Armed Conflict Data to state-year data frame",
         "add_ucdp_onsets", "S", "Add UCDP onsets to state-year data",
@@
 This assumes at least some familiarity with the R programming language.
-Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.2.
+Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.3.
 `{peacesciencer}` is designed to be as user-friendly as possible.
@@
 
-[^poundsign]: R users will recognize that pound signs (`#`) allow for commenting code.
+[^poundsign]: R users will recognize that hash signs (`#`) allow for commenting code.
 Users new to R should understand its use here is primarily for exposition of what the code snippets in this manuscript are doing.
@@
 To start, assume a new user without much familiarity with the R programming language installed RStudio, `{tidyverse}`, and `{peacesciencer}` with the idea of using `{peacesciencer}` to help them start a new research project that seeks to explain civil conflict onset across state-years.
-Toward that end, they have identified their unit of analysis is state-year and can be created with the `create_stateyears()`.
+Toward that end, they have identified their unit of analysis is state-year and can be created with the `create_stateyears()` function.
 To get started, I encourage entering the following command in the console in RStudio.
@@
 A user who approaches `{peacesciencer}` with a project in mind will see they can better tailor this function to what they want.
-Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the UCDP armed conflict data.
+Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the Uppsala Conflict Data Program (UCDP) armed conflict data.
 They will also know they have no use for pre-1946 observations since civil conflict data typically have coverage only from 1946 forward.
@@
 Archigos often has leader entries prior to state system entry in a few cases).
-Thus, the user can standardize leader-year data to Gleditsch-Ward state system dates of CoW state system dates.
-Finally, the user can subset the leader-year data to a more narrow temporal domain that may interest them with the `subset_years` function.
+Thus, the user can standardize leader-year data to Gleditsch-Ward state system dates or CoW state system dates.
+Finally, the user can subset the leader-year data to a more narrow temporal domain with the `subset_years` function.
 If a user is interested in creating a data set of leader-years for an analysis of inter-state conflict initiation with the GML MID data, they can create the data that interests them with the following function.
@@
 
-Researchers using `{peacesciencer}` to create data for their research project should start with one of these "create" functions.
-Whether state-year, dyad-year, or a leader-level analysis, these functions will create the full universe of cases of interest to a researcher for the unit of analysis that serves as the core of their project.
+Researchers using `{peacesciencer}` to create data for their research project should start with one of these "create" functions, which will create the full universe of cases of interest to a researcher.
 
@@
 Researchers should see that `{peacesciencer}`'s functionality can scale up nicely from there.
-For example, the following would round out the kind of information necessary to replicate Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis by adding information about national material capabilities for both states in the dyad (`add_nmc()`), estimates of democracy for both states in the dyad (`add_democracy()`), information about alliance commitments in the dyad-year (`add_cow_alliance()`, by way of @gibler2009ima), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by @andersetal2020bbgb.
+For example, the following would round out the kind of information necessary to replicate Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis by adding information about national material capabilities (e.g.
+the composite index of national capabilities [CINC]) for both states in the dyad (`add_nmc()`), estimates of democracy for both states in the dyad (`add_democracy()`), information about alliance commitments in the dyad-year (`add_cow_alliance()`, by way of @gibler2009ima), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by @andersetal2020bbgb.
 Whereas `add_sdp_gdp()` is the last command in the pipe-based workflow, the `{peacesciencer}` call ends by assigning to an object called `Data`.
@@
 If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like `haven::write_dta(Data, "my-data.dta")` and import it into Stata when they are done.
-No matter, `{peacesciencer}` has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding `-9s` in these data to be `NA` in the capabilities data.
+No matter, `{peacesciencer}` has pre-processed, cleaned, recoded, and merged the desired data that have greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding `-9s` in these data to be `NA` in the capabilities data.
 In this particular application, it has already created the main data required for a replication of @bremer1992ddc.
-There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some "weak-link" indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the @marquez2016qme method for extending the Unified Democracy Scores data), and the GDP per capita in the dyad.
-Table \@ref(tab:tab-dd) is a formatted version of the results of a logistic regression model of conflict onset using these "dangerous dyads" indicators.
+There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land contiguity, the presence of a major power in the dyad, and some "weak-link" indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the @marquez2016qme method for extending the Unified Democracy Scores [UDS] data), and the GDP per capita in the dyad.
+Table \@ref(tab:tab-dd) is a formatted version of the results of a logistic regression model of conflict onset using these "dangerous dyads" indicators and temporal adjustment variables (*t*).
 Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but `{peacesciencer}` greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.
@@
 `type = c("intrastate", "II")`).
-`issue` is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict.
+`issue` is another optional argument for what issue types of conflicts the user wants.
 Options include "territory", "government", and "both".
@@
 The argument defaults to `FALSE`, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths.
-In this application, `add_ucdp_acd(type = "intrastate", only_wars = FALSE))` returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.[^adducdpacdalso]
+In this application, `add_ucdp_acd(type = "intrastate", only_wars = FALSE)` returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.[^adducdpacdalso]
 
@@
 `add_spells()` calculates peace spells between ongoing conflicts in the data generated by `add_ucdp_acd()`.
-`add_democracy()` adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' [-@marquez2016qme] extension of Pemstein et al's [-@pemsteinetal2010dc] Unified Democracy Scores).
+`add_democracy()` adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' [-@marquez2016qme] extension of Pemstein et al's [-@pemsteinetal2010dc] UDS data).
 `add_creg_fractionalization()` adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois.
@@
 If the researcher wants a new set of conflicts with a new set of peace years, it would be a simple matter of repeating the pipe-based workflow, but altering the argument in `add_ucdp_acd()` to be `only_wars = TRUE`.
-`{peacesciencer}` would then calculate the peace years for those (`add_peace_years()`).
+`{peacesciencer}` would then calculate the peace years for those (`add_spells()`).
 To avoid confusion with the overlapping column names, the researcher can use some `{tidyverse}` verbs to rename all those conflict variables to have a distinct prefix of `war_` (i.e.
@@
                           "l1_wbpopest" = "Population Size (Lagged)", 
-                          "l1_xm_qudsest" = "Extended UDS Democracy Score (Lagged)",
-                          "I(l1_xm_qudsest^2)" = "Extended UDS Democracy Score^2 (Lagged)", 
+                          "l1_xm_qudsest" = "Extended UDS (Lagged)",
+                          "I(l1_xm_qudsest^2)" = "Extended UDS^2 (Lagged)", 
                           "newlmtnest" = "% Mountainous Terrain (Logged)", 
@@
 `{peacesciencer}` also has support for newer levels of analysis in the peace science community, prominently leader-levels of analysis.
-There has been considerable emphasis in peace science research to emphasize that state leaders, not "states", make foreign policy decisions that may lead to war.
+There has been considerable emphasis in peace science research that state leaders, not "states", make foreign policy decisions that may lead to war.
 Understanding the attributes of leaders themselves are critical to the core research questions of the community [@ellisetal2015lead; @goemansetal2009ia; @horowitzstam2014hpme] and `{peacesciencer}` wants to help researchers toward that end.
@@
 estimated leader age, year in office, and leader gender).
-From there, they would add information about leader conflict behavior in the year with `add_gml_mids()` and calculate peace-years with the `add_spells()` function.[^oninitiation] `add_lead()` will add a battery of leader experience and attributes variables from data created by @ellisetal2015lead, including whether the leader had military experience, combat experience, was a rebel fighter, and many more.
-Finally, `add_nmc()` and `add_democracy()` will add state-year estimates of national capabilities and democracy for assisting in making state-to-state comparisons, even for leader-level analyses.
+From there, they would add information about leader conflict behavior in the year with `add_gml_mids()` and calculate peace-years with the `add_spells()` function.[^oninitiation] `add_lead()` will add a battery of leader experience and attributes variables from data created by @ellisetal2015lead, including whether the leader had military experience, combat experience, was a rebel fighter, and more.
+Finally, `add_nmc()` and `add_democracy()` will add state-year estimates of national capabilities and democracy for making state-to-state comparisons, even for leader-level analyses.
 Some light recoding of the data created in `{peacesciencer}` and some regression modeling reproduces the results presented in Table \@ref(tab:tab-ly), itself an approximation of the kind of leader-year analysis exemplified in @horowitzstam2014hpme.
@@
 EUGene's user manual was amply informative about what it was calculating and why it was doing what it was doing for some defaults.
-Likewise, `{peacesciencer}`'s documentation strives to be as informative as possible as to what it is doing and why it is doing what it is doing for some defaults.
+Likewise, `{peacesciencer}`'s documentation strives to be as informative as possible as to what it is doing and why it is doing what it is doing.
 A scholar who remembers EUGene well will ideally think of `{peacesciencer}` as the most faithful approximation of what that software did for the community at the time.
@@
 This is not to say `{peacesciencer}` does everything, but `{peacesciencer}` can only evolve and expand on what it already does well.
-This package can only evolve to meet new analytical demands for the peace science community.
 Users are free to request new features as "issues" on the project's Github.
@@
 For example, `add_cow_alliance()` returns information about various types of alliance pledges in the dyad-year---should one exist---but leaves it to the researcher to say whether they want to define the presence of an alliance to be just a defense pledge or any type of alliance pledge.
-`add_contiguity()` returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity threshold.
+`add_contiguity()` returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity relationship.
 The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g.

### 2023-03-22 README.Rmd
update README

@@
 `peacesciencer` is an R package including various functions and data sets to allow easier analyses in the field of quantitative peace science.
-The goal is to provide an R package that reasonably approximates what made [EUGene](https://journals.sagepub.com/doi/abs/10.1177/0738894211413055) so attractive to scholars working in the field of quantitative peace science in the early 2000s.
+The goal is to provide an R package that reasonably approximates what made [EUGene](https://doi.org/10.1177/0738894211413055) so attractive to scholars working in the field of quantitative peace science in the early 2000s.
 EUGene shined because it encouraged replications of conflict models while having the user also generate data from scratch.

### 2023-03-24 README.Rmd
fix

@@
 `peacesciencer` is an R package including various functions and data sets to allow easier analyses in the field of quantitative peace science.
-The goal is to provide an R package that reasonably approximates what made [EUGene](https://doi.org/10.1177/0738894211413055) so attractive to scholars working in the field of quantitative peace science in the early 2000s.
+The goal is to provide an R package that reasonably approximates what made EUGene so attractive to scholars working in the field of quantitative peace science in the early 2000s.
 EUGene shined because it encouraged replications of conflict models while having the user also generate data from scratch.
@@
 not requiring Windows for installation).
+
+<!-- [EUGene](https://journals.sagepub.com/doi/10.1177/0738894211413055) -->
 

### 2023-03-24 tests/testthat/test_nodamnduplicates.R

@@ -1,49 +1,0 @@
-context("No duplicates! What are duplicates doing in your peace science data when I told you NO. DUPLICATES. EVER! WHY!")
-
-library(dplyr)
-library(magrittr)
-
-test_that("Dyad-year additions do not create duplicates", {
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_contiguity()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_gwcode_to_cow()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_capital_distance()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_majors()))
-#   expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% download_extdata() %>% add_cow_trade()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_democracy()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_gwcode_to_cow()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_gml_mids()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_mids()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_mids() %>% add_gml_mids() %>% add_peace_years()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_alliance()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_nmc()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_archigos()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_minimum_distance()))
-  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_strategic_rivalries()))
-  expect_equal(nrow(create_dyadyears(system = "gw")), nrow(create_dyadyears(system = "gw") %>% add_minimum_distance()))
-  expect_equal(nrow(create_dyadyears(system = "gw")), nrow(create_dyadyears(system = "gw") %>% add_ccode_to_gw()))
-  expect_equal(nrow(create_dyadyears(system = "gw")), nrow(create_dyadyears(system = "gw") %>% add_capital_distance()))
[…]

### 2025-06-23 R/add_democracy.R
add

@@ -1,6 +1,8 @@
 #' Add democracy information to a data frame
 #'
-#' @description \code{add_democracy()} allows you to add estimates of democracy to your data.
+#' @description
+#'
+#' \code{add_democracy()} allows you to add estimates of democracy to your data.
 #'
 #'
 #' @return
@@ -19,9 +21,28 @@
 #'
 #' @details
 #'
+#' As of version 1.2, this function leans on the information made available in
+#' the \pkg{isard} package. This is a spin-off package I maintain for data that
+#' require periodic updates for the functionality in this package. As of writing,
+#' \pkg{peacesciencer} only requires that you have the \pkg{isard} package
+#' installed. It does not require you to have any particular version of the
+#' package installed. Thus, what exactly this function returns may depend on the
+#' particular version of \pkg{isard} you have installed. This will assuredly
+#' concern the right-bound of the temporal domain of data you get.
+#'
+#' You can read more about the data in the documentation for \pkg{isard}.
+#'
+#' - [https://svmiller.com/isard/reference/cw_democracy.html](https://svmiller.com/isard/reference/cw_democracy.html)
+#' - [https://svmiller.com/isard/reference/gw_democracy.html](https://svmiller.com/isard/reference/gw_democracy.html)
+#'
 #' Be mindful that the data are fundamentally state-year and that extensions to
 #' leader-level data should be understood as approximations for leaders in a
 #' given state-year.
+#'
+#" The `keep` argument must include one or more of the democracy estimates
+#' included in the `cw_democracy` or `gw_democracy` data in the \pkg{isard}
+#' data. Otherwise, it will return an error that it cannot subset columns that
+#' do not exist.
 #'
 #' A vignette on the package's website talks about how these data are here
 #' primarily to encourage you to maximize the number of observations in the
@@ -41,28 +62,15 @@
 #' @author Steven V. Miller
 #'
 #' @param data a data frame with appropriate \pkg{peacesciencer} attributes
[…]

### 2025-06-23 R/data-cow.R

@@ -61,3 +61,184 @@ NULL
 #' Management and Peace Science 19 (2):58-66.
 #'
 "cow_contdir"
+
+
+#' Correlates of War National Trade Data Set (v. 4.0)
+#'
+#' These are state-year-level data for national trade from the Correlates of War
+#' project.
+#'
+#'
+#' @format A data frame with 14410 observations on the following four variables.
+#' \describe{
+#' \item{\code{ccode}}{the Correlates of War state system code}
+#' \item{\code{year}}{the year}
+#' \item{\code{imports}}{total imports of the state in current million USD}
+#' \item{\code{exports}}{total exports of the state in current million USD}
+#' }
+#'
+#' @details
+#'
+#' The \code{data-raw} directory on the project's Github shows how the data were
+#' processed.
+#'
+#' @references
+#'
+#' Barbieri, Katherine and Omar M.G. Keshk. 2016. Correlates of War Project
+#' Trade Data Set Codebook, Version 4.0. Online: \url{https://correlatesofwar.org}
+#'
+#' Barbieri, Katherine, Omar M.G. Keshk, and Brian Pollins. 2009. "TRADING DATA:
+#' Evaluating Our Assumptions and Coding Rules."
+#' *Conflict Management and Peace Science*, 26(5): 471-491.
+#'
+"cow_trade_sy"
+
+#' Directed Dyadic Dispute-Year Data (CoW-MID, v. 5.0)
+#'
+#' These are directed dyadic dispute year data derived from the Correlates of War
+#' (CoW) Militarized Interstate Dispute (MID) project. Data are from version 5.0.
+#'
+#'
+#' @format A data frame with 11390 observations on the following 18 variables.
+#' \describe{
+#' \item{\code{dispnum}}{a numeric vector for the CoW-MID dispute number}
+#' \item{\code{ccode1}}{a numeric vector for the focal state in the dyad}
+#' \item{\code{ccode2}}{a numeric vector for the target state in the dyad}
+#' \item{\code{year}}{a numeric vector for the dispute-year}
+#' \item{\code{dispongoing}}{a numeric vector for whether there was a dispute ongoing in that year}
+#' \item{\code{disponset}}{a numeric vector for whether it was the onset of a new dispute (or new participant-entry into a recurring dispute)}
+#' \item{\code{sidea1}}{is \code{ccode1} on side A of the dispute?}
+#' \item{\code{sidea2}}{is \code{ccode2} on side A of the dispute?}
[…]

### 2025-06-23 R/data.R

@@ -3,47 +3,7 @@ NULL
 
 
 
-#' Democracy data for all Correlates of War states
-#'
-#' These are democracy data for all Correlates of War state system members.
-#'
-#' @format A data frame with 16536 observations on the following 5 variables.
-#' \describe{
-#' \item{\code{ccode}}{the Correlates of War system code}
-#' \item{\code{year}}{a numeric vector for the year}
-#' \item{\code{v2x_polyarchy}}{the Varieties of Democracy "polyarchy" estimate}
-#' \item{\code{polity2}}{the the \code{polity2} score from the Polity project}
-#' \item{\code{xm_qudsest}}{an extension of the Unified Democracy Scores (UDS) estimates, made possibly by the \code{QuickUDS} package from Xavier Marquez.}
-#' }
-#'
-#' @details Missing data connote data that are unavailable for various reasons. Either there is no democracy data to code or, in the case of the Polity project, the state
-#' system member is outright not evaluated for the variable.
-#'
-#' The Polity data are from 2017. The Varieties of Democracy data are version 10. Xavier Marquez' \code{QuickUDS} estimates (i.e. extensions of Pemstein et al. (2010)) come from a package Marquez makes available on his Github (\url{https://github.com/xmarquez/QuickUDS}).
-#'
-#' @references
-#'
-#' Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg,
-#' Jan Teorell, David Altman, Michael Bernhard, M. Steven Fish, Adam Glynn,
-#' Allen Hicken, Anna Luhrmann, Kyle L. Marquardt, Kelly McMann, Pamela
-#' Paxton, Daniel Pemstein, Brigitte Seim, Rachel Sigman, Svend-Erik
-#' Skaaning, Jeffrey Staton, Agnes Cornell, Lisa Gastaldi, Haakon Gjerlow,
-#' Valeriya Mechkova, Johannes von Romer, Aksel Sundtrom, Eitan Tzelgov,
-#' Luca Uberti, Yi-ting Wang, Tore Wig, and Daniel Ziblatt. 2020.
-#' "V-Dem Codebook v10" Varieties of Democracy (V-Dem) Project.
-#'
-#' Marshall, Monty G., Ted Robert Gurr, and Keith Jaggers. 2017.
-#' "Polity IV Project: Political Regime Characteristics and Transitions,
[…]

### 2025-06-23 R/helpymchelperson.R

@@ -9,3 +9,15 @@
   return(x)
 
 }
+
+#' Extract months from date objects
+#' @param date a date
+#' @keywords internal
+#' @export
+#' @noRd
+
+.pshf_month <- function(date) {
+  x <- as.POSIXlt(date)$mon + 1 # Importantly, Jan = 0 and Dec = 11
+  return(x)
+
+}

### 2025-06-30 R/create_dyadyears.R
dsfsd

@@ -35,7 +35,7 @@
 #' Data With Just a Few Lines in R''
 #' \url{https://svmiller.com/blog/2019/01/create-country-year-dyad-year-from-country-data/}
 #'
-#' Miller, Steven V. 2025. \pgk{isard}: Overflow Data for Quantitative Peace
+#' Miller, Steven V. 2025. \pkg{isard}: Overflow Data for Quantitative Peace
 #' Science Research. \url{https://CRAN.R-project.org/package=isard}
 #'
 #' @param system a character specifying whether the user wants Correlates of War

### 2025-07-10 R/create_stateyears.R
batch updates

@@ -99,6 +99,12 @@ create_stateyears <- function(system = "cow", mry = TRUE, subset_years) {
 
     data <- data[order(data$ccode, data$year), ]
 
+    if (!missing(subset_years)) {
+      data <- subset(data, data$year %in% subset_years)
+    } else {
+      data <- data
+    }
+
     attr(data, "ps_data_type") = "state_year"
     attr(data, "ps_system") = "cow"
 
@@ -153,16 +159,15 @@ create_stateyears <- function(system = "cow", mry = TRUE, subset_years) {
 
     data <- data[order(data$gwcode, data$year), ]
 
+    if (!missing(subset_years)) {
+      data <- subset(data, data$year %in% subset_years)
+    } else {
+      data <- data
+    }
+
     attr(data, "ps_data_type") = "state_year"
     attr(data, "ps_system") = "gw"
 
-  }
-
-
-  if (!missing(subset_years)) {
-    data <- subset(data, data$year %in% subset_years)
-  } else {
-    data <- data
   }
 
 

### 2025-07-10 R/data.R

@@ -298,12 +298,13 @@ NULL
 
 "ucdp_onsets"
 
-#' UCDP Armed Conflict Data (ACD) (v. 20.1)
+#' UCDP Armed Conflict Data (ACD) (v. 25.1)
 #'
-#' These are (kind of) dyadic, but mostly state-level data, used internally for doing stuff with the UCDP armed conflict data
+#' These are (kind of) dyadic, but mostly state-level data, used internally for
+#' doing stuff with the UCDP armed conflict data
 #'
 #'
-#' @format A data frame with 4164 observations on the following 15 variables.
+#' @format A data frame with 5652 observations on the following 15 variables.
 #' \describe{
 #' \item{\code{conflict_id}}{a conflict identifier, not to be confused with an episode identifier (which I don't think UCDP offers)}
 #' \item{\code{year}}{a numeric vector for the year}
@@ -321,16 +322,22 @@ NULL
 #' \item{\code{ep_end}}{a dummy variable for whether the conflict episode ended in the calendar year of observation}
 #' \item{\code{ep_end_date}}{the episode end date, if applicable}
 #' }
-#' @details The \code{data-raw} directory on the project's Github will show how I processed the multiple strings for when there are multiple states on a given side.
+#'
+#' @details
+#'
+#' The \code{data-raw} directory on the project's Github will show how I processed
+#' the multiple strings for when there are multiple states on a given side.
 #'
 #'
 #' @references
 #'
-#' Gleditsch, Nils Petter; Peter Wallensteen, Mikael Eriksson, Margareta Sollenberg & Havard Strand (2002)
-#' Armed Conflict 1946–2001: A New Dataset. \emph{Journal of Peace Research} 39(5): 615–637.
+#' Gleditsch, Nils Petter; Peter Wallensteen, Mikael Eriksson, Margareta
+#' Sollenberg, and Havard Strand. 2002. "Armed Conflict 1946–2001: A New
+#' Dataset." *Journal of Peace Research* 39(5): 615–637.
 #'
-#' Pettersson, Therese; Stina Hogbladh & Magnus Oberg (2019). Organized violence, 1989-2018 and peace
-#' agreements. Journal of Peace Research 56(4): 589-603.
+#' Davies, Shawn, Therése PEttersson, Margareta Sollenberg, and Magnus Öberg.
+#' 2025. "Organized violence 1989–2024, and the challenges of identifying
+#' civilian victims." *Journal of Peace Research* 62(4): 1223--1240.
 
 "ucdp_acd"
 

### 2025-07-10 README.Rmd
update

@@
 
-Here's what this will look like in operation.
-Assume you want to create some data for something analogous to a "dangerous dyads" design for all non-directed dyad-years.
-Here's how you'd do it in `{peacesciencer}`, which is going to be lifted from the source R scripts for the user's guide.
-The first part of this code chunk will lean on core `{peacesciencer}` functionality whereas the other stuff is some post-processing and, as a bonus, some modeling.
-
-```{r, cache=F, message=F, eval=T}
-
-# library(tidyverse) # load this first for most/all things
-# library(peacesciencer) # the package of interest
-# library(stevemisc) # a dependency, but also used for standardizing variables for better interpretation
-library(tictoc)
-
-tic()
-create_dyadyears(directed = FALSE, mry = FALSE) %>%
-  filter_prd() %>%
-  add_gml_mids(keep = NULL) %>%
-  add_peace_years() %>%
-  add_nmc() %>%
-  add_democracy() %>%
-  add_cow_alliance() %>%
-  add_sdp_gdp() -> Data
-
-Data %>%
-  mutate(landcontig = ifelse(conttype == 1, 1, 0)) %>%
-  mutate(cowmajdyad = ifelse(cowmaj1 == 1 | cowmaj2 == 1, 1, 0)) %>%
-  # Create estimate of militarization as milper/tpop
-  # Then make a weak-link
-  mutate(milit1 = milper1/tpop1,
-         milit2 = milper2/tpop2,
-         minmilit = ifelse(milit1 > milit2,
-                           milit2, milit1)) %>%
-  # create CINC proportion (lower over higher)
-  mutate(cincprop = ifelse(cinc1 > cinc2,
-                           cinc2/cinc1, cinc1/cinc2)) %>%
-  # create weak-link specification using Quick UDS data
-  mutate(mindemest = ifelse(xm_qudsest1 > xm_qudsest2,
-                            xm_qudsest2, xm_qudsest1)) %>%
-  # Create "weak-link" measure of jointly advanced economies
-  mutate(minwbgdppc = ifelse(wbgdppc2011est1 > wbgdppc2011est2,
-                             wbgdppc2011est2, wbgdppc2011est1)) -> Data
-
-# r2sd() is in {stevemisc}, a {peacesciencer} dependency.
-# This is just for a more readable regression output.
-Data %>%
-  mutate_at(vars("cincprop", "mindemest", "minwbgdppc", "minmilit"),
-            ~r2sd(.)) -> Data
-
-broom::tidy(modDD <- glm(gmlmidonset ~ landcontig + cincprop + cowmajdyad + cow_defense +
-               mindemest + minwbgdppc + minmilit +
-               gmlmidspell + I(gmlmidspell^2) + I(gmlmidspell^3), data= Data,
-             family=binomial(link="logit")))
-toc()
-
-```
-
-Here is how you might do a standard civil conflict analysis using Gleditsch-Ward states and UCDP conflict data.
-
-```{r, cache=F, message=F, eval=T}
-
-tic()
-create_stateyears(system = 'gw') %>%
-  filter(year %in% c(1946:2019)) %>%
-  add_ucdp_acd(type=c("intrastate"), only_wars = FALSE) %>%
-  add_peace_years() %>%
-  add_democracy() %>%
-  add_creg_fractionalization() %>%
-  add_sdp_gdp() %>%
-  add_rugged_terrain() -> Data
-
-create_stateyears(system = 'gw') %>%
-  filter(year %in% c(1946:2019)) %>%
-  add_ucdp_acd(type=c("intrastate"), only_wars = TRUE) %>%
-  add_peace_years() %>%
-  rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .)) %>%
-  left_join(Data, .) -> Data
-
-Data %>%
-  arrange(gwcode, year) %>%
-  group_by(gwcode) %>%
-  mutate_at(vars("xm_qudsest", "wbgdppc2011est",
-                 "wbpopest"), list(l1 = ~lag(., 1))) %>%
-  rename_at(vars(contains("_l1")),
-            ~paste("l1", gsub("_l1", "", .), sep = "_") ) -> Data
-
-modCW <- list()
-broom::tidy(modCW$"All UCDP Conflicts" <- glm(ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
-                    l1_xm_qudsest + I(l1_xm_qudsest^2) +
-                    newlmtnest + ethfrac + relfrac +
-                    ucdpspell + I(ucdpspell^2) + I(ucdpspell^3), data=subset(Data),
-                  family = binomial(link="logit")))
-
-broom::tidy(modCW$"Wars Only"  <- glm(war_ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
-                    l1_xm_qudsest + I(l1_xm_qudsest^2) +
-                    newlmtnest + ethfrac + relfrac +
-                    war_ucdpspell + I(war_ucdpspell^2) + I(war_ucdpspell^3), data=subset(Data),
-                  family = binomial(link="logit")))
-
-toc()
-```
-
 # Citing What You Do in `{peacesciencer}`

### 2025-07-16 R/data.R
getting there

@@ -22,46 +22,6 @@ NULL
 #'
 "cow_ddy"
 
-
-
-
-
-#' Correlates of War State System Membership Data (1816-2016)
-#'
-#' These are the Correlates of War state system membership data.
-#'
-#' @format A data frame with 243 observations on the following 10 variables.
-#' \describe{
-#' \item{\code{stateabb}}{a character vector for the state abbreviation}
-#' \item{\code{ccode}}{a numeric vector for the Correlates of War country code}
-#' \item{\code{statenme}}{a character vector for the state name}
-#' \item{\code{styear}}{the start year in the system}
-#' \item{\code{stmonth}}{the start month in the system}
-#' \item{\code{stday}}{the start day in the system}
-#' \item{\code{endyear}}{the end year in the system}
-#' \item{\code{endmonth}}{the end month in the system}
-#' \item{\code{endday}}{the end day in the system}
-#' \item{\code{version}}{a version identifier}
-#' }
-#'
-#' @details
-#'
-#' Data are provided "as-is" with no additional re-cleaning before inclusion
-#' into this data set.
-#'
-#' The functions that previously used these data no longer use these data.
-#' They instead use a copy of the data in the \pkg{isard} package I also
-#' maintain.
-#'
-#' @references Correlates of War Project. 2017. "State System Membership
-#' List, v2016." Online,
-#' \url{https://correlatesofwar.org/data-sets/state-system-membership/}
-#'
-"cow_states"
-
-
-
-
 #' A directed dyad-year data frame of Gleditsch-Ward state system members
 #'
 #' This is a complete directed dyad-year data frame of Gleditsch-Ward
@@ -120,35 +80,6 @@ NULL
 #' *International Interactions* 25(4): 393--413.
 #'
 "gw_states"
-
-
-#' Zeev Maoz' Regional/Global Power Data
-#'
-#' These are Zeev Maoz' data for what states are regional or global powers at a
-#' given point time. They are extensions of the Correlates of War major power
-#' data, which only codes "major" power without consideration of regional or
-#' global distinctions. Think of Austria-Hungary as intuitive of the issue here.
-#' Austria-Hungary is a major power in the Correlates of War data, but there is
-#' good reason to treat Austria-Hungary as a major power only within Europe.
-#' That is what Zeev Maoz tries to do here.
-#'
-#' @format A data frame with 20 observations on the following 5 variables.
-#' \describe{
[…]
